### 2010-12-17 Methods.tex
Formed the preamble section and document layout

\section{Computational Methods}

\subsection{Molecular Dynamics Simulations}

\subsection{Steered Molecular Dynamics Trajectories}

### 2010-12-17 Background.tex
Fleshing out with new sections

\section{Background}

\subsection{Analysis of Orientation}

\subsubsection{Water Orientation}
\subsubsection{\suldiox Orientation}

\subsection{Bonding Analysis}
\subsubsection{Nearest Neighbor}
\subsubsection{Hydrate Complex Formation}

### 2010-12-17 Article.tex
new packages and formatting

\documentclass{article}

\usepackage {amsmath}
\usepackage{setspace}

% some formatting tags
\oddsidemargin  0.0in
\evensidemargin 0.0in
\textwidth      6.5in

\title{Sulfur Dioxide at Water's Surface}
\author{Eric S. Shamay \and Kevin E. Johnson \and Geraldine L. Richmond}

\begin{document}
\newcommand{\suldiox}{SO$_2$}
\newcommand{\ang}{\,$\textrm{\AA}$}
\newcommand{\angs}{\ang}
\newcommand{\wat}{H$_2$O}

\maketitle

\onehalfspacing

\begin{abstract}
	...Abstract goes here...
\end{abstract}

\include{Introduction}
\include{Background}
\include{Methods}
\include{SteeredMDResults}
\include{EquilibratedResults}

\end{document}

### 2010-12-17 clean.sh
new cleaning script clean.sh

@@ -1,0 +1,3 @@
+rm *.aux
+rm *.log
+rm *.pdf

### 2010-12-17 Background.tex
New image for the spherical angle definition, fleshed out the section describing the angles theta and phi for orientation analysis. Added bibliography for citations

@@
 
+	Analyses of the simulated systems were performed to characterize the orientation of \suldiox in various environments above, within, and below the surface water region.
+The orientation of the surface molecules as \suldiox adsorbs and interacts with water is a property of any hydrate complex formation that takes places.
+Better understanding of the molecular orientational distributions will help elucidate the chemistry occurring during the adsorption process.
+
+	The two molecules studied, \wat~and \suldiox, and similarly shaped with a C2$_v$ axis along their bisectors, and a plane is defined by the three atoms in the molecules.
+The orientational analyses presented herein focus on two angles used to define the molecular orientation in space.
+In each analysis a reference axis is used to define molecular angles, and is the z-axis unless otherwise specified.
+A ``tilt'' angle, $\theta$, defines the angle formed between a vector (generally the bisector vector pointing from the central atom in the direction of the other two atoms) and the positive reference axis.
+Thus the value of $\theta$ falls within a range of $[0,\pi]$.
+A second angle, $\phi$, defines the molecular ``twist'' around the bisector.
+$\phi$ is formed between the projection of a vector onto the x-y plane and the x-axis.
+$\phi$ is thus defined relative to the x-axis, and its values are in the interval $[-\pi,\pi]$.
+Figure \ref{fig:spherical-angle} shows the angles as related to the \suldiox~molecule's bisector and using the z-axis as the reference.
+
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[scale=1.0]{images/sphericalanglegraphic.png}
+		\caption{The visual representation of the angles $\theta$ and $\phi$ used to define molecular orientation of \suldiox~and \wat~relative to the reference axis.
+$\theta$ is the value of the ``tilt'' of the molecular bisector, and $\phi$ is the ``twist'' of the vector about the reference axis, starting at the x-axis.}
+		\label{fig:spherical-angle}
+	\end{center}
+\end{figure}
+
 \subsubsection{Water Orientation}
-\subsubsection{\suldiox Orientation}
+The orientation of water molecules at the 
+\subsubsection{\suldiox~Orientation}
 

### 2010-12-17 Methods.tex
Fleshing out with new sections
New image for the spherical angle definition, fleshed out the section describing the angles theta and phi for orientation analysis. Added bibliography for citations
Explaining the MD methods for SMD and pre-quilibrated systems

@@
 
-\subsection{Molecular Dynamics Simulations}
+Molecular dynamics simulations were performed using the Amber 11 software suite.\cite{Case2010} In this study we incorporate polarizable models for the \wat~and \suldiox~molecules.
+Fully polarizable models have been used previously in studies on interfacial systems because they are known to more accurately reproduce interfacial structure and free energy profiles.\cite{Wick2007,Rivera2006,Dang1998} The \wat~model used is the POL3 model,\cite{Caldwell1995} and for \suldiox we used the model of Dang, et al.
+that places a single polarizable center on the sulfur atom.\cite{Baer2010}
 
-\subsection{Steered Molecular Dynamics Trajectories}
+All simulations began with an equilibrated box of 900 \wat~molecules, with sides of length 30\angs.
+The simulations all employed periodic boundaries to create an ``infinite-slab'' geometry for the water phase.
+One set of simulations involved adding a single \suldiox~to the center of the water box and then allowing the system to equilibrate over 1 ns of simulation.
+The ensuing simulations are herein referred to as the ``surface equilibrated'' simulations.
+
+A second set of simulations involved preparing a water slab as explained above.
+A single \suldiox~was then introduced 20\angs above the slab, with the sulfur atom tethered to its initial position.
+The system was evolved for 1 ns, taking coordinate snapshots every 20 ps to create 50 starting points for further simulations.
+Steered molecular dynamics (SMD) were then performed to guide the \suldiox~down towards a tethered water 15\angs under the water slab surface by applying a small steering force.\cite{Isralewitz2001} The \suldiox~thus passed through the continuum of environments from gas phase to hydrate complex interaction, water surface adsorption, and absorption into the bulk of the neat-\wat~slab.
+
+The third set of simulations were performed similarly to the neat-\wat~SMD simulations, but the water slab and gas phase above it were first loaded with \suldiox.
+The water slab was loaded with 22 \suldiox molecules, which represent a saturation level coinciding with the Henry's law constant for \suldiox~in \wat.
+The gas phase was populated with 50 \suldiox~molecules representing approx.
+50 atm of \suldiox~pressure.
+
+\subsection{Steered Molecular Dynamics}

### 2010-12-20 Methods.tex
Adding in the starting configuration images

@@
 
-All simulations began with an equilibrated box of 900 \wat~molecules, with sides of length 30\angs.
+All simulations began with an equilibrated cube of 900 \wat~molecules, with sides of length 30\angs.
 The simulations all employed periodic boundaries to create an ``infinite-slab'' geometry for the water phase.
 One set of simulations involved adding a single \suldiox~to the center of the water box and then allowing the system to equilibrate over 1 ns of simulation.
-The ensuing simulations are herein referred to as the ``surface equilibrated'' simulations.
+The system was then evolved for a further 2.5 ns with atomic coordinates written every 100 fs.
+The simulation is herein referred to as the ``surface equilibrated'' simulations.
 
-A second set of simulations involved preparing a water slab as explained above.
+\subsection{Steered Molecular Dynamics}
+
+A second set of simulations involved preparing an equilibrated water slab as in the surface equilibrated method.
 A single \suldiox~was then introduced 20\angs above the slab, with the sulfur atom tethered to its initial position.
@@
 
-The third set of simulations were performed similarly to the neat-\wat~SMD simulations, but the water slab and gas phase above it were first loaded with \suldiox.
-The water slab was loaded with 22 \suldiox molecules, which represent a saturation level coinciding with the Henry's law constant for \suldiox~in \wat.
+The third set of simulations were performed similarly to the neat-\wat~and single \suldiox~SMD simulations, but the water slab and gas phase above it were first saturated with \suldiox.
+The water slab was loaded internally with 22 \suldiox molecules, which represent a saturation level coinciding with the Henry's law constant for \suldiox~in \wat.
 The gas phase was populated with 50 \suldiox~molecules representing approx.
 50 atm of \suldiox~pressure.
+As in the neat-\wat~SMD simulations, a single \suldiox~was introduced 20\angs above the saturated water slab surface and then steered into the bulk of the water.
+Figure \ref{fig:starting-configurations} illustrates two sample starting configurations for the SMD simulations, showing both the neat-\wat~slab and the saturated-\wat~slab configurations before steering the \suldiox~into the water bulk.
 
-\subsection{Steered Molecular Dynamics}
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[scale=1.0]{images/startingconfigurations.png}
+		\caption{Sample starting configurations for the two types of SMD simulations.
+The neat-\wat~slab simulation introduces a single \suldiox~molecule that is then guided into the surface of the water and further into the bulk (left).
+The saturated-\wat~slab simulation begins with different configurations of a water system that has been loaded with \suldiox~to saturate the water phase, and also with a high pressure of \suldiox~gas (right).
+The single \suldiox~(shown at the top) is then steered through the surface region saturated with \suldiox~molecules, and into the water bulk.}
+		\label{fig:starting-configurations}
+	\end{center}
+\end{figure}

### 2010-12-20 Methods.tex
new image size for startingconfigurations, and reworded the text for the MD methods.

@@
 All simulations began with an equilibrated cube of 900 \wat~molecules, with sides of length 30\angs.
-The simulations all employed periodic boundaries to create an ``infinite-slab'' geometry for the water phase.
-One set of simulations involved adding a single \suldiox~to the center of the water box and then allowing the system to equilibrate over 1 ns of simulation.
-The system was then evolved for a further 2.5 ns with atomic coordinates written every 100 fs.
-The simulation is herein referred to as the ``surface equilibrated'' simulations.
+The long axis of each simulation cell (the axis normal to the water surface) was set to 120\angs.
+The simulations all employed periodic boundaries to create an ``infinite-slab'' geometry.
+After equilibrating the neat-\wat~slabs for 1 ns, two types of systems were created by introducing \suldiox.
 
-\subsection{Steered Molecular Dynamics}
+One system involved the addition of a single \suldiox~molecule either within the bulk of the water slab, or above the slab surface.
+The single-\suldiox~system was then evolved for 2 ns to produce an equilibrated starting configuration.
+The second system introduced 22 \suldiox~molecules into the water slab in order to saturate it to a level coinciding with the Henry's law constant for \suldiox~in water.
+Additionally, 50 \suldiox~molecules were introduced into the gas phase outside of the water slab to simulate an added 50 atm of \suldiox~gas pressure.
+The saturated system was then evolved for 2 ns to produce a starting configuration for further simulations.
 
-A second set of simulations involved preparing an equilibrated water slab as in the surface equilibrated method.
-A single \suldiox~was then introduced 20\angs above the slab, with the sulfur atom tethered to its initial position.
+\subsection{Equilibrium Simulations}
+
+Equilibrium simulations involved adding \suldiox~to a water slab and equilibrating as outlined above.
+One system had a single \suldiox~added to the center of the water box, representing a concentration of 0.06 M.
+The complementary high-concentration system with 22 \suldiox~molecule in the bulk represents a system of concentration 1.35 M, with an additional 50 \suldiox~in the gas phase to model 50 atm of pressure above the water surface.
+Both the low and high concentration systems were then evolved for a further 2.5 ns with atomic coordinates written every 100 fs.
+
+\subsection{Steered Molecular Dynamics Simulations}
+
+% need to add in the specifics about the SMD routine, force constants, calculations, etc.
+A second set of simulations involved preparing an equilibrated water slab as in the surface equilibrated method above.
+However, for both the low and high concentration systems, a single \suldiox~was introduced 20\angs above the slab, with the sulfur atom tethered to its initial position.
 The system was evolved for 1 ns, taking coordinate snapshots every 20 ps to create 50 starting points for further simulations.
-Steered molecular dynamics (SMD) were then performed to guide the \suldiox~down towards a tethered water 15\angs under the water slab surface by applying a small steering force.\cite{Isralewitz2001} The \suldiox~thus passed through the continuum of environments from gas phase to hydrate complex interaction, water surface adsorption, and absorption into the bulk of the neat-\wat~slab.
+Steered molecular dynamics (SMD) were then performed on the 50 system configurations (in both the low and high concentrations) to guide the \suldiox~down towards a tethered water 15\angs under the water slab surface by applying a small steering force.\cite{Isralewitz2001} The \suldiox~thus passed through the continuum of environments from gas phase to hydrate complex formation or interaction with surface-bound \suldiox, water surface adsorption, and absorption into the bulk of the \wat~slab.
 
-The third set of simulations were performed similarly to the neat-\wat~and single \suldiox~SMD simulations, but the water slab and gas phase above it were first saturated with \suldiox.
-The water slab was loaded internally with 22 \suldiox molecules, which represent a saturation level coinciding with the Henry's law constant for \suldiox~in \wat.
-The gas phase was populated with 50 \suldiox~molecules representing approx.
-50 atm of \suldiox~pressure.
-As in the neat-\wat~SMD simulations, a single \suldiox~was introduced 20\angs above the saturated water slab surface and then steered into the bulk of the water.
 Figure \ref{fig:starting-configurations} illustrates two sample starting configurations for the SMD simulations, showing both the neat-\wat~slab and the saturated-\wat~slab configurations before steering the \suldiox~into the water bulk.

### 2010-12-21 Background.tex
Reworded background, added new subsection to results

@@
 
-	The two molecules studied, \wat~and \suldiox, and similarly shaped with a C2$_v$ axis along their bisectors, and a plane is defined by the three atoms in the molecules.
+	The two molecules studied, \wat~and \suldiox, and similarly shaped with a C2$_v$ axis along their bisectors, and a molecular plane is defined by their three atoms.
 The orientational analyses presented herein focus on two angles used to define the molecular orientation in space.
@@
 Thus the value of $\theta$ falls within a range of $[0,\pi]$.
-A second angle, $\phi$, defines the molecular ``twist'' around the bisector.
-$\phi$ is formed between the projection of a vector onto the x-y plane and the x-axis.
+A second angle, $\phi$, defines the molecular ``twist'' around the reference axis.
+$\phi$ is formed between the projection of a vector onto the x-y reference plane, and the x-axis.
 $\phi$ is thus defined relative to the x-axis, and its values are in the interval $[-\pi,\pi]$.
-Figure \ref{fig:spherical-angle} shows the angles as related to the \suldiox~molecule's bisector and using the z-axis as the reference.
+However, because of the symmetry of the \wat~and \suldiox~molecules and the equivalence of the two H or O atoms (in the \wat~or\suldiox, respectively), $\phi=-\pi$ is equivalent to $\phi=\pi$, and both situations are equivalent to $\phi=0$.
+Because $\phi=-\phi$, the value is reported within the range of $[0,\frac \pi 2]$.
+Figure \ref{fig:spherical-angle} shows the angle definitions using the z-axis as the reference axis.
 

### 2010-12-21 EquilibratedResults.tex
Reworded background, added new subsection to results

\section{Equilibrated MD}

\subsection{Water Orientation}

\subsection{\suldiox~Orientation}

### 2010-12-22 Background.tex
New text on molecular orientation angles and body-fixed frames

--- 2010-12-17 Background.tex
@@
 
+	Analyses of the simulated systems were performed to characterize the orientation of \suldiox in various environments above, within, and below the surface water region.
+The orientation of the surface molecules as \suldiox adsorbs and interacts with water is a property of any hydrate complex formation that takes places.
+Better understanding of the molecular orientational distributions will help elucidate the chemistry occurring during the adsorption process.
+
+	The two molecules studied, \wat~and \suldiox, and similarly shaped with a C2$_v$ axis along their bisectors, and a molecular plane is defined by their three atoms.
+An body-fixed frame is defined for \wat~and \suldiox~as shown in figure \ref{fig:molecular-frame}.
+In each analysis a world-fixed reference axis is used that is the long axis of the system's periodic cell (normal to the plane of the water surface) unless otherwise noted.
+The orientational analyses presented herein focus on two angles used to define molecular orientation in space.
+
+	A ``tilt'' angle, $\theta$, defines the angle formed between the bisector vector (the molecular z-axis, pointing from the central atom in the direction of the other two atoms) and the positive system reference axis.
+Thus the value of $\theta$ falls within a range of $[0,\pi]$.
+An angle of $\theta=0$ indicates a molecule with its bisector aligned with the reference axis, while $\theta=\pi$ results from an anti-aligned configuration.
+
+	A second angle, $\phi$, defines the molecular ``twist''.
+$\phi$ is the angle formed between the vector normal to the plane of the molecule (molecular y-axis) and the system reference axis.
+The values of $\phi$ fall in the interval $[-\pi,\pi]$.
+Given certain values of $\theta$, $\phi$ provides additional information about whether the molecular orientation is ``flat'' to the surface (e.g.
+the plane of the molecule is aligned with the plane of the surface), or if it is perpendicular.
+%However, the symmetry of the \wat~and \suldiox~molecules create equivalence between the two H, or O, atoms (in the \wat~or\suldiox, respectively).
+$\phi=-\pi$ is equivalent to $\phi=\pi$, and both situations are equivalent to $\phi=0$.
+Because $\phi=-\phi$, the value is reported within the range of $[0,\frac \pi 2]$.
+	Figure \ref{fig:water-angles} shows the angle definitions relating the body-fixed axes to a system-fixed reference axis.
+
+%	The two molecules studied, \wat~and \suldiox, and similarly shaped with a C2$_v$ axis along their bisectors, and a molecular plane is defined by their three atoms.
+The orientational analyses presented herein focus on two angles used to define the molecular orientation in space.
+In each analysis a reference axis is used to define molecular angles, and is the z-axis unless otherwise specified.
+A ``tilt'' angle, $\theta$, defines the angle formed between a vector (generally the bisector vector pointing from the central atom in the direction of the other two atoms) and the positive reference axis.
+Thus the value of $\theta$ falls within a range of $[0,\pi]$.
+A second angle, $\phi$, defines the molecular ``twist'' around the reference axis.
+$\phi$ is formed between the projection of a vector onto the x-y reference plane, and the x-axis.
+$\phi$ is thus defined relative to the x-axis, and its values are in the interval $[-\pi,\pi]$.
+However, because of the symmetry of the \wat~and \suldiox~molecules and the equivalence of the two H or O atoms (in the \wat~or\suldiox, respectively), $\phi=-\pi$ is equivalent to $\phi=\pi$, and both situations are equivalent to $\phi=0$.
+Because $\phi=-\phi$, the value is reported within the range of $[0,\frac \pi 2]$.
+Figure \ref{fig:spherical-angle} shows the angle definitions using the z-axis as the reference axis.
+
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[scale=1.0]{images/molecularframesmall.png}
+		\caption{The body-fixed are defined with the molecular plane in the x-z plane, and the z-axis aligned to the molecular bisector.
+This sets the y-axis normal to the molecular plane, and one of the molecular bonds in the positive x direction.}
+		\label{fig:molecular-frame}
+	\end{center}
+\end{figure}
+
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[scale=1.0]{images/wateranglessmall.png}
+		\caption{The visual representation of the angles $\theta$ and $\phi$ used to define molecular orientation of \suldiox~and \wat~relative to the reference axis.
+$\theta$ is the value of the ``tilt'' of the molecular bisector, and $\phi$ is the ``twist'' that indicates how flat the molecule is relative to the interfacial plane.}
+		\label{fig:water-angles}
+	\end{center}
+\end{figure}
+
 \subsubsection{Water Orientation}
-\subsubsection{\suldiox Orientation}
+The orientation of water molecules at the
+
+\subsubsection{\suldiox~Orientation}
 

### 2010-12-22 Background.tex
talking about how the surface location is determined

@@
 \section{Background}
+
+\subsection {Aqueous Surface Location}
+
+	The analyses performed in this study often relate results to the location of the gas-liquid interface.
+An interface, however, is a dynamic surface that is neither flat nor stationary.
+Previous studies have used the technique of fitting a lineshape to the density profile of the water or solute to extract interfacial shape and location parameters.\cite{find something for this} Often a hyperbolic tangent function is fit, and values for the ``Gibb's dividing surface'' location, and interfacial width are determined.\cite{couple papers about the tangent function} However, in long simulations the location of the interface will change, and the net motion of surface waters will alter the interfacial width at any given timestep.
+Thus, the averaged density profile fitting will capture average widths and locations, not instantaneous ones.
+This problem has been addressed previously by an algorithmic method of finding an instantaneous interface for each timestep.\cite{the paper that addresses this} In this work we attack the problem by finding an instantaneous water surface location for each timestep by averaging the locations of the topmost 20 water molecules.
+This provides us with a consistent measure of the top water surface's location for relating our analyses.
 
@@
 
-	Analyses of the simulated systems were performed to characterize the orientation of \suldiox in various environments above, within, and below the surface water region.
-The orientation of the surface molecules as \suldiox adsorbs and interacts with water is a property of any hydrate complex formation that takes places.
+	The simulated systems were analyzed to characterize the orientation of \wat~and \suldiox~in various environments above, within, and below the surface region.
+The orientation of the surface molecules as \suldiox~adsorbs and interacts with water is a property of any hydrate complex formation that takes places.
 Better understanding of the molecular orientational distributions will help elucidate the chemistry occurring during the adsorption process.
@@
 
-\subsubsection{Water Orientation}
-The orientation of water molecules at the
-
-\subsubsection{\suldiox~Orientation}
-
 \subsection{Bonding Analysis}

### 2010-12-22 Background.tex
Wrote all about the water orientation at the surface for both of the equilibrated systems

--- 2010-12-22 Background.tex
@@
 \section{Background}
+
+\subsection {Aqueous Surface Location}
+
+	The analyses performed in this study relate results to the location of the gas-liquid interface.
+An interface, however, is a dynamic surface that is neither flat nor stationary.
+Previous studies have used the technique of fitting a lineshape to the density profile of the water or solute to extract interfacial shape and location parameters.\cite{Shamay2010,Wick2006c,Chowdhary2006} Hyperbolic tangent functions have been used for fitting, and values for the ``Gibb's dividing surface'' location, and interfacial width are determined.\cite{Matsumoto1988} However, in long simulations the location of the interface will change, and the net motion of surface waters will alter the interfacial width at any given timestep.
+Thus, the averaged density profile fitting will capture average widths and locations, not instantaneous values.
+The interface can be characterized by density profiles, molecular orientational profiles, or interfacial intermolecular correlations.
+In this work we approach the problem by finding a reference location for each timestep by averaging the positions of the waters contained in the topmost monolayer.
+This provides us with a consistent reference point in the simulations to which we relate our analyses.
+This method is independent of the various measures of the interface, but still provides an intuitive point of reference.
 

### 2010-12-22 EquilibratedResults.tex
talking about how the surface location is determined
Wrote all about the water orientation at the surface for both of the equilibrated systems

@@
 \section{Equilibrated MD}
+
+Angular analyses were performed to characterize the net molecular orientation of \wat~and \suldiox~molecules.
+Histograms were created to show the angles $\theta$ and $\phi$ (see figure \ref{fig:water-angles}) relative to a positions from the top surface of the aqueous slab.
+Both \wat~and \suldiox~orientations are considered for both the neat-\wat~and saturated systems.
 
@@
 
+The results of the analyses for \wat~are shown in figure \ref{fig:water-orientation} for both of the surface equilibrated systems.
+In both systems an orientational preference is found at both the top and bottom slab surfaces.
+Because the bottom slab is inverted with respect to the top, it is expected that the orientational distributions will reflect the symmetry.
+The bisector tilt concentrates around $\cos(\theta)=0$ within the first few\angs of the surface, and then the distribution becomes isotropic further into the water bulk.
+As the tilt nears $\cos(\theta)=0$ the \wat~bisector lies within the plane of the surface indicating a water orientation either flat on the surface, or with some amount of ``twist'' sending the OH bonds in towards, or out of the bulk.
+The value of $\phi$ determines the ``twist'' in this case.
+Both systems show a peak in the distributions around $\cos(\phi)=1$ at the water surfaces.
+This results from an orientation of the water's y-axis (normal to the molecular plane) aligned perpendicular to the plane of the water surface.
+
+Peaks in the distributions of the neat-\wat~system are more clearly pronounced as their intensities are more concentrated and larger than the surrounding area of the profiles.
+This difference indicates that the transition from the preferred orientation at the water surface has a sharper distinction from the isotropic bulk than in the system with the saturated \suldiox~surface.
+It appears that the same orientation trend is present in both systems, but the presence of the \suldiox~at the water surface decreases the degree of water orientation at the interface.
+
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[scale=1.0]{images/waterorientationsmall.png}
+		\caption{Molecular orientation histograms of \wat~throughout the surface equilibrated systems.
+The top surface is located at a distance of 0 with negative positions in the bulk of the slab.
+The bottom slab surface is approx.
+30\angs below the top surface.
+Shown are the angle distributions for $\theta$ (left column) and $\phi$ (right column) in both the neat-\wat~system (top row) and the saturated system (bottom row).}
+		\label{fig:water-orientation}
+	\end{center}
+\end{figure}
+
 \subsection{\suldiox~Orientation}

### 2010-12-23 EquilibratedResults.tex
so2 orientation

@@
 30\angs below the top surface.
-Shown are the angle distributions for $\theta$ (left column) and $\phi$ (right column) in both the neat-\wat~system (top row) and the saturated system (bottom row).}
+Shown are the angle distributions for $\theta$ (left column) and $\phi$ (right column) in both the neat-\wat~system (top row) and the saturated system (bottom row).
+Dark blue regions are void of water molecules above the top and below the bottom surfaces.}
 		\label{fig:water-orientation}
@@
 \subsection{\suldiox~Orientation}
+
+Orientation distributions of the adsorbed \suldiox~molecules were created for both the neat-\wat~and saturated \suldiox~systems.
+Figure \ref{fig:so2-orientation} shows the distributions of $\cos(\theta)$ and $\cos(\phi)$.
+As with the water orientation (figure \ref{fig:water-orientation}), the \suldiox~orients similarly in both of the surface equilibrated systems.
+
+The location of \suldiox~within the water surface is slightly different between the two systems.
+The distribution of the single \suldiox~molecule is located just below the 0\angs position, and the distribution of the concentrated system is centered above the surface.
+This indicates that in both systems the \suldiox~is highly surface active, but the locations of the peak \suldiox~densities differ.
+The single molecule spends most of its time bound within the top water monolayer, and often traversing below that location by several\angs.
+The saturated system's surface molecules sit on top of the monolayer.
+The angular distribution of the single \suldiox~is concentrated primarily in $\cos(\theta)>0$ taking an orientation of the bisector pointing out of the water surface.
+This same distribution occurs in the saturated system for positions below 0\angs, but the orientation becomes isotropic for molecules located above 0\angs, on top of the water surface.
+
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[scale=1.0]{images/so2orientationsmall.png}
+		\caption{Molecular orientation distributions for \suldiox~molecules adsorbed to the water slab surface.
+Distributions are shown for $\cos(\theta)$ (left column) and $\cos(\phi)$ (right column) for both the neat-\wat~(top) and high \suldiox~concentration (bottom) systems.
+For both systems the $\theta$ distributions suggest \suldiox~bound to the water surface with the sulfur pointing towards the water slab, and the oxygens pointing to the gas phase.
+In this configuration the $\phi$ distribution is isotropic because of the water slab's in-plane symmetry.}
+		\label{fig:so2-orientation}
+	\end{center}
+\end{figure}

### 2010-12-29 TransitResults.tex
New transit so2 angle analysis files - not yet added

\section{Steered MD}

\subsection {Neat-\wat~Single \suldiox System}

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/so2transitangles.png}
		\caption{}
		\label{fig:so2-transit-angles}
	\end{center}
\end{figure}

\subsection {Saturated \wat~System}

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/so2transitsaturatedangles.png}
		\caption{}
		\label{fig:so2-transit-saturated-angles}
	\end{center}
\end{figure}

### 2010-12-29 Background.tex
New section on the so2 transit orientation analysis + pictures

@@
 		\includegraphics[scale=1.0]{images/molecularframesmall.png}
-		\caption{The body-fixed are defined with the molecular plane in the x-z plane, and the z-axis aligned to the molecular bisector.
-This sets the y-axis normal to the molecular plane, and one of the molecular bonds in the positive x direction.}
+		\caption{The body-fixed axes are defined with the x-z plane formed by the three atoms, and the z-axis aligned to the molecular bisector.
+This sets the y-axis normal to the molecular plane, and one of the bonds in the positive x direction.}
 		\label{fig:molecular-frame}
@@
 		\includegraphics[scale=1.0]{images/wateranglessmall.png}
-		\caption{The visual representation of the angles $\theta$ and $\phi$ used to define molecular orientation of \suldiox~and \wat~relative to the reference axis.
-$\theta$ is the value of the ``tilt'' of the molecular bisector, and $\phi$ is the ``twist'' that indicates how flat the molecule is relative to the interfacial plane.}
+		\caption{The definition of the angles $\theta$ and $\phi$ used to define molecular orientation of \suldiox~and \wat~relative to a reference axis.
+$\theta$ is the value of the ``tilt'' of the molecular bisector from the reference axis.
+$\phi$ is the ``twist'' angle that indicates how flat the molecule is relative to the interfacial plane, and is formed between the vector normal to the molecular plane (the body-fixed y-axis) and the reference axis.}
 		\label{fig:water-angles}

### 2010-12-29 TransitResults.tex
New section on the so2 transit orientation analysis + pictures

@@
 
-\subsection {Neat-\wat~Single \suldiox System}
+\subsection {\suldiox~Orientation}
+
+	The orientation of a \suldiox~molecule adsorbing to an aqueous surface was determined through an analysis process to find the angles $\theta$ and $\phi$ (figure \ref{fig:water-angles}).
+For each timestep of the SMD simulations the orientation angles were determined for a single \suldiox~as it was pulled into the water slab, both in the neat-\wat~and saturated slab systems.
+The resulting values were collected for the 50 simulations for both systems, and the angle values were binned indexed with the position of the \suldiox~relative to the water slab surface location.
+The resulting 2-dimensional histograms shown in figure \ref{fig:so2-transit-angles} show the orientation trend for the \suldiox~in the gas phase, near the aqueous surface, and after entering the water bulk.
 
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/so2transitangles.png}
+		\includegraphics[scale=1.0]{images/so2transitorientationsmall.png}
 		\caption{}
@@
 
-\subsection {Saturated \wat~System}
+	From its starting position 20\angs above the water surface, until the \suldiox~moves to within 5\angs of the surface of both systems, the orientation is isotropic in both $\theta$ and $\phi$.
+Upon approaching the saturated surface at approx.
+5\angs, the \suldiox~$\phi$ orientation distribution begins to cluster near $\cos(\phi)=1$, with the $\theta$ distribution remaining more isotropic, but with a trend of $\cos(\theta)>0$.
+At 5\angs the \suldiox~begins interacting with the other \suldiox~molecules that sit atop the water, and the resulting orientation distribution suggests that contact with the \suldiox~layer causes the adsorbing molecule to lie more flat to the surface.
+Within this same region from 0-5\angs above the surface, the $\phi$ distribution of the neat-\wat~system remains isotropic.
+The intensity of the distribution (darker blue) also indicates that the \suldiox~spends little time in this region, as it is quickly adsorbed further into the water bulk.
 
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[scale=1.0]{images/so2transitsaturatedangles.png}
-		\caption{}
-		\label{fig:so2-transit-saturated-angles}
-	\end{center}
-\end{figure}
+	Just below the water surface as the distance moves into negative values, both systems show clear orientational preferences.
+In both systems the bisector orientation clusters such that $\cos(\theta)<0.5$, with most of the distribution intensity around $\cos(\theta)=1$.
+This suggests that \suldiox~in a water surface orients with the sulfur pointing into the water bulk, and the oxygens pointing towards the surface water molecules.
+
+	Differences occurr between the system orientation profiles once the \suldiox~has moved further than 10\angs below the surface.
+While both $\theta$ and $\phi$ hold the same trend from just under the surface through to the bulk of the slab in the saturated system, the orientation profile becomes much more isotropic past 10\angs under the surface of the neat-\wat~system.
+The orientation effect is shallower under the neat-\wat~surface.
+However, the orientation distribution peaks most intensely, and more tightly clustered with the \suldiox~bisector aligning with the surface normal, just underneath the neat-\wat~surface than the saturated one.
+The distributions thus show an orienting force extending deeper into the surface of the saturated system than the neat-\wat~system.
+In terms of how deeply the \suldiox~molecule is oriented, the interfacial region of the saturated system is larger and extends further into the aqueous bulk.

### 2011-01-12 Background.tex
New image and text for the background on finding cycles using the graph technique

@@
 
-\subsection{Bonding Analysis}
-\subsubsection{Nearest Neighbor}
-\subsubsection{Hydrate Complex Formation}
+\subsection{Hydrate Complex Analysis}
+
+	During the adsorption process of a gaseous \suldiox~moving in towards a water surface hydrate complexes are formed as the waters bind to the molecule.
+Computationally, there are various methods to determine hydrate complex formation.
+The most simple metric is to count the number of hydrogen bonds formed from the \suldiox~to neighboring molecules.
+This method can be furthered extended to also include the strength of those bonding interactions by additionally reporting the bondlengths to the hydrogen bonding partners.
+Longer bondlengths are typical of weaker bonds, and this method is used in this work to identify the bonding environment of the \suldiox~during solvation.
+
+	A simple count of hydrogen bonds is telling of the interactions and environments of an adsorbing \suldiox.
+However, a more sophisticated analysis is needed to fully characterize the geometry of any complexes that form.
+In this work a graph theoretical method has been used to identify hydrate complexes that form cyclic structures in addition to those that form more simple hydrogen bonding geometries.
+Previously published works have detailed graph theory for use in molecular systems, and have used graph theoretical methods to describe ice and water clusters, structural energies, as well as in the study of relationships in biological systems.\cite{Huber2007,Shi2005,Radhakrishnan1991,Anick2002} As such, we will not present here a review of the basic graph theoretical details that are well documented in the referenced works.
+As shown in figure \ref{fig:so2-complex-graph}, the atoms of a \suldiox~and its nearest neighbors are represented as a set of verrtices, with the covalent and hydrogen bonds between them forming the edges of a graph.
+
+	We present here a technique for determining bonding cycles in a hydrate complex, but this technique is general enough to be applied to any graph representation of simple molecular systems.
+The first step in the determination of cycles is translating the atoms (vertices) and bonds (edges) to a graph.
+Figure \ref{fig:so2-complex-graph} shows the procedure visually, where here the root vertex of the graph is the \suldiox-sulfur, and each atom is numbered by the distance, in atoms, to the root vertex.
+We use only the closest neighboring water molecules that may take part in a cyclic complex in building the graph.
+After forming the graph, a breadth-first search (BFS) is performed starting at the root vertex.\cite{Knuth1997} The BFS algorithm uses a coloring technique where all unexamined vertices are colored white, queued vertices are gray, and examined vertices are black.
+During the BFS, encountering a gray vertex as a target of an adjacent edge during examination indicates that a cycle has been found in the graph.
+At each vertex dequeue for examination, the parent of every vertex is recorded such that the lineage from gray target to root vertex can be reconstructed to determine the participating atoms in the cycle.
+
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[scale=1.0]{images/so2complexfigure.png}
+		\caption{A \suldiox~and its nearest neighbors are represented as a graph of vertices and edges.
+The bonds are all given equal weights and the graph is undirected.
+The numbering shows the relative distance from any atom to the root node (\suldiox-sulfur 1), and also enumerates the iteration through the BFS process at which the atom is discovered.
+A BFS on this graph would result in discovering the doubly-bound hydrogen (5) as a gray target, and the gray source would be either of the connected oxygens (4).
+At each iteration the predecessor in the BFS is recorded in order to reconstruct the connectivity of the cycle.
+This cycle is composed of 8 atoms, and 3 molecules}
+		\label{fig:so2-complex-graph}
+	\end{center}
+\end{figure}

### 2011-02-18 Background.tex
Rearranging the filenames and working on the h2o-orientation results

@@
 
-	The analyses performed in this study relate results to the location of the gas-liquid interface.
-An interface, however, is a dynamic surface that is neither flat nor stationary.
-Previous studies have used the technique of fitting a lineshape to the density profile of the water or solute to extract interfacial shape and location parameters.\cite{Shamay2010,Wick2006c,Chowdhary2006} Hyperbolic tangent functions have been used for fitting, and values for the ``Gibb's dividing surface'' location, and interfacial width are determined.\cite{Matsumoto1988} However, in long simulations the location of the interface will change, and the net motion of surface waters will alter the interfacial width at any given timestep.
+	The results of analyses performed in this study are related to the location of the liquid water surface.
+A surface, however, is a dynamic boundary that is neither flat or stationary.
+Several previous studies have used the technique of fitting a lineshape to the density profile of the water or solute to extract interfacial shape and location parameters.\cite{Shamay2010,Wick2006c,Chowdhary2006} Hyperbolic tangent functions have been used to extract fitting parameters, and values for the ``Gibb's dividing surface'' location, and interfacial width have thus been determined.\cite{Matsumoto1988} However, in long simulations the location and shape of the interface will change, and the motion of surface waters alters the interfacial width at any given timestep.
 Thus, the averaged density profile fitting will capture average widths and locations, not instantaneous values.
-The interface can be characterized by density profiles, molecular orientational profiles, or interfacial intermolecular correlations.
-In this work we approach the problem by finding a reference location for each timestep by averaging the positions of the waters contained in the topmost monolayer.
+Similarly, the averaged values of location and width will obscure information about any drift or deformation the surface undergoes.
+In this work we approach the problem of surface loation by calculating a reference location for each timestep by averaging the long-axis positions of the waters contained in the topmost monolayer.
 This provides us with a consistent reference point in the simulations to which we relate our analyses.
-This method is independent of the various measures of the interface, but still provides an intuitive point of reference.
+This method provides an intuitive point of reference, but does not increase the computational burden of the analyses.
 
-\subsection{Analysis of Orientation}
+  The location of the water surface was thus determined by the averaged positions of a set number of the topmost waters in the simulated aqueous slabs.
+The number of waters counted was decided by taking a few issues into account.
+First, a visual inspection of the top 30x30 \angs$^2$ cross-section of the water slab over several timesteps indicated between 65-75 waters established a full monolayer.
+Alternatively, assuming a spherical model of water with a radius of 2.4\angs (the typical length of a bulk-water hydrogen bond), two layers of hexagonally tight-packed spheres would yield a maximum of 52 molecules.
+Increasing the number of waters used in calculating the surface location diminishes the effects of the few waters that briefly rise above the surface into the gas phase, stabilizing both the surface position and thickness.
+Taking the close-packed model as a minimum number of waters fit into a flat surface, the authors decided to use 70 water molecules for calculating surface location during the simulation analyses.
 
-	The simulated systems were analyzed to characterize the orientation of \wat~and \suldiox~in various environments above, within, and below the surface region.
-The orientation of the surface molecules as \suldiox~adsorbs and interacts with water is a property of any hydrate complex formation that takes places.
-Better understanding of the molecular orientational distributions will help elucidate the chemistry occurring during the adsorption process.
+  %Taking one of the simulations of the neat-water slabs, a series of calculations were performed to find the averaged positions and the standard deviations of varying numbers of the topmost waters.
+It was found that using 70 waters produced a standard deviation of the water positions corresponding to a width of 3.2\angs.
+as reported by Dang et al.when fitting the density profile with a hyperbolic tangent lineshape.
+To match the interfacial width of Dang using the standard deviation of our simulations, 70 water molecules would have to be included in the top monolayer.
+Thus, the calculations and analyses presented in this work incorporate 70 water molecules when calculating the surface location.
 
-	The two molecules studied, \wat~and \suldiox, and similarly shaped with a C2$_v$ axis along their bisectors, and a molecular plane is defined by their three atoms.
-An body-fixed frame is defined for \wat~and \suldiox~as shown in figure \ref{fig:molecular-frame}.
-In each analysis a world-fixed reference axis is used that is the long axis of the system's periodic cell (normal to the plane of the water surface) unless otherwise noted.
+\subsection{Analysis of Molecular Orientation}
+
+	The simulated systems were analyzed to characterize the orientation of \wat~and \suldiox~in various environments above, within, and below the aqueous surface region.
+In order to better understand the chemistry occurring during the \suldiox~adsorption process, the molecular orientation of both the \was~and \suldiox~must be understood.
+The two molecules studied are similarly shaped with a C2$_v$ axis along their bisectors, and a molecular plane defined by three atoms.
+For analysis, a body-fixed frame is defined for \wat~and \suldiox~as shown in figure \ref{fig:molecular-frame}.
+In each analysis a space-fixed reference axis is used that corresponds to the long axis of the system's periodic cell (normal to the plane of the water surface) unless otherwise noted.
 The orientational analyses presented herein focus on two angles used to define molecular orientation in space.
 
-	A ``tilt'' angle, $\theta$, defines the angle formed between the bisector vector (the molecular z-axis, pointing from the central atom in the direction of the other two atoms) and the positive system reference axis.
+	A ``tilt'' angle, $\theta$, defines the angle formed between the molecular bisector vector (the molecular z-axis, pointing from the central atom in the direction of the other two atoms) and the positive system reference axis.
 Thus the value of $\theta$ falls within a range of $[0,\pi]$.
-An angle of $\theta=0$ indicates a molecule with its bisector aligned with the reference axis, while $\theta=\pi$ results from an anti-aligned configuration.
+An angle of $\cos(\theta)=1$ indicates a molecule with its bisector aligned with the reference axis, while $\cos(\theta)=-1$ results from an anti-aligned configuration.
 
-	A second angle, $\phi$, defines the molecular ``twist''.
+	A second angle, $\phi$, defines the molecular ``twist'' for certain configurations of the molecule.
 $\phi$ is the angle formed between the vector normal to the plane of the molecule (molecular y-axis) and the system reference axis.

### 2011-02-18 EquilibratedOrientation.tex
Rearranging the filenames and working on the h2o-orientation results

--- 2010-12-23 EquilibratedResults.tex
@@
 
-Angular analyses were performed to characterize the net molecular orientation of \wat~and \suldiox~molecules.
-Histograms were created to show the angles $\theta$ and $\phi$ (see figure \ref{fig:water-angles}) relative to a positions from the top surface of the aqueous slab.
+Molecular orientation analyses were performed to characterize the net molecular orientation of \wat~and \suldiox~molecules in the different locations near and within the water surface.
+Histograms were created to show the angles $\theta$ and $\phi$ relative to positions from the top surface of the aqueous slab.
 Both \wat~and \suldiox~orientations are considered for both the neat-\wat~and saturated systems.
@@
 
-The results of the analyses for \wat~are shown in figure \ref{fig:water-orientation} for both of the surface equilibrated systems.
-In both systems an orientational preference is found at both the top and bottom slab surfaces.
-Because the bottom slab is inverted with respect to the top, it is expected that the orientational distributions will reflect the symmetry.
-The bisector tilt concentrates around $\cos(\theta)=0$ within the first few\angs of the surface, and then the distribution becomes isotropic further into the water bulk.
+The results of the analyses for \wat~are shown in figure \ref{fig:water-orientation} for both the neat-water and saturated systems.
+In both systems an orientational preference is found at the slab surfaces where the water is in contact with the gas phase, or the surface \suldiox~molecules.
+%The bisector tilt concentrates around $\cos(\theta)=0$ within the first few\angs of the surface, and then the distribution becomes isotropic further into the water bulk.
 As the tilt nears $\cos(\theta)=0$ the \wat~bisector lies within the plane of the surface indicating a water orientation either flat on the surface, or with some amount of ``twist'' sending the OH bonds in towards, or out of the bulk.

### 2011-02-18 Methods.tex
Rearranging the filenames and working on the h2o-orientation results

@@
 
-Molecular dynamics simulations were performed using the Amber 11 software suite.\cite{Case2010} In this study we incorporate polarizable models for the \wat~and \suldiox~molecules.
-Fully polarizable models have been used previously in studies on interfacial systems because they are known to more accurately reproduce interfacial structure and free energy profiles.\cite{Wick2007,Rivera2006,Dang1998} The \wat~model used is the POL3 model,\cite{Caldwell1995} and for \suldiox we used the model of Dang, et al.
+Molecular dynamics simulations were all performed using the Amber 11 software suite.\cite{Case2010} Polarizable models for the \wat~and \suldiox~molecules were used in the simulations, and have been used previously in studies on interfacial systems because they are known to more accurately reproduce interfacial structure and free energy profiles.\cite{Wick2007,Rivera2006,Dang1998} The \wat~model used is the POL3 model,\cite{Caldwell1995} and for \suldiox we used the model of Dang, et al.
 that places a single polarizable center on the sulfur atom.\cite{Baer2010}
@@
 All simulations began with an equilibrated cube of 900 \wat~molecules, with sides of length 30\angs.
-The long axis of each simulation cell (the axis normal to the water surface) was set to 120\angs.
+The long axis of each simulation cell (the axis normal to the water surface) was then lengthened to 120\angs, and the systems were further equilibrated for 10 ns.
 The simulations all employed periodic boundaries to create an ``infinite-slab'' geometry.
-After equilibrating the neat-\wat~slabs for 1 ns, two types of systems were created by introducing \suldiox.
+After equilibrating the neat-\wat~slabs two types of systems were created by introducing \suldiox: a single-\suldiox~system, herein referred to as the ``neat-water'' system, and a saturated \suldiox~system.
 
-One system involved the addition of a single \suldiox~molecule either within the bulk of the water slab, or above the slab surface.
+One system involved the addition of a single \suldiox~molecule either within the bulk of the water slab (for equilibrium MD), or above the slab surface (in the SMD simulations).
 The single-\suldiox~system was then evolved for 2 ns to produce an equilibrated starting configuration.
 The second system introduced 22 \suldiox~molecules into the water slab in order to saturate it to a level coinciding with the Henry's law constant for \suldiox~in water.
-Additionally, 50 \suldiox~molecules were introduced into the gas phase outside of the water slab to simulate an added 50 atm of \suldiox~gas pressure.
-The saturated system was then evolved for 2 ns to produce a starting configuration for further simulations.
+Additionally, 50 \suldiox~molecules were introduced into the gas phase outside of the water slab, on both surfaces, to simulate an added 50 atm of \suldiox~gas pressure.
+The saturated system was then evolved for 2 ns to produce a starting configuration for further saturated simulations.
 
@@
 One system had a single \suldiox~added to the center of the water box, representing a concentration of 0.06 M.
-The complementary high-concentration system with 22 \suldiox~molecule in the bulk represents a system of concentration 1.35 M, with an additional 50 \suldiox~in the gas phase to model 50 atm of pressure above the water surface.
-Both the low and high concentration systems were then evolved for a further 2.5 ns with atomic coordinates written every 100 fs.
+The complementary saturated system with 22 \suldiox~molecule in the bulk represents a system of concentration 1.35 M, with an additional 50 \suldiox~in the gas phase to model 50 atm of pressure above the water surface.
+Both the low and high concentration systems were then evolved for a further 10 ns using a timestep of 0.5fs, with atomic coordinates written every 100 fs.
 
@@
 A second set of simulations involved preparing an equilibrated water slab as in the surface equilibrated method above.
-However, for both the low and high concentration systems, a single \suldiox~was introduced 20\angs above the slab, with the sulfur atom tethered to its initial position.
-The system was evolved for 1 ns, taking coordinate snapshots every 20 ps to create 50 starting points for further simulations.
-Steered molecular dynamics (SMD) were then performed on the 50 system configurations (in both the low and high concentrations) to guide the \suldiox~down towards a tethered water 15\angs under the water slab surface by applying a small steering force.\cite{Isralewitz2001} The \suldiox~thus passed through the continuum of environments from gas phase to hydrate complex formation or interaction with surface-bound \suldiox, water surface adsorption, and absorption into the bulk of the \wat~slab.
+However, for both the neat-water and saturated systems, a single \suldiox~was introduced 20\angs above the water slab surface, with the sulfur atom tethered to its initial position.
+The systems were evolved for 1 ns, taking coordinate snapshots every 20 ps to create 50 starting points for further simulations.
+Steered molecular dynamics (SMD) were then performed on the 50 system configurations (in both the low and high concentrations) to guide the \suldiox~down towards a tethered water 15\angs under the water slab surface by applying a small steering force.\cite{Isralewitz2001} The \suldiox~thus passed through the continuum of environments from gas phase to (neat- and saturated) water surface adsorption, and finally absorption into the bulk of the \wat~slab.
 
-Figure \ref{fig:starting-configurations} illustrates two sample starting configurations for the SMD simulations, showing both the neat-\wat~slab and the saturated-\wat~slab configurations before steering the \suldiox~into the water bulk.
+Figure \ref{fig:starting-configurations} illustrates two sample starting configurations for the SMD simulations, showing both the neat-\wat~slab and the saturated-\wat~slab configurations before steering the \suldiox~towards the water bulk.
 

### 2011-02-18 EquilibratedOrientation.tex
more on water orientation results

@@
 In both systems an orientational preference is found at the slab surfaces where the water is in contact with the gas phase, or the surface \suldiox~molecules.
+The histograms are arranged with the plots of $\theta$ on the left and $\phi$ on the right columns, and the results for the neat-water and saturated systems are in the top and bottom rows, respectively.
 %The bisector tilt concentrates around $\cos(\theta)=0$ within the first few\angs of the surface, and then the distribution becomes isotropic further into the water bulk.
@@
 This results from an orientation of the water's y-axis (normal to the molecular plane) aligned perpendicular to the plane of the water surface.
+The histograms show overall similarities in their shapes and distributions indicating that the presence of a saturated layer of adsorbed \suldiox~molecules alters the water orientation, but not necessarily the depth of the interface.
 
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/waterorientationsmall.png}
+		\includegraphics[scale=1.0]{images/h2o-angles/h2oangles.png}
 		\caption{Molecular orientation histograms of \wat~throughout the surface equilibrated systems.

### 2011-02-18 EquilibratedOrientation.tex
rounding out the h2o-angles section and starting more in the so2-angles for the equilibrium simulations

@@
 
-The results of the analyses for \wat~are shown in figure \ref{fig:water-orientation} for both the neat-water and saturated systems.
+The results of the analyses for \wat~are shown in figure \ref{fig:water-orientation} for both the neat-water and saturated systems during the equilibrated MD simulations.
 In both systems an orientational preference is found at the slab surfaces where the water is in contact with the gas phase, or the surface \suldiox~molecules.
 The histograms are arranged with the plots of $\theta$ on the left and $\phi$ on the right columns, and the results for the neat-water and saturated systems are in the top and bottom rows, respectively.
+The distance axis was shifted such that the top surface is located at 0\angs, and the bottom surface is approx.
+30\angs below.
+It is also noted that $\theta$ values of the bottom surface are inverted with respect to the top surface and should mirror those results.
 %The bisector tilt concentrates around $\cos(\theta)=0$ within the first few\angs of the surface, and then the distribution becomes isotropic further into the water bulk.
@@
 This results from an orientation of the water's y-axis (normal to the molecular plane) aligned perpendicular to the plane of the water surface.
-The histograms show overall similarities in their shapes and distributions indicating that the presence of a saturated layer of adsorbed \suldiox~molecules alters the water orientation, but not necessarily the depth of the interface.
 
-Peaks in the distributions of the neat-\wat~system are more clearly pronounced as their intensities are more concentrated and larger than the surrounding area of the profiles.
+The histograms show overall similarities for both systems in their shapes and intensities.
+The presence of a saturated layer of adsorbed \suldiox~molecules alters the water orientation, but not necessarily the location depth of the interface.
+In both $\theta$ distributions, the orientations of waters in the bulk region are isotropic until 5\angs below the surface location.
+At 5\angs below the surface the water molecules begin to orient with their bisectors within the plane of the interface (perpendicular to the surface normal, $\cos(\theta)\approx 0$).
+The corresponding location in the plot of $\phi$ shows that those molecules are also mostly flat to the surface ($\cos(\phi)\approx 1$).
+Moving further out from the bulk and into the gas phase, the distributions show $\cos(\theta)$ increasing.
+Waters further out from the bulk have fewer bonding interactions, and orient with their hydrogens more towards the gas phase.
+The bisector pointing further into the gas phase leads to isotropy in the values of $\phi$.
+
+The change in $\cos(\theta)$ is not as rapid at the neat-water surface as in the saturated system.
+Furthermore, the noise in the distribution above 5\angs (and below approx -34\angs) in the neat-water system is a result of fewer waters venturing beyond those extents and thus less data far from the surface.
+This is one indication that waters near a layer of adsorbed \suldiox~venture further above the interface and interact with the \suldiox~molecules.
+The interactions with neighboring \suldiox~molecules allow the waters above the surface to orient more perpendicular to the interface.
+Recent SFG studies by our group indicate that ***<something goes here>***.
+
+The distribution of $\cos(\phi)$ is more sharply defined (i.e.
+less isotropic) for the neat-water system than for the saturated one.
+Waters on the neat surface lie flatter, whereas the presence of the \suldiox~allows a greater range of ``twist'' for those waters in the plane of the interface.
+The $\phi$ distributions quickly become isotropic above a couple angstroms from the surface as the bisectors orient perpendicularly, and below the surface as the bulk water loses any orientational preference.
+
+%Peaks in the distributions of the neat-\wat~system are more clearly pronounced as their intensities are more concentrated and larger than the surrounding area of the profiles.
 This difference indicates that the transition from the preferred orientation at the water surface has a sharper distinction from the isotropic bulk than in the system with the saturated \suldiox~surface.
@@
 Shown are the angle distributions for $\theta$ (left column) and $\phi$ (right column) in both the neat-\wat~system (top row) and the saturated system (bottom row).
-Dark blue regions are void of water molecules above the top and below the bottom surfaces.}
+The distributions are normalized to account for the changing number of water molecules at different positions in the system.}
 		\label{fig:water-orientation}
@@
 Orientation distributions of the adsorbed \suldiox~molecules were created for both the neat-\wat~and saturated \suldiox~systems.
-Figure \ref{fig:so2-orientation} shows the distributions of $\cos(\theta)$ and $\cos(\phi)$.
-As with the water orientation (figure \ref{fig:water-orientation}), the \suldiox~orients similarly in both of the surface equilibrated systems.
+Figure \ref{fig:so2-orientation} shows the distributions of $\cos(\theta)$ and $\cos(\phi)$ (arranged similar to the water orientation distributions in figure \ref{fig:water-orientation}).
+The data set for the neat-\wat~system is much smaller as only a single \suldiox~molecule was simulated.
+The resulting distribution plots are noisier than the corresponding saturated plots, but comparisons can still be drawn.
 

### 2011-02-22 EquilibratedOrientation.tex
Some changes to the SO2 equilibrated MD results (orientation), and added the two plots of density for the equilibrium MD.

@@
 
-\subsection{Water Orientation}
+\subsection{\wat~Orientation}
 
@@
 
-Orientation distributions of the adsorbed \suldiox~molecules were created for both the neat-\wat~and saturated \suldiox~systems.
-Figure \ref{fig:so2-orientation} shows the distributions of $\cos(\theta)$ and $\cos(\phi)$ (arranged similar to the water orientation distributions in figure \ref{fig:water-orientation}).
-The data set for the neat-\wat~system is much smaller as only a single \suldiox~molecule was simulated.
-The resulting distribution plots are noisier than the corresponding saturated plots, but comparisons can still be drawn.
+Orientation distributions of the adsorbed \suldiox~molecules were created during the equilibrium simulations for both the neat-\wat~and saturated \suldiox~systems.
+Figure \ref{fig:so2-orientation} shows the distributions of $\cos(\theta)$ and $\cos(\phi)$ (arranged similar to the water orientation distributions plots in figure \ref{fig:water-orientation}).
+The data set for the neat-\wat~system is much smaller as only a single \suldiox~molecule was simulated in the bulk.
+The resulting distribution plots are thus noisier than the corresponding saturated plots, but effective comparisons can still be drawn.
 
-The location of \suldiox~within the water surface is slightly different between the two systems.
-The distribution of the single \suldiox~molecule is located just below the 0\angs position, and the distribution of the concentrated system is centered above the surface.
+The location of \suldiox~within the water surface is slightly different between the two equilibrium systems.
+The single \suldiox~molecule is generally located just below the 0\angs position of the surface, and the distribution of the concentrated system \suldiox~molecules is centered above the surface.
 This indicates that in both systems the \suldiox~is highly surface active, but the locations of the peak \suldiox~densities differ.

### 2011-02-22 DensityResults.tex
New density plots and wrapped up most of the density results section.

\section{Surface Density Distributions}

One measure of surface activity is the spatial distribution of molecules around the interfacial region. The density distribution of both \wat~and\suldiox~were calculated for the equilibrium MD simulations, and the results are presented in figure \ref{fig:density} for both the neat-water and saturated simulations. The \suldiox~in the neat-water system remained at the top slab surface. While the \suldiox~equilibrated in the saturated system, both in the bulk and gas phases accumulated at both the top and bottom surfaces.

The density distributions of the \suldiox~were fit to gaussian lineshapes as shown in equation \ref{eq:gaussian} using the common definitions of the variables $A$, $\mu$, and $\sigma$, and given as a function of the distance to the surface, $z$.

\begin{equation}
  \rho(z)=\frac{A}{\sqrt{2 \pi \sigma^2}} e^{-\frac{(z-\mu)^2}{2\sigma^2}}
  \label{eq:gaussian}
\end{equation}

In the neat-water \suldiox~distribution, $\mu=0.781$, $A=0.109$, and $\sigma=1.684$\angs. Thus, the \suldiox~tends to locate just beneath the surface when in a low-concentration water environment, and it did not fall further into the bulk or move out of the surface into the gas phase. This surface active behavior was found experimentally by a previous SFG study by our group,\cite{Tarbuck2005,Tarbuck2006} and also verified by the computational simulations and spectral calculations of Dang et al.\cite{Baer2010} It was concluded that a surface \suldiox~hydrate complex forms and modifies the structure of water above the Gibbs dividing surface (GDS) shifting \was~free-OH oscillators in towards the bulk, creating a stronger hydrogen-bonding network. In a later section of this article we look at the effect of \suldiox~on surface water orientation.

Both saturated \suldiox~surfaces exhibit accumulations of \suldiox. However, unlike the neat-water slab, the saturated slab has a non-zero bulk concentration of \suldiox~that does not adsorb to the surface. The coefficient values for the top surface from eq \ref{eq:gaussian} are $\mu_{top}=2.695$, $A_{top}=7.001$, and $\sigma_{top}=3.203$. The bottom surface values are $\mu_{bottom}=-33.846$, $A_{bottom}=6.618$, and $\sigma_{bottom}=3.151$\angs. The added concentration of \suldiox~creates a layer of molecules adsorbed on top of the water surface. The center of this distribution is further into the gas phase than for the single-\suldiox~molecule surface. The accumulation of \suldiox~in this classical simulation coincides with the experimental conclusions of a surface of hydrate-complexed \suldiox, but without breakble bonds in the simulation the chemistry that may pull \suldiox~into the bulk water phase is absent. We have confidence in the surface complexation that is exhibited, the effect of the \suldiox~on the surface water molecules, and the accumulation of \suldiox at the interface. However, we do not yet conclude that the specific locations of \suldiox~on top of the water surface are as accurate as those found in the DFT ab initio calculations of Dang et al.\cite{Baer2010}

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/density/density.png}
		\caption{Molecular density distributions of \wat~and \suldiox~calculated along the long-axis of the simulated cells. The distance axis shows the distance of the given molecule from the top surface of the \wat~slab, with positive values located above the slab towards the gas phase, and negative values located in the water bulk. The lower slab surface is located at approx. -30\angs. Distributions of both the neat-water simulation with a single \suldiox~(left, scaled 10x for clarity), and the saturated system (right) are shown.}
		\label{fig:density}
	\end{center}
\end{figure}

### 2011-02-22 Background.tex
resized images to fit in the PDF

@@
 	The simulated systems were analyzed to characterize the orientation of \wat~and \suldiox~in various environments above, within, and below the aqueous surface region.
-In order to better understand the chemistry occurring during the \suldiox~adsorption process, the molecular orientation of both the \was~and \suldiox~must be understood.
+In order to better understand the chemistry occurring during the \suldiox~adsorption process, the molecular orientation of both the \wat~and \suldiox~must be understood.
 The two molecules studied are similarly shaped with a C2$_v$ axis along their bisectors, and a molecular plane defined by three atoms.
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/molecularframesmall.png}
+		\includegraphics[scale=1.0]{images/angle-cartoons/molecularframesmall.png}
 		\caption{The body-fixed axes are defined with the x-z plane formed by the three atoms, and the z-axis aligned to the molecular bisector.
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/wateranglessmall.png}
+		\includegraphics[scale=1.0]{images/angle-cartoons/wateranglessmall.png}
 		\caption{The definition of the angles $\theta$ and $\phi$ used to define molecular orientation of \suldiox~and \wat~relative to a reference axis.
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/so2complexfigure.png}
+		\includegraphics[scale=1.0]{images/Cycles/so2complexfigure.png}
 		\caption{A \suldiox~and its nearest neighbors are represented as a graph of vertices and edges.

### 2011-02-22 DensityResults.tex
resized images to fit in the PDF

@@
 Thus, the \suldiox~tends to locate just beneath the surface when in a low-concentration water environment, and it did not fall further into the bulk or move out of the surface into the gas phase.
-This surface active behavior was found experimentally by a previous SFG study by our group,\cite{Tarbuck2005,Tarbuck2006} and also verified by the computational simulations and spectral calculations of Dang et al.\cite{Baer2010} It was concluded that a surface \suldiox~hydrate complex forms and modifies the structure of water above the Gibbs dividing surface (GDS) shifting \was~free-OH oscillators in towards the bulk, creating a stronger hydrogen-bonding network.
+This surface active behavior was found experimentally by a previous SFG study by our group,\cite{Tarbuck2005,Tarbuck2006} and also verified by the computational simulations and spectral calculations of Dang et al.\cite{Baer2010} It was concluded that a surface \suldiox~hydrate complex forms and modifies the structure of water above the Gibbs dividing surface (GDS) shifting \wat~free-OH oscillators in towards the bulk, creating a stronger hydrogen-bonding network.
 In a later section of this article we look at the effect of \suldiox~on surface water orientation.

### 2011-02-24 EquilibratedOrientation.tex
resized images to fit in the PDF
not sure

@@
 
-The location of \suldiox~within the water surface is slightly different between the two equilibrium systems.
-The single \suldiox~molecule is generally located just below the 0\angs position of the surface, and the distribution of the concentrated system \suldiox~molecules is centered above the surface.
+The location of \suldiox~within the water surface is slightly different between the two equilibrium MD systems (see figure \ref{fig:density}).
+The single \suldiox~molecule is generally located just below the 0\angs position of the surface, and the distribution of the concentrated system \suldiox~molecules is centered above the surface location.
 This indicates that in both systems the \suldiox~is highly surface active, but the locations of the peak \suldiox~densities differ.
-The single molecule spends most of its time bound within the top water monolayer, and often traversing below that location by several\angs.
+The single molecule spends most of its time bound within the top water monolayer, and often traversing below that location by only a few\angs.
 The saturated system's surface molecules sit on top of the monolayer.
-The angular distribution of the single \suldiox~is concentrated primarily in $\cos(\theta)>0$ taking an orientation of the bisector pointing out of the water surface.
+The angular distribution of the single \suldiox~is concentrated primarily in $\cos(\theta)>0$, such that the bisector points out of the water surface.
 This same distribution occurs in the saturated system for positions below 0\angs, but the orientation becomes isotropic for molecules located above 0\angs, on top of the water surface.
+Promixity to the water highly orients the \suldiox~with the sulfur atom pointing in towards the water bulk.
 
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/so2orientationsmall.png}
+		\includegraphics[scale=1.0]{images/so2-angles/so2-angles.png}
 		\caption{Molecular orientation distributions for \suldiox~molecules adsorbed to the water slab surface.

### 2011-02-24 Methods.tex
not sure

@@
 % need to add in the specifics about the SMD routine, force constants, calculations, etc.
-A second set of simulations involved preparing an equilibrated water slab as in the surface equilibrated method above.
+A second set of simulations involved an equilibrated water slab as in the surface equilibrated method above.
 However, for both the neat-water and saturated systems, a single \suldiox~was introduced 20\angs above the water slab surface, with the sulfur atom tethered to its initial position.
 The systems were evolved for 1 ns, taking coordinate snapshots every 20 ps to create 50 starting points for further simulations.
-Steered molecular dynamics (SMD) were then performed on the 50 system configurations (in both the low and high concentrations) to guide the \suldiox~down towards a tethered water 15\angs under the water slab surface by applying a small steering force.\cite{Isralewitz2001} The \suldiox~thus passed through the continuum of environments from gas phase to (neat- and saturated) water surface adsorption, and finally absorption into the bulk of the \wat~slab.
+Steered molecular dynamics (SMD) were then performed on the 50 system configurations (in both the low and high concentrations) to guide the \suldiox~down towards a tethered water-oxygen 15\angs under the water slab surface by applying a small steering force to the \suldiox-sulfur atom.\cite{Isralewitz2001} The \suldiox~thus passed through the continuum of environments from gas phase to (neat- and saturated) water surface adsorption, and finally absorption into the bulk of the \wat~slab.
+Each of the SMD simulations were performed for a total of 200 ps, using a timestep of 1 fs, and taking snapshots of the system every 25 fs.
+Figure \ref{fig:starting-configurations} illustrates two sample starting configurations for the SMD simulations, showing both the neat-water slab and the saturated-\wat~slab configurations before steering the \suldiox~towards the water bulk.
 
-Figure \ref{fig:starting-configurations} illustrates two sample starting configurations for the SMD simulations, showing both the neat-\wat~slab and the saturated-\wat~slab configurations before steering the \suldiox~towards the water bulk.
+A separate set of SMD simulations were performed similarly, but with the tethering to one of the \suldiox-oxygens.
+This was done to ensure that the orientation of the \suldiox~during the adsorption transit was not an artifact of the choice of atom used for tethering.
+The simulations produced the same results (not shown) for the orientational analyses, so the data from the original tethering scheme was used.
 

### 2011-02-24 TransitOrientation.tex
Rearranging the filenames and working on the h2o-orientation results
resized images to fit in the PDF
not sure

--- 2010-12-29 TransitResults.tex
@@
-\section{Steered MD}
+\section{Steered MD Transit Simulations}
 
-\subsection {Neat-\wat~Single \suldiox System}
+\subsection {\suldiox~Orientation}
+
+	The orientation of a \suldiox~molecule adsorbing to an aqueous surface was determined through an analysis process to find the angles $\theta$ and $\phi$ (figure \ref{fig:water-angles}).
+For each timestep of the SMD simulations the orientation angles were determined for the transiting \suldiox~as it was pulled into the water slab from the gas phase, both in the neat-water and saturated slab systems.
+The angle cosine values were collected for the 50 simulations for both systems, and the binned with the position of the \suldiox~relative to the water slab surface location.
+The resulting 2-dimensional histograms shown in figure \ref{fig:so2-transit-angles} show the orientation trend for the \suldiox~in the gas phase (above 0\angs), near the aqueous surface (near 0\angs), and after entering the water bulk (below 0\angs).
+Only the surface into which the transiting \suldiox~was steered is shown for the saturated system.
 
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/so2transitangles.png}
-		\caption{}
+		\includegraphics[scale=1.0]{images/so2-angles/so2-angles-transit.png}
+		\caption{\suldiox~molecular orientation distributions during SMD transit simulations into an aqueous slab.
+Both the neat-water (top row) and saturated (bottom row) data sets were analyzed to determine the angles $\theta$ (left column), and $\phi$ (right column) of the transiting \suldiox.
+The distributions show angle cosines plotted against the distance of the \suldiox~to the location of the water surface.
+The data sets were averaged at each distance for the respective 50 simulations.}
 		\label{fig:so2-transit-angles}
@@
 
-\subsection {Saturated \wat~System}
+	%From its starting position 20\angs above the water surface, until the \suldiox~moves to within 5\angs of the surface of both systems, the orientation is isotropic in both $\theta$ and $\phi$.
+Upon approaching the saturated surface at approx.
+5\angs, the \suldiox~$\phi$ orientation distribution begins to cluster near $\cos(\phi)=1$, with the $\theta$ distribution remaining more isotropic, but with a trend of $\cos(\theta)>0$.
+At 5\angs the \suldiox~begins interacting with the other \suldiox~molecules that sit atop the water, and the resulting orientation distribution suggests that contact with the \suldiox~layer causes the adsorbing molecule to lie more flat to the surface.
+Within this same region from 0-5\angs above the surface, the $\phi$ distribution of the neat-\wat~system remains isotropic.
+The intensity of the distribution (darker blue) also indicates that the \suldiox~spends little time in this region, as it is quickly adsorbed further into the water bulk.
 
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[scale=1.0]{images/so2transitsaturatedangles.png}
-		\caption{}
-		\label{fig:so2-transit-saturated-angles}
-	\end{center}
-\end{figure}
+	%Just below the water surface as the distance moves into negative values, both systems show clear orientational preferences.
+In both systems the bisector orientation clusters such that $\cos(\theta)<0.5$, with most of the distribution intensity around $\cos(\theta)=1$.
+This suggests that \suldiox~in a water surface orients with the sulfur pointing into the water bulk, and the oxygens pointing towards the surface water molecules.
+
+	%Differences occurr between the system orientation profiles once the \suldiox~has moved further than 10\angs below the surface.
+While both $\theta$ and $\phi$ hold the same trend from just under the surface through to the bulk of the slab in the saturated system, the orientation profile becomes much more isotropic past 10\angs under the surface of the neat-\wat~system.
+The orientation effect is shallower under the neat-\wat~surface.
+However, the orientation distribution peaks most intensely, and more tightly clustered with the \suldiox~bisector aligning with the surface normal, just underneath the neat-\wat~surface than the saturated one.
+The distributions thus show an orienting force extending deeper into the surface of the saturated system than the neat-\wat~system.
+In terms of how deeply the \suldiox~molecule is oriented, the interfacial region of the saturated system is larger and extends further into the aqueous bulk.

### 2011-02-24 clean.sh

@@ -1,3 +1,5 @@
 rm *.aux
 rm *.log
 rm *.pdf
+rm *.blg
+rm *.bbl

### 2011-02-24 TransitOrientation.tex
Added new transit orientation plots, and the Tarbuck references. About to work on the transit orientation text.
New transit images for orientation using new normalization scheme for 2d histograms

@@
 
-	The orientation of a \suldiox~molecule adsorbing to an aqueous surface was determined through an analysis process to find the angles $\theta$ and $\phi$ (figure \ref{fig:water-angles}).
-For each timestep of the SMD simulations the orientation angles were determined for the transiting \suldiox~as it was pulled into the water slab from the gas phase, both in the neat-water and saturated slab systems.
-The angle cosine values were collected for the 50 simulations for both systems, and the binned with the position of the \suldiox~relative to the water slab surface location.
-The resulting 2-dimensional histograms shown in figure \ref{fig:so2-transit-angles} show the orientation trend for the \suldiox~in the gas phase (above 0\angs), near the aqueous surface (near 0\angs), and after entering the water bulk (below 0\angs).
-Only the surface into which the transiting \suldiox~was steered is shown for the saturated system.
+	The orientation of a \suldiox~molecule during adsorption to, and absorbtion in an aqueous surface was determined through an analysis of the transit SMD simulations.
+The angles $\theta$ and $\phi$ (figure \ref{fig:water-angles}) were calculated for each timestep of the SMD simulations as the transiting \suldiox~was pulled into the water slab from the gas phase, both in the neat-water and saturated slab systems.
+The angle cosine values were collected for the 50 simulations of both systems for each distance from the water surface, resulting in the 2-dimensional histograms shown in figure \ref{fig:so2-transit-angles}.
+Only the surface into which the transiting \suldiox~was steered into is shown for the saturated system's analysis.
 
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/so2-angles/so2-angles-transit.png}
+		\includegraphics[scale=1.0]{images/transit-so2-angles/so2-angles-transit.png}
 		\caption{\suldiox~molecular orientation distributions during SMD transit simulations into an aqueous slab.
@@
 
-	%From its starting position 20\angs above the water surface, until the \suldiox~moves to within 5\angs of the surface of both systems, the orientation is isotropic in both $\theta$ and $\phi$.
+	From its starting position 20\angs above the water surface, until the \suldiox~moves to within 10\angs of the surface of both systems, the orientation is isotropic in $\phi$.
+The trend for the bisector angle, $\theta$, upon approaching the surface is to become more perpendicular to the surface with the \suldiox~sulfur pointing into the water phase.
+At the point when the \suldiox~reaches the surface location (distance$=0$), the bisector is mostly perpendicular in both the neat-water and saturated systems.
+In the distribution of $\phi$, both systems show isotropy in the orientation
+
   Upon approaching the saturated surface at approx.

### 2011-02-24 TransitOrientation.tex
added bit about the normalization for 2d histogrms... about to start more on the transit orientation of the so2
rewrote the so2 transit orientation section with the new images

@@
 
-	The orientation of a \suldiox~molecule during adsorption to, and absorbtion in an aqueous surface was determined through an analysis of the transit SMD simulations.
+	The orientation of a \suldiox~molecule throughout the aqueous adsorption process was monitored during the transit SMD simulations.
 The angles $\theta$ and $\phi$ (figure \ref{fig:water-angles}) were calculated for each timestep of the SMD simulations as the transiting \suldiox~was pulled into the water slab from the gas phase, both in the neat-water and saturated slab systems.
@@
 Only the surface into which the transiting \suldiox~was steered into is shown for the saturated system's analysis.
+In all the data sets, the populations of the angle histograms at each distance from the surface were normalized to aid in comparison of regions with differing \suldiox~residence times.
 
@@
 The distributions show angle cosines plotted against the distance of the \suldiox~to the location of the water surface.
-The data sets were averaged at each distance for the respective 50 simulations.}
+The data sets were averaged at each distance for the respective 50 simulations, and every distribution at each distance was normalized such that each histogram has a max population of 1 for purposes of comparison.}
 		\label{fig:so2-transit-angles}
@@
 
-	From its starting position 20\angs above the water surface, until the \suldiox~moves to within 10\angs of the surface of both systems, the orientation is isotropic in $\phi$.
-The trend for the bisector angle, $\theta$, upon approaching the surface is to become more perpendicular to the surface with the \suldiox~sulfur pointing into the water phase.
-At the point when the \suldiox~reaches the surface location (distance$=0$), the bisector is mostly perpendicular in both the neat-water and saturated systems.
-In the distribution of $\phi$, both systems show isotropy in the orientation
+	From its starting position 20\angs above the water surface, until the \suldiox~moves to within 10\angs of the surface of both systems, the orientation is isotropic in $\theta$ and $\phi$.
+The bisector angle $\theta$, near to the surface becomes more perpendicular ($\cos(\theta)\approx1$) with the \suldiox~sulfur pointing into the water phase.
+At the point when the \suldiox~reaches the surface location (distance$=0$), the bisector is perpendicular in both the neat-water and saturated systems.
+There are differences between the neat-water and saturated systems, however, in the onset points of the orientational preferences.
 
-  Upon approaching the saturated surface at approx.
+  % talk about theta differences
+  The $\theta$ distribution clearly shows that the \suldiox~is isotropically oriented in the gas-phase and then orients with its bisector perpendicular to the water surface within a certain distance above the water phase.
+On approach to the neat-water surface, that transition occurs once just above 10\angs from the surface, and again at 5\angs from the surface and below.
+The onset of orientation at 10\angs appears to be an artifact of the small data set and the normalization scheme, as the isotropic behavior resumes between 5-10\angs.
+Below 5\angs, however, the orientation persists and the \suldiox~begins to interact with the topmost surface waters, and orients accordingly.
+In the saturated system, the same trend occurs, however the onset of the perpendicular orientation begins at approx.
+8\angs above the water surface.
+The layer of adsorbed \suldiox~already present in the saturated system most likely interacts with the transiting \suldiox~molecule.
+Also, topmost water molecules from the surface move up to a few\angs within the \suldiox~layer and interact with the transiting \suldiox~further from the surface than those in the neat-water system.
+
+  % talk about phi differences
+  With a mostly perpendicular bisector angle, it is expected that the values of $\phi$ would be isotropic relative to the reference axis.
+This is the case in both systems with one exception in the neat-water system.
+At the water surface and just below from 0\angs to 5\angs in the water phase, the $\theta$ profile broadens, and a clear trend appears in the profile of $\phi$.
+This corresponds to a \suldiox~orienting more flat to the surface (inclined up to 60 degrees from the surface normal).
+However, below this 5\angs region, the \suldiox~resumes an isotropic orientation as it becomes fully solvated by bulk waters.
+
+  %Upon approaching the saturated surface at approx.
 5\angs, the \suldiox~$\phi$ orientation distribution begins to cluster near $\cos(\phi)=1$, with the $\theta$ distribution remaining more isotropic, but with a trend of $\cos(\theta)>0$.

### 2011-03-17 Introduction.tex
New introduction!

\section{Introduction}

Interactions between sulfur species and aqueous systems are the current topic of an active area of research. Sulfur dioxide remains an important industrial product, and also enters the environment naturally through terrestrial processes. \suldiox~acts as a major component of atmospheric pollution and is a precursor to acid rain formation, and cloud nucleation. Its high solubility in water makes \suldiox~ an integral compound in many atmospheric reactions and surface reaction mechanisms. We seek a fuller understanding of the interactions between \suldiox~and water interfaces because of the role \suldiox~takes in atmospheric particle and aerosol chemistry.

Our recent experimental works provided insight into the adsorption and reaction of gases in atmospheric aerosols. We concluded that an \suldiox~surface complex forms with water when an aqueous surface is exposed to \suldiox~gas. A computational study by Dang et al.\cite{Baer2010} then made a series of predictions of the specific nature of the surface complex through classical and ab initio simulations. That work probed the cause of changes in the surface water OH vibrational spectra, and developed a detailed picture of the nature of the \suldiox~surface complex with water.

Our latest experimental work on \suldiox~behavior further expands our understanding of \suldiox~in atmospheric processes.\cite{Ota2011} A temperature study showed that atmospherically relevant cold temperatures greatly enhance the binding of gaseous \suldiox~to water interfaces, and facilitates the reaction with water to produce dissolved sulfur species. However, the surface binding of unreacted \suldiox~complexes was found to be a completely reversible process. The same study also conluded that low pH aqueous solutions, as often found in atmospheric aerosol composition, inhibit the reaction of \suldiox, but does not affect the binding of the gaseous molecules. The analysis of the resultant experimental spectra suggested that the surface waters reorient in response to the presence of gaseous \suldiox~complexes, with the interfacial water ``free-OH'' bonds orienting more perpendicularly to the plane of the aqueous surface. Because of the geometry of the experiment the \suldiox~modes were not possible to analyze.

In this study we look at the orientational depth profiles of water and \suldiox~during, and after, the adsorption process. Spectroscopic experiments can not probe depth or orientation profiles of surface species in the same detailed manner as computational simulations. Thus we hope to complement the experimental results of previous studies. We determine the effect of introducing \suldiox~to a \wat~surface on both the \suldiox~and \wat~molecules using classical molecular dynamics simulations. The results of this study are necessary to understand the behavior of adsorbing gaseous complexes, and we help verify the ascertions of our recent experiments. Along with other ongoing studies, we develop a more complete and detailed picture of gaseous adsorption and complex formation on aqueous surface.

### 2011-03-18 Background.tex
Writing about the density distro/profile, and the way to claculate the surface layer

@@
 
-	The results of analyses performed in this study are related to the location of the liquid water surface.
-A surface, however, is a dynamic boundary that is neither flat or stationary.
-Several previous studies have used the technique of fitting a lineshape to the density profile of the water or solute to extract interfacial shape and location parameters.\cite{Shamay2010,Wick2006c,Chowdhary2006} Hyperbolic tangent functions have been used to extract fitting parameters, and values for the ``Gibb's dividing surface'' location, and interfacial width have thus been determined.\cite{Matsumoto1988} However, in long simulations the location and shape of the interface will change, and the motion of surface waters alters the interfacial width at any given timestep.
+	The results of analyses performed in this study are related to the location of the liquid water surface to relate depth profiles of the orientation distributions.
+A liquid surface, however, is a dynamic boundary that is neither flat nor stationary.
+To define a water surface location, several previous studies have used the technique of fitting a lineshape to the density profile of the water, and extracting interfacial shape and location parameters.\cite{Shamay2010,Wick2006c,Chowdhary2006} Hyperbolic tangent functions have been used often, and values for the ``Gibb's dividing surface'' location, and interfacial width have thus been determined.\cite{Matsumoto1988} However, in long simulations the location and shape of the interface changes, and the motion of surface waters alters the interfacial width at any given timestep.
 Thus, the averaged density profile fitting will capture average widths and locations, not instantaneous values.
-Similarly, the averaged values of location and width will obscure information about any drift or deformation the surface undergoes.
-In this work we approach the problem of surface loation by calculating a reference location for each timestep by averaging the long-axis positions of the waters contained in the topmost monolayer.
+Similarly, the averaged values of location and width will obscure information about any drift or deformations the surface undergoes.
+In this work we approach the problem of defining surface location by calculating a reference location for each timestep.
+We perform an averaging of the positions of the waters contained in the topmost monolayer.
 This provides us with a consistent reference point in the simulations to which we relate our analyses.
@@
 
-  The location of the water surface was thus determined by the averaged positions of a set number of the topmost waters in the simulated aqueous slabs.
-The number of waters counted was decided by taking a few issues into account.
+  The location of the water surface was determined by the averaged positions of a set number of the outter-most waters in the simulated aqueous slabs.
+The number of waters included in the averaging was decided by taking a few issues into account.
 First, a visual inspection of the top 30x30 \angs$^2$ cross-section of the water slab over several timesteps indicated between 65-75 waters established a full monolayer.
-Alternatively, assuming a spherical model of water with a radius of 2.4\angs (the typical length of a bulk-water hydrogen bond), two layers of hexagonally tight-packed spheres would yield a maximum of 52 molecules.
-Increasing the number of waters used in calculating the surface location diminishes the effects of the few waters that briefly rise above the surface into the gas phase, stabilizing both the surface position and thickness.
+Alternatively, assuming a spherical model of water with a radius of 2.2\angs (the typical length of a bulk-water hydrogen bond), two layers of hexagonally tight-packed spheres would yield a similar number of molecules.
+Increasing the number of waters used in calculating the surface location diminishes the effects of the few waters that briefly rise above the surface into the gas phase, stabilizing both the surface position and thickness values.
 Taking the close-packed model as a minimum number of waters fit into a flat surface, the authors decided to use 70 water molecules for calculating surface location during the simulation analyses.
+
+Figure \ref{fig:density-flaw} demonstrates the problem of surface location drift during a simulation.
+The figure shows a density profile of water and \suldiox~over the course of one of the 10 ns trajectories used in this work.
+The water density profile (dashed line) is produced by averaging the density profile of each timestep in the simulation.
+The water profile was then fit with a $\tanh$ function lineshape to extract the position and width parameters of the water surface.
+The fitted Gibbs dividing surface has a ``90-10'' width of 8.28 $\AA$.
+A bulge in the gas-phase (left) side of the water profile is indicative of the drift of the surface waters, or the entire slab over the course of the trajectory.
+Thus the calculated location and width using the fitting routine are not accurate over long trajectories.
+
+The new procedure described above for finding the outter monolayer location was implemented, and the distribution of that location is also plotted in figure \ref{fig:density-flaw} (blue).
+It is apparent from the surface location distribution that the water slab drifts over the 10 ns, spanning approx.
+12 $\AA$.
+However, the maximum standard deviation of the positions of the waters comprising the surface layer at each timestep is only 1.85 $\AA$.
 
@@
 
-\subsection{Hydrate Complex Analysis}
+%\subsection{Hydrate Complex Analysis}
 
-	During the adsorption process of a gaseous \suldiox~moving in towards a water surface hydrate complexes are formed as the waters bind to the molecule.
+%During the adsorption process of a gaseous \suldiox~moving in towards a water surface hydrate complexes are formed as the waters bind to the molecule.
 Computationally, there are various methods to determine hydrate complex formation.
@@
 
-	A simple count of hydrogen bonds is telling of the interactions and environments of an adsorbing \suldiox.
+%A simple count of hydrogen bonds is telling of the interactions and environments of an adsorbing \suldiox.
 However, a more sophisticated analysis is needed to fully characterize the geometry of any complexes that form.
@@
 
-	We present here a technique for determining bonding cycles in a hydrate complex, but this technique is general enough to be applied to any graph representation of simple molecular systems.
+%We present here a technique for determining bonding cycles in a hydrate complex, but this technique is general enough to be applied to any graph representation of simple molecular systems.
 The first step in the determination of cycles is translating the atoms (vertices) and bonds (edges) to a graph.
@@
 
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[scale=1.0]{images/Cycles/so2complexfigure.png}
-		\caption{A \suldiox~and its nearest neighbors are represented as a graph of vertices and edges.
+%\begin{figure}[h!]
+%\begin{center}
+%\includegraphics[scale=1.0]{images/Cycles/so2complexfigure.png}
+%\caption{A \suldiox~and its nearest neighbors are represented as a graph of vertices and edges.
 The bonds are all given equal weights and the graph is undirected.
@@
 This cycle is composed of 8 atoms, and 3 molecules}
-		\label{fig:so2-complex-graph}
-	\end{center}
-\end{figure}
+%\label{fig:so2-complex-graph}
+%\end{center}
+%\end{figure}

### 2011-03-19 Background.tex
wrote the conclusions, added some material in the density results.

@@
 Thus the calculated location and width using the fitting routine are not accurate over long trajectories.
+
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[scale=1.0]{images/density/Density+SurfaceLocation-flawed.png}
+		\caption{Density profiles of \wat~(dashed) and\suldiox~(red) from a 10 ns simulation of the neat-water system with a single sulfur dioxide.
+The \wat~density profile was fit using a hyperbolic tangent function to extract location and width data.
+The location of the outter \wat~monolayer was calculated, and the corresponding location distribution (blue) is also plotted.
+Long simulations of liquid slabs result in drift of the slab location, and the consequent broadening of the water density profile if the surface location is not determined repeatedly throughout the simulation.}
+		\label{fig:density-flaw}
+	\end{center}
+\end{figure}
 

### 2011-03-19 Conclusions.tex
wrote the conclusions, added some material in the density results.

\section {Conclusions}

We have presented the results of several classical simulations of \suldiox~interacting with an aqueous surface. We set out to find the orientational behavior of \wat~and \suldiox~through depth-profiling of molecular orientational distributions. In previous experimental studies we posited that water reorients in the presence of \suldiox, and this behavior is one of the steps in formation of gaseous hydrate complexes on water's surface. Spectroscopic techniques can not precisely define a depth profile or orientational distribution of adsorbed species, so we have used molecular dynamics simulations to gain that information.

Our simulations recreated the experimental setup of a recent study by our group,\ref{Ota2011} by introducing a concentrated gas phase of \suldiox~to a water surface. It was found that gaseous \suldiox~quickly adsorbs to the water surface, and continues to bind until a complete surface coverage is reached. Surface waters reorient in the presence of adsorbed \suldiox. The waters at the interface of a neat-water surface tend to lay flatter to the surface than when a saturating layer of \suldiox~is present. The waters interacting with the layer of adsorbed \suldiox~orient more perpendicularly to the interface, and further expose their ``free-OH'' uncoupled bonds for interactions with \suldiox, and hydrate complex formation.

The surface behavior of the \suldiox~molecules was also determined. Experimental surface spectroscopy on the \suldiox~vibrational spectra is not possible because of the nature of the experimental geometry. The spectra of the surface \suldiox~vibrational modes are still unreported. However, through the simulations we are able to characterize the orientational behavior of \suldiox~during and after adsorption. Our equilibrium simulations show that a single \suldiox~molecule, representing a low concentration, has a high surface affinity. This was previously reported by Dang.\ref{Baer2010} At a high \suldiox~concentration, the unreacted \suldiox~moves further out of the water phase as more molecules bind on the surface. The orientation of \suldiox~on the water surface was found to be similar for both low and high concentrations. Those \suldiox~molecules at or below the surface strongly orient with the sulfur atom pointed in towards the water bulk, and the oxygen atoms out towards the gas phase. The \suldiox~slightly above the water surface loses the orientational preference within 10 $\AA$ and those further from the water are more isotropically oriented.

Steered molecular dynamics simulations were used to model the behavior of an adsorbing \suldiox~as it moves from the gas phase above the water down through the surface and into the bulk. The results for the transit through the interface show that in both systems of low and high \suldiox~concentration an adsorbing \suldiox has very similar orientation to those already bound to the water surface. As the \suldiox~approaches the water surface the first contact with water causes it to reorient. Within 5 $\AA$ of the surface the \suldiox~is mostly oriented with its sulfur towards the water phase. The \suldiox~pulled further into the water bulk retains its orientation until it is past the interfacial region and then isotropically orients with the bulk water.

This study is one of several needed to characterize \suldiox~adsorption and behavior on aqueous surfaces. The atmospheric chemistry driving many important reactions occurs on aerosol and small particle surfaces. To further our understanding of gaseous adsorption we plan to report further simulation results of temperature and chemical constituent effects on atmospherically relevant water surfaces.

### 2011-03-19 DensityResults.tex
wrote the conclusions, added some material in the density results.

@@
 In the neat-water \suldiox~distribution, $\mu=0.781$, $A=0.109$, and $\sigma=1.684$\angs.
-Thus, the \suldiox~tends to locate just beneath the surface when in a low-concentration water environment, and it did not fall further into the bulk or move out of the surface into the gas phase.
+Thus, the \suldiox~has a surface affinity when in a low-concentration water environment, and it does not fall further into the bulk or move out of the surface into the gas phase.
 This surface active behavior was found experimentally by a previous SFG study by our group,\cite{Tarbuck2005,Tarbuck2006} and also verified by the computational simulations and spectral calculations of Dang et al.\cite{Baer2010} It was concluded that a surface \suldiox~hydrate complex forms and modifies the structure of water above the Gibbs dividing surface (GDS) shifting \wat~free-OH oscillators in towards the bulk, creating a stronger hydrogen-bonding network.
@@
 The center of this distribution is further into the gas phase than for the single-\suldiox~molecule surface.
-The accumulation of \suldiox~in this classical simulation coincides with the experimental conclusions of a surface of hydrate-complexed \suldiox, but without breakble bonds in the simulation the chemistry that may pull \suldiox~into the bulk water phase is absent.
+The accumulation of \suldiox~in this classical simulation coincides with the experimental conclusions of \suldiox~surface hydrate complex formation.
+However, without simulating breakble bonds the chemistry that may pull \suldiox~into the bulk water phase, and the subsequent reaction to form ions is absent.
 We have confidence in the surface complexation that is exhibited, the effect of the \suldiox~on the surface water molecules, and the accumulation of \suldiox at the interface.
-However, we do not yet conclude that the specific locations of \suldiox~on top of the water surface are as accurate as those found in the DFT ab initio calculations of Dang et al.\cite{Baer2010}
+However, we do not yet conclude that the specific locations of \suldiox~on top of the water surface are as accurate as those found in the DFT ab initio calculations of Dang et al.\cite{Baer2010} As was pointed out in their computations, the classical potential does not reproduce the bonding of the first hydration shell around \suldiox, and so conclusions about specific complex formation or geometries are best made from the ab initio simulation results reported.
 

### 2011-03-21 Background.tex
Finalized changes to beta draft. Ready to hand over for a read-through.

@@
 
-	The results of analyses performed in this study are related to the location of the liquid water surface to relate depth profiles of the orientation distributions.
+	In this study we create orientational depth-profiles of \suldiox~and \wat.
+The reference point used for depth calculations is the location of the water surface.
 A liquid surface, however, is a dynamic boundary that is neither flat nor stationary.
-To define a water surface location, several previous studies have used the technique of fitting a lineshape to the density profile of the water, and extracting interfacial shape and location parameters.\cite{Shamay2010,Wick2006c,Chowdhary2006} Hyperbolic tangent functions have been used often, and values for the ``Gibb's dividing surface'' location, and interfacial width have thus been determined.\cite{Matsumoto1988} However, in long simulations the location and shape of the interface changes, and the motion of surface waters alters the interfacial width at any given timestep.
-Thus, the averaged density profile fitting will capture average widths and locations, not instantaneous values.
+To define a water surface location, several previous studies have used the technique of fitting a lineshape to the averaged density profile of the water, and extracting interfacial shape and location parameters.\cite{Shamay2010,Wick2006c,Chowdhary2006} Hyperbolic tangent functions have been used often, and values for the ``Gibb's dividing surface'' location, and interfacial width have thus been determined.\cite{Matsumoto1988} However, in long simulations the location and shape of the interface changes, and the motion of surface waters alters the interfacial width at any given timestep.
+Thus, the density profile fitting will capture averaged widths and locations, not instantaneous values.
 Similarly, the averaged values of location and width will obscure information about any drift or deformations the surface undergoes.
-In this work we approach the problem of defining surface location by calculating a reference location for each timestep.
-We perform an averaging of the positions of the waters contained in the topmost monolayer.
-This provides us with a consistent reference point in the simulations to which we relate our analyses.
-This method provides an intuitive point of reference, but does not increase the computational burden of the analyses.
+In this work we approach the problem of defining surface location by calculating a reference location at each timestep by averaging the positions of the waters contained in the topmost monolayer.
+This provides us with a consistent and intuitive reference point in the simulations to which we relate our analyses, but does not increase the computational burden.
 
@@
 The number of waters included in the averaging was decided by taking a few issues into account.
-First, a visual inspection of the top 30x30 \angs$^2$ cross-section of the water slab over several timesteps indicated between 65-75 waters established a full monolayer.
-Alternatively, assuming a spherical model of water with a radius of 2.2\angs (the typical length of a bulk-water hydrogen bond), two layers of hexagonally tight-packed spheres would yield a similar number of molecules.
+First, a visual inspection of the top cross-section of the water slab over several timesteps indicated between 65-75 waters established a full monolayer.
+Alternatively, assuming a spherical model of water with a radius of 2.2\angs~(the typical length of a bulk-water hydrogen bond), two layers of hexagonally tight-packed spheres would yield a similar number of molecules to the visual inspection.
 Increasing the number of waters used in calculating the surface location diminishes the effects of the few waters that briefly rise above the surface into the gas phase, stabilizing both the surface position and thickness values.
-Taking the close-packed model as a minimum number of waters fit into a flat surface, the authors decided to use 70 water molecules for calculating surface location during the simulation analyses.
+Taking the close-packed model as a maximum number of waters fit into a flat surface, the authors decided to use 70 water molecules for calculating surface location during the simulation analyses.
 
@@
 The figure shows a density profile of water and \suldiox~over the course of one of the 10 ns trajectories used in this work.
-The water density profile (dashed line) is produced by averaging the density profile of each timestep in the simulation.
-The water profile was then fit with a $\tanh$ function lineshape to extract the position and width parameters of the water surface.
+The water density profile (dashed line) is produced by averaging the instantaneous density profile at each timestep in the simulation.
+The water profile was then fit to a $\tanh$ function (black) to extract the position and width parameters of the water surface.
 The fitted Gibbs dividing surface has a ``90-10'' width of 8.28 $\AA$.
-A bulge in the gas-phase (left) side of the water profile is indicative of the drift of the surface waters, or the entire slab over the course of the trajectory.
-Thus the calculated location and width using the fitting routine are not accurate over long trajectories.
+A bulge in the gas-phase (left) side of the water profile is indicative of the drift of the water slab over the course of the trajectory.
+Thus the calculated location and width from the $\tanh$ line fit are not accurate over long trajectories for defining a stationary reference point.
+All depth locations in our analyses are calculated relative to the surface location of each timestep of the trajectories.
 
@@
 The two molecules studied are similarly shaped with a C2$_v$ axis along their bisectors, and a molecular plane defined by three atoms.
-For analysis, a body-fixed frame is defined for \wat~and \suldiox~as shown in figure \ref{fig:molecular-frame}.
-In each analysis a space-fixed reference axis is used that corresponds to the long axis of the system's periodic cell (normal to the plane of the water surface) unless otherwise noted.
+A body-fixed frame is defined for both \wat~and \suldiox~as shown in figure \ref{fig:molecular-frame}.
+In each analysis a space-fixed reference axis is used that corresponds to the long axis of the system's periodic cell normal to the plane of the water surface.
 The orientational analyses presented herein focus on two angles used to define molecular orientation in space.
@@
 Because $\phi=-\phi$, the value is reported within the range of $[0,\frac \pi 2]$.
+	However, with $\cos(\theta)$ close to 1 or -1 the symmetry of the plane of the surface results in an isotropic distribution in $\cos(\phi)$.
 Figure \ref{fig:water-angles} shows the angle definitions relating the body-fixed axes to a system-fixed reference axis.
@@
 		\includegraphics[scale=1.0]{images/angle-cartoons/molecularframesmall.png}
-		\caption{The body-fixed axes are defined with the x-z plane formed by the three atoms, and the z-axis aligned to the molecular bisector.
-This sets the y-axis normal to the molecular plane, and one of the bonds in the positive x direction.}
+		\caption{Molecular body-fixed axes are defined with the x-z plane formed by the three atoms, and the z-axis aligned to the molecular bisector.
+The y-axis is normal to the molecular plane, and one of the bonds points in the positive x direction.}
 		\label{fig:molecular-frame}

### 2011-03-21 Conclusions.tex
Finalized changes to beta draft. Ready to hand over for a read-through.

@@
 
-Our simulations recreated the experimental setup of a recent study by our group,\ref{Ota2011} by introducing a concentrated gas phase of \suldiox~to a water surface.
+Our simulations recreated the experimental setup of a recent study by our group,\cite{Ota2011} by introducing a concentrated gas phase of \suldiox~to a water surface.
 It was found that gaseous \suldiox~quickly adsorbs to the water surface, and continues to bind until a complete surface coverage is reached.
@@
 The surface behavior of the \suldiox~molecules was also determined.
-Experimental surface spectroscopy on the \suldiox~vibrational spectra is not possible because of the nature of the experimental geometry.
-The spectra of the surface \suldiox~vibrational modes are still unreported.
-However, through the simulations we are able to characterize the orientational behavior of \suldiox~during and after adsorption.
+Experimental surface spectroscopy on the \suldiox~vibrational spectra is not possible because of the nature of the experimental geometry, so the spectra of the surface \suldiox~vibrational modes are still unreported.
+However, through these simulations we are able to characterize the orientational behavior of \suldiox~during and after adsorption.
 Our equilibrium simulations show that a single \suldiox~molecule, representing a low concentration, has a high surface affinity.
-This was previously reported by Dang.\ref{Baer2010} At a high \suldiox~concentration, the unreacted \suldiox~moves further out of the water phase as more molecules bind on the surface.
+This was previously reported by Dang.\cite{Baer2010} At a high \suldiox~concentration, the unreacted \suldiox~moves further out of the water phase as more molecules bind on the surface.
 The orientation of \suldiox~on the water surface was found to be similar for both low and high concentrations.
@@
 
-This study is one of several needed to characterize \suldiox~adsorption and behavior on aqueous surfaces.
+This study is one of several aiming to characterize \suldiox~adsorption and behavior on aqueous surfaces.
 The atmospheric chemistry driving many important reactions occurs on aerosol and small particle surfaces.
-To further our understanding of gaseous adsorption we plan to report further simulation results of temperature and chemical constituent effects on atmospherically relevant water surfaces.
+To gain a better understanding of gaseous adsorption we plan to report further computational simulation and experimental results of temperature and chemical constituent effects on atmospherically relevant water surfaces.

### 2011-03-21 DensityResults.tex
Finalized changes to beta draft. Ready to hand over for a read-through.

@@
 One measure of surface activity is the spatial distribution of molecules around the interfacial region.
-The density distribution of both \wat~and\suldiox~were calculated for the equilibrium MD simulations, and the results are presented in figure \ref{fig:density} for both the neat-water and saturated simulations.
+The density distribution of both \wat~and \suldiox~were calculated for the equilibrium MD simulations, and the results are presented in figure \ref{fig:density} for both the neat-water and saturated simulations.
 The \suldiox~in the neat-water system remained at the top slab surface.
@@
 
-The density distributions of the \suldiox~were fit to gaussian lineshapes as shown in equation \ref{eq:gaussian} using the common definitions of the variables $A$, $\mu$, and $\sigma$, and given as a function of the distance to the surface, $z$.
+The density distributions of the \suldiox~were fit to gaussian lineshapes as shown in equation \ref{eq:gaussian} using the common definitions of the variables $A$, $\mu$, and $\sigma$, and given as a function of the distance to the top surface, $z$.
 
@@
 The bottom surface values are $\mu_{bottom}=-33.846$, $A_{bottom}=6.618$, and $\sigma_{bottom}=3.151$\angs.
-The added concentration of \suldiox~creates a layer of molecules adsorbed on top of the water surface.
+The added concentration of \suldiox~creates a layer of molecules bound to the top of the water surface.
 The center of this distribution is further into the gas phase than for the single-\suldiox~molecule surface.

### 2011-03-21 EquilibratedOrientation.tex
Finalized changes to beta draft. Ready to hand over for a read-through.

@@
 
-Molecular orientation analyses were performed to characterize the net molecular orientation of \wat~and \suldiox~molecules in the different locations near and within the water surface.
-Histograms were created to show the angles $\theta$ and $\phi$ relative to positions from the top surface of the aqueous slab.
-Both \wat~and \suldiox~orientations are considered for both the neat-\wat~and saturated systems.
+Molecular orientation analyses were performed to characterize the net molecular orientation of \wat~and \suldiox~molecules in the different locations of the water surface.
+Histograms were created to show depth-profiles of the angles $\theta$ and $\phi$ in the neat-water and saturated equilibrium systems.
 
@@
 
-The results of the analyses for \wat~are shown in figure \ref{fig:water-orientation} for both the neat-water and saturated systems during the equilibrated MD simulations.
-In both systems an orientational preference is found at the slab surfaces where the water is in contact with the gas phase, or the surface \suldiox~molecules.
-The histograms are arranged with the plots of $\theta$ on the left and $\phi$ on the right columns, and the results for the neat-water and saturated systems are in the top and bottom rows, respectively.
-The distance axis was shifted such that the top surface is located at 0\angs, and the bottom surface is approx.
+The orientation profiles for \wat~are shown in figure \ref{fig:water-orientation} for both the neat-water (top) and saturated (bottom) systems during the equilibrium MD simulations.
+In both systems an orientational preference is found at the slab surfaces where the water is furthest towards the gas phase.
+The histograms are arranged with the plots of $\cos(\theta)$ on the left and $\cos(\phi)$ on the right columns.
+The top surface is located at 0\angs, and the bottom surface is approx.
 30\angs below.
-It is also noted that $\theta$ values of the bottom surface are inverted with respect to the top surface and should mirror those results.
+It is also noted that $\theta$ values of the bottom surface are inverted with respect to the top surface, mirroring those results.
+
 %The bisector tilt concentrates around $\cos(\theta)=0$ within the first few\angs of the surface, and then the distribution becomes isotropic further into the water bulk.
@@
 The histograms show overall similarities for both systems in their shapes and intensities.
-The presence of a saturated layer of adsorbed \suldiox~molecules alters the water orientation, but not necessarily the location depth of the interface.
-In both $\theta$ distributions, the orientations of waters in the bulk region are isotropic until 5\angs below the surface location.
-At 5\angs below the surface the water molecules begin to orient with their bisectors within the plane of the interface (perpendicular to the surface normal, $\cos(\theta)\approx 0$).
+The presence of a saturated layer of adsorbed \suldiox~molecules alters the water orientation, but not necessarily the orientational depth of the interface.
+In both $\theta$ distributions, the orientations of waters in the bulk region are isotropic until 5\angs~below the surface location.
+At 5\angs~below the surface the water molecules begin to orient with their bisectors within the plane of the interface (perpendicular to the surface normal, $\cos(\theta)\approx 0$).
 The corresponding location in the plot of $\phi$ shows that those molecules are also mostly flat to the surface ($\cos(\phi)\approx 1$).
@@
 The change in $\cos(\theta)$ is not as rapid at the neat-water surface as in the saturated system.
-Furthermore, the noise in the distribution above 5\angs (and below approx -34\angs) in the neat-water system is a result of fewer waters venturing beyond those extents and thus less data far from the surface.
-This is one indication that waters near a layer of adsorbed \suldiox~venture further above the interface and interact with the \suldiox~molecules.
+Furthermore, the noise in the distribution above 5\angs~(and below approx -34\angs) in the neat-water system is a result of fewer waters venturing beyond those extents and thus less data far from the surface.
+This is one indication that waters near a layer of adsorbed \suldiox~venture further above the interface and interact with the adsorbed gas molecules.
 The interactions with neighboring \suldiox~molecules allow the waters above the surface to orient more perpendicular to the interface.
-Recent SFG studies by our group indicate that ***<something goes here>***.
+Our experimental SFG studies also predicted this reorienting behavior due to the \suldiox~interactions with the topmost surface waters.\cite{Ota2011}
 
@@
 Waters on the neat surface lie flatter, whereas the presence of the \suldiox~allows a greater range of ``twist'' for those waters in the plane of the interface.
-The $\phi$ distributions quickly become isotropic above a couple angstroms from the surface as the bisectors orient perpendicularly, and below the surface as the bulk water loses any orientational preference.
+The $\phi$ distributions quickly become isotropic above a couple angstroms from the surface as the bisectors orient more perpendicularly, and below the surface as the bulk water loses any orientational preference.
 
@@
 
-Orientation distributions of the adsorbed \suldiox~molecules were created during the equilibrium simulations for both the neat-\wat~and saturated \suldiox~systems.
+Orientation distributions of the adsorbed \suldiox~molecules were created during the equilibrium simulations for both the neat-water and saturated systems.
 Figure \ref{fig:so2-orientation} shows the distributions of $\cos(\theta)$ and $\cos(\phi)$ (arranged similar to the water orientation distributions plots in figure \ref{fig:water-orientation}).
-The data set for the neat-\wat~system is much smaller as only a single \suldiox~molecule was simulated in the bulk.
+Only a single surface is shown in those plots with \suldiox~data from the second surface averaged with the first.
+The \suldiox~orientation data set for the neat-water system is much smaller as only a single \suldiox~molecule was simulated in the bulk.
 The resulting distribution plots are thus noisier than the corresponding saturated plots, but effective comparisons can still be drawn.
 
-The location of \suldiox~within the water surface is slightly different between the two equilibrium MD systems (see figure \ref{fig:density}).
-The single \suldiox~molecule is generally located just below the 0\angs position of the surface, and the distribution of the concentrated system \suldiox~molecules is centered above the surface location.
-This indicates that in both systems the \suldiox~is highly surface active, but the locations of the peak \suldiox~densities differ.
-The single molecule spends most of its time bound within the top water monolayer, and often traversing below that location by only a few\angs.
-The saturated system's surface molecules sit on top of the monolayer.
 The angular distribution of the single \suldiox~is concentrated primarily in $\cos(\theta)>0$, such that the bisector points out of the water surface.
@@
 		\caption{Molecular orientation distributions for \suldiox~molecules adsorbed to the water slab surface.
-Distributions are shown for $\cos(\theta)$ (left column) and $\cos(\phi)$ (right column) for both the neat-\wat~(top) and high \suldiox~concentration (bottom) systems.
-For both systems the $\theta$ distributions suggest \suldiox~bound to the water surface with the sulfur pointing towards the water slab, and the oxygens pointing to the gas phase.
+Distributions are shown for $\cos(\theta)$ (left column) and $\cos(\phi)$ (right column) for both the neat-water (top) and saturated (bottom) systems.
+For both systems the $\theta$ distributions show \suldiox~bound to the water surface with the sulfur pointing towards the water slab, and the oxygens pointing to the gas phase.
 In this configuration the $\phi$ distribution is isotropic because of the water slab's in-plane symmetry.}

### 2011-03-21 Introduction.tex
Finalized changes to beta draft. Ready to hand over for a read-through.

@@
 Sulfur dioxide remains an important industrial product, and also enters the environment naturally through terrestrial processes.
-\suldiox~acts as a major component of atmospheric pollution and is a precursor to acid rain formation, and cloud nucleation.
+\suldiox~acts as a major component of atmospheric pollution, and is a precursor to acid rain formation, and cloud nucleation.
 Its high solubility in water makes \suldiox~ an integral compound in many atmospheric reactions and surface reaction mechanisms.
-We seek a fuller understanding of the interactions between \suldiox~and water interfaces because of the role \suldiox~takes in atmospheric particle and aerosol chemistry.
+We seek a fuller understanding of the interactions between \suldiox~and water interfaces because of the central role \suldiox~takes in many atmospheric particle and aerosol reactions.
 
-Our recent experimental works provided insight into the adsorption and reaction of gases in atmospheric aerosols.
-We concluded that an \suldiox~surface complex forms with water when an aqueous surface is exposed to \suldiox~gas.
-A computational study by Dang et al.\cite{Baer2010} then made a series of predictions of the specific nature of the surface complex through classical and ab initio simulations.
-That work probed the cause of changes in the surface water OH vibrational spectra, and developed a detailed picture of the nature of the \suldiox~surface complex with water.
+Our recent experimental works provided insight into the adsorption and reaction of gases in atmospheric aerosols.\cite{Ota2011} We concluded that an \suldiox~surface hydrate complex forms when an aqueous surface is exposed to \suldiox~gas.
+A computational study by Dang et al.\cite{Baer2010} then made a series of predictions of the specific nature of the hydrated complex through classical and ab initio simulations.
+That work developed a detailed picture of the nature of the \suldiox~surface complex with water, and related it to the surface water OH vibrational IR spectra.
 
-Our latest experimental work on \suldiox~behavior further expands our understanding of \suldiox~in atmospheric processes.\cite{Ota2011} A temperature study showed that atmospherically relevant cold temperatures greatly enhance the binding of gaseous \suldiox~to water interfaces, and facilitates the reaction with water to produce dissolved sulfur species.
+Our latest experimental work on \suldiox~behavior further expands our understanding of \suldiox~in atmospheric processes.\cite{Ota2011} A temperature study showed that the binding of gaseous \suldiox~to water interfaces at atmospherically relevant cold temperatures is greatly enhanced.
+The increased binding facilitates the reaction with water to produce dissolved sulfur species.
 However, the surface binding of unreacted \suldiox~complexes was found to be a completely reversible process.
-The same study also conluded that low pH aqueous solutions, as often found in atmospheric aerosol composition, inhibit the reaction of \suldiox, but does not affect the binding of the gaseous molecules.
-The analysis of the resultant experimental spectra suggested that the surface waters reorient in response to the presence of gaseous \suldiox~complexes, with the interfacial water ``free-OH'' bonds orienting more perpendicularly to the plane of the aqueous surface.
-Because of the geometry of the experiment the \suldiox~modes were not possible to analyze.
+The same study also showed that low pH aqueous solutions, as often found in atmospheric aerosol composition, inhibit the reaction of \suldiox, but do not affect the surface binding of the gas molecules.
+The resultant experimental spectra suggest that the surface waters reorient in response to the presence of gaseous \suldiox, with the interfacial water ``free-OH'' bonds orienting more perpendicularly to the plane of the aqueous surface.
 
-In this study we look at the orientational depth profiles of water and \suldiox~during, and after, the adsorption process.
+In this study we look at the orientational depth profiles of water and \suldiox~during and after the adsorption process.
 Spectroscopic experiments can not probe depth or orientation profiles of surface species in the same detailed manner as computational simulations.
 Thus we hope to complement the experimental results of previous studies.
-We determine the effect of introducing \suldiox~to a \wat~surface on both the \suldiox~and \wat~molecules using classical molecular dynamics simulations.
-The results of this study are necessary to understand the behavior of adsorbing gaseous complexes, and we help verify the ascertions of our recent experiments.
-Along with other ongoing studies, we develop a more complete and detailed picture of gaseous adsorption and complex formation on aqueous surface.
+We determine the effect of introducing \suldiox~to a \wat~surface, and analyze the orientational response of both the \suldiox~and \wat~molecules using equilibrium and steered (SMD) classical molecular dynamics simulations.
+The results of this study will help to understand the behavior of adsorbing gaseous complexes, and will verify predictions of our recent experiments.
+Along with other ongoing experiments and computations, we develop a more complete picture of gaseous adsorption and complex formation on an aqueous surface.

### 2011-03-21 Methods.tex
Finalized changes to beta draft. Ready to hand over for a read-through.

@@
 
-Molecular dynamics simulations were all performed using the Amber 11 software suite.\cite{Case2010} Polarizable models for the \wat~and \suldiox~molecules were used in the simulations, and have been used previously in studies on interfacial systems because they are known to more accurately reproduce interfacial structure and free energy profiles.\cite{Wick2007,Rivera2006,Dang1998} The \wat~model used is the POL3 model,\cite{Caldwell1995} and for \suldiox we used the model of Dang, et al.
+Molecular dynamics simulations were performed using the Amber 11 software suite.\cite{Case2010} Polarizable models for the \wat~and \suldiox~molecules were used in the simulations, and have been used previously in studies on interfacial systems because they are known to more accurately reproduce interfacial structure and free energy profiles.\cite{Wick2007,Rivera2006,Dang1998} The \wat~model used is the POL3 water model,\cite{Caldwell1995} and for \suldiox~we used the model of Dang, et al.
 that places a single polarizable center on the sulfur atom.\cite{Baer2010}
@@
 The simulations all employed periodic boundaries to create an ``infinite-slab'' geometry.
-After equilibrating the neat-\wat~slabs two types of systems were created by introducing \suldiox: a single-\suldiox~system, herein referred to as the ``neat-water'' system, and a saturated \suldiox~system.
+After equilibrating the neat-\wat~slabs two types of systems were created by introducing \suldiox: a single-\suldiox~system, herein referred to as the ``neat-water'' system, and a saturated \suldiox~system with multiple gaseous \suldiox~molecules.
 
-One system involved the addition of a single \suldiox~molecule either within the bulk of the water slab (for equilibrium MD), or above the slab surface (in the SMD simulations).
+The neat-water simulation involved the addition of a single \suldiox~molecule either within the bulk of the water slab (for equilibrium MD), or above the slab surface (in the SMD simulations).
 The single-\suldiox~system was then evolved for 2 ns to produce an equilibrated starting configuration.
-The second system introduced 22 \suldiox~molecules into the water slab in order to saturate it to a level coinciding with the Henry's law constant for \suldiox~in water.
+The saturated system had 22 \suldiox~molecules in the water slab in order to saturate it to a level coinciding with the Henry's law constant for \suldiox~in water.
 Additionally, 50 \suldiox~molecules were introduced into the gas phase outside of the water slab, on both surfaces, to simulate an added 50 atm of \suldiox~gas pressure.
@@
 One system had a single \suldiox~added to the center of the water box, representing a concentration of 0.06 M.
-The complementary saturated system with 22 \suldiox~molecule in the bulk represents a system of concentration 1.35 M, with an additional 50 \suldiox~in the gas phase to model 50 atm of pressure above the water surface.
+The saturated system with 22 \suldiox~molecules in the bulk represents a system of concentration 1.35 M, with an additional 50 \suldiox~in the gas phase to model 50 atm of pressure above the water surface.
 Both the low and high concentration systems were then evolved for a further 10 ns using a timestep of 0.5fs, with atomic coordinates written every 100 fs.
@@
 % need to add in the specifics about the SMD routine, force constants, calculations, etc.
-A second set of simulations involved an equilibrated water slab as in the surface equilibrated method above.
-However, for both the neat-water and saturated systems, a single \suldiox~was introduced 20\angs above the water slab surface, with the sulfur atom tethered to its initial position.
+A second set of simulations began with an equilibrated water slab as in the surface equilibrated method above.
+However, in both the neat-water and saturated systems equilibrated systems, a single \suldiox~was introduced 20\angs~above the water slab surface, with the sulfur atom tethered to its initial position.
 The systems were evolved for 1 ns, taking coordinate snapshots every 20 ps to create 50 starting points for further simulations.
-Steered molecular dynamics (SMD) were then performed on the 50 system configurations (in both the low and high concentrations) to guide the \suldiox~down towards a tethered water-oxygen 15\angs under the water slab surface by applying a small steering force to the \suldiox-sulfur atom.\cite{Isralewitz2001} The \suldiox~thus passed through the continuum of environments from gas phase to (neat- and saturated) water surface adsorption, and finally absorption into the bulk of the \wat~slab.
+Steered molecular dynamics (SMD) were then performed on the 50 system configurations (in both the neat-water and saturated configurations) to guide the \suldiox~down towards a tethered water 15\angs under the water slab surface by applying a small steering force to the \suldiox-sulfur atom.\cite{Isralewitz2001} The \suldiox~thus passed through the continuum of environments from gas phase to (neat- and saturated) water surface adsorption, and finally absorption into the bulk of the \wat~slab.
 Each of the SMD simulations were performed for a total of 200 ps, using a timestep of 1 fs, and taking snapshots of the system every 25 fs.
-Figure \ref{fig:starting-configurations} illustrates two sample starting configurations for the SMD simulations, showing both the neat-water slab and the saturated-\wat~slab configurations before steering the \suldiox~towards the water bulk.
+Figure \ref{fig:starting-configurations} illustrates two sample starting configurations for the SMD simulations, showing both the neat-water slab and the saturated slab configurations before steering the \suldiox~towards the water bulk.
 
-A separate set of SMD simulations were performed similarly, but with the tethering to one of the \suldiox-oxygens.
+A separate set of SMD simulations were performed with tethering to one of the \suldiox-oxygens.
 This was done to ensure that the orientation of the \suldiox~during the adsorption transit was not an artifact of the choice of atom used for tethering.
@@
 		\caption{Sample starting configurations for the two types of SMD simulations.
-The neat-\wat~slab simulation introduces a single \suldiox~molecule that is then guided into the surface of the water and further into the bulk (left).
-The saturated-\wat~slab simulation begins with different configurations of a water system that has been loaded with \suldiox~to saturate the water phase, and also with a high pressure of \suldiox~gas (right).
+The neat-water slab simulation introduces a single \suldiox~molecule that is then guided into the surface of the water and further into the bulk (left).
+The saturated slab simulation begins with a water system that has been loaded with \suldiox~to saturate the water phase, and also with a high pressure of \suldiox~gas (right).
 The single \suldiox~(shown at the top) is then steered through the surface region saturated with \suldiox~molecules, and into the water bulk.}

### 2011-03-21 TransitOrientation.tex
Finalized changes to beta draft. Ready to hand over for a read-through.

@@
 
-	The orientation of a \suldiox~molecule throughout the aqueous adsorption process was monitored during the transit SMD simulations.
+	The orientation of \suldiox~molecules throughout the aqueous adsorption process was monitored during the transit SMD simulations.
 The angles $\theta$ and $\phi$ (figure \ref{fig:water-angles}) were calculated for each timestep of the SMD simulations as the transiting \suldiox~was pulled into the water slab from the gas phase, both in the neat-water and saturated slab systems.
 The angle cosine values were collected for the 50 simulations of both systems for each distance from the water surface, resulting in the 2-dimensional histograms shown in figure \ref{fig:so2-transit-angles}.
-Only the surface into which the transiting \suldiox~was steered into is shown for the saturated system's analysis.
+Only the surface into which the transiting \suldiox~was steered is used in the analyses.
 In all the data sets, the populations of the angle histograms at each distance from the surface were normalized to aid in comparison of regions with differing \suldiox~residence times.
@@
 
-	From its starting position 20\angs above the water surface, until the \suldiox~moves to within 10\angs of the surface of both systems, the orientation is isotropic in $\theta$ and $\phi$.
+	From its starting position 20\angs~above the water surface, until the \suldiox~moves to within 10\angs~of the surface of both systems, the orientation is isotropic in $\theta$ and $\phi$.
 The bisector angle $\theta$, near to the surface becomes more perpendicular ($\cos(\theta)\approx1$) with the \suldiox~sulfur pointing into the water phase.
-At the point when the \suldiox~reaches the surface location (distance$=0$), the bisector is perpendicular in both the neat-water and saturated systems.
-There are differences between the neat-water and saturated systems, however, in the onset points of the orientational preferences.
+At the point when the \suldiox~reaches the surface location (distance $=0$\angs), the bisector is perpendicular in both the neat-water and saturated systems.
+There are differences between where the orientation transitions from isotropic in the neat-water and saturated systems.
+In the absence of simulated ionic species that form through \suldiox-\wat~chemistry at the surface, it is clear that the adsorbing \suldiox~in the gas phase takes on a preferred orientation to adsorb on a water surface.
 
@@
   The $\theta$ distribution clearly shows that the \suldiox~is isotropically oriented in the gas-phase and then orients with its bisector perpendicular to the water surface within a certain distance above the water phase.
-On approach to the neat-water surface, that transition occurs once just above 10\angs from the surface, and again at 5\angs from the surface and below.
-The onset of orientation at 10\angs appears to be an artifact of the small data set and the normalization scheme, as the isotropic behavior resumes between 5-10\angs.
-Below 5\angs, however, the orientation persists and the \suldiox~begins to interact with the topmost surface waters, and orients accordingly.
+On approach to the neat-water surface, that transition occurs once just above 10\angs~from the surface, and again at 5\angs~from the surface and below.
+The onset of a preferred orientation at 10\angs~occurrs at the cutoff length of the simulated intermolecular bonding interactions.
+Between 5-10\angs~the molecule appears to orient more isotropically.
+Below 5\angs, the orientation persists and the \suldiox~begins to interact with the topmost surface waters, and orients accordingly.
 In the saturated system, the same trend occurs, however the onset of the perpendicular orientation begins at approx.
-8\angs above the water surface.
+8\angs~above the surface location.
 The layer of adsorbed \suldiox~already present in the saturated system most likely interacts with the transiting \suldiox~molecule.
-Also, topmost water molecules from the surface move up to a few\angs within the \suldiox~layer and interact with the transiting \suldiox~further from the surface than those in the neat-water system.
+Also, topmost water molecules from the surface move up to a few\angs~within the \suldiox~layer and interact with the transiting \suldiox~further from the surface than those in the neat-water system.
 
@@
 This is the case in both systems with one exception in the neat-water system.
-At the water surface and just below from 0\angs to 5\angs in the water phase, the $\theta$ profile broadens, and a clear trend appears in the profile of $\phi$.
+At the water surface and just below from 0\angs~to 5\angs~in the water phase, the $\theta$ profile broadens, and a clear trend appears in the profile of $\phi$.
 This corresponds to a \suldiox~orienting more flat to the surface (inclined up to 60 degrees from the surface normal).
-However, below this 5\angs region, the \suldiox~resumes an isotropic orientation as it becomes fully solvated by bulk waters.
+However, below this 5\angs~region, the \suldiox~resumes an isotropic orientation as it becomes fully solvated by bulk waters.
 

### 2011-04-04 Background.tex
Lots of new angle images. Only showing one surface for angle plots instead of both slab surfaces. rewrote the first draft incorporating Steph's comments. Still need an image for the conclusions.

@@
 Similarly, the averaged values of location and width will obscure information about any drift or deformations the surface undergoes.
-In this work we approach the problem of defining surface location by calculating a reference location at each timestep by averaging the positions of the waters contained in the topmost monolayer.
+In this work we define the surface location by calculating a reference location at each timestep by averaging the positions of the waters contained in the topmost monolayer.
 This provides us with a consistent and intuitive reference point in the simulations to which we relate our analyses, but does not increase the computational burden.
@@
 Thus the calculated location and width from the $\tanh$ line fit are not accurate over long trajectories for defining a stationary reference point.
-All depth locations in our analyses are calculated relative to the surface location of each timestep of the trajectories.
+All depth locations in our analyses are calculated relative to the surface location at each timestep of the trajectories.
 
@@
 		\includegraphics[scale=1.0]{images/density/Density+SurfaceLocation-flawed.png}
-		\caption{Density profiles of \wat~(dashed) and\suldiox~(red) from a 10 ns simulation of the neat-water system with a single sulfur dioxide.
+		\caption{Density profiles of \wat~(dashed) and \suldiox~(red) from a 10 ns simulation of the neat-water system with a single sulfur dioxide.
 The \wat~density profile was fit using a hyperbolic tangent function to extract location and width data.
-The location of the outter \wat~monolayer was calculated, and the corresponding location distribution (blue) is also plotted.
+The location of the outer \wat~monolayer was calculated, and the corresponding location distribution (blue) is also plotted.
 Long simulations of liquid slabs result in drift of the slab location, and the consequent broadening of the water density profile if the surface location is not determined repeatedly throughout the simulation.}
@@
 
-The new procedure described above for finding the outter monolayer location was implemented, and the distribution of that location is also plotted in figure \ref{fig:density-flaw} (blue).
+The new procedure described above for finding the outer monolayer location was implemented, and the distribution of that location is also plotted in figure \ref{fig:density-flaw} (blue).
 It is apparent from the surface location distribution that the water slab drifts over the 10 ns, spanning approx.
@@
 The values of $\phi$ fall in the interval $[-\pi,\pi]$.
-Given certain values of $\theta$, $\phi$ provides additional information about whether the molecular orientation is ``flat'' to the surface (e.g.
+For certain values of $\theta$, $\phi$ provides additional information about whether the molecular orientation is ``flat'' to the surface (e.g.
 the plane of the molecule is aligned with the plane of the surface), or if it is perpendicular.
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/angle-cartoons/wateranglessmall.png}
-		\caption{The definition of the angles $\theta$ and $\phi$ used to define molecular orientation of \suldiox~and \wat~relative to a reference axis.
-$\theta$ is the value of the ``tilt'' of the molecular bisector from the reference axis.
-$\phi$ is the ``twist'' angle that indicates how flat the molecule is relative to the interfacial plane, and is formed between the vector normal to the molecular plane (the body-fixed y-axis) and the reference axis.}
+		\includegraphics[scale=1.0]{images/angle-cartoons/molecular-angles.png}
+		\caption{(A) The definition of the angles $\theta$ and $\phi$ used to define molecular orientation of \suldiox~and \wat~relative to a reference axis.
+(B) $\theta$ is the value of the ``tilt'' of the molecular bisector from the reference axis, with values ranging from -1 to +1 indicating different bisector orientations.
+(C) $\phi$ is the ``twist'' angle formed between the vector normal to the molecular plane (the body-fixed y-axis) and the reference axis.
+For certain values of $\theta$, $\phi$ shows how flat the molecule is relative to the interfacial plane.}
 		\label{fig:water-angles}

### 2011-04-04 DensityResults.tex
Lots of new angle images. Only showing one surface for angle plots instead of both slab surfaces. rewrote the first draft incorporating Steph's comments. Still need an image for the conclusions.

@@
 
-One measure of surface activity is the spatial distribution of molecules around the interfacial region.
+One measure of surface activity is the spatial distribution of molecules in the interfacial region.
 The density distribution of both \wat~and \suldiox~were calculated for the equilibrium MD simulations, and the results are presented in figure \ref{fig:density} for both the neat-water and saturated simulations.
 The \suldiox~in the neat-water system remained at the top slab surface.
-While the \suldiox~equilibrated in the saturated system, both in the bulk and gas phases accumulated at both the top and bottom surfaces.
+The \suldiox~in the saturated system accumulated mostly at the surface, but some residual \suldiox~remained in the bulk.
 
-The density distributions of the \suldiox~were fit to gaussian lineshapes as shown in equation \ref{eq:gaussian} using the common definitions of the variables $A$, $\mu$, and $\sigma$, and given as a function of the distance to the top surface, $z$.
+%The density distributions of the \suldiox~were fit to gaussian lineshapes as shown in equation \ref{eq:gaussian} using the common definitions of the variables $A$, $\mu$, and $\sigma$, and given as a function of the distance to the top surface, $z$.
 
-\begin{equation}
-  \rho(z)=\frac{A}{\sqrt{2 \pi \sigma^2}} e^{-\frac{(z-\mu)^2}{2\sigma^2}}
-  \label{eq:gaussian}
-\end{equation}
+%\begin{equation}
+%  \rho(z)=\frac{A}{\sqrt{2 \pi \sigma^2}} e^{-\frac{(z-\mu)^2}{2\sigma^2}}
+%  \label{eq:gaussian}
+%\end{equation}
 
-In the neat-water \suldiox~distribution, $\mu=0.781$, $A=0.109$, and $\sigma=1.684$\angs.
-Thus, the \suldiox~has a surface affinity when in a low-concentration water environment, and it does not fall further into the bulk or move out of the surface into the gas phase.
-This surface active behavior was found experimentally by a previous SFG study by our group,\cite{Tarbuck2005,Tarbuck2006} and also verified by the computational simulations and spectral calculations of Dang et al.\cite{Baer2010} It was concluded that a surface \suldiox~hydrate complex forms and modifies the structure of water above the Gibbs dividing surface (GDS) shifting \wat~free-OH oscillators in towards the bulk, creating a stronger hydrogen-bonding network.
+In the neat-water slab, the \suldiox~distribution shows a surface affinity when in a low-concentration environment.
+It does not reside in the bulk or move out of the surface into the gas phase.
+This surface active behavior was found experimentally by the SFG study by our group,\cite{Tarbuck2005,Tarbuck2006} and also verified by the computational simulations and spectral calculations of Baer et al.\cite{Baer2010} The water behavior hypothesized from the experimental analysis is that a surface \suldiox~hydrate complex forms, and it modifies the structure of water above the Gibbs dividing surface (GDS).
+\wat~free-OH oscillators are thus shifted in towards the bulk, creating a stronger hydrogen-bonding network.
 In a later section of this article we look at the effect of \suldiox~on surface water orientation.
@@
 However, unlike the neat-water slab, the saturated slab has a non-zero bulk concentration of \suldiox~that does not adsorb to the surface.
-The coefficient values for the top surface from eq \ref{eq:gaussian} are $\mu_{top}=2.695$, $A_{top}=7.001$, and $\sigma_{top}=3.203$.
-The bottom surface values are $\mu_{bottom}=-33.846$, $A_{bottom}=6.618$, and $\sigma_{bottom}=3.151$\angs.
 The added concentration of \suldiox~creates a layer of molecules bound to the top of the water surface.
-The center of this distribution is further into the gas phase than for the single-\suldiox~molecule surface.
+The center of the \suldiox~density distribution is further into the gas phase than for the neat-water surface with a single-\suldiox~molecule.
 The accumulation of \suldiox~in this classical simulation coincides with the experimental conclusions of \suldiox~surface hydrate complex formation.
 However, without simulating breakble bonds the chemistry that may pull \suldiox~into the bulk water phase, and the subsequent reaction to form ions is absent.
-We have confidence in the surface complexation that is exhibited, the effect of the \suldiox~on the surface water molecules, and the accumulation of \suldiox at the interface.
+We have confidence in the surface complexation that is exhibited, the effect of the \suldiox~on the surface water molecules, and the accumulation of \suldiox~at the interface.
 However, we do not yet conclude that the specific locations of \suldiox~on top of the water surface are as accurate as those found in the DFT ab initio calculations of Dang et al.\cite{Baer2010} As was pointed out in their computations, the classical potential does not reproduce the bonding of the first hydration shell around \suldiox, and so conclusions about specific complex formation or geometries are best made from the ab initio simulation results reported.

### 2011-04-04 EquilibratedOrientation.tex
Lots of new angle images. Only showing one surface for angle plots instead of both slab surfaces. rewrote the first draft incorporating Steph's comments. Still need an image for the conclusions.

@@
 Molecular orientation analyses were performed to characterize the net molecular orientation of \wat~and \suldiox~molecules in the different locations of the water surface.
-Histograms were created to show depth-profiles of the angles $\theta$ and $\phi$ in the neat-water and saturated equilibrium systems.
+At each distance from the surface location, an orientation profile was created for both the \wat~and \suldiox~molecules.
+The orientation distribution for the angles $\theta$ and $\phi$ at all depths were combined to form 2D intensity plots that show how the molecular orientation distributions changed with distance to the surface location.
+These plots allow for a visual interpretation of how the net orientations are affected when moving from the gas phase to the surface, and then further into the aqueous bulk.
 
@@
 
-The orientation profiles for \wat~are shown in figure \ref{fig:water-orientation} for both the neat-water (top) and saturated (bottom) systems during the equilibrium MD simulations.
-In both systems an orientational preference is found at the slab surfaces where the water is furthest towards the gas phase.
+The orientation depth-profiles for \wat~are shown in figure \ref{fig:water-orientation} for both the neat-water (top) and saturated (bottom) systems during the equilibrium MD simulations.
+In both systems the strongest orientational preference is found at the slab surfaces where the water is furthest towards the gas phase.
 The histograms are arranged with the plots of $\cos(\theta)$ on the left and $\cos(\phi)$ on the right columns.
-The top surface is located at 0\angs, and the bottom surface is approx.
-30\angs below.
-It is also noted that $\theta$ values of the bottom surface are inverted with respect to the top surface, mirroring those results.
+The surface is located at 0\angs.
+The angle distributions from the top and bottom slab surfaces were averaged for all the orientation analyses.
 
-%The bisector tilt concentrates around $\cos(\theta)=0$ within the first few\angs of the surface, and then the distribution becomes isotropic further into the water bulk.
+The bisector tilt, $\cos(\theta)$, concentrates around $\cos(\theta)=0$ within the first few\angs~of the surface, and then the distribution becomes isotropic further into the water bulk.
 As the tilt nears $\cos(\theta)=0$ the \wat~bisector lies within the plane of the surface indicating a water orientation either flat on the surface, or with some amount of ``twist'' sending the OH bonds in towards, or out of the bulk.
@@
 This results from an orientation of the water's y-axis (normal to the molecular plane) aligned perpendicular to the plane of the water surface.
+Thus, both the neat-water and saturated systems have waters lying mostly flat to the plane of the interface at a distance of 0\angs.
 
-The histograms show overall similarities for both systems in their shapes and intensities.
-The presence of a saturated layer of adsorbed \suldiox~molecules alters the water orientation, but not necessarily the orientational depth of the interface.
+Although the plots show overall similarities for both the neat-water and saturated systems, the presence of a layer of adsorbed \suldiox~molecules alters the water orientation of the waters furthest from the aqueous bulk.
+The waters located above 0\angs~are further into the gas phase, and interact with the layer of adsorbed \suldiox~molecules.
+The \suldiox~layer does not provide the same amount of strong hydrogen bonding and solvation that the waters experience further in to the aqueous phase.
+The resulting orientation of waters above 0\angs, shown in the saturated plot of $\cos(\theta)$ of figure \ref{h2o-angles}, is with a bisector pointing further into the adsorbed \suldiox~gas layer, and both hydrogens pointing outward from the aqueous bulk.
+The effect is more pronounced further from the water phase, and above 5\angs~the $\cos(\theta)$ distribution is completely centered around $\cos(\theta)=+1$ (see figure \ref{water-angles}).
+
+%The histograms show overall similarities for both systems in their shapes and intensities from approx.
+0\angs~and below.
+The presence of a saturated layer of adsorbed \suldiox~molecules, however, alters the water orientation, but not necessarily the orientational depth of the interface.
 In both $\theta$ distributions, the orientations of waters in the bulk region are isotropic until 5\angs~below the surface location.
@@
 The change in $\cos(\theta)$ is not as rapid at the neat-water surface as in the saturated system.
-Furthermore, the noise in the distribution above 5\angs~(and below approx -34\angs) in the neat-water system is a result of fewer waters venturing beyond those extents and thus less data far from the surface.
+Furthermore, the noise in the distribution above 5\angs~in the neat-water system is a result of fewer waters venturing beyond those extents and thus less data far from the surface.
 This is one indication that waters near a layer of adsorbed \suldiox~venture further above the interface and interact with the adsorbed gas molecules.
@@
 Waters on the neat surface lie flatter, whereas the presence of the \suldiox~allows a greater range of ``twist'' for those waters in the plane of the interface.
-The $\phi$ distributions quickly become isotropic above a couple angstroms from the surface as the bisectors orient more perpendicularly, and below the surface as the bulk water loses any orientational preference.
+The $\phi$ distributions quickly become isotropic above the surface as the bisectors orient more perpendicularly, and below the surface as the bulk water loses any orientational preference.
 
@@
 The bottom slab surface is approx.
-30\angs below the top surface.
+30\angs~below the top surface.
 Shown are the angle distributions for $\theta$ (left column) and $\phi$ (right column) in both the neat-\wat~system (top row) and the saturated system (bottom row).
@@
 
-The angular distribution of the single \suldiox~is concentrated primarily in $\cos(\theta)>0$, such that the bisector points out of the water surface.
-This same distribution occurs in the saturated system for positions below 0\angs, but the orientation becomes isotropic for molecules located above 0\angs, on top of the water surface.
+Within 5\angs~of the water surface (the approximate distance of interaction with surface water molecules) the angular distribution of the single \suldiox~(in the neat-water system) is concentrated primarily in $\cos(\theta)>0$.
+The peak of the distribution occurs at $\cos(\theta)=1$.
+The \suldiox~bisector points out of the water surface, with the sulfur atom pointing towards the aquoues bulk, and the two oxygens pointing into the gas phase.
+This same distribution occurs in the saturated system for positions below 5\angs, but the orientation becomes isotropic for molecules located above 0\angs, on top of the water surface.
+After 5\angs~above the surface both distributions become isotropic.
 Promixity to the water highly orients the \suldiox~with the sulfur atom pointing in towards the water bulk.
+Moving further away from the water surface, and interacting less with \wat~molcules allows for greater orientational freedom as exhibited in the isotropy of the distributions above 5\angs.
+Because both the neat-water and saturated systems show rather similar distributions, it is possible that the concentration of \suldiox~does not strongly affect the orientation of the \suldiox, unlike the orientations of waters.
+
+The plots of $\cos(\phi)$ are both isotropic, although the neat-water system plot is quite noisy from the small data set.
+Because the \suldiox~near the surface is oriented perpendicularly to the interface, the $\cos(\phi)$ orientation is expected to be isotropic.
+Further from the water surface where the bisector orientation becomes isotropic, the $\cos(\phi)$ behavior is expected to behave similarly.
+For the \suldiox~orientation, the $\phi$ angle does not provide further information regarding the surface behavior.
 

### 2011-04-04 Introduction.tex
Lots of new angle images. Only showing one surface for angle plots instead of both slab surfaces. rewrote the first draft incorporating Steph's comments. Still need an image for the conclusions.

@@
 Its high solubility in water makes \suldiox~ an integral compound in many atmospheric reactions and surface reaction mechanisms.
-We seek a fuller understanding of the interactions between \suldiox~and water interfaces because of the central role \suldiox~takes in many atmospheric particle and aerosol reactions.
+We seek a fuller understanding of the interactions between \suldiox~and water at interfaces because of the central role \suldiox~takes in many atmospheric particle and aerosol reactions.
 
-Our recent experimental works provided insight into the adsorption and reaction of gases in atmospheric aerosols.\cite{Ota2011} We concluded that an \suldiox~surface hydrate complex forms when an aqueous surface is exposed to \suldiox~gas.
+Our recent experimental works provided insight into the adsorption and reaction of gases in atmospheric aerosols.\cite{Tarbuck2005,Tarbuck2006} We concluded that an \suldiox~surface hydrate complex forms when an aqueous surface is exposed to \suldiox~gas.
 A computational study by Dang et al.\cite{Baer2010} then made a series of predictions of the specific nature of the hydrated complex through classical and ab initio simulations.
@@
 
-Our latest experimental work on \suldiox~behavior further expands our understanding of \suldiox~in atmospheric processes.\cite{Ota2011} A temperature study showed that the binding of gaseous \suldiox~to water interfaces at atmospherically relevant cold temperatures is greatly enhanced.
+Our latest experimental work on \suldiox~behavior further expands our understanding of \suldiox~in atmospheric processes.\cite{Ota2011} A temperature study showed that the binding of gaseous \suldiox~to water surfaces at atmospherically relevant cold temperatures is greatly enhanced.
 The increased binding facilitates the reaction with water to produce dissolved sulfur species.
@@
 The same study also showed that low pH aqueous solutions, as often found in atmospheric aerosol composition, inhibit the reaction of \suldiox, but do not affect the surface binding of the gas molecules.
-The resultant experimental spectra suggest that the surface waters reorient in response to the presence of gaseous \suldiox, with the interfacial water ``free-OH'' bonds orienting more perpendicularly to the plane of the aqueous surface.
+The resultant experimental spectra suggest that the surface waters reorient in response to the presence of gaseous \suldiox.
+From the fitting analysis, one possible geometry of water was posited with the interfacial water ``free-OH'' bonds orienting more perpendicularly to the plane of the aqueous surface.
 

### 2011-04-04 Methods.tex
Lots of new angle images. Only showing one surface for angle plots instead of both slab surfaces. rewrote the first draft incorporating Steph's comments. Still need an image for the conclusions.

@@
 A second set of simulations began with an equilibrated water slab as in the surface equilibrated method above.
-However, in both the neat-water and saturated systems equilibrated systems, a single \suldiox~was introduced 20\angs~above the water slab surface, with the sulfur atom tethered to its initial position.
+However, in both the equilibrated neat-water and saturated systems, a single \suldiox~was introduced 20\angs~above the water slab surface, with the sulfur atom tethered to its initial position.
 The systems were evolved for 1 ns, taking coordinate snapshots every 20 ps to create 50 starting points for further simulations.
-Steered molecular dynamics (SMD) were then performed on the 50 system configurations (in both the neat-water and saturated configurations) to guide the \suldiox~down towards a tethered water 15\angs under the water slab surface by applying a small steering force to the \suldiox-sulfur atom.\cite{Isralewitz2001} The \suldiox~thus passed through the continuum of environments from gas phase to (neat- and saturated) water surface adsorption, and finally absorption into the bulk of the \wat~slab.
+Steered molecular dynamics (SMD) were then performed on the 50 system configurations (in both the neat-water and saturated configurations) to guide the \suldiox~down towards a tethered water 15\angs~under the water slab surface by applying a small steering force to the \suldiox-sulfur atom.\cite{Isralewitz2001} The \suldiox~thus passed through the continuum of environments from gas phase to (neat- and saturated) water surface adsorption, and finally absorption into the bulk of the \wat~slab.
 Each of the SMD simulations were performed for a total of 200 ps, using a timestep of 1 fs, and taking snapshots of the system every 25 fs.
@@
 This was done to ensure that the orientation of the \suldiox~during the adsorption transit was not an artifact of the choice of atom used for tethering.
-The simulations produced the same results (not shown) for the orientational analyses, so the data from the original tethering scheme was used.
+The simulations produced the same results (not shown) for the orientational analyses, so the data from the original tethering scheme were used.
 

### 2011-04-04 TransitOrientation.tex
Lots of new angle images. Only showing one surface for angle plots instead of both slab surfaces. rewrote the first draft incorporating Steph's comments. Still need an image for the conclusions.

@@
 The angles $\theta$ and $\phi$ (figure \ref{fig:water-angles}) were calculated for each timestep of the SMD simulations as the transiting \suldiox~was pulled into the water slab from the gas phase, both in the neat-water and saturated slab systems.
-The angle cosine values were collected for the 50 simulations of both systems for each distance from the water surface, resulting in the 2-dimensional histograms shown in figure \ref{fig:so2-transit-angles}.
+The angle cosine values were collected for the 50 simulations of both systems for each distance from the water surface, resulting in the 2-dimensional angle and depth-profile histograms shown in figure \ref{fig:so2-transit-angles}.
 Only the surface into which the transiting \suldiox~was steered is used in the analyses.
@@
 	From its starting position 20\angs~above the water surface, until the \suldiox~moves to within 10\angs~of the surface of both systems, the orientation is isotropic in $\theta$ and $\phi$.
-The bisector angle $\theta$, near to the surface becomes more perpendicular ($\cos(\theta)\approx1$) with the \suldiox~sulfur pointing into the water phase.
+Near the surface the bisector angle $\theta$ becomes more perpendicular ($\cos(\theta)\approx1$) with the \suldiox~sulfur pointing into the water phase.
 At the point when the \suldiox~reaches the surface location (distance $=0$\angs), the bisector is perpendicular in both the neat-water and saturated systems.
-There are differences between where the orientation transitions from isotropic in the neat-water and saturated systems.
+Figure \ref{fig:so2-surface-angles}a depicts \suldiox~near the water surface with the bisector aligned perpendicular to the interface.
+There are differences between the two systems as to where the point of transition from isotropic to preferred orientations happens.
 In the absence of simulated ionic species that form through \suldiox-\wat~chemistry at the surface, it is clear that the adsorbing \suldiox~in the gas phase takes on a preferred orientation to adsorb on a water surface.
@@
   % talk about theta differences
-  The $\theta$ distribution clearly shows that the \suldiox~is isotropically oriented in the gas-phase and then orients with its bisector perpendicular to the water surface within a certain distance above the water phase.
-On approach to the neat-water surface, that transition occurs once just above 10\angs~from the surface, and again at 5\angs~from the surface and below.
-The onset of a preferred orientation at 10\angs~occurrs at the cutoff length of the simulated intermolecular bonding interactions.
-Between 5-10\angs~the molecule appears to orient more isotropically.
+  The $\cos(\theta)$ distribution above 10\angs~from the surface is isotropic.
+There is a transition below 10\angs~where the \suldiox~orients with its bisector perpendicular to the water surface.
+On approach to the neat-water surface, that transition occurs at approximately 5\angs~from the surface.
+This coincides with the approximate distance of bonds formed to the outer-most water molecules.
 Below 5\angs, the orientation persists and the \suldiox~begins to interact with the topmost surface waters, and orients accordingly.
-In the saturated system, the same trend occurs, however the onset of the perpendicular orientation begins at approx.
-8\angs~above the surface location.
+In the saturated system, the same trend occurs, however the onset of the perpendicular orientation begins at approximately 8\angs~above the surface.
 The layer of adsorbed \suldiox~already present in the saturated system most likely interacts with the transiting \suldiox~molecule.
-Also, topmost water molecules from the surface move up to a few\angs~within the \suldiox~layer and interact with the transiting \suldiox~further from the surface than those in the neat-water system.
+Also, topmost water molecules from the surface move up to a few\angs~inside the adsorbed \suldiox~layer and interact with the transiting \suldiox~further from the surface than those in the neat-water system.
+
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[scale=1.0]{images/angle-cartoons/so2-surface.png}
+		\caption{(A) \suldiox~molecules above and below both neat-water and saturated surfaces orient with the bisectors perpendicular to the interface.
+The \suldiox~sulfur points down towards the aqueous phase.
+(B) In the neat-water system, a transiting \suldiox~molecule located below the surface briefly reorients with one S-O bond more towards the interface above, and the other bond pointing further down into the water bulk.}
+		\label{fig:so2-surface-angles}
+	\end{center}
+\end{figure}
 
@@
   With a mostly perpendicular bisector angle, it is expected that the values of $\phi$ would be isotropic relative to the reference axis.
-This is the case in both systems with one exception in the neat-water system.
-At the water surface and just below from 0\angs~to 5\angs~in the water phase, the $\theta$ profile broadens, and a clear trend appears in the profile of $\phi$.
-This corresponds to a \suldiox~orienting more flat to the surface (inclined up to 60 degrees from the surface normal).
-However, below this 5\angs~region, the \suldiox~resumes an isotropic orientation as it becomes fully solvated by bulk waters.
+This is the case in both systems, with one exception in the neat-water system.
+At the neat-water surface and just below from 0\angs~to 5\angs~in the water phase, the $\theta$ profile broadens, and a clear trend appears in the profile of $\phi$.
+This corresponds to a \suldiox~orienting less flat to the surface (inclined up to 60 degrees from the surface normal).
+The result is that \suldiox~in this region shows a preference for one of the S-O bonds to point more towards the aqueous bulk and the other to point away towards the interface.
+Figure \ref{so2-surface-angles}b depicts the orientation of the \suldiox~slightly below the water surface.
+This is in contrast to \suldiox~molecules above the surface, oriented more perpendicularly.
+The behavior is likely due to the interaction between the waters and one of the S-O bonds leading to a higher solvation than above the water surface.
+As the \suldiox~is solvated by more highly-coordinate bulk water, the S-O bonds experience less equal interaction environments.
+Baer et al.
+noted that their force field model for the \suldiox~does not reproduce well the first hydration shell geometries,\cite{Baer2010} so conclusions regarding the specific interactions and hydrate geometries between the \suldiox~and \wat~cannot be made here.
+It is notable that the same reorientation does not occur as strongly in the saturated system.
+The presence of the adsorbed \suldiox~layer apparently decreases the reorienting behavior likely because of the disrupting effect the higher \suldiox~concentration has on the water interactions at the surface.
+Below this 5\angs~region under the surface, the \suldiox~resumes an isotropic orientation as it becomes fully solvated by bulk waters.
 
@@
 
-	%Differences occurr between the system orientation profiles once the \suldiox~has moved further than 10\angs below the surface.
+	%Differences ocurr between the system orientation profiles once the \suldiox~has moved further than 10\angs below the surface.
 While both $\theta$ and $\phi$ hold the same trend from just under the surface through to the bulk of the slab in the saturated system, the orientation profile becomes much more isotropic past 10\angs under the surface of the neat-\wat~system.

### 2011-04-05 Conclusions.tex
Finishing up the 2nd draft. Added new so2/h2o molecule cartoons, and a cartoon for the molecular layout near the interface for both systems.

@@
 
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[scale=1.0]{images/angle-cartoons/system-surface.png}
+		\caption{\wat~and \suldiox~both exhibit preferred orientations in the region near the liquid-gas interface.
+The neat-water system (A) with a single \suldiox~molecule (low-concentration) has surface waters orienting mostly flat to the interface.
+When the \suldiox~concentration is increased, as in the saturated system (B), the waters at the surface behave similar to the neat-water interface, but waters that venture into the adsorbed \suldiox~gas layer orient strongly with their bisectors pointing out from the aqueous phase towards the gas.
+In both cases, the \suldiox~orients with its molecular bisector pointing out to the gas phase when it is near the surface.}
+		\label{fig:so2-surface-cartoon}
+	\end{center}
+\end{figure}
+
 Steered molecular dynamics simulations were used to model the behavior of an adsorbing \suldiox~as it moves from the gas phase above the water down through the surface and into the bulk.

### 2011-04-05 EquilibratedOrientation.tex
Finishing up the 2nd draft. Added new so2/h2o molecule cartoons, and a cartoon for the molecular layout near the interface for both systems.

@@
 The \suldiox~layer does not provide the same amount of strong hydrogen bonding and solvation that the waters experience further in to the aqueous phase.
-The resulting orientation of waters above 0\angs, shown in the saturated plot of $\cos(\theta)$ of figure \ref{h2o-angles}, is with a bisector pointing further into the adsorbed \suldiox~gas layer, and both hydrogens pointing outward from the aqueous bulk.
-The effect is more pronounced further from the water phase, and above 5\angs~the $\cos(\theta)$ distribution is completely centered around $\cos(\theta)=+1$ (see figure \ref{water-angles}).
+The resulting orientation of waters above 0\angs, shown in the saturated plot of $\cos(\theta)$ of figure \ref{fig:water-orientation}, is with a bisector pointing further into the adsorbed \suldiox~gas layer, and both hydrogens pointing outward from the aqueous bulk.
+The effect is more pronounced further from the water phase, and above 5\angs~the $\cos(\theta)$ distribution is completely centered around $\cos(\theta)=+1$ (see figure \ref{fig:water-angles}).
 

### 2011-04-05 Conclusions.tex
Draft 2

@@
 Our equilibrium simulations show that a single \suldiox~molecule, representing a low concentration, has a high surface affinity.
-This was previously reported by Dang.\cite{Baer2010} At a high \suldiox~concentration, the unreacted \suldiox~moves further out of the water phase as more molecules bind on the surface.
+This was previously reported by Baer et al.\cite{Baer2010} At a high \suldiox~concentration, the unreacted \suldiox~moves further out of the water phase as more molecules bind on the surface.
 The orientation of \suldiox~on the water surface was found to be similar for both low and high concentrations.

### 2011-04-05 DensityResults.tex
Draft 2

@@
 We have confidence in the surface complexation that is exhibited, the effect of the \suldiox~on the surface water molecules, and the accumulation of \suldiox~at the interface.
-However, we do not yet conclude that the specific locations of \suldiox~on top of the water surface are as accurate as those found in the DFT ab initio calculations of Dang et al.\cite{Baer2010} As was pointed out in their computations, the classical potential does not reproduce the bonding of the first hydration shell around \suldiox, and so conclusions about specific complex formation or geometries are best made from the ab initio simulation results reported.
+However, we do not yet conclude that the specific locations of \suldiox~on top of the water surface are as accurate as those found in the DFT ab initio calculations of Baer et al.\cite{Baer2010} As was pointed out in their computations, the classical potential does not reproduce the bonding of the first hydration shell around \suldiox, and so conclusions about specific complex formation or geometries are best made from the ab initio simulation results reported.
 

### 2011-04-05 Introduction.tex
Draft 2

@@
 \suldiox~acts as a major component of atmospheric pollution, and is a precursor to acid rain formation, and cloud nucleation.
-Its high solubility in water makes \suldiox~ an integral compound in many atmospheric reactions and surface reaction mechanisms.
+Atmospheric dust particles and gases have also been implicated in the oxidation of \suldiox, and act as reaction surfaces for chemical mechanisms that are still poorly understood.\cite{Baltrusaitis2011,Rubasinghege2010,Li2007} Its high solubility in water makes \suldiox~ an integral compound in many atmospheric reactions and surface reaction mechanisms.
 We seek a fuller understanding of the interactions between \suldiox~and water at interfaces because of the central role \suldiox~takes in many atmospheric particle and aerosol reactions.
@@
 Our recent experimental works provided insight into the adsorption and reaction of gases in atmospheric aerosols.\cite{Tarbuck2005,Tarbuck2006} We concluded that an \suldiox~surface hydrate complex forms when an aqueous surface is exposed to \suldiox~gas.
-A computational study by Dang et al.\cite{Baer2010} then made a series of predictions of the specific nature of the hydrated complex through classical and ab initio simulations.
+A computational study by Baer et al.\cite{Baer2010} then made a series of predictions of the specific nature of the hydrated complex through classical and ab initio simulations.
 That work developed a detailed picture of the nature of the \suldiox~surface complex with water, and related it to the surface water OH vibrational IR spectra.

### 2011-04-05 Methods.tex
Draft 2

@@
 
-Molecular dynamics simulations were performed using the Amber 11 software suite.\cite{Case2010} Polarizable models for the \wat~and \suldiox~molecules were used in the simulations, and have been used previously in studies on interfacial systems because they are known to more accurately reproduce interfacial structure and free energy profiles.\cite{Wick2007,Rivera2006,Dang1998} The \wat~model used is the POL3 water model,\cite{Caldwell1995} and for \suldiox~we used the model of Dang, et al.
+Molecular dynamics simulations were performed using the Amber 11 software suite.\cite{Case2010} Polarizable models for the \wat~and \suldiox~molecules were used in the simulations, and have been used previously in studies on interfacial systems because they are known to more accurately reproduce interfacial structure and free energy profiles.\cite{Wick2007,Rivera2006,Dang1998} The \wat~model used is the POL3 water model,\cite{Caldwell1995} and for \suldiox~we used the model of Baer et al.
 that places a single polarizable center on the sulfur atom.\cite{Baer2010}

### 2011-04-07 Background.tex
Draft 2
Incorporating Stephanie's comments from draft 2 and reworking the surface-location image

@@
 
-  The location of the water surface was determined by the averaged positions of a set number of the outter-most waters in the simulated aqueous slabs.
+  The location of the water surface was determined by the averaged positions of a set number of the outer-most waters in the simulated aqueous slabs.
 The number of waters included in the averaging was decided by taking a few issues into account.
@@
 Increasing the number of waters used in calculating the surface location diminishes the effects of the few waters that briefly rise above the surface into the gas phase, stabilizing both the surface position and thickness values.
-Taking the close-packed model as a maximum number of waters fit into a flat surface, the authors decided to use 70 water molecules for calculating surface location during the simulation analyses.
+Taking the close-packed model as a maximum number of waters fit into a flat surface, the authors decided to use 70 water molecules for calculating instantaneous water surface locations.
 
 Figure \ref{fig:density-flaw} demonstrates the problem of surface location drift during a simulation.
-The figure shows a density profile of water and \suldiox~over the course of one of the 10 ns trajectories used in this work.
+This figure shows the density profile of water and \suldiox~over the course of one of the 10 ns trajectories used in this work.
 The water density profile (dashed line) is produced by averaging the instantaneous density profile at each timestep in the simulation.
@@
 Thus the calculated location and width from the $\tanh$ line fit are not accurate over long trajectories for defining a stationary reference point.
-All depth locations in our analyses are calculated relative to the surface location at each timestep of the trajectories.
+All depth locations in our analyses are calculated relative to the surface location at the corresponding timestep of the trajectories.
 
@@
 It was found that using 70 waters produced a standard deviation of the water positions corresponding to a width of 3.2\angs.
-as reported by Dang et al.when fitting the density profile with a hyperbolic tangent lineshape.
-To match the interfacial width of Dang using the standard deviation of our simulations, 70 water molecules would have to be included in the top monolayer.
+as reported by Baer et al.when fitting the density profile with a hyperbolic tangent lineshape.
+To match the interfacial width of Baer using the standard deviation of our simulations, 70 water molecules would have to be included in the top monolayer.
 Thus, the calculations and analyses presented in this work incorporate 70 water molecules when calculating the surface location.
@@
 In each analysis a space-fixed reference axis is used that corresponds to the long axis of the system's periodic cell normal to the plane of the water surface.
-The orientational analyses presented herein focus on two angles used to define molecular orientation in space.
+The orientational analyses presented herein focus on two angles used to define molecular orientation.
+The molecular orientation angles $\theta$ and $\phi$ are determined from a set reference axis as shown in figure \ref{fig:water-angles}a.
 
-	A ``tilt'' angle, $\theta$, defines the angle formed between the molecular bisector vector (the molecular z-axis, pointing from the central atom in the direction of the other two atoms) and the positive system reference axis.
+	The ``tilt'' angle, $\theta$, defines the angle formed between the molecular bisector vector (the molecular z-axis, pointing from the central atom in the direction of the other two atoms) and the positive system reference axis.
 Thus the value of $\theta$ falls within a range of $[0,\pi]$.
 An angle of $\cos(\theta)=1$ indicates a molecule with its bisector aligned with the reference axis, while $\cos(\theta)=-1$ results from an anti-aligned configuration.
+Sample representations of different values of $\cos(\theta)$ are shown in figure \ref{fig:water-angles}b.
 
@@
 $\phi$ is the angle formed between the vector normal to the plane of the molecule (molecular y-axis) and the system reference axis.
-The values of $\phi$ fall in the interval $[-\pi,\pi]$.
-For certain values of $\theta$, $\phi$ provides additional information about whether the molecular orientation is ``flat'' to the surface (e.g.
+The values of $\phi$ fall in the interval $[0,\frac{\pi}{2}]$ because of the symmetry of \wat~and \suldiox~molecules with respect to twist about their bisector axes.
+For values of $\cos(\theta) \approx 0$, $\phi$ provides additional information about whether the molecular orientation is ``flat'' to the surface (e.g.
 the plane of the molecule is aligned with the plane of the surface), or if it is perpendicular.
+The values of $\cos(\phi)$ for different molecular orientations are shown in figure \ref{fig:water-angles}c.
 %However, the symmetry of the \wat~and \suldiox~molecules create equivalence between the two H, or O, atoms (in the \wat~or\suldiox, respectively).
@@
 Because $\phi=-\phi$, the value is reported within the range of $[0,\frac \pi 2]$.
-	However, with $\cos(\theta)$ close to 1 or -1 the symmetry of the plane of the surface results in an isotropic distribution in $\cos(\phi)$.
-Figure \ref{fig:water-angles} shows the angle definitions relating the body-fixed axes to a system-fixed reference axis.
+	However, values of $\cos(\theta)$ close to 1 or -1 result in an isotropic distribution in $\cos(\phi)$ because of the symmetry of the plane of the surface in directions perpendicular to the reference axis.
 
@@
 		\caption{(A) The definition of the angles $\theta$ and $\phi$ used to define molecular orientation of \suldiox~and \wat~relative to a reference axis.
-(B) $\theta$ is the value of the ``tilt'' of the molecular bisector from the reference axis, with values ranging from -1 to +1 indicating different bisector orientations.
+(B) $\theta$ is the value of the ``tilt'' of the molecular bisector from the reference axis, with $\cos(\theta)$ values ranging from -1 to +1 indicating different bisector orientations.
 (C) $\phi$ is the ``twist'' angle formed between the vector normal to the molecular plane (the body-fixed y-axis) and the reference axis.
-For certain values of $\theta$, $\phi$ shows how flat the molecule is relative to the interfacial plane.}
+For values of $\cos(\theta) \approx 0$, $\phi$ shows how flat the molecule is relative to the interfacial plane.}
 		\label{fig:water-angles}

### 2011-04-07 Conclusions.tex
Incorporating Stephanie's comments from draft 2 and reworking the surface-location image

--- 2011-04-05 Conclusions.tex
@@
 The surface behavior of the \suldiox~molecules was also determined.
-Experimental surface spectroscopy on the \suldiox~vibrational spectra is not possible because of the nature of the experimental geometry, so the spectra of the surface \suldiox~vibrational modes are still unreported.
+Vibrational sum-frequency surface spectroscopy on the \suldiox~vibrational spectra is not possible because of the nature of the experimental geometry.
+At aqueous interfaces, the spectra of the surface \suldiox~vibrational modes are still unreported.
 However, through these simulations we are able to characterize the orientational behavior of \suldiox~during and after adsorption.
 Our equilibrium simulations show that a single \suldiox~molecule, representing a low concentration, has a high surface affinity.
-This was previously reported by Dang.\cite{Baer2010} At a high \suldiox~concentration, the unreacted \suldiox~moves further out of the water phase as more molecules bind on the surface.
+This was previously reported by Baer et al.\cite{Baer2010} At a high \suldiox~concentration, unreacted \suldiox~molecules are found further out of the water phase as a bound layer of \suldiox~crowds the surface and binds all the surface waters.
 The orientation of \suldiox~on the water surface was found to be similar for both low and high concentrations.
@@
 The \suldiox~slightly above the water surface loses the orientational preference within 10 $\AA$ and those further from the water are more isotropically oriented.
+Figure \ref{fig:so2-surface-cartoon} depicts what the neat-water and saturated surfaces might look like for both \suldiox~and \wat~orientations and locations.
 
@@
 Steered molecular dynamics simulations were used to model the behavior of an adsorbing \suldiox~as it moves from the gas phase above the water down through the surface and into the bulk.
-The results for the transit through the interface show that in both systems of low and high \suldiox~concentration an adsorbing \suldiox has very similar orientation to those already bound to the water surface.
+The results for the transit through the interface show that in both systems of low and high \suldiox~concentration an adsorbing \suldiox~has very similar orientation to those already bound to the water surface.
 As the \suldiox~approaches the water surface the first contact with water causes it to reorient.

### 2011-04-07 DensityResults.tex
Incorporating Stephanie's comments from draft 2 and reworking the surface-location image

@@
 One measure of surface activity is the spatial distribution of molecules in the interfacial region.
-The density distribution of both \wat~and \suldiox~were calculated for the equilibrium MD simulations, and the results are presented in figure \ref{fig:density} for both the neat-water and saturated simulations.
+The density distributions of both \wat~and \suldiox~were calculated for the equilibrium MD simulations, and the results are presented in figure \ref{fig:density} for both the neat-water and saturated simulations.
 The \suldiox~in the neat-water system remained at the top slab surface.
@@
 
-In the neat-water slab, the \suldiox~distribution shows a surface affinity when in a low-concentration environment.
+In the neat-water slab, the \suldiox~molecular density distribution shows a surface affinity when in a low-concentration environment.
 It does not reside in the bulk or move out of the surface into the gas phase.
-This surface active behavior was found experimentally by the SFG study by our group,\cite{Tarbuck2005,Tarbuck2006} and also verified by the computational simulations and spectral calculations of Baer et al.\cite{Baer2010} The water behavior hypothesized from the experimental analysis is that a surface \suldiox~hydrate complex forms, and it modifies the structure of water above the Gibbs dividing surface (GDS).
-\wat~free-OH oscillators are thus shifted in towards the bulk, creating a stronger hydrogen-bonding network.
+This surface active behavior was found experimentally by the SFG study by our group,\cite{Tarbuck2005,Tarbuck2006} and also verified by the computational simulations and spectral calculations of Baer et al.\cite{Baer2010} The water behavior hypothesized from the experimental analysis is that a surface \suldiox~hydrate complex forms, modifying the structure of water above the Gibbs dividing surface (GDS).
+\wat~free-OH oscillators are thus shifted to stronger net orientation, creating a stronger hydrogen-bonding network, and resulting in an increase of the ``free-OH'' contribution to SFG signal intensity from the surface waters.
 In a later section of this article we look at the effect of \suldiox~on surface water orientation.
@@
 However, without simulating breakble bonds the chemistry that may pull \suldiox~into the bulk water phase, and the subsequent reaction to form ions is absent.
-We have confidence in the surface complexation that is exhibited, the effect of the \suldiox~on the surface water molecules, and the accumulation of \suldiox~at the interface.
-However, we do not yet conclude that the specific locations of \suldiox~on top of the water surface are as accurate as those found in the DFT ab initio calculations of Baer et al.\cite{Baer2010} As was pointed out in their computations, the classical potential does not reproduce the bonding of the first hydration shell around \suldiox, and so conclusions about specific complex formation or geometries are best made from the ab initio simulation results reported.
+We have confidence in the presence of surface complexation, the orienting effect of the \suldiox~on the surface water molecules, and the accumulation of \suldiox~at the interface.
+The specific geometries of the surface complexes, however, are not as well reproduced with the classical force-field as would be with ab initio methods.\cite{Baer2010} %However, we do not yet conclude that the specific locations of \suldiox~on top of the water surface are as accurate as those found in the DFT ab initio calculations of Baer et al.\cite{Baer2010} As was pointed out in their computations, the classical potential does not reproduce the bonding of the first hydration shell around \suldiox, and so conclusions about specific complex formation or geometries are best made from the ab initio simulation results reported.
 
@@
 		\caption{Molecular density distributions of \wat~and \suldiox~calculated along the long-axis of the simulated cells.
-The distance axis shows the distance of the given molecule from the top surface of the \wat~slab, with positive values located above the slab towards the gas phase, and negative values located in the water bulk.
-The lower slab surface is located at approx.
--30\angs.
+The position axis shows the distance of the given molecule from the top surface of the \wat~slab, with positive values located above the slab towards the gas phase, and negative values located in the water bulk.
 Distributions of both the neat-water simulation with a single \suldiox~(left, scaled 10x for clarity), and the saturated system (right) are shown.}

### 2011-04-07 EquilibratedOrientation.tex
Incorporating Stephanie's comments from draft 2 and reworking the surface-location image

@@
 
-Molecular orientation analyses were performed to characterize the net molecular orientation of \wat~and \suldiox~molecules in the different locations of the water surface.
+Geometric analyses were performed to characterize the net molecular orientation of \wat~and \suldiox~molecules at different depths from the water surface.
 At each distance from the surface location, an orientation profile was created for both the \wat~and \suldiox~molecules.
@@
 The surface is located at 0\angs.
-The angle distributions from the top and bottom slab surfaces were averaged for all the orientation analyses.
+The angle distributions from both simulated slab surfaces were averaged for all the orientation analyses.
 
@@
 
-Although the plots show overall similarities for both the neat-water and saturated systems, the presence of a layer of adsorbed \suldiox~molecules alters the water orientation of the waters furthest from the aqueous bulk.
+Although the plots show overall similarities for both the neat-water and saturated systems, the presence of a layer of adsorbed \suldiox~molecules alters the water orientation of the waters furthest into the gas phase.
 The waters located above 0\angs~are further into the gas phase, and interact with the layer of adsorbed \suldiox~molecules.
-The \suldiox~layer does not provide the same amount of strong hydrogen bonding and solvation that the waters experience further in to the aqueous phase.
+The \suldiox~layer does not provide the same amount of strong hydrogen bonding and solvation that the waters experience further in the aqueous phase.
 The resulting orientation of waters above 0\angs, shown in the saturated plot of $\cos(\theta)$ of figure \ref{fig:water-orientation}, is with a bisector pointing further into the adsorbed \suldiox~gas layer, and both hydrogens pointing outward from the aqueous bulk.
@@
 Orientation distributions of the adsorbed \suldiox~molecules were created during the equilibrium simulations for both the neat-water and saturated systems.
-Figure \ref{fig:so2-orientation} shows the distributions of $\cos(\theta)$ and $\cos(\phi)$ (arranged similar to the water orientation distributions plots in figure \ref{fig:water-orientation}).
-Only a single surface is shown in those plots with \suldiox~data from the second surface averaged with the first.
+Figure \ref{fig:so2-orientation} shows the distributions of $\cos(\theta)$ and $\cos(\phi)$ (arranged similarly to the water orientation distributions plots in figure \ref{fig:water-orientation}).
 The \suldiox~orientation data set for the neat-water system is much smaller as only a single \suldiox~molecule was simulated in the bulk.
@@
 
-Within 5\angs~of the water surface (the approximate distance of interaction with surface water molecules) the angular distribution of the single \suldiox~(in the neat-water system) is concentrated primarily in $\cos(\theta)>0$.
+Within 5\angs~of the water surface (the approximate distance from the surface that \suldiox~begins interacting with the outer-most waters) the angular distribution of the single \suldiox~(in the neat-water system) is concentrated primarily in $\cos(\theta)>0$.
 The peak of the distribution occurs at $\cos(\theta)=1$.
-The \suldiox~bisector points out of the water surface, with the sulfur atom pointing towards the aquoues bulk, and the two oxygens pointing into the gas phase.
-This same distribution occurs in the saturated system for positions below 5\angs, but the orientation becomes isotropic for molecules located above 0\angs, on top of the water surface.
-After 5\angs~above the surface both distributions become isotropic.
+The \suldiox~bisector points out of the water surface, with the sulfur atom pointing towards the aqueous bulk, and the two oxygens pointing into the gas phase.
+This same distribution occurs in the saturated system for positions below 5\angs.
+Beyond 5\angs~above the surface both distributions become isotropic.
 Promixity to the water highly orients the \suldiox~with the sulfur atom pointing in towards the water bulk.
 Moving further away from the water surface, and interacting less with \wat~molcules allows for greater orientational freedom as exhibited in the isotropy of the distributions above 5\angs.
-Because both the neat-water and saturated systems show rather similar distributions, it is possible that the concentration of \suldiox~does not strongly affect the orientation of the \suldiox, unlike the orientations of waters.
+%Because both the neat-water and saturated systems show rather similar distributions, it is possible that the concentration of \suldiox~does not strongly affect the orientation of the \suldiox, unlike the orientations of waters.
 
@@
 Because the \suldiox~near the surface is oriented perpendicularly to the interface, the $\cos(\phi)$ orientation is expected to be isotropic.
-Further from the water surface where the bisector orientation becomes isotropic, the $\cos(\phi)$ behavior is expected to behave similarly.
+Further from the water surface where the bisector orientation becomes isotropic, the $\cos(\phi)$ distribution also becomes isotropic.
 For the \suldiox~orientation, the $\phi$ angle does not provide further information regarding the surface behavior.

### 2011-04-07 Introduction.tex
Incorporating Stephanie's comments from draft 2 and reworking the surface-location image

@@
 
-Interactions between sulfur species and aqueous systems are the current topic of an active area of research.
-Sulfur dioxide remains an important industrial product, and also enters the environment naturally through terrestrial processes.
-\suldiox~acts as a major component of atmospheric pollution, and is a precursor to acid rain formation, and cloud nucleation.
-Atmospheric dust particles and gases have also been implicated in the oxidation of \suldiox, and act as reaction surfaces for chemical mechanisms that are still poorly understood.\cite{Baltrusaitis2011,Rubasinghege2010,Li2007} Its high solubility in water makes \suldiox~ an integral compound in many atmospheric reactions and surface reaction mechanisms.
-We seek a fuller understanding of the interactions between \suldiox~and water at interfaces because of the central role \suldiox~takes in many atmospheric particle and aerosol reactions.
+The interactions between sulfur species and aqueous systems are currently an active area of chemical research.
+Sulfur dioxide enters the environment as an important industrial product, and also naturally through terrestrial processes.
+Atmospheric dust particles and gases have been implicated in the oxidation of \suldiox, and act as reaction surfaces for chemical mechanisms that are still poorly understood.\cite{Baltrusaitis2011,Rubasinghege2010,Li2007} \suldiox~acts as a major component of atmospheric pollution, and is a precursor to acid rain formation, and cloud nucleation.
+Its high solubility in water makes \suldiox~an integral compound in many aqueous atmospheric reactions, as well.
+We seek a fuller understanding of the interactions between \suldiox~and water at interfaces because they are central to many atmospheric particle and aerosol reactions.
 
@@
 
-Our latest experimental work on \suldiox~behavior further expands our understanding of \suldiox~in atmospheric processes.\cite{Ota2011} A temperature study showed that the binding of gaseous \suldiox~to water surfaces at atmospherically relevant cold temperatures is greatly enhanced.
+Our latest experiments on \suldiox~behavior further expand our understanding of \suldiox~in atmospheric processes.\cite{Ota2011} A temperature study showed that the binding of gaseous \suldiox~to water surfaces is greatly enhanced at atmospherically relevant cold temperatures.
 The increased binding facilitates the reaction with water to produce dissolved sulfur species.
 However, the surface binding of unreacted \suldiox~complexes was found to be a completely reversible process.
-The same study also showed that low pH aqueous solutions, as often found in atmospheric aerosol composition, inhibit the reaction of \suldiox, but do not affect the surface binding of the gas molecules.
+The same study also showed that low pH aqueous solutions, as often found in atmospheric aerosol composition, inhibit the bulk reaction of \suldiox, but do not affect the surface binding of the gas molecules.
 The resultant experimental spectra suggest that the surface waters reorient in response to the presence of gaseous \suldiox.
-From the fitting analysis, one possible geometry of water was posited with the interfacial water ``free-OH'' bonds orienting more perpendicularly to the plane of the aqueous surface.
+From the fitting analysis, a geometry of surface waters was posited with the interfacial water ``free-OH'' bonds orienting more perpendicularly to the plane of the aqueous surface.
 
-In this study we look at the orientational depth profiles of water and \suldiox~during and after the adsorption process.
+In this study we look at computed orientational depth profiles of water and \suldiox~before and after the adsorption process.
 Spectroscopic experiments can not probe depth or orientation profiles of surface species in the same detailed manner as computational simulations.
-Thus we hope to complement the experimental results of previous studies.
+Thus we hope to augment the experimental results of previous studies.
 We determine the effect of introducing \suldiox~to a \wat~surface, and analyze the orientational response of both the \suldiox~and \wat~molecules using equilibrium and steered (SMD) classical molecular dynamics simulations.
 The results of this study will help to understand the behavior of adsorbing gaseous complexes, and will verify predictions of our recent experiments.
+Computational simulations like this are necessary to further understand the microscopic nature of the gaseous adsorption process, as many of the molecular-level subtleties are lost in experiment.
 Along with other ongoing experiments and computations, we develop a more complete picture of gaseous adsorption and complex formation on an aqueous surface.

### 2011-04-07 Methods.tex
Incorporating Stephanie's comments from draft 2 and reworking the surface-location image

@@
 Equilibrium simulations involved adding \suldiox~to a water slab and equilibrating as outlined above.
-One system had a single \suldiox~added to the center of the water box, representing a concentration of 0.06 M.
+The neat-water system had a single \suldiox~added to the center of the water box, representing a concentration of 0.06 M.
 The saturated system with 22 \suldiox~molecules in the bulk represents a system of concentration 1.35 M, with an additional 50 \suldiox~in the gas phase to model 50 atm of pressure above the water surface.
@@
 The systems were evolved for 1 ns, taking coordinate snapshots every 20 ps to create 50 starting points for further simulations.
-Steered molecular dynamics (SMD) were then performed on the 50 system configurations (in both the neat-water and saturated configurations) to guide the \suldiox~down towards a tethered water 15\angs~under the water slab surface by applying a small steering force to the \suldiox-sulfur atom.\cite{Isralewitz2001} The \suldiox~thus passed through the continuum of environments from gas phase to (neat- and saturated) water surface adsorption, and finally absorption into the bulk of the \wat~slab.
+Steered molecular dynamics (SMD) were then performed on the 50 system configurations (in both the neat-water and saturated configurations) to guide the \suldiox~down towards a tethered water 15\angs~under the water slab surface by applying a small steering force to the \suldiox-sulfur atom.
+This steering technique has been previously developed and used to successfuly model chemical events.\cite{Isralewitz2001} The \suldiox~thus passed through the continuum of environments from gas phase to (neat- and saturated) water surface adsorption, and finally absorption into the bulk of the \wat~slab.
 Each of the SMD simulations were performed for a total of 200 ps, using a timestep of 1 fs, and taking snapshots of the system every 25 fs.

### 2011-04-07 TransitOrientation.tex
Finishing up the 2nd draft. Added new so2/h2o molecule cartoons, and a cartoon for the molecular layout near the interface for both systems.
Incorporating Stephanie's comments from draft 2 and reworking the surface-location image

@@
 Figure \ref{fig:so2-surface-angles}a depicts \suldiox~near the water surface with the bisector aligned perpendicular to the interface.
-There are differences between the two systems as to where the point of transition from isotropic to preferred orientations happens.
 In the absence of simulated ionic species that form through \suldiox-\wat~chemistry at the surface, it is clear that the adsorbing \suldiox~in the gas phase takes on a preferred orientation to adsorb on a water surface.
+The main difference between the two systems is where the point of the transition from isotropic to preferred orientations is found.
 
@@
 This is the case in both systems, with one exception in the neat-water system.
-At the neat-water surface and just below from 0\angs~to 5\angs~in the water phase, the $\theta$ profile broadens, and a clear trend appears in the profile of $\phi$.
+At the neat-water surface and just below, from 0\angs~to 5\angs~in the water phase, the $\theta$ profile broadens, and a clear trend appears in the profile of $\phi$.
 This corresponds to a \suldiox~orienting less flat to the surface (inclined up to 60 degrees from the surface normal).
 The result is that \suldiox~in this region shows a preference for one of the S-O bonds to point more towards the aqueous bulk and the other to point away towards the interface.
-Figure \ref{so2-surface-angles}b depicts the orientation of the \suldiox~slightly below the water surface.
+Figure \ref{fig:so2-surface-angles}b depicts the orientation of the \suldiox~slightly below the water surface.
 This is in contrast to \suldiox~molecules above the surface, oriented more perpendicularly.
 The behavior is likely due to the interaction between the waters and one of the S-O bonds leading to a higher solvation than above the water surface.
-As the \suldiox~is solvated by more highly-coordinate bulk water, the S-O bonds experience less equal interaction environments.
+As the \suldiox~is solvated by more highly-coordinated bulk water, the S-O bonds experience less equal interaction environments.
 Baer et al.

### 2011-04-07 Background.tex
New image for showing system density and surface-location drift

@@
 This figure shows the density profile of water and \suldiox~over the course of one of the 10 ns trajectories used in this work.
-The water density profile (dashed line) is produced by averaging the instantaneous density profile at each timestep in the simulation.
+The water density profile and location (dashed line) is produced by averaging the instantaneous density profile at each timestep in the simulation.
 The water profile was then fit to a $\tanh$ function (black) to extract the position and width parameters of the water surface.
 The fitted Gibbs dividing surface has a ``90-10'' width of 8.28 $\AA$.
-A bulge in the gas-phase (left) side of the water profile is indicative of the drift of the water slab over the course of the trajectory.
+A bulge in the gas-phase (left) side of the water density profile is indicative of the drift of the water slab over the course of the trajectory.
 Thus the calculated location and width from the $\tanh$ line fit are not accurate over long trajectories for defining a stationary reference point.
-All depth locations in our analyses are calculated relative to the surface location at the corresponding timestep of the trajectories.
+All depth locations in our analyses are calculated relative to the surface location (blue) at the corresponding timestep of the trajectories.
 
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/density/Density+SurfaceLocation-flawed.png}
-		\caption{Density profiles of \wat~(dashed) and \suldiox~(red) from a 10 ns simulation of the neat-water system with a single sulfur dioxide.
-The \wat~density profile was fit using a hyperbolic tangent function to extract location and width data.
-The location of the outer \wat~monolayer was calculated, and the corresponding location distribution (blue) is also plotted.
+		\includegraphics[scale=1.0]{images/density/density-location.png}
+		\caption{(A) Density profiles of \wat~(dashed) and \suldiox~(red) from a 10 ns simulation of the neat-water system with a single sulfur dioxide.
+The \wat~density profile was fit (black) using a hyperbolic tangent function to extract location and width data for the water surface.
+(B) The location of the outer \wat~monolayer was calculated for each simulation timestep (blue), and the surface location from the density fitting was also determined (dashed).
 Long simulations of liquid slabs result in drift of the slab location, and the consequent broadening of the water density profile if the surface location is not determined repeatedly throughout the simulation.}

### 2011-05-19 Introduction.tex
Rewrote the introduction to ask questions and emphasize the importance of the work, as per Geri's request

@@
 \section{Introduction}
+
+The physics of gaseous adsorption on water surfaces is a mostly unexplored field.
+Our current molecular-level understanding of adsorption phenomena relies on experiments and theory of model systems, and on an increasing number of computational simulations of molecular processes.
+What species form during gas adsorption onto liquid surfaces, and what are the intermediary steps?
+To what extent does an adsorbing gas affect liquid water surface molecules, or the structure of the interface throughout an adsorption process?
+Are specific gas or liquid molecular orientations necessary for gaseous adsorption?
+Very little work has been done to characterize the molecular processes that take place during the transit of a molecule from the gas phase onto a liquid interfacial region.
+Furthermore, experiment does not yet provide us with the resolution necessary to determine the geometries of adsorbing gases, or to determine the orientations of the molecules at the liquid surface near the adsorption site.
+In this work we look at the orientational properties of sulfur dioxide gas adsorbing to a water surface as a model for the transit of gases to liquid interfaces.
 
@@
 
-Our recent experimental works provided insight into the adsorption and reaction of gases in atmospheric aerosols.\cite{Tarbuck2005,Tarbuck2006} We concluded that an \suldiox~surface hydrate complex forms when an aqueous surface is exposed to \suldiox~gas.
+Our recent vibational sum frequency spectroscopy (vsfs) experiments provided insight into the adsorption and reaction of gases in atmospheric aerosols.\cite{Tarbuck2005,Tarbuck2006} We concluded that an \suldiox~surface hydrate complex forms when an aqueous surface is exposed to \suldiox~gas.
 A computational study by Baer et al.\cite{Baer2010} then made a series of predictions of the specific nature of the hydrated complex through classical and ab initio simulations.
@@
 
-Our latest experiments on \suldiox~behavior further expand our understanding of \suldiox~in atmospheric processes.\cite{Ota2011} A temperature study showed that the binding of gaseous \suldiox~to water surfaces is greatly enhanced at atmospherically relevant cold temperatures.
-The increased binding facilitates the reaction with water to produce dissolved sulfur species.
-However, the surface binding of unreacted \suldiox~complexes was found to be a completely reversible process.
-The same study also showed that low pH aqueous solutions, as often found in atmospheric aerosol composition, inhibit the bulk reaction of \suldiox, but do not affect the surface binding of the gas molecules.
-The resultant experimental spectra suggest that the surface waters reorient in response to the presence of gaseous \suldiox.
-From the fitting analysis, a geometry of surface waters was posited with the interfacial water ``free-OH'' bonds orienting more perpendicularly to the plane of the aqueous surface.
+Our latest vsfs experiments on \suldiox~behavior further expand our understanding of \suldiox~in atmospheric processes.\cite{Ota2011} A temperature study showed that the binding of gaseous \suldiox~to water surfaces is greatly enhanced at atmospherically relevant cold temperatures.
+The surface binding of unreacted \suldiox~complexes was found to be a completely reversible process.
+The same study also showed that low pH aqueous environments inhibit the bulk reaction of \suldiox, but do not affect the surface binding of the gas molecules.
+The resultant experimental spectra suggest that the surface waters reorient in response to the presence of adsorbed \suldiox.
+Knowing that reorientation may occur due to \suldiox~adsorption, how can we confirm it?
+what is the nature of the reorienting waters?
+How do they reorient, and to what extent?
+Conclusive experimental evidence of the adsorbing gas or surface water molecular orientation is still missing, so we are moved to use computational methods for further understanding this phenomena.
 
@@
 Spectroscopic experiments can not probe depth or orientation profiles of surface species in the same detailed manner as computational simulations.
-Thus we hope to augment the experimental results of previous studies.
 We determine the effect of introducing \suldiox~to a \wat~surface, and analyze the orientational response of both the \suldiox~and \wat~molecules using equilibrium and steered (SMD) classical molecular dynamics simulations.
-The results of this study will help to understand the behavior of adsorbing gaseous complexes, and will verify predictions of our recent experiments.
+We use a technique of steering a gas molecule into the aqueous phase, and characterize the transit through the interfacial region.
+This unique method gives us much greater insight into the behaviors of gas molecules as they move near to liquid water.
+We also simulate the behavior of adsorbed \suldiox~at equilibrium to show how an already-adsorbed \suldiox~behaves on a water surface.
+The results of this study will help us to understand the behavior of adsorbing gases.
 Computational simulations like this are necessary to further understand the microscopic nature of the gaseous adsorption process, as many of the molecular-level subtleties are lost in experiment.
-Along with other ongoing experiments and computations, we develop a more complete picture of gaseous adsorption and complex formation on an aqueous surface.
+Along with other ongoing experiments and computations, we develop a more complete picture of gaseous adsorption on aqueous surfaces.

### 2011-05-19 Introduction.tex
rearranged sections per notes from chat with Geri. New Computational Approach section has methods for MD, issues of surface location, and molecular orientation introduction. Got rid of background.tex altogether.

@@
 Furthermore, experiment does not yet provide us with the resolution necessary to determine the geometries of adsorbing gases, or to determine the orientations of the molecules at the liquid surface near the adsorption site.
-In this work we look at the orientational properties of sulfur dioxide gas adsorbing to a water surface as a model for the transit of gases to liquid interfaces.
+In this work we look at the orientational properties of sulfur dioxide gas adsorbing to a water surface as a model for the transit of gases on to liquid interfaces.
 

### 2011-05-19 Methods.tex
rearranged sections per notes from chat with Geri. New Computational Approach section has methods for MD, issues of surface location, and molecular orientation introduction. Got rid of background.tex altogether.

@@
-\section{Computational Methods}
+\section{Computational Approach}
 
@@
 \end{figure}
+
+\subsection {Aqueous Surface Location}
+
+	In this study we create orientational depth-profiles of \suldiox~and \wat.
+The reference point used for depth calculations is the location of the water surface.
+A liquid surface, however, is a dynamic boundary that is neither flat nor stationary.
+To define a water surface location, several previous studies have used the technique of fitting a lineshape to the averaged density profile of the water, and extracting interfacial shape and location parameters.\cite{Shamay2010,Wick2006c,Chowdhary2006} Hyperbolic tangent functions have been used often, and values for the ``Gibb's dividing surface'' location, and interfacial width have thus been determined.\cite{Matsumoto1988} However, in long simulations the location and shape of the interface changes, and the motion of surface waters alters the interfacial width at any given timestep.
+Thus, the density profile fitting will capture averaged widths and locations, not instantaneous values.
+Similarly, the averaged values of location and width will obscure information about any drift or deformations the surface undergoes.
+In this work we define the surface location by calculating a reference location at each timestep by averaging the positions of the waters contained in the topmost monolayer.
+This provides us with a consistent and intuitive reference point in the simulations to which we relate our analyses, but does not increase the computational burden.
+
+        The location of the water surface was determined by the averaged positions of a set number of the outer-most waters in the simulated aqueous slabs.
+The number of waters included in the averaging was decided by taking a few issues into account.
+First, a visual inspection of the top cross-section of the water slab over several timesteps indicated between 65-75 waters established a full monolayer.
+Alternatively, assuming a spherical model of water with a radius of 2.2\angs~(the typical length of a bulk-water hydrogen bond), two layers of hexagonally tight-packed spheres would yield a similar number of molecules to the visual inspection.
+Increasing the number of waters used in calculating the surface location diminishes the effects of the few waters that briefly rise above the surface into the gas phase, stabilizing both the surface position and thickness values.
+Taking the close-packed model as a maximum number of waters fit into a flat surface, the authors decided to use 70 water molecules for calculating instantaneous water surface locations.
+
+Figure \ref{fig:density-flaw} demonstrates the problem of surface location drift during a simulation.
+This figure shows the density profile of water and \suldiox~over the course of one of the 10 ns trajectories used in this work.
+The water density profile and location (dashed line) is produced by averaging the instantaneous density profile at each timestep in the simulation.
+The water profile was then fit to a $\tanh$ function (black) to extract the position and width parameters of the water surface.
+The fitted Gibbs dividing surface has a ``90-10'' width of 8.28 $\AA$.
+A bulge in the gas-phase (left) side of the water density profile is indicative of the drift of the water slab over the course of the trajectory.
+Thus the calculated location and width from the $\tanh$ line fit are not accurate over long trajectories for defining a stationary reference point.
+All depth locations in our analyses are calculated relative to the surface location (blue) at the corresponding timestep of the trajectories.
+
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[scale=1.0]{images/density/density-location.png}
+		\caption{(A) Density profiles of \wat~(dashed) and \suldiox~(red) from a 10 ns simulation of the neat-water system with a single sulfur dioxide.
+The \wat~density profile was fit (black) using a hyperbolic tangent function to extract location and width data for the water surface.
+(B) The location of the outer \wat~monolayer was calculated for each simulation timestep (blue), and the surface location from the density fitting was also determined (dashed).
+Long simulations of liquid slabs result in drift of the slab location, and the consequent broadening of the water density profile if the surface location is not determined repeatedly throughout the simulation.}
+		\label{fig:density-flaw}
+	\end{center}
+\end{figure}
+
+The new procedure described above for finding the outer monolayer location was implemented, and the distribution of that location is also plotted in figure \ref{fig:density-flaw} (blue).
+It is apparent from the surface location distribution that the water slab drifts over the 10 ns, spanning approx.
+12 $\AA$.
+However, the maximum standard deviation of the positions of the waters comprising the surface layer at each timestep is only 1.85 $\AA$.
+
+  %Taking one of the simulations of the neat-water slabs, a series of calculations were performed to find the averaged positions and the standard deviations of varying numbers of the topmost waters.
+It was found that using 70 waters produced a standard deviation of the water positions corresponding to a width of 3.2\angs.
+as reported by Baer et al.when fitting the density profile with a hyperbolic tangent lineshape.
+To match the interfacial width of Baer using the standard deviation of our simulations, 70 water molecules would have to be included in the top monolayer.
+Thus, the calculations and analyses presented in this work incorporate 70 water molecules when calculating the surface location.
+
+\subsection{Molecular Orientation}
+
+	The simulated systems were analyzed to characterize the orientation of \wat~and \suldiox~in various environments above, within, and below the aqueous surface region.
+In order to better understand the chemistry occurring during the \suldiox~adsorption process, the molecular orientation of both the \wat~and \suldiox~must be understood.
+The two molecules studied are similarly shaped with a C2$_v$ axis along their bisectors, and a molecular plane defined by three atoms.
+A body-fixed frame is defined for both \wat~and \suldiox~as shown in figure \ref{fig:molecular-frame}.
+In each analysis a space-fixed reference axis is used that corresponds to the long axis of the system's periodic cell normal to the plane of the water surface.
+The orientational analyses presented herein focus on two angles used to define molecular orientation.
+The molecular orientation angles $\theta$ and $\phi$ are determined from a set reference axis as shown in figure \ref{fig:water-angles}a.
+
+	The ``tilt'' angle, $\theta$, defines the angle formed between the molecular bisector vector (the molecular z-axis, pointing from the central atom in the direction of the other two atoms) and the positive system reference axis.
+Thus the value of $\theta$ falls within a range of $[0,\pi]$.
+An angle of $\cos(\theta)=1$ indicates a molecule with its bisector aligned with the reference axis, while $\cos(\theta)=-1$ results from an anti-aligned configuration.
+Sample representations of different values of $\cos(\theta)$ are shown in figure \ref{fig:water-angles}b.
+
+	A second angle, $\phi$, defines the molecular ``twist'' for certain configurations of the molecule.
+$\phi$ is the angle formed between the vector normal to the plane of the molecule (molecular y-axis) and the system reference axis.
+The values of $\phi$ fall in the interval $[0,\frac{\pi}{2}]$ because of the symmetry of \wat~and \suldiox~molecules with respect to twist about their bisector axes.
+For values of $\cos(\theta) \approx 0$, $\phi$ provides additional information about whether the molecular orientation is ``flat'' to the surface (e.g.
+the plane of the molecule is aligned with the plane of the surface), or if it is perpendicular.
+The values of $\cos(\phi)$ for different molecular orientations are shown in figure \ref{fig:water-angles}c.
+%However, the symmetry of the \wat~and \suldiox~molecules create equivalence between the two H, or O, atoms (in the \wat~or\suldiox, respectively).
+$\phi=-\pi$ is equivalent to $\phi=\pi$, and both situations are equivalent to $\phi=0$.
+Because $\phi=-\phi$, the value is reported within the range of $[0,\frac \pi 2]$.
+	However, values of $\cos(\theta)$ close to 1 or -1 result in an isotropic distribution in $\cos(\phi)$ because of the symmetry of the plane of the surface in directions perpendicular to the reference axis.
+
+%	The two molecules studied, \wat~and \suldiox, and similarly shaped with a C2$_v$ axis along their bisectors, and a molecular plane is defined by their three atoms.
+The orientational analyses presented herein focus on two angles used to define the molecular orientation in space.
+In each analysis a reference axis is used to define molecular angles, and is the z-axis unless otherwise specified.
+A ``tilt'' angle, $\theta$, defines the angle formed between a vector (generally the bisector vector pointing from the central atom in the direction of the other two atoms) and the positive reference axis.
+Thus the value of $\theta$ falls within a range of $[0,\pi]$.
+A second angle, $\phi$, defines the molecular ``twist'' around the reference axis.
+$\phi$ is formed between the projection of a vector onto the x-y reference plane, and the x-axis.
+$\phi$ is thus defined relative to the x-axis, and its values are in the interval $[-\pi,\pi]$.
+However, because of the symmetry of the \wat~and \suldiox~molecules and the equivalence of the two H or O atoms (in the \wat~or\suldiox, respectively), $\phi=-\pi$ is equivalent to $\phi=\pi$, and both situations are equivalent to $\phi=0$.
+Because $\phi=-\phi$, the value is reported within the range of $[0,\frac \pi 2]$.
+Figure \ref{fig:spherical-angle} shows the angle definitions using the z-axis as the reference axis.
+
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[scale=1.0]{images/angle-cartoons/molecularframesmall.png}
+		\caption{Molecular body-fixed axes are defined with the x-z plane formed by the three atoms, and the z-axis aligned to the molecular bisector.
+The y-axis is normal to the molecular plane, and one of the bonds points in the positive x direction.}
+		\label{fig:molecular-frame}
+	\end{center}
+\end{figure}
+
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[scale=1.0]{images/angle-cartoons/molecular-angles.png}
+		\caption{(A) The definition of the angles $\theta$ and $\phi$ used to define molecular orientation of \suldiox~and \wat~relative to a reference axis.
+(B) $\theta$ is the value of the ``tilt'' of the molecular bisector from the reference axis, with $\cos(\theta)$ values ranging from -1 to +1 indicating different bisector orientations.
+(C) $\phi$ is the ``twist'' angle formed between the vector normal to the molecular plane (the body-fixed y-axis) and the reference axis.
+For values of $\cos(\theta) \approx 0$, $\phi$ shows how flat the molecule is relative to the interfacial plane.}
+		\label{fig:water-angles}
+	\end{center}
+\end{figure}
+
+%\subsection{Hydrate Complex Analysis}
+
+%During the adsorption process of a gaseous \suldiox~moving in towards a water surface hydrate complexes are formed as the waters bind to the molecule.
+Computationally, there are various methods to determine hydrate complex formation.
+The most simple metric is to count the number of hydrogen bonds formed from the \suldiox~to neighboring molecules.
+This method can be furthered extended to also include the strength of those bonding interactions by additionally reporting the bondlengths to the hydrogen bonding partners.
+Longer bondlengths are typical of weaker bonds, and this method is used in this work to identify the bonding environment of the \suldiox~during solvation.
+
+%A simple count of hydrogen bonds is telling of the interactions and environments of an adsorbing \suldiox.
+However, a more sophisticated analysis is needed to fully characterize the geometry of any complexes that form.
+In this work a graph theoretical method has been used to identify hydrate complexes that form cyclic structures in addition to those that form more simple hydrogen bonding geometries.
+Previously published works have detailed graph theory for use in molecular systems, and have used graph theoretical methods to describe ice and water clusters, structural energies, as well as in the study of relationships in biological systems.\cite{Huber2007,Shi2005,Radhakrishnan1991,Anick2002} As such, we will not present here a review of the basic graph theoretical details that are well documented in the referenced works.
+As shown in figure \ref{fig:so2-complex-graph}, the atoms of a \suldiox~and its nearest neighbors are represented as a set of verrtices, with the covalent and hydrogen bonds between them forming the edges of a graph.
+
+%We present here a technique for determining bonding cycles in a hydrate complex, but this technique is general enough to be applied to any graph representation of simple molecular systems.
+The first step in the determination of cycles is translating the atoms (vertices) and bonds (edges) to a graph.
+Figure \ref{fig:so2-complex-graph} shows the procedure visually, where here the root vertex of the graph is the \suldiox-sulfur, and each atom is numbered by the distance, in atoms, to the root vertex.
+We use only the closest neighboring water molecules that may take part in a cyclic complex in building the graph.
+After forming the graph, a breadth-first search (BFS) is performed starting at the root vertex.\cite{Knuth1997} The BFS algorithm uses a coloring technique where all unexamined vertices are colored white, queued vertices are gray, and examined vertices are black.
+During the BFS, encountering a gray vertex as a target of an adjacent edge during examination indicates that a cycle has been found in the graph.
+At each vertex dequeue for examination, the parent of every vertex is recorded such that the lineage from gray target to root vertex can be reconstructed to determine the participating atoms in the cycle.
+
+%\begin{figure}[h!]
+%\begin{center}
+%\includegraphics[scale=1.0]{images/Cycles/so2complexfigure.png}
+%\caption{A \suldiox~and its nearest neighbors are represented as a graph of vertices and edges.
+The bonds are all given equal weights and the graph is undirected.
+The numbering shows the relative distance from any atom to the root node (\suldiox-sulfur 1), and also enumerates the iteration through the BFS process at which the atom is discovered.
+A BFS on this graph would result in discovering the doubly-bound hydrogen (5) as a gray target, and the gray source would be either of the connected oxygens (4).
+At each iteration the predecessor in the BFS is recorded in order to reconstruct the connectivity of the cycle.
+This cycle is composed of 8 atoms, and 3 molecules}
+%\label{fig:so2-complex-graph}
+%\end{center}
+%\end{figure}

### 2011-05-19 DensityResults.tex
reworked the results section. One overall results section with 3 different subsections.
Rewrote the surface density results section to take out the ambiguity regarding the model used

@@
-\section{Surface Density Distributions}
+\subsection{Surface Density Distributions}
 
@@
 
-In the neat-water slab, the \suldiox~molecular density distribution shows a surface affinity when in a low-concentration environment.
-It does not reside in the bulk or move out of the surface into the gas phase.
-This surface active behavior was found experimentally by the SFG study by our group,\cite{Tarbuck2005,Tarbuck2006} and also verified by the computational simulations and spectral calculations of Baer et al.\cite{Baer2010} The water behavior hypothesized from the experimental analysis is that a surface \suldiox~hydrate complex forms, modifying the structure of water above the Gibbs dividing surface (GDS).
-\wat~free-OH oscillators are thus shifted to stronger net orientation, creating a stronger hydrogen-bonding network, and resulting in an increase of the ``free-OH'' contribution to SFG signal intensity from the surface waters.
+In the simulated neat-water slab, the \suldiox~molecular density distribution indicates an affinity for the water surface.
+During the course of the simulation, the \suldiox~does not venture into the water bulk.
+Also, the \suldiox~does not escape the water surface into the gas phase, but remains located within 5-\angs~of the surface.
+This is consistent with what is found experimentally in our laboratory using VSFS,\cite{Tarbuck2005,Tarbuck2006} and also verified by the computational simulations and spectral calculations of Baer et al.\cite{Baer2010} The experimental studies indicated that upon exposure of \suldiox~to a \wat~surface, a layer of \suldiox~hydrate complex forms, modifying the structure of water above the Gibbs dividing surface (GDS).
+The \suldiox~bonding interaction with the free OH oscillators is manifested in a red-shift of the free OH frequency indicative of a bonding interaction.
 In a later section of this article we look at the effect of \suldiox~on surface water orientation.
 
-Both saturated \suldiox~surfaces exhibit accumulations of \suldiox.
-However, unlike the neat-water slab, the saturated slab has a non-zero bulk concentration of \suldiox~that does not adsorb to the surface.
+The saturated surface simulation results indicate that under the high concentration conditions, \suldiox~accumulation at the surface is increased.
+However, unlike the neat-water slab, the saturated slab has a non-zero bulk concentration of \suldiox.
 The added concentration of \suldiox~creates a layer of molecules bound to the top of the water surface.
 The center of the \suldiox~density distribution is further into the gas phase than for the neat-water surface with a single-\suldiox~molecule.
-The accumulation of \suldiox~in this classical simulation coincides with the experimental conclusions of \suldiox~surface hydrate complex formation.
-However, without simulating breakble bonds the chemistry that may pull \suldiox~into the bulk water phase, and the subsequent reaction to form ions is absent.
-We have confidence in the presence of surface complexation, the orienting effect of the \suldiox~on the surface water molecules, and the accumulation of \suldiox~at the interface.
-The specific geometries of the surface complexes, however, are not as well reproduced with the classical force-field as would be with ab initio methods.\cite{Baer2010} %However, we do not yet conclude that the specific locations of \suldiox~on top of the water surface are as accurate as those found in the DFT ab initio calculations of Baer et al.\cite{Baer2010} As was pointed out in their computations, the classical potential does not reproduce the bonding of the first hydration shell around \suldiox, and so conclusions about specific complex formation or geometries are best made from the ab initio simulation results reported.
+Because of the limitations of the classical model in accurately reproducing the first hydration shell around the \suldiox,\cite{Baer2010} we do not draw specific conclusions about the hydrate complex's geometry.
+However, the surface affinity of the hydrated \suldiox~is well reproduced by both the classical and ab initio methods, and is furthermore verified by our experimental results.
 

### 2011-05-19 Methods.tex
Rewrote the surface density results section to take out the ambiguity regarding the model used

@@
 
-        The location of the water surface was determined by the averaged positions of a set number of the outer-most waters in the simulated aqueous slabs.
-The number of waters included in the averaging was decided by taking a few issues into account.
+        The location of the water surface is determined by the averaged positions of a set number of the outer-most waters in the simulated aqueous slabs.
+The number of waters included in the averaging is determined by taking a few issues into account.
 First, a visual inspection of the top cross-section of the water slab over several timesteps indicated between 65-75 waters established a full monolayer.
@@
 	The simulated systems were analyzed to characterize the orientation of \wat~and \suldiox~in various environments above, within, and below the aqueous surface region.
-In order to better understand the chemistry occurring during the \suldiox~adsorption process, the molecular orientation of both the \wat~and \suldiox~must be understood.
+The molecular orientation of both the \wat~and \suldiox~is a prerequisite for understanding the chemistry occurring during the \suldiox~adsorption process.
 The two molecules studied are similarly shaped with a C2$_v$ axis along their bisectors, and a molecular plane defined by three atoms.

### 2011-06-06 EquilibratedOrientation.tex
reworked the results section. One overall results section with 3 different subsections.
Rewrote some of draft 4, but still need to finish up results and conclusions section.

@@
-\section{Equilibrated MD}
+\subsection{Equilibrated MD}
 
@@
 At each distance from the surface location, an orientation profile was created for both the \wat~and \suldiox~molecules.
-The orientation distribution for the angles $\theta$ and $\phi$ at all depths were combined to form 2D intensity plots that show how the molecular orientation distributions changed with distance to the surface location.
+The orientation distribution for the angles $\theta$ and $\phi$ at all depths were combined to form 2D intensity plots show how the molecular orientation distributions change with distance to the surface location.
 These plots allow for a visual interpretation of how the net orientations are affected when moving from the gas phase to the surface, and then further into the aqueous bulk.
+Both the neat-water, with only a single \suldiox~introduced, and the high-concentration saturated system were analyzed.
+In the case of the neat-water system, the introduction of a single \suldiox~does not greatly affect the water surface, and the results of the water orientation are very similar to a neat-water system without any adsorbed solutes.
 
-\subsection{\wat~Orientation}
+\subsubsection{\wat~Orientation}
 
@@
 
-\subsection{\suldiox~Orientation}
+\subsubsection{\suldiox~Orientation}
 

### 2011-06-06 Introduction.tex
only removed the first sentence of the intro
Rewrote some of draft 4, but still need to finish up results and conclusions section.

@@
 
-The physics of gaseous adsorption on water surfaces is a mostly unexplored field.
-Our current molecular-level understanding of adsorption phenomena relies on experiments and theory of model systems, and on an increasing number of computational simulations of molecular processes.
+Very little work has been done to characterize the molecular processes that take place during the transit of a molecule from the gas phase onto a liquid interfacial region, and how such a surface process is related to uptake.
 What species form during gas adsorption onto liquid surfaces, and what are the intermediary steps?
@@
 Are specific gas or liquid molecular orientations necessary for gaseous adsorption?
-Very little work has been done to characterize the molecular processes that take place during the transit of a molecule from the gas phase onto a liquid interfacial region.
-Furthermore, experiment does not yet provide us with the resolution necessary to determine the geometries of adsorbing gases, or to determine the orientations of the molecules at the liquid surface near the adsorption site.
+Experimental studies to address such questions are valuable but do not provide the resolution necessary to determine the geometries of adsorbing gases, or to determine the orientations of the molecules at the liquid surface near the adsorption site.
+This type of information can be determined computationally, and when coupled to the experimental studies can provide important new insight into the gas-liquid suface adsorption process.
+
 In this work we look at the orientational properties of sulfur dioxide gas adsorbing to a water surface as a model for the transit of gases on to liquid interfaces.
-
 The interactions between sulfur species and aqueous systems are currently an active area of chemical research.
@@
 The same study also showed that low pH aqueous environments inhibit the bulk reaction of \suldiox, but do not affect the surface binding of the gas molecules.
-The resultant experimental spectra suggest that the surface waters reorient in response to the presence of adsorbed \suldiox.
-Knowing that reorientation may occur due to \suldiox~adsorption, how can we confirm it?
+The resultant experimental spectra suggest that the surface waters reorient in response to the presence of adsorbed \suldiox, but the details about this reorientation can not be garnered from the experiments.
+Such details include knowing how surface water orientation changes upon approach of the gaseous \suldiox, how this varies upon \suldiox~surface bonding, and the orientation of \suldiox~throughout the surface bonding process.
+%Knowing that reorientation may occur due to \suldiox~adsorption, how can we confirm it?
 what is the nature of the reorienting waters?
@@
 In this study we look at computed orientational depth profiles of water and \suldiox~before and after the adsorption process.
-Spectroscopic experiments can not probe depth or orientation profiles of surface species in the same detailed manner as computational simulations.
-We determine the effect of introducing \suldiox~to a \wat~surface, and analyze the orientational response of both the \suldiox~and \wat~molecules using equilibrium and steered (SMD) classical molecular dynamics simulations.
-We use a technique of steering a gas molecule into the aqueous phase, and characterize the transit through the interfacial region.
-This unique method gives us much greater insight into the behaviors of gas molecules as they move near to liquid water.
-We also simulate the behavior of adsorbed \suldiox~at equilibrium to show how an already-adsorbed \suldiox~behaves on a water surface.
-The results of this study will help us to understand the behavior of adsorbing gases.
-Computational simulations like this are necessary to further understand the microscopic nature of the gaseous adsorption process, as many of the molecular-level subtleties are lost in experiment.
-Along with other ongoing experiments and computations, we develop a more complete picture of gaseous adsorption on aqueous surfaces.
+The studies augment the spectroscopic experiments that can not probe depth or orientation profiles of surface species in the same detailed manner as computational simulations.
+We determine the molecular effects that occur upon introduction of \suldiox~to a \wat~surface, and analyze the orientational response of both the \suldiox~and \wat~molecules using equilibrium and steered (SMD) classical molecular dynamics simulations.
+With this approach we steer a gas molecule into the aqueous phase, and characterize its molecular orientation as it transits through the interfacial region.
+This unique approach provides insight into the behaviors of gas molecules as they move near to liquid water.
+We also simulate \suldiox~at equilibrium to show how an already-adsorbed \suldiox~behaves on a water surface.
+The results of this study will help us to understand the behavior of adsorbing gases, allowing a more complete picture of gaseous adsorption on aqueous surfaces.

### 2011-06-06 Methods.tex
Rewrote some of draft 4, but still need to finish up results and conclusions section.

@@
 The neat-water system had a single \suldiox~added to the center of the water box, representing a concentration of 0.06 M.
-The saturated system with 22 \suldiox~molecules in the bulk represents a system of concentration 1.35 M, with an additional 50 \suldiox~in the gas phase to model 50 atm of pressure above the water surface.
+The second, more concentrated system examined consisted of 22 \suldiox~molecules in the bulk corresponding to a concentration 1.35 M.
+This saturated system was exposed to an additional 50 \suldiox~in the gas phase above the water surface.
 Both the low and high concentration systems were then evolved for a further 10 ns using a timestep of 0.5fs, with atomic coordinates written every 100 fs.
@@
 The systems were evolved for 1 ns, taking coordinate snapshots every 20 ps to create 50 starting points for further simulations.
-Steered molecular dynamics (SMD) were then performed on the 50 system configurations (in both the neat-water and saturated configurations) to guide the \suldiox~down towards a tethered water 15\angs~under the water slab surface by applying a small steering force to the \suldiox-sulfur atom.
+Steered molecular dynamics (SMD) were then performed on the 50 system configurations (in both the neat-water and saturated configurations) to guide the \suldiox~down towards a tethered water near the water slab's center of mass by applying a small steering force to the \suldiox-sulfur atom.
 This steering technique has been previously developed and used to successfuly model chemical events.\cite{Isralewitz2001} The \suldiox~thus passed through the continuum of environments from gas phase to (neat- and saturated) water surface adsorption, and finally absorption into the bulk of the \wat~slab.
@@
 
-A separate set of SMD simulations were performed with tethering to one of the \suldiox-oxygens.
+A separate set of SMD simulations were performed with tethering of one of the \suldiox-oxygens to the water slab center of mass.
 This was done to ensure that the orientation of the \suldiox~during the adsorption transit was not an artifact of the choice of atom used for tethering.
@@
 
-	In this study we create orientational depth-profiles of \suldiox~and \wat.
+	In this portion of the study we create orientational depth-profiles of \suldiox~and \wat~upon exposure and adsorption of the gas.
 The reference point used for depth calculations is the location of the water surface.
@@
 The number of waters included in the averaging is determined by taking a few issues into account.
-First, a visual inspection of the top cross-section of the water slab over several timesteps indicated between 65-75 waters established a full monolayer.
-Alternatively, assuming a spherical model of water with a radius of 2.2\angs~(the typical length of a bulk-water hydrogen bond), two layers of hexagonally tight-packed spheres would yield a similar number of molecules to the visual inspection.
+First, counting the waters found in the topmost cross-section of the water slab over several timesteps indicated between 65-75 waters that established a full monolayer.
+This was done by a visual inspection of the slab using the VMD MD visualization package.\cite{HUMP96} Alternatively, assuming a spherical model of water with a radius of 2.2\angs~(the typical length of a bulk-water hydrogen bond), two layers of hexagonally tight-packed spheres yielded a similar number of surface water molecules.
 Increasing the number of waters used in calculating the surface location diminishes the effects of the few waters that briefly rise above the surface into the gas phase, stabilizing both the surface position and thickness values.
@@
 Figure \ref{fig:density-flaw} demonstrates the problem of surface location drift during a simulation.
-This figure shows the density profile of water and \suldiox~over the course of one of the 10 ns trajectories used in this work.
-The water density profile and location (dashed line) is produced by averaging the instantaneous density profile at each timestep in the simulation.
-The water profile was then fit to a $\tanh$ function (black) to extract the position and width parameters of the water surface.
-The fitted Gibbs dividing surface has a ``90-10'' width of 8.28 $\AA$.
-A bulge in the gas-phase (left) side of the water density profile is indicative of the drift of the water slab over the course of the trajectory.
+Figure \ref{fig:density-flaw}a shows the density profile of water and \suldiox~over the course of one of the 10 ns trajectories used in this work using the original uncorrected coordinates of the system.
+The water density profile and location (dashed line in figure \ref{fig:density-flaw}a) is produced by averaging the instantaneous density profile at each timestep in the simulation over all the timesteps.
+The water profile was then fit to a $\tanh$ function (black line) to extract the position and width parameters of the water surface.
+The fitted Gibbs dividing surface has a width of 3.77\angs, which is comparable to values reported for similar neat-water systems.\cite{Dang1997,Hore2008} A bulge in the gas-phase ($\leq 52\AA$) side of the water density profile is indicative of the drift of the water slab over the course of the trajectory.
 Thus the calculated location and width from the $\tanh$ line fit are not accurate over long trajectories for defining a stationary reference point.
-All depth locations in our analyses are calculated relative to the surface location (blue) at the corresponding timestep of the trajectories.
+
+The new procedure described previously for finding the outer monolayer location was implemented, and the surface location is plotted as a function of simulation time in figure \ref{fig:density-flaw}b.
+It is apparent from the surface location plot that the water slab location, and thus the surface location, drifts over the 10 ns, spanning approx.
+12 $\AA$.
+However, the maximum standard deviation of the positions of the waters comprising the surface layer at each timestep is only 1.85 $\AA$.
+All depth locations in our analyses are calculated relative to the instantaneous surface location at the corresponding timestep of the trajectories.
 
@@
 		\caption{(A) Density profiles of \wat~(dashed) and \suldiox~(red) from a 10 ns simulation of the neat-water system with a single sulfur dioxide.
-The \wat~density profile was fit (black) using a hyperbolic tangent function to extract location and width data for the water surface.
-(B) The location of the outer \wat~monolayer was calculated for each simulation timestep (blue), and the surface location from the density fitting was also determined (dashed).
+The \wat~density profile was fit using a hyperbolic tangent function (black) to extract location and width data for the water surface.
+(B) The instantaneous location of the outer \wat~monolayer was calculated for each simulation timestep (blue), and the surface location extracted from the density fitting was also shown for reference (horizontal dashed line).
 Long simulations of liquid slabs result in drift of the slab location, and the consequent broadening of the water density profile if the surface location is not determined repeatedly throughout the simulation.}
@@
 \end{figure}
-
-The new procedure described above for finding the outer monolayer location was implemented, and the distribution of that location is also plotted in figure \ref{fig:density-flaw} (blue).
-It is apparent from the surface location distribution that the water slab drifts over the 10 ns, spanning approx.
-12 $\AA$.
-However, the maximum standard deviation of the positions of the waters comprising the surface layer at each timestep is only 1.85 $\AA$.
 

### 2011-06-07 Conclusions.tex
Finished first pass of fixing up draft 4.

@@
 \section {Conclusions}
+
+% intro-type paragraph here
+The aqueous chemistry taking place in the Earth's atmosphere is goverened by the physics of the water surfaces found there.
+Many recent experiments have revealed specific pathways for reaction of \suldiox, but the molecular-level nature of those reactions are still poorly understood.
+Through computer simulations we are able to probe the microscopic environments in which those reactions occur, and better understand the conditions in which \suldiox~reacts.
 
@@
 Our simulations recreated the experimental setup of a recent study by our group,\cite{Ota2011} by introducing a concentrated gas phase of \suldiox~to a water surface.
-It was found that gaseous \suldiox~quickly adsorbs to the water surface, and continues to bind until a complete surface coverage is reached.
+Through simulation it was found that gaseous \suldiox~quickly adsorbs to the water surface, and continues to bind until a complete surface coverage is reached.
 Surface waters reorient in the presence of adsorbed \suldiox.
@@
 
-The surface behavior of the \suldiox~molecules was also determined.
-Vibrational sum-frequency surface spectroscopy on the \suldiox~vibrational spectra is not possible because of the nature of the experimental geometry.
-At aqueous interfaces, the spectra of the surface \suldiox~vibrational modes are still unreported.
-However, through these simulations we are able to characterize the orientational behavior of \suldiox~during and after adsorption.
+The surface behavior of the \suldiox~molecules was also determined, information that is difficult to obtain with vibrational sum-frequency spectroscopy.
+Through these simulations we are able to characterize the orientational behavior of \suldiox~during and after adsorption.
 Our equilibrium simulations show that a single \suldiox~molecule, representing a low concentration, has a high surface affinity.
-This was previously reported by Baer et al.\cite{Baer2010} At a high \suldiox~concentration, unreacted \suldiox~molecules are found further out of the water phase as a bound layer of \suldiox~crowds the surface and binds all the surface waters.
+At a high \suldiox~concentration, unreacted \suldiox~molecules are found further out of the water phase as a bound layer of \suldiox~crowds the surface and binds all the surface waters.
 The orientation of \suldiox~on the water surface was found to be similar for both low and high concentrations.

### 2011-06-07 EquilibratedOrientation.tex
Finished first pass of fixing up draft 4.

@@
 
-\subsubsection{\wat~Orientation}
-
-The orientation depth-profiles for \wat~are shown in figure \ref{fig:water-orientation} for both the neat-water (top) and saturated (bottom) systems during the equilibrium MD simulations.
-In both systems the strongest orientational preference is found at the slab surfaces where the water is furthest towards the gas phase.
+The 2D orientational depth profile plots show areas of low intensity in dark blue, and highest intensity in dark red.
+Regions of the plots where the intensity (coloration) is equally distributed from top to bottom along the entire orientational range are considered isotropic.
+Likewise, areas of the plot with high intensity over a small orientational range are considered to exhibit an orientational preference at the given depth.
 The histograms are arranged with the plots of $\cos(\theta)$ on the left and $\cos(\phi)$ on the right columns.
@@
 
-The bisector tilt, $\cos(\theta)$, concentrates around $\cos(\theta)=0$ within the first few\angs~of the surface, and then the distribution becomes isotropic further into the water bulk.
+\subsubsection{\wat~Orientation}
+
+The orientation depth-profiles for \wat~are shown in figure \ref{fig:water-orientation} for both the neat-water (top) and saturated (bottom) systems during the equilibrium MD simulations.
+In both systems the strongest orientational preference is found at the slab surfaces (distances greater than 0\angs) where the water is furthest towards the gas phase.
+The narrow "penninsula" of high intensity that extends in the $\cos(\theta)$ plots to the right of the -5\angs~location shows the overall preference of water to orient at the surface.
+The $\cos(\phi)$ plots are similar to each other with a narrow region of reorientation, but the effect of the surface is greater in the neat-water system as evidenced by the sharper transition in intensity from blue to red, compared to the saturated system that has a less pronounced intensity change.
+
+The bisector tilt of the water molecules, $\cos(\theta)$, concentrates around $\cos(\theta)=0$ within the first few\angs~above the surface, and then the distribution becomes isotropic further into the water bulk of both systems.
 As the tilt nears $\cos(\theta)=0$ the \wat~bisector lies within the plane of the surface indicating a water orientation either flat on the surface, or with some amount of ``twist'' sending the OH bonds in towards, or out of the bulk.
 The value of $\phi$ determines the ``twist'' in this case.
-Both systems show a peak in the distributions around $\cos(\phi)=1$ at the water surfaces.
+Both systems show a defined intensity concentration (dark red) in the distributions around $\cos(\phi)=1$ at the water surfaces.
 This results from an orientation of the water's y-axis (normal to the molecular plane) aligned perpendicular to the plane of the water surface.
@@
 The waters located above 0\angs~are further into the gas phase, and interact with the layer of adsorbed \suldiox~molecules.
-The \suldiox~layer does not provide the same amount of strong hydrogen bonding and solvation that the waters experience further in the aqueous phase.
 The resulting orientation of waters above 0\angs, shown in the saturated plot of $\cos(\theta)$ of figure \ref{fig:water-orientation}, is with a bisector pointing further into the adsorbed \suldiox~gas layer, and both hydrogens pointing outward from the aqueous bulk.
@@
 
-The change in $\cos(\theta)$ is not as rapid at the neat-water surface as in the saturated system.
-Furthermore, the noise in the distribution above 5\angs~in the neat-water system is a result of fewer waters venturing beyond those extents and thus less data far from the surface.
+The noise in the neat-water plots of figure \ref{fig:water-orientation} above 5\angs~(manifested as disconnected points of intensity throughout the range of orientations) is a result of fewer waters venturing beyond those extents and thus less data far from the surface.
 This is one indication that waters near a layer of adsorbed \suldiox~venture further above the interface and interact with the adsorbed gas molecules.
-The interactions with neighboring \suldiox~molecules allow the waters above the surface to orient more perpendicular to the interface.
+The interactions with neighboring \suldiox~molecules allow the waters above the surface to orient more perpendicularly to the interface.
 Our experimental SFG studies also predicted this reorienting behavior due to the \suldiox~interactions with the topmost surface waters.\cite{Ota2011}
@@
 		\caption{Molecular orientation histograms of \wat~throughout the surface equilibrated systems.
-The top surface is located at a distance of 0 with negative positions in the bulk of the slab.
-The bottom slab surface is approx.
-30\angs~below the top surface.
+The top surface is located at a distance of 0\angs~with negative distance values located in the bulk of the slab.
 Shown are the angle distributions for $\theta$ (left column) and $\phi$ (right column) in both the neat-\wat~system (top row) and the saturated system (bottom row).
-The distributions are normalized to account for the changing number of water molecules at different positions in the system.}
+The distributions are normalized to account for the changing number of water molecules at different positions in the system.
+Regions of high intensity are dark red, and low intensity are dark blue.
+The scattered points of coloration to the far right of each plot indicates that few waters were located at those positions, and thus few data points were collected.}
 		\label{fig:water-orientation}
@@
 The \suldiox~orientation data set for the neat-water system is much smaller as only a single \suldiox~molecule was simulated in the bulk.
-The resulting distribution plots are thus noisier than the corresponding saturated plots, but effective comparisons can still be drawn.
+The resulting distribution plots are thus noisier than the corresponding saturated plots with more scattered points of high intensity, but effective comparisons can still be drawn.
 
-Within 5\angs~of the water surface (the approximate distance from the surface that \suldiox~begins interacting with the outer-most waters) the angular distribution of the single \suldiox~(in the neat-water system) is concentrated primarily in $\cos(\theta)>0$.
+Within 0-5\angs~of the water surface (the approximate distance from the surface that \suldiox~begins interacting with the outer-most waters) the angular distribution of the single \suldiox~(in the neat-water system) is concentrated primarily in $\cos(\theta)>0$.
 The peak of the distribution occurs at $\cos(\theta)=1$.
-The \suldiox~bisector points out of the water surface, with the sulfur atom pointing towards the aqueous bulk, and the two oxygens pointing into the gas phase.
+This indicates that the \suldiox~bisector points out of the water surface, with the sulfur atom pointing towards the aqueous bulk, and the two oxygens pointing into the gas phase.
 This same distribution occurs in the saturated system for positions below 5\angs.
@@
 
+The plots indicate that \suldiox~orients similarly at both the low and high concentration water surface environments.
+The plots of $\cos(\phi)$ for both concentrations show a isotropic behavior.
+However, the $\theta$ values follow similar trends at both concentrations indicating that the \suldiox~sulfur orients down towards the water phase, with both oxygens pointing away from the surface once the \suldiox~is within approx.
+5\angs~of the surface.
+This behavior continues down to at least 5\angs~below the surface, notwithstanding any chemical reactions that may occur that are not simulated using the classical MD techniques utilized in this work.
+
 \begin{figure}[h!]

### 2011-06-07 TransitOrientation.tex
reworked the results section. One overall results section with 3 different subsections.
Finished first pass of fixing up draft 4.

@@
-\section{Steered MD Transit Simulations}
+\subsection{Steered MD Transit Simulations}
 
-\subsection {\suldiox~Orientation}
+\subsubsection {\suldiox~Orientation}
 
@@
 	From its starting position 20\angs~above the water surface, until the \suldiox~moves to within 10\angs~of the surface of both systems, the orientation is isotropic in $\theta$ and $\phi$.
+This can be seen in the plots as mostly uniform coloration at a given distance from the surface independent of $\cos$ of $\phi$.
 Near the surface the bisector angle $\theta$ becomes more perpendicular ($\cos(\theta)\approx1$) with the \suldiox~sulfur pointing into the water phase.
@@
   % talk about theta differences
-  The $\cos(\theta)$ distribution above 10\angs~from the surface is isotropic.
-There is a transition below 10\angs~where the \suldiox~orients with its bisector perpendicular to the water surface.
-On approach to the neat-water surface, that transition occurs at approximately 5\angs~from the surface.
+  Looking more closely at the approach, the $\cos(\theta)$ distribution above 10\angs~from the surface is isotropic.
+For the neat-water surface, a transition occurs at approximately 5\angs~from the surface where the \suldiox~orients with its bisector perpendicular to the water surface.
 This coincides with the approximate distance of bonds formed to the outer-most water molecules.
@@
 The layer of adsorbed \suldiox~already present in the saturated system most likely interacts with the transiting \suldiox~molecule.
-Also, topmost water molecules from the surface move up to a few\angs~inside the adsorbed \suldiox~layer and interact with the transiting \suldiox~further from the surface than those in the neat-water system.
+Also, topmost water molecules from the surface move up to a few \angs~inside the adsorbed \suldiox~layer and interact with the transiting \suldiox~further from the surface than those in the neat-water system.
+It is remarkable that the orientational trend appears so strong in the $\cos(\theta)$ plots even with so little data as was collected from the single \suldiox~molecule of each simulation.
+From onset of orientation above the surface until 10\angs~below, the \suldiox~holds a preferred orientation.
 
@@
   % talk about phi differences
-  With a mostly perpendicular bisector angle, it is expected that the values of $\phi$ would be isotropic relative to the reference axis.
-This is the case in both systems, with one exception in the neat-water system.
-At the neat-water surface and just below, from 0\angs~to 5\angs~in the water phase, the $\theta$ profile broadens, and a clear trend appears in the profile of $\phi$.
-This corresponds to a \suldiox~orienting less flat to the surface (inclined up to 60 degrees from the surface normal).
+	%Indeed, the $\phi$ orientations appear mostly isotropic without any clear trends for preferred orientation in that angle.
+This appears as a mostly uniform color across the entire plot, with minor variations that are disjointed throughout the distance range.
+
+ With a mostly perpendicular bisector angle, it is expected that the values of $\phi$ for \suldiox~would be isotropic relative to the reference axis.
+This is the case in both systems, with only a few exceptions.
+At the neat-water surface and just below, from 0\angs~to 5\angs~in the water phase, the $\theta$ profile broadens down to $\cos(\theta)=0.5$, and a clear trend appears in the profile of $\phi$ near $\cos(\phi)=0$.
+This corresponds to \suldiox~inclined up to 60 degrees from the surface normal.
 The result is that \suldiox~in this region shows a preference for one of the S-O bonds to point more towards the aqueous bulk and the other to point away towards the interface.
@@
 The presence of the adsorbed \suldiox~layer apparently decreases the reorienting behavior likely because of the disrupting effect the higher \suldiox~concentration has on the water interactions at the surface.
-Below this 5\angs~region under the surface, the \suldiox~resumes an isotropic orientation as it becomes fully solvated by bulk waters.
+%Below this 5\angs~region under the surface, the \suldiox~resumes an isotropic orientation as it becomes fully solvated by bulk waters.
 

### 2011-06-07 EquilibratedOrientation.tex
Draft 5

@@
 At each distance from the surface location, an orientation profile was created for both the \wat~and \suldiox~molecules.
-The orientation distribution for the angles $\theta$ and $\phi$ at all depths were combined to form 2D intensity plots show how the molecular orientation distributions change with distance to the surface location.
+The orientation distribution for the angles $\theta$ and $\phi$ at all depths were combined to form 2D intensity plots that show how the molecular orientation distributions change with distance to the surface location.
 These plots allow for a visual interpretation of how the net orientations are affected when moving from the gas phase to the surface, and then further into the aqueous bulk.
 Both the neat-water, with only a single \suldiox~introduced, and the high-concentration saturated system were analyzed.
-In the case of the neat-water system, the introduction of a single \suldiox~does not greatly affect the water surface, and the results of the water orientation are very similar to a neat-water system without any adsorbed solutes.
+In the case of the neat-water system, the introduction of a single \suldiox~does not greatly affect the water surface, and the results of the water orientation are very similar to a neat-water system without any adsorbed solutes (not shown).
 
@@
 
-The plots of $\cos(\phi)$ are both isotropic, although the neat-water system plot is quite noisy from the small data set.
+The plots of $\phi$ are both isotropic, although the neat-water system plot is quite noisy from the small data set.
 Because the \suldiox~near the surface is oriented perpendicularly to the interface, the $\cos(\phi)$ orientation is expected to be isotropic.
@@
 The plots indicate that \suldiox~orients similarly at both the low and high concentration water surface environments.
-The plots of $\cos(\phi)$ for both concentrations show a isotropic behavior.
+The plots of $\phi$ for both concentrations exhibit isotropic distributions.
 However, the $\theta$ values follow similar trends at both concentrations indicating that the \suldiox~sulfur orients down towards the water phase, with both oxygens pointing away from the surface once the \suldiox~is within approx.

### 2011-06-07 Methods.tex
Draft 5

@@
 The water profile was then fit to a $\tanh$ function (black line) to extract the position and width parameters of the water surface.
-The fitted Gibbs dividing surface has a width of 3.77\angs, which is comparable to values reported for similar neat-water systems.\cite{Dang1997,Hore2008} A bulge in the gas-phase ($\leq 52\AA$) side of the water density profile is indicative of the drift of the water slab over the course of the trajectory.
+The fitted Gibbs dividing surface has a width of 3.77\angs, which is comparable to values reported for similar neat-water systems.\cite{Dang1997,Hore2008} A bulge in the gas-phase ($\leq 52$\angs) side of the water density profile is indicative of the drift of the water slab over the course of the trajectory.
 Thus the calculated location and width from the $\tanh$ line fit are not accurate over long trajectories for defining a stationary reference point.

### 2011-06-07 TransitOrientation.tex
Draft 5

@@
 	The orientation of \suldiox~molecules throughout the aqueous adsorption process was monitored during the transit SMD simulations.
-The angles $\theta$ and $\phi$ (figure \ref{fig:water-angles}) were calculated for each timestep of the SMD simulations as the transiting \suldiox~was pulled into the water slab from the gas phase, both in the neat-water and saturated slab systems.
+The angles $\theta$ and $\phi$ of the transiting \suldiox~(figure \ref{fig:water-angles}) were calculated for each timestep of the SMD simulations as the \suldiox~was pulled into the water slab from the gas phase, both in the neat-water and saturated slab systems.
 The angle cosine values were collected for the 50 simulations of both systems for each distance from the water surface, resulting in the 2-dimensional angle and depth-profile histograms shown in figure \ref{fig:so2-transit-angles}.
@@
 This can be seen in the plots as mostly uniform coloration at a given distance from the surface independent of $\cos$ of $\phi$.
-Near the surface the bisector angle $\theta$ becomes more perpendicular ($\cos(\theta)\approx1$) with the \suldiox~sulfur pointing into the water phase.
+Near the surface the bisector angle $\theta$ becomes more perpendicular to the water surface ($\cos(\theta)\approx1$) with the \suldiox~sulfur pointing into the water phase.
 At the point when the \suldiox~reaches the surface location (distance $=0$\angs), the bisector is perpendicular in both the neat-water and saturated systems.
@@
   % talk about theta differences
-  Looking more closely at the approach, the $\cos(\theta)$ distribution above 10\angs~from the surface is isotropic.
+  Looking more closely at the approach, the $\theta$ distribution above 10\angs~from the surface is isotropic.
 For the neat-water surface, a transition occurs at approximately 5\angs~from the surface where the \suldiox~orients with its bisector perpendicular to the water surface.
@@
 Also, topmost water molecules from the surface move up to a few \angs~inside the adsorbed \suldiox~layer and interact with the transiting \suldiox~further from the surface than those in the neat-water system.
-It is remarkable that the orientational trend appears so strong in the $\cos(\theta)$ plots even with so little data as was collected from the single \suldiox~molecule of each simulation.
+It is remarkable that the orientational trend appears so strongly in the $\theta$ plots even with so little data as was collected from the single \suldiox~molecule of each simulation.
 From onset of orientation above the surface until 10\angs~below, the \suldiox~holds a preferred orientation.

### 2011-06-22 Introduction.tex
fixed draft 5 Introduction

@@
 
-Very little work has been done to characterize the molecular processes that take place during the transit of a molecule from the gas phase onto a liquid interfacial region, and how such a surface process is related to uptake.
+The doorway to the uptake of a gas by an aqueous solution is the water surface.
+Although we know much about the behavior of a gas on either side of that entrance, far less is known about how that surface acts to attract, facilitate, or thwart the transit of the gas between the two bulk phases.
 What species form during gas adsorption onto liquid surfaces, and what are the intermediary steps?
@@
 
-In this work we look at the orientational properties of sulfur dioxide gas adsorbing to a water surface as a model for the transit of gases on to liquid interfaces.
+In this work we look at the orientational properties of sulfur dioxide gas adsorbing to a water surface as a model for the transit of gases across an aqueous surface.
 The interactions between sulfur species and aqueous systems are currently an active area of chemical research.
@@
 
-Our recent vibational sum frequency spectroscopy (vsfs) experiments provided insight into the adsorption and reaction of gases in atmospheric aerosols.\cite{Tarbuck2005,Tarbuck2006} We concluded that an \suldiox~surface hydrate complex forms when an aqueous surface is exposed to \suldiox~gas.
+Our recent vibational sum frequency spectroscopy (VSFS) experiments have provided insight into the adsorption and reaction of gases in atmospheric aerosols.\cite{Tarbuck2005,Tarbuck2006} These studies showed that an \suldiox~surface hydrate complex forms when an aqueous surface is exposed to \suldiox~gas.
 A computational study by Baer et al.\cite{Baer2010} then made a series of predictions of the specific nature of the hydrated complex through classical and ab initio simulations.
@@
 
-Our latest vsfs experiments on \suldiox~behavior further expand our understanding of \suldiox~in atmospheric processes.\cite{Ota2011} A temperature study showed that the binding of gaseous \suldiox~to water surfaces is greatly enhanced at atmospherically relevant cold temperatures.
-The surface binding of unreacted \suldiox~complexes was found to be a completely reversible process.
+Our latest VSFS experiments on \suldiox~behavior further expand our understanding of \suldiox~in atmospheric processes.\cite{Ota2011} A temperature study has shown that the binding of gaseous \suldiox~to water surfaces is greatly enhanced at atmospherically relevant cold temperatures.
+The surface binding of unreacted \suldiox~complexes is found to be a completely reversible process.
 The same study also showed that low pH aqueous environments inhibit the bulk reaction of \suldiox, but do not affect the surface binding of the gas molecules.
-The resultant experimental spectra suggest that the surface waters reorient in response to the presence of adsorbed \suldiox, but the details about this reorientation can not be garnered from the experiments.
+Furthermore, the resultant experimental spectra suggest that the surface waters reorient in response to the presence of adsorbed \suldiox, but the details about this reorientation can not be garnered from the experiments.
 Such details include knowing how surface water orientation changes upon approach of the gaseous \suldiox, how this varies upon \suldiox~surface bonding, and the orientation of \suldiox~throughout the surface bonding process.
@@
 
-In this study we look at computed orientational depth profiles of water and \suldiox~before and after the adsorption process.
-The studies augment the spectroscopic experiments that can not probe depth or orientation profiles of surface species in the same detailed manner as computational simulations.
-We determine the molecular effects that occur upon introduction of \suldiox~to a \wat~surface, and analyze the orientational response of both the \suldiox~and \wat~molecules using equilibrium and steered (SMD) classical molecular dynamics simulations.
-With this approach we steer a gas molecule into the aqueous phase, and characterize its molecular orientation as it transits through the interfacial region.
-This unique approach provides insight into the behaviors of gas molecules as they move near to liquid water.
-We also simulate \suldiox~at equilibrium to show how an already-adsorbed \suldiox~behaves on a water surface.
-The results of this study will help us to understand the behavior of adsorbing gases, allowing a more complete picture of gaseous adsorption on aqueous surfaces.
+We provide such detailed informatino to complement and augment these experimental studies.
+Our studies involve computation of the orientational depth profiles of water and \suldiox~before and after the adsorption process.
+They augment the spectroscopic experiments that can not probe depth or orientation profiles of surface species in the same detailed manner as computational simulations.
+The molecular behaviors that occur upon introduction of \suldiox~to a \wat~surface are determined, and we analyze the orientational response of both the \suldiox~and \wat~molecules using equilibrium and steered (SMD) classical molecular dynamics simulations.
+The latter approach involves steering a gas molecule into the aqueous phase, and characterizing its molecular orientation as it transits through the interfacial region.
+This unique approach enables new insights into the behaviors of gas molecules as they move near to liquid water to be obtained.
+We also simulate how \suldiox~adsorption occurs on a water surface saturated with adsorbed \suldiox, analogous to the cold \suldiox~experimental studies recently performed.\cite{Ota2011} The results of this study provide an intimate perspective on the adsorption of \suldiox~at an aqueous surface, and a more complete picture of gaseous adsorption to liquid interfaces.

### 2011-06-23 DensityResults.tex
Rewrote some of draft 4, but still need to finish up results and conclusions section.
Draft 5 revisions - finished Geri's edits. Time for a re-read.

@@
 One measure of surface activity is the spatial distribution of molecules in the interfacial region.
-The density distributions of both \wat~and \suldiox~were calculated for the equilibrium MD simulations, and the results are presented in figure \ref{fig:density} for both the neat-water and saturated simulations.
-The \suldiox~in the neat-water system remained at the top slab surface.
-The \suldiox~in the saturated system accumulated mostly at the surface, but some residual \suldiox~remained in the bulk.
+The density distributions of both \wat~and \suldiox~were calculated for the equilibrium MD simulations.
+The results presented in figure \ref{fig:density} show both the water and \suldiox~density distributions across the two simulated slabs.
+As shown, the \suldiox~exposed to the neat-water system remained at the water slab surface.
+The \suldiox~in the saturated system accumulated mostly at the surface, but some residual \suldiox~remained well into the bulk water.
 
@@
 
-In the simulated neat-water slab, the \suldiox~molecular density distribution indicates an affinity for the water surface.
-During the course of the simulation, the \suldiox~does not venture into the water bulk.
+In the simulated neat-water slab, the \suldiox~molecular density distribution concentrates near the water surface, indicating an affinity for the interfacial region.
+This is consistent with what is found experimentally in our laboratory using VSFS,\cite{Tarbuck2005,Tarbuck2006} and also verified by the computational simulations and spectral calculations of Baer et al.\cite{Baer2010} The experimental studies indicated that upon exposure of \suldiox~to a \wat~surface, a layer of \suldiox~hydrate complex forms, modifying the structure of water above the Gibbs dividing surface (GDS).
+In the experimental system, the \suldiox~bonding interaction with the free OH oscillators is manifested in a red-shift of the free OH frequency indicative of a bonding interaction.
+In a later section of this article we look at the effect of \suldiox~on surface water orientation.
+During the course of the simulation, the \suldiox~does not venture into the bulk water.
 Also, the \suldiox~does not escape the water surface into the gas phase, but remains located within 5-\angs~of the surface.
-This is consistent with what is found experimentally in our laboratory using VSFS,\cite{Tarbuck2005,Tarbuck2006} and also verified by the computational simulations and spectral calculations of Baer et al.\cite{Baer2010} The experimental studies indicated that upon exposure of \suldiox~to a \wat~surface, a layer of \suldiox~hydrate complex forms, modifying the structure of water above the Gibbs dividing surface (GDS).
-The \suldiox~bonding interaction with the free OH oscillators is manifested in a red-shift of the free OH frequency indicative of a bonding interaction.
-In a later section of this article we look at the effect of \suldiox~on surface water orientation.
 
-The saturated surface simulation results indicate that under the high concentration conditions, \suldiox~accumulation at the surface is increased.
+The saturated solution simulation results indicate that under the high concentration of \suldiox, the \suldiox~accumulation at the surface is increased.
 However, unlike the neat-water slab, the saturated slab has a non-zero bulk concentration of \suldiox.
@@
 Because of the limitations of the classical model in accurately reproducing the first hydration shell around the \suldiox,\cite{Baer2010} we do not draw specific conclusions about the hydrate complex's geometry.
-However, the surface affinity of the hydrated \suldiox~is well reproduced by both the classical and ab initio methods, and is furthermore verified by our experimental results.
+However, the surface affinity of the hydrated \suldiox~is well reproduced by both the classical and ab initio methods.
 
@@
 		\includegraphics[scale=1.0]{images/density/density.png}
-		\caption{Molecular density distributions of \wat~and \suldiox~calculated along the long-axis of the simulated cells.
-The position axis shows the distance of the given molecule from the top surface of the \wat~slab, with positive values located above the slab towards the gas phase, and negative values located in the water bulk.
-Distributions of both the neat-water simulation with a single \suldiox~(left, scaled 10x for clarity), and the saturated system (right) are shown.}
+		\caption{Molecular density distributions of \wat~(black) and \suldiox~(red) calculated along the long-axis of the simulated cells.
+The position axis shows the distance from the instantaneous surface location of the \wat~slab, with positive values located above the slab towards the gas phase, and negative values located in the water bulk.
+Distributions of both the neat-water simulation with a single \suldiox~(left, with the \suldiox~density scaled 10x for clarity), and the saturated system (right) are shown.}
 		\label{fig:density}

### 2011-06-23 EquilibratedOrientation.tex
Draft 5 revisions - finished Geri's edits. Time for a re-read.

@@
 In both systems the strongest orientational preference is found at the slab surfaces (distances greater than 0\angs) where the water is furthest towards the gas phase.
-The narrow "penninsula" of high intensity that extends in the $\cos(\theta)$ plots to the right of the -5\angs~location shows the overall preference of water to orient at the surface.
+Previous work on orientational preference of water at air surfaces shows the same trend as our neat-water results.\cite{Walker2006b,Hore2008} The narrow "penninsula" of high intensity that extends in the $\cos(\theta)$ plots to the right of the -5\angs~location shows the overall preference of water to orient at the surface.
 The $\cos(\phi)$ plots are similar to each other with a narrow region of reorientation, but the effect of the surface is greater in the neat-water system as evidenced by the sharper transition in intensity from blue to red, compared to the saturated system that has a less pronounced intensity change.
 
-The bisector tilt of the water molecules, $\cos(\theta)$, concentrates around $\cos(\theta)=0$ within the first few\angs~above the surface, and then the distribution becomes isotropic further into the water bulk of both systems.
+The bisector tilt of the water molecules, $\cos(\theta)$, concentrates around $\cos(\theta)=0$ within the first few\angs~above and below the surface, becoming progressively isotropic further into the water bulk of both systems.
 As the tilt nears $\cos(\theta)=0$ the \wat~bisector lies within the plane of the surface indicating a water orientation either flat on the surface, or with some amount of ``twist'' sending the OH bonds in towards, or out of the bulk.
@@
 This results from an orientation of the water's y-axis (normal to the molecular plane) aligned perpendicular to the plane of the water surface.
-Thus, both the neat-water and saturated systems have waters lying mostly flat to the plane of the interface at a distance of 0\angs.
+Thus, for both the neat-water and saturated systems, the preferred orientation of water molecules at a distance of 0\angs~ is to lie mostly flat to the plane of the interface.
 

### 2011-06-23 Methods.tex
Draft 5 revisions - finished Geri's edits. Time for a re-read.

@@
 
-The neat-water simulation involved the addition of a single \suldiox~molecule either within the bulk of the water slab (for equilibrium MD), or above the slab surface (in the SMD simulations).
-The single-\suldiox~system was then evolved for 2 ns to produce an equilibrated starting configuration.
-The saturated system had 22 \suldiox~molecules in the water slab in order to saturate it to a level coinciding with the Henry's law constant for \suldiox~in water.
-Additionally, 50 \suldiox~molecules were introduced into the gas phase outside of the water slab, on both surfaces, to simulate an added 50 atm of \suldiox~gas pressure.
+The low and high concentration simulated systems, ``neat-water'' and saturated, respectively, were created as follows: the ``neat-water'' simulation involved the addition of a single \suldiox~molecule either within the bulk of the water slab (for equilibrium MD), or above the slab surface (in the SMD simulations).
+The single-\suldiox~neat-water system was then evolved for 2 ns to produce an equilibrated starting configuration.
+Because of the extremely low concentration of \suldiox~in the neat-water system the surface waters behave similarly to a true ``neat-water'' air-liquid interface.
+The neat-water system orientational results shown later in this work reproduce well the results of our previous orientational studies of surface water behavior.\cite{Walker2006b,Hore2008} The saturated system had 22 \suldiox~molecules introduced to the water slab bulk in order to saturate it to a level coinciding with the Henry's law constant for \suldiox~in water.
+Additionally, 50 \suldiox~molecules were introduced into the gas phase outside of the saturated water slab to simulate an \suldiox~gas pressure.
 The saturated system was then evolved for 2 ns to produce a starting configuration for further saturated simulations.
@@
 The neat-water system had a single \suldiox~added to the center of the water box, representing a concentration of 0.06 M.
-The second, more concentrated system examined consisted of 22 \suldiox~molecules in the bulk corresponding to a concentration 1.35 M.
+The more concentrated ``saturated'' system consisted of 22 \suldiox~molecules in the bulk corresponding to a concentration 1.35 M.
 This saturated system was exposed to an additional 50 \suldiox~in the gas phase above the water surface.
-Both the low and high concentration systems were then evolved for a further 10 ns using a timestep of 0.5fs, with atomic coordinates written every 100 fs.
+Both the low and high concentration systems were then evolved for a further 10 ns using a timestep of 0.5fs, with atomic coordinates recorded every 100 fs.
 
@@
 Similarly, the averaged values of location and width will obscure information about any drift or deformations the surface undergoes.
+
+Figure \ref{fig:density-flaw} demonstrates the problem of surface location drift during a simulation.
+Figure \ref{fig:density-flaw}a shows the density profile of water and \suldiox~over the course of one of the 10 ns trajectories used in this work using the original uncorrected coordinates of the system.
+The water density profile and location (dashed line in figure \ref{fig:density-flaw}a) is produced by averaging the instantaneous density profile at each timestep in the simulation over all the timesteps.
+The water profile was then fit to a $\tanh$ function (black line) to extract the position and width parameters of the water surface.
+The fitted Gibbs dividing surface has a width of 3.77\angs, which is comparable to values reported for similar neat-water systems.\cite{Dang1997,Hore2008} A bulge in the gas-phase ($\leq 52$\angs) side of the water density profile is indicative of the drift of the water slab over the course of the trajectory.
+Thus the calculated location and width from the $\tanh$ line fit are not accurate over long trajectories for defining a stationary reference point.
+
 	In this work we define the surface location by calculating a reference location at each timestep by averaging the positions of the waters contained in the topmost monolayer.
@@
 
-Figure \ref{fig:density-flaw} demonstrates the problem of surface location drift during a simulation.
-Figure \ref{fig:density-flaw}a shows the density profile of water and \suldiox~over the course of one of the 10 ns trajectories used in this work using the original uncorrected coordinates of the system.
-The water density profile and location (dashed line in figure \ref{fig:density-flaw}a) is produced by averaging the instantaneous density profile at each timestep in the simulation over all the timesteps.
-The water profile was then fit to a $\tanh$ function (black line) to extract the position and width parameters of the water surface.
-The fitted Gibbs dividing surface has a width of 3.77\angs, which is comparable to values reported for similar neat-water systems.\cite{Dang1997,Hore2008} A bulge in the gas-phase ($\leq 52$\angs) side of the water density profile is indicative of the drift of the water slab over the course of the trajectory.
-Thus the calculated location and width from the $\tanh$ line fit are not accurate over long trajectories for defining a stationary reference point.
-
-The new procedure described previously for finding the outer monolayer location was implemented, and the surface location is plotted as a function of simulation time in figure \ref{fig:density-flaw}b.
+Using this alternative method for finding the outer monolayer location was implemented, and the surface location is plotted as a function of simulation time in figure \ref{fig:density-flaw}b.
 It is apparent from the surface location plot that the water slab location, and thus the surface location, drifts over the 10 ns, spanning approx.
@@
 However, the maximum standard deviation of the positions of the waters comprising the surface layer at each timestep is only 1.85 $\AA$.
-All depth locations in our analyses are calculated relative to the instantaneous surface location at the corresponding timestep of the trajectories.
+Consequently, all depth locations in our analyses are calculated relative to the instantaneous surface location at the corresponding timestep of the trajectories.
 

### 2011-06-23 EquilibratedOrientation.tex
Draft 6

@@
 In both systems the strongest orientational preference is found at the slab surfaces (distances greater than 0\angs) where the water is furthest towards the gas phase.
-Previous work on orientational preference of water at air surfaces shows the same trend as our neat-water results.\cite{Walker2006b,Hore2008} The narrow "penninsula" of high intensity that extends in the $\cos(\theta)$ plots to the right of the -5\angs~location shows the overall preference of water to orient at the surface.
+Previous work on orientational preference of water at air surfaces shows the same trend as our neat-water results.\cite{Walker2006b,Hore2008} The narrow "peninsula" of high intensity that extends in the $\cos(\theta)$ plots to the right of the -5\angs~location shows the overall preference of water to orient at the surface.
 The $\cos(\phi)$ plots are similar to each other with a narrow region of reorientation, but the effect of the surface is greater in the neat-water system as evidenced by the sharper transition in intensity from blue to red, compared to the saturated system that has a less pronounced intensity change.
@@
 Promixity to the water highly orients the \suldiox~with the sulfur atom pointing in towards the water bulk.
-Moving further away from the water surface, and interacting less with \wat~molcules allows for greater orientational freedom as exhibited in the isotropy of the distributions above 5\angs.
+Moving further away from the water surface, and interacting less with \wat~molecules allows for greater orientational freedom as exhibited in the isotropy of the distributions above 5\angs.
 %Because both the neat-water and saturated systems show rather similar distributions, it is possible that the concentration of \suldiox~does not strongly affect the orientation of the \suldiox, unlike the orientations of waters.

### 2011-06-23 Introduction.tex
Draft 6

@@
 Experimental studies to address such questions are valuable but do not provide the resolution necessary to determine the geometries of adsorbing gases, or to determine the orientations of the molecules at the liquid surface near the adsorption site.
-This type of information can be determined computationally, and when coupled to the experimental studies can provide important new insight into the gas-liquid suface adsorption process.
+This type of information can be determined computationally, and when coupled to the experimental studies can provide important new insight into the gas-liquid surface adsorption process.
 
@@
 
-Our recent vibational sum frequency spectroscopy (VSFS) experiments have provided insight into the adsorption and reaction of gases in atmospheric aerosols.\cite{Tarbuck2005,Tarbuck2006} These studies showed that an \suldiox~surface hydrate complex forms when an aqueous surface is exposed to \suldiox~gas.
+Our recent vibrational sum frequency spectroscopy (VSFS) experiments have provided insight into the adsorption and reaction of gases in atmospheric aerosols.\cite{Tarbuck2005,Tarbuck2006} These studies showed that an \suldiox~surface hydrate complex forms when an aqueous surface is exposed to \suldiox~gas.
 A computational study by Baer et al.\cite{Baer2010} then made a series of predictions of the specific nature of the hydrated complex through classical and ab initio simulations.
@@
 
-We provide such detailed informatino to complement and augment these experimental studies.
+We provide such detailed information to complement and augment these experimental studies.
 Our studies involve computation of the orientational depth profiles of water and \suldiox~before and after the adsorption process.

### 2011-06-23 Methods.tex
Draft 6

@@
 The simulations all employed periodic boundaries to create an ``infinite-slab'' geometry.
-After equilibrating the neat-\wat~slabs two types of systems were created by introducing \suldiox: a single-\suldiox~system, herein referred to as the ``neat-water'' system, and a saturated \suldiox~system with multiple gaseous \suldiox~molecules.
+After equilibrating the neat-\wat~slabs two types of systems were created by introducing \suldiox: a single-\suldiox~system, herein referred to as the ``neat-water'' system, and a saturated \suldiox~system with many gaseous surface and bulk-water \suldiox~molecules.
 
@@
 The neat-water system orientational results shown later in this work reproduce well the results of our previous orientational studies of surface water behavior.\cite{Walker2006b,Hore2008} The saturated system had 22 \suldiox~molecules introduced to the water slab bulk in order to saturate it to a level coinciding with the Henry's law constant for \suldiox~in water.
-Additionally, 50 \suldiox~molecules were introduced into the gas phase outside of the saturated water slab to simulate an \suldiox~gas pressure.
-The saturated system was then evolved for 2 ns to produce a starting configuration for further saturated simulations.
+Additionally, 50 \suldiox~molecules were introduced into the gas phase outside of the saturated water slab to simulate an added \suldiox~gas pressure.
+The saturated system with both bulk and gaseous \suldiox~was then evolved for 2 ns to produce a starting configuration for further saturated simulations.
 
@@
 This saturated system was exposed to an additional 50 \suldiox~in the gas phase above the water surface.
-Both the low and high concentration systems were then evolved for a further 10 ns using a timestep of 0.5fs, with atomic coordinates recorded every 100 fs.
+After equilibration for 2 ns, both the low and high concentration systems were then evolved for a further 10 ns data collection using a time step of 0.5fs, with atomic coordinates recorded every 100 fs.
 
@@
 Steered molecular dynamics (SMD) were then performed on the 50 system configurations (in both the neat-water and saturated configurations) to guide the \suldiox~down towards a tethered water near the water slab's center of mass by applying a small steering force to the \suldiox-sulfur atom.
-This steering technique has been previously developed and used to successfuly model chemical events.\cite{Isralewitz2001} The \suldiox~thus passed through the continuum of environments from gas phase to (neat- and saturated) water surface adsorption, and finally absorption into the bulk of the \wat~slab.
-Each of the SMD simulations were performed for a total of 200 ps, using a timestep of 1 fs, and taking snapshots of the system every 25 fs.
+This steering technique has been previously developed and used to successfully model chemical events.\cite{Isralewitz2001} The \suldiox~thus passed through the continuum of environments from gas phase to (neat- and saturated) water surface adsorption, and finally absorption into the bulk of the \wat~slab.
+Each of the SMD simulations were performed for a total of 200 ps, using a time step of 1 fs, and taking snapshots of the system every 25 fs.
 Figure \ref{fig:starting-configurations} illustrates two sample starting configurations for the SMD simulations, showing both the neat-water slab and the saturated slab configurations before steering the \suldiox~towards the water bulk.
@@
 A liquid surface, however, is a dynamic boundary that is neither flat nor stationary.
-To define a water surface location, several previous studies have used the technique of fitting a lineshape to the averaged density profile of the water, and extracting interfacial shape and location parameters.\cite{Shamay2010,Wick2006c,Chowdhary2006} Hyperbolic tangent functions have been used often, and values for the ``Gibb's dividing surface'' location, and interfacial width have thus been determined.\cite{Matsumoto1988} However, in long simulations the location and shape of the interface changes, and the motion of surface waters alters the interfacial width at any given timestep.
+To define a water surface location, several previous studies have used the technique of fitting a line shape to the averaged density profile of the water, and extracting interfacial shape and location parameters.\cite{Shamay2010,Wick2006c,Chowdhary2006} Hyperbolic tangent functions have been used often, and values for the ``Gibb's dividing surface'' location, and interfacial width have thus been determined.\cite{Matsumoto1988} However, in long simulations the location and shape of the interface changes, and the motion of surface waters alters the interfacial width at any given time step.
 Thus, the density profile fitting will capture averaged widths and locations, not instantaneous values.
@@
 Figure \ref{fig:density-flaw}a shows the density profile of water and \suldiox~over the course of one of the 10 ns trajectories used in this work using the original uncorrected coordinates of the system.
-The water density profile and location (dashed line in figure \ref{fig:density-flaw}a) is produced by averaging the instantaneous density profile at each timestep in the simulation over all the timesteps.
+The water density profile and location (dashed line in figure \ref{fig:density-flaw}a) is produced by averaging the instantaneous density profile at each time step in the simulation over all the time steps.
 The water profile was then fit to a $\tanh$ function (black line) to extract the position and width parameters of the water surface.
@@
 
-	In this work we define the surface location by calculating a reference location at each timestep by averaging the positions of the waters contained in the topmost monolayer.
+	In this work we define the surface location by calculating a reference location at each time step by averaging the positions of the waters contained in the topmost monolayer.
 This provides us with a consistent and intuitive reference point in the simulations to which we relate our analyses, but does not increase the computational burden.
@@
 The number of waters included in the averaging is determined by taking a few issues into account.
-First, counting the waters found in the topmost cross-section of the water slab over several timesteps indicated between 65-75 waters that established a full monolayer.
+First, counting the waters found in the topmost cross-section of the water slab over several time steps indicated between 65-75 waters that established a full monolayer.
 This was done by a visual inspection of the slab using the VMD MD visualization package.\cite{HUMP96} Alternatively, assuming a spherical model of water with a radius of 2.2\angs~(the typical length of a bulk-water hydrogen bond), two layers of hexagonally tight-packed spheres yielded a similar number of surface water molecules.
@@
 12 $\AA$.
-However, the maximum standard deviation of the positions of the waters comprising the surface layer at each timestep is only 1.85 $\AA$.
-Consequently, all depth locations in our analyses are calculated relative to the instantaneous surface location at the corresponding timestep of the trajectories.
+However, the maximum standard deviation of the positions of the waters comprising the surface layer at each time step is only 1.85 $\AA$.
+Consequently, all depth locations in our analyses are calculated relative to the instantaneous surface location at the corresponding time step of the trajectories.
 
@@
 The \wat~density profile was fit using a hyperbolic tangent function (black) to extract location and width data for the water surface.
-(B) The instantaneous location of the outer \wat~monolayer was calculated for each simulation timestep (blue), and the surface location extracted from the density fitting was also shown for reference (horizontal dashed line).
+(B) The instantaneous location of the outer \wat~monolayer was calculated for each simulation time step (blue), and the surface location extracted from the density fitting was also shown for reference (horizontal dashed line).
 Long simulations of liquid slabs result in drift of the slab location, and the consequent broadening of the water density profile if the surface location is not determined repeatedly throughout the simulation.}

### 2011-06-24 Conclusions.tex
Draft 7

@@
 % intro-type paragraph here
-The aqueous chemistry taking place in the Earth's atmosphere is goverened by the physics of the water surfaces found there.
+The aqueous chemistry taking place in the Earth's atmosphere is governed by the physics and chemistry of the water surfaces found there.
 Many recent experiments have revealed specific pathways for reaction of \suldiox, but the molecular-level nature of those reactions are still poorly understood.
@@
 The waters interacting with the layer of adsorbed \suldiox~orient more perpendicularly to the interface, and further expose their ``free-OH'' uncoupled bonds for interactions with \suldiox, and hydrate complex formation.
+Furthermore, we have found that surface waters underneath a blanket layer of adsorbed \suldiox~will penetrate further into the gas phase, allowing for greater mobility of waters away from the aqueous bulk.
 
@@
 The results for the transit through the interface show that in both systems of low and high \suldiox~concentration an adsorbing \suldiox~has very similar orientation to those already bound to the water surface.
-As the \suldiox~approaches the water surface the first contact with water causes it to reorient.
+The \suldiox~reorients as it makes its first contact with the water surface.
 Within 5 $\AA$ of the surface the \suldiox~is mostly oriented with its sulfur towards the water phase.
@@
 
+Having now examined this passageway of \suldiox~from the gas to adsorbed aqueous phases, we may now turn to further filling in the details of adsorption and interfacial chemistry of gas molecules.
+What species form during the adsorption transition, and how do molecular properties affect the process?
+Can we form a theory that will more fully explain the transit of many small molecules of environmental and industrial interest?
 This study is one of several aiming to characterize \suldiox~adsorption and behavior on aqueous surfaces.
-The atmospheric chemistry driving many important reactions occurs on aerosol and small particle surfaces.
-To gain a better understanding of gaseous adsorption we plan to report further computational simulation and experimental results of temperature and chemical constituent effects on atmospherically relevant water surfaces.
+We plan to report further results of computational simulation and experimental studies regarding temperature and chemical constituent effects on atmospherically relevant water surfaces.

### 2011-06-24 DensityResults.tex
Draft 7

@@
 The density distributions of both \wat~and \suldiox~were calculated for the equilibrium MD simulations.
-The results presented in figure \ref{fig:density} show both the water and \suldiox~density distributions across the two simulated slabs.
+The results presented in figure \ref{fig:density} show both the water (black) and \suldiox~(red) density distributions across the two simulated slabs.
 As shown, the \suldiox~exposed to the neat-water system remained at the water slab surface.
@@
 
-In the simulated neat-water slab, the \suldiox~molecular density distribution concentrates near the water surface, indicating an affinity for the interfacial region.
-This is consistent with what is found experimentally in our laboratory using VSFS,\cite{Tarbuck2005,Tarbuck2006} and also verified by the computational simulations and spectral calculations of Baer et al.\cite{Baer2010} The experimental studies indicated that upon exposure of \suldiox~to a \wat~surface, a layer of \suldiox~hydrate complex forms, modifying the structure of water above the Gibbs dividing surface (GDS).
+In the simulated neat-water slab, the result that the \suldiox~molecular density distribution concentrates near the water surface indicates an affinity for the interfacial region.
+This is consistent with what is found experimentally in our laboratory using VSFS,\cite{Tarbuck2005,Tarbuck2006} and also supported by the computational simulations and spectral calculations of Baer et al.\cite{Baer2010} The experimental studies indicated that upon exposure of \suldiox~to a \wat~surface, a layer of \suldiox~hydrate complex forms, modifying the structure of water above the Gibbs dividing surface (GDS).
 In the experimental system, the \suldiox~bonding interaction with the free OH oscillators is manifested in a red-shift of the free OH frequency indicative of a bonding interaction.
 In a later section of this article we look at the effect of \suldiox~on surface water orientation.
-During the course of the simulation, the \suldiox~does not venture into the bulk water.
-Also, the \suldiox~does not escape the water surface into the gas phase, but remains located within 5-\angs~of the surface.
+During the course of the simulation, the \suldiox~does not venture into the bulk water nor does it escape the water surface into the gas phase, but remains located within 5-\angs~of the surface region.
 

### 2011-06-24 EquilibratedOrientation.tex
Draft 7

@@
 This results from an orientation of the water's y-axis (normal to the molecular plane) aligned perpendicular to the plane of the water surface.
-Thus, for both the neat-water and saturated systems, the preferred orientation of water molecules at a distance of 0\angs~ is to lie mostly flat to the plane of the interface.
+Thus, for both the neat-water and saturated systems, the preferred orientation of water molecules at a distance of 0\angs~ is to lie mostly flat to the plane of the interface with a slight ``twist'' sending one OH bond further into the gas phase, and the other OH towards the water bulk.
+This result agrees somewhat with a recent air/water study by Fan et al.
+in which the surface orientations of several water models were analyzed.\cite{Fan2009} They simulated water using non-polarizable models, however, which alters the behavior of water at interfacial regions when compared to the polarizable POL3 model used in our work.
+The main conclusions are similar in that one OH tends to point further into the air phase than the other, but differs in that this effect is less pronounced with the polarizable model.
 
 Although the plots show overall similarities for both the neat-water and saturated systems, the presence of a layer of adsorbed \suldiox~molecules alters the water orientation of the waters furthest into the gas phase.
-The waters located above 0\angs~are further into the gas phase, and interact with the layer of adsorbed \suldiox~molecules.
-The resulting orientation of waters above 0\angs, shown in the saturated plot of $\cos(\theta)$ of figure \ref{fig:water-orientation}, is with a bisector pointing further into the adsorbed \suldiox~gas layer, and both hydrogens pointing outward from the aqueous bulk.
+For the saturated solution the resulting orientation of waters above 0\angs, shown in the saturated plot of $\cos(\theta)$ of figure \ref{fig:water-orientation}, is with a bisector pointing further into the adsorbed \suldiox~gas layer, and both hydrogens pointing outward from the aqueous bulk.
 The effect is more pronounced further from the water phase, and above 5\angs~the $\cos(\theta)$ distribution is completely centered around $\cos(\theta)=+1$ (see figure \ref{fig:water-angles}).
@@
 The noise in the neat-water plots of figure \ref{fig:water-orientation} above 5\angs~(manifested as disconnected points of intensity throughout the range of orientations) is a result of fewer waters venturing beyond those extents and thus less data far from the surface.
-This is one indication that waters near a layer of adsorbed \suldiox~venture further above the interface and interact with the adsorbed gas molecules.
-The interactions with neighboring \suldiox~molecules allow the waters above the surface to orient more perpendicularly to the interface.
-Our experimental SFG studies also predicted this reorienting behavior due to the \suldiox~interactions with the topmost surface waters.\cite{Ota2011}
+This is one indication that waters near a layer of adsorbed \suldiox~venture further above the interface relative to the low \suldiox~concentration, where they can have more \suldiox~interaction with the adsorbed gas molecules.
+These results show that the interactions with neighboring \suldiox~molecules allow the waters above the surface to orient more perpendicularly to the interface.
+This is consistent with our recent experimental VSFS studies which showed evidence for the reorienting behavior of water due to the \suldiox~interactions with the topmost surface waters.\cite{Ota2011}
 

### 2011-06-24 Introduction.tex
Draft 7

@@
 Are specific gas or liquid molecular orientations necessary for gaseous adsorption?
-Experimental studies to address such questions are valuable but do not provide the resolution necessary to determine the geometries of adsorbing gases, or to determine the orientations of the molecules at the liquid surface near the adsorption site.
-This type of information can be determined computationally, and when coupled to the experimental studies can provide important new insight into the gas-liquid surface adsorption process.
+Experimental studies to address such questions are valuable but do not provide the full resolution necessary to determine the geometries of adsorbing gases, or to determine the orientations of the molecules at the liquid surface near the adsorption site.
+This type of information can be determined computationally, and when coupled to the experimental studies can provide a more comprehensive picture of the gas-liquid surface adsorption process.
 
 In this work we look at the orientational properties of sulfur dioxide gas adsorbing to a water surface as a model for the transit of gases across an aqueous surface.
-The interactions between sulfur species and aqueous systems are currently an active area of chemical research.
-Sulfur dioxide enters the environment as an important industrial product, and also naturally through terrestrial processes.
+The interactions between sulfur species and aqueous systems are currently an active area of chemical research and general interest.\cite{Lattanzi2010,Shah2011,Tzivian2011,Johns2011,Faloona2009,Jurkat2010,Wu2011} Sulfur dioxide enters the environment as an important industrial product, and also naturally through terrestrial processes.
 Atmospheric dust particles and gases have been implicated in the oxidation of \suldiox, and act as reaction surfaces for chemical mechanisms that are still poorly understood.\cite{Baltrusaitis2011,Rubasinghege2010,Li2007} \suldiox~acts as a major component of atmospheric pollution, and is a precursor to acid rain formation, and cloud nucleation.
@@
 
-Our latest VSFS experiments on \suldiox~behavior further expand our understanding of \suldiox~in atmospheric processes.\cite{Ota2011} A temperature study has shown that the binding of gaseous \suldiox~to water surfaces is greatly enhanced at atmospherically relevant cold temperatures.
-The surface binding of unreacted \suldiox~complexes is found to be a completely reversible process.
+Our latest VSFS experiments on \suldiox~behavior further expand our understanding of \suldiox~in atmospheric processes.\cite{Ota2011} A temperature dependent study has shown that the binding of gaseous \suldiox~to water surfaces is greatly enhanced at atmospherically relevant cold temperatures.
+The surface binding of unreacted \suldiox~complexes is found to be a completely reversible process when at cold temperatures.
 The same study also showed that low pH aqueous environments inhibit the bulk reaction of \suldiox, but do not affect the surface binding of the gas molecules.
@@
 
-We provide such detailed information to complement and augment these experimental studies.
+Here we provide such detailed information to complement and augment these experimental studies.
 Our studies involve computation of the orientational depth profiles of water and \suldiox~before and after the adsorption process.
-They augment the spectroscopic experiments that can not probe depth or orientation profiles of surface species in the same detailed manner as computational simulations.
-The molecular behaviors that occur upon introduction of \suldiox~to a \wat~surface are determined, and we analyze the orientational response of both the \suldiox~and \wat~molecules using equilibrium and steered (SMD) classical molecular dynamics simulations.
+The molecular behaviors that occur upon introduction of \suldiox~to a \wat~surface are determined, and the orientational response of both the \suldiox~and \wat~molecules using equilibrium and steered (SMD) classical molecular dynamics simulations are analyzed.
 The latter approach involves steering a gas molecule into the aqueous phase, and characterizing its molecular orientation as it transits through the interfacial region.

### 2011-06-24 Methods.tex
Draft 7

@@
 
-The low and high concentration simulated systems, ``neat-water'' and saturated, respectively, were created as follows: the ``neat-water'' simulation involved the addition of a single \suldiox~molecule either within the bulk of the water slab (for equilibrium MD), or above the slab surface (in the SMD simulations).
+The low and high concentration simulated systems, ``neat-water'' and saturated water, respectively, were created as follows: the neat-water simulation involved the addition of a single \suldiox~molecule either within the bulk of the water slab (for equilibrium MD), or above the slab surface (in the SMD simulations).
 The single-\suldiox~neat-water system was then evolved for 2 ns to produce an equilibrated starting configuration.
-Because of the extremely low concentration of \suldiox~in the neat-water system the surface waters behave similarly to a true ``neat-water'' air-liquid interface.
+Because of the extremely low concentration of \suldiox~in the neat-water system the surface waters behave similarly to a true neat-water air-liquid interface.
 The neat-water system orientational results shown later in this work reproduce well the results of our previous orientational studies of surface water behavior.\cite{Walker2006b,Hore2008} The saturated system had 22 \suldiox~molecules introduced to the water slab bulk in order to saturate it to a level coinciding with the Henry's law constant for \suldiox~in water.
@@
 
-\subsection{Equilibrium Simulations}
+\subsection{Equilibrium Molecular Dynamics Simulations}
 
@@
 This was done to ensure that the orientation of the \suldiox~during the adsorption transit was not an artifact of the choice of atom used for tethering.
-The simulations produced the same results (not shown) for the orientational analyses, so the data from the original tethering scheme were used.
+The simulations produced the same results (not shown) for the orientational analyses, so the data from the original tethering scheme was used.
 
@@
 
-	In this portion of the study we create orientational depth-profiles of \suldiox~and \wat~upon exposure and adsorption of the gas.
+	The first portion of the studies involved creating orientational depth-profiles of \suldiox~and \wat~upon exposure and adsorption of the gas.
 The reference point used for depth calculations is the location of the water surface.
@@
 
-	In this work we define the surface location by calculating a reference location at each time step by averaging the positions of the waters contained in the topmost monolayer.
+	To overcome this problem we define the surface location by calculating a reference location at each time step by averaging the positions of the waters contained in the topmost monolayer.
 This provides us with a consistent and intuitive reference point in the simulations to which we relate our analyses, but does not increase the computational burden.
@@
 
-	The simulated systems were analyzed to characterize the orientation of \wat~and \suldiox~in various environments above, within, and below the aqueous surface region.
 The molecular orientation of both the \wat~and \suldiox~is a prerequisite for understanding the chemistry occurring during the \suldiox~adsorption process.
+With the surface location as defined above, the simulated systems were analyzed to characterize the orientation of \wat~and \suldiox~in various environments above, within, and below the aqueous surface region.
 The two molecules studied are similarly shaped with a C2$_v$ axis along their bisectors, and a molecular plane defined by three atoms.

### 2011-06-24 TransitOrientation.tex
Draft 7

@@
 
-	From its starting position 20\angs~above the water surface, until the \suldiox~moves to within 10\angs~of the surface of both systems, the orientation is isotropic in $\theta$ and $\phi$.
-This can be seen in the plots as mostly uniform coloration at a given distance from the surface independent of $\cos$ of $\phi$.
-Near the surface the bisector angle $\theta$ becomes more perpendicular to the water surface ($\cos(\theta)\approx1$) with the \suldiox~sulfur pointing into the water phase.
+	From its starting position above the water surface, until the \suldiox~moves to within 10\angs~of the surface of both systems, the orientation is isotropic in $\theta$ and $\phi$ (not shown).
+Isotropic orientation is manifested in the plots as mostly uniform coloration at a given distance from the surface independent of $\cos$ of $\phi$.
+Near and into the surface the bisector angle $\theta$ becomes more perpendicular to the water interface ($\cos(\theta)\approx1$) with the \suldiox~sulfur pointing into the water phase, consistent with the equilibrium MD simulation results above.
 At the point when the \suldiox~reaches the surface location (distance $=0$\angs), the bisector is perpendicular in both the neat-water and saturated systems.
@@
 In the absence of simulated ionic species that form through \suldiox-\wat~chemistry at the surface, it is clear that the adsorbing \suldiox~in the gas phase takes on a preferred orientation to adsorb on a water surface.
-The main difference between the two systems is where the point of the transition from isotropic to preferred orientations is found.
+The main difference between the neat and saturated water systems is where the point of the transition from isotropic to preferred orientations is found.
 
   % talk about theta differences
-  Looking more closely at the approach, the $\theta$ distribution above 10\angs~from the surface is isotropic.
-For the neat-water surface, a transition occurs at approximately 5\angs~from the surface where the \suldiox~orients with its bisector perpendicular to the water surface.
+  Comparing in more detail the difference in these two systems upon \suldiox~approach, for the neat-water surface, a transition occurs at approximately 5\angs~from the surface.
 This coincides with the approximate distance of bonds formed to the outer-most water molecules.
-Below 5\angs, the orientation persists and the \suldiox~begins to interact with the topmost surface waters, and orients accordingly.
+Below 5\angs above the surface, the orientation persists and the \suldiox~begins to interact with the topmost surface waters, and orients accordingly.
 In the saturated system, the same trend occurs, however the onset of the perpendicular orientation begins at approximately 8\angs~above the surface.
@@
 This is the case in both systems, with only a few exceptions.
-At the neat-water surface and just below, from 0\angs~to 5\angs~in the water phase, the $\theta$ profile broadens down to $\cos(\theta)=0.5$, and a clear trend appears in the profile of $\phi$ near $\cos(\phi)=0$.
+In both systems the $\cos(\phi)$ profiles exhibit mostly isotropic distributions above 0\angs, with several regions of lighter coloration interspersed, but without a clearly formed orientational trend.
+At the neat-water surface and just below, from 0\angs~to -5\angs~in the water phase, the $\theta$ profile broadens down to $\cos(\theta)=0.5$, coinciding with a clear trend in the profile of $\phi$ near $\cos(\phi)=0$ shown in dark red at the bottom of the plot.
 This corresponds to \suldiox~inclined up to 60 degrees from the surface normal.

### 2011-06-28 Conclusions.tex
Not much

@@
 
-Having now examined this passageway of \suldiox~from the gas to adsorbed aqueous phases, we may now turn to further filling in the details of adsorption and interfacial chemistry of gas molecules.
+Having now examined this passageway of \suldiox~from the gas to adsorbed aqueous phases, we may turn to further filling in the details of adsorption and interfacial chemistry of gas molecules.
 What species form during the adsorption transition, and how do molecular properties affect the process?
@@
 This study is one of several aiming to characterize \suldiox~adsorption and behavior on aqueous surfaces.
-We plan to report further results of computational simulation and experimental studies regarding temperature and chemical constituent effects on atmospherically relevant water surfaces.
+We plan to report further results of ongoing computational simulation and experimental studies regarding temperature and chemical constituent effects on atmospherically relevant water surfaces.

### 2011-06-28 Article.tex
New image for the spherical angle definition, fleshed out the section describing the angles theta and phi for orientation analysis. Added bibliography for citations
New transit so2 angle analysis files - not yet added
New cycle results
Rearranging the filenames and working on the h2o-orientation results
New density plots and wrapped up most of the density results section.
changed the title
wrote the conclusions, added some material in the density results.
Rewrote the introduction to ask questions and emphasize the importance of the work, as per Geri's request
rearranged sections per notes from

@@
 \usepackage {amsmath}
-\usepackage{setspace}
+\usepackage {setspace}
+\usepackage [pdftex]{graphicx}
+\usepackage[sort&compress]{achemso}
+\usepackage{lineno}
 
@@
 
-\title{Sulfur Dioxide at Water's Surface}
+\title{Sulfur Dioxide on Water's Surface}
 \author{Eric S. Shamay \and Kevin E. Johnson \and Geraldine L. Richmond}
@@
 \begin{document}
+
 \newcommand{\suldiox}{SO$_2$}
@@
 
-\maketitle
+%\maketitle
 
-\onehalfspacing
+%\onehalfspacing
+\linenumbers 
+\doublespacing
 
 \begin{abstract}
-	...Abstract goes here...
+Sulfur dioxide binding to water surfaces is a model system for the study of gaseous adsorption processes on aqueous systems.
+Equilibrium and adsorbing \suldiox~molecules in a pure water interfacial system are modeled by classical molecular dynamics simulations using fully polarizable models for both \suldiox~and \wat.
+Neat-water systems with a single \suldiox, and high \suldiox~concentration saturated water systems are modeled at equilibrium and using steered molecular dynamics (SMD) to ascertain the orientational profile of a transiting \suldiox~molecule.
+Orientational depth profiling reveals the preferred orientations of \suldiox~during adsorption, and the complementary behaviors of surface waters adsorbing the gaseous solute.
 \end{abstract}
@@
 \include{Introduction}
-\include{Background}
 \include{Methods}
-\include{SteeredMDResults}
-\include{EquilibratedResults}
+\include{Results}
+\include{Conclusions}
+
+\bibliography{bib/SulfurDioxideText}
 

### 2011-06-28 Conclusions.tex
Draft 7 and a new abstract!

@@
 When the \suldiox~concentration is increased, as in the saturated system (B), the waters at the surface behave similar to the neat-water interface, but waters that venture into the adsorbed \suldiox~gas layer orient strongly with their bisectors pointing out from the aqueous phase towards the gas.
-In both cases, the \suldiox~orients with its molecular bisector pointing out to the gas phase when it is near the surface.}
+In both cases, the \suldiox~orients with its molecular bisector pointing out to the gas phase when it is near the surface, and isotropically further into the gas phase.
+Sulfur, oxygen and hydrogen atoms are colord yellow, red, and white, respectively.}
 		\label{fig:so2-surface-cartoon}

### 2011-06-30 Article.tex
New title and abstract as written by Geri

--- 2010-12-17 Article.tex
@@
 
-\title{Sulfur Dioxide at Water's Surface}
+\title{Dancing on Water: The Choreography of Sulfur Dioxide Adsorption to Aqueous Surfaces}
 \author{Eric S. Shamay \and Kevin E. Johnson \and Geraldine L. Richmond}
@@
 
-\maketitle
+%\maketitle
 
-\onehalfspacing
+%\onehalfspacing
+\linenumbers 
+\doublespacing
 
 \begin{abstract}
-	...Abstract goes here...
+One might expect that the high surface tension of water to be a barrier to adsorption of a gas into the liquid phase, but we know that gaseous adsorption on and into a water surface is a common phenomenon on this planet.
+What is not commonly known is how this happens and the interplay between an atmospheric gas such as SO2 and molecules at the water surface that can overcome the barrier created by the strong water-water surface bonding interactions.
+What this interplay looks like, the distances from the water surface at which these attractive interactions begin and how they influence the orientational nature of both SO2 and surface water molecules is the focus of this computational study.
+The results fill a void in the information about this system existing from previous experimental studies by providing information about the dimensional nature the gas-surface interactions and the details of how the two species twist and turn orientationally with increased surface interactions.
+Classical molecular dynamics simulations have been employed in these equilibrium and steered molecular dynamics simulations for SO2 at a neat water surface and at a surface with high interfacial SO2 concentrations.
+The results provide new molecular insights for understanding the interaction of this prevalent gas on aerosol and other aqueous surfaces in the environment.
 \end{abstract}

### 2011-06-30 Article.tex
minor changes to abstract for formatting and grammar.

--- 2010-12-17 Article.tex
@@
 
-\maketitle
+%\maketitle
 
-\onehalfspacing
+%\onehalfspacing
+\linenumbers 
+\doublespacing
 
 \begin{abstract}
-	...Abstract goes here...
+One might expect the high surface tension of water to be a barrier to adsorption of a gas into the liquid phase, but we know that gaseous adsorption on and into a water surface is a common phenomenon on this planet.
+What is not commonly known is how this happens and the interplay between an atmospheric gas such as \suldiox~and molecules at the water surface that can overcome the barrier created by the strong water-water surface bonding interactions.
+What this interplay looks like, the distances from the water surface at which these attractive interactions begin and how they influence the orientational nature of both SO2 and surface water molecules is the focus of this computational study.
+The results fill a void in the information about this system existing from previous experimental studies by providing information about the dimensional nature of the gas-surface interactions and the details of how the two species twist and turn orientationally with increased surface interactions.
+Classical molecular dynamics have been employed in both equilibrium and steered molecular dynamics (SMD) simulations for \suldiox~at a neat-water surface and at a surface with high interfacial SO2 concentrations.
+The results provide new molecular insights for understanding the interaction of this prevalent gas on aerosol and other aqueous surfaces in the environment.
 \end{abstract}

### 2011-06-30 Conclusions.tex
Draft 8

@@
 
-% intro-type paragraph here
-The aqueous chemistry taking place in the Earth's atmosphere is governed by the physics and chemistry of the water surfaces found there.
-Many recent experiments have revealed specific pathways for reaction of \suldiox, but the molecular-level nature of those reactions are still poorly understood.
-Through computer simulations we are able to probe the microscopic environments in which those reactions occur, and better understand the conditions in which \suldiox~reacts.
+Gaseous adsorption on solid surfaces has been extensively studied over the past few decades with much learned about how molecular geometry and orientation of the adsorbate are influenced by the proximity of the solid slab.
+For a liquid surface where the surface slab is no longer rigid but has molecules with considerable freedom of movement, the surface and approaching gas molecules can be active partners in attaining the optimal geometry and orientation necessary for adsorption and subsequent uptake.
+And unlike the solid surface defined by a sharp plane, the interfacial region for the liquid-gas system is much broader, extends on either side of a defined center plane, and is host to a broad distribution of gas-liquid molecular geometries and orientations that change as the gas molecules transit through the interfacial region into the bulk liquid.
+Although  our current molecular level understanding of  the complex dances that these molecules play in this fluid interfacial region is in its infancy, emerging studies such as these are beginning to provide unique new insights that are key to understanding many  environmentally important processes at aqueous surfaces.
 
-We have presented the results of several classical simulations of \suldiox~interacting with an aqueous surface.
-We set out to find the orientational behavior of \wat~and \suldiox~through depth-profiling of molecular orientational distributions.
-In previous experimental studies we posited that water reorients in the presence of \suldiox, and this behavior is one of the steps in formation of gaseous hydrate complexes on water's surface.
-Spectroscopic techniques can not precisely define a depth profile or orientational distribution of adsorbed species, so we have used molecular dynamics simulations to gain that information.
+Presented herein are the results of several classical molecular dynamic simulations that focus on understanding how surface water molecules and adsorbing \suldiox~gas molecules twist and turn as the gas transits the interfacial region.
+The computational studies emulate and expand on the experimental spectroscopic studies from this laboratory which have found \suldiox~surface complexation at a water surface.\cite{Tarbuck2005,Tarbuck2006,Ota2011}  These spectroscopic studies show clear evidence of \suldiox-water surface complexation but details about this surface complex could only be inferred from spectral changes in the surface water spectrum since \suldiox~could not be monitored directly.
+These simulations do not have that limitation and hence can provide information about the behavior of both surface partners and in particular, how their proximity influences the orientation behavior of each other.
+The orientational information obtained in these simulations are provided via calculated depth profiles which show the molecular distribution of orientations of the two different interfacial molecules throughout the dimensions of the interfacial region.
 
-Our simulations recreated the experimental setup of a recent study by our group,\cite{Ota2011} by introducing a concentrated gas phase of \suldiox~to a water surface.
-Through simulation it was found that gaseous \suldiox~quickly adsorbs to the water surface, and continues to bind until a complete surface coverage is reached.
+The simulations described herein show that gaseous \suldiox~quickly adsorbs to the water surface and continues to bind until a complete surface coverage is reached.
 Surface waters reorient in the presence of adsorbed \suldiox.
@@
 The waters interacting with the layer of adsorbed \suldiox~orient more perpendicularly to the interface, and further expose their ``free-OH'' uncoupled bonds for interactions with \suldiox, and hydrate complex formation.
-Furthermore, we have found that surface waters underneath a blanket layer of adsorbed \suldiox~will penetrate further into the gas phase, allowing for greater mobility of waters away from the aqueous bulk.
+Furthermore, we have found that surface waters underneath a blanket layer of adsorbed \suldiox~will penetrate further into the gas phase, allowing for greater mobility of waters away from the aqueous bulk in the presence of \suldiox.
 
-The surface behavior of the \suldiox~molecules was also determined, information that is difficult to obtain with vibrational sum-frequency spectroscopy.
 Through these simulations we are able to characterize the orientational behavior of \suldiox~during and after adsorption.
@@
 In both cases, the \suldiox~orients with its molecular bisector pointing out to the gas phase when it is near the surface, and isotropically further into the gas phase.
-Sulfur, oxygen and hydrogen atoms are colord yellow, red, and white, respectively.}
+Sulfur, oxygen and hydrogen atoms are colored yellow, red, and white, respectively.}
 		\label{fig:so2-surface-cartoon}

### 2011-06-30 DensityResults.tex
Draft 8

@@
 The density distributions of both \wat~and \suldiox~were calculated for the equilibrium MD simulations.
-The results presented in figure \ref{fig:density} show both the water (black) and \suldiox~(red) density distributions across the two simulated slabs.
-As shown, the \suldiox~exposed to the neat-water system remained at the water slab surface.
+The results presented in Figure \ref{fig:density} show both the water (black) and \suldiox~(red) density distributions across the two simulated slabs.
+As shown, the single \suldiox~in the neat-water system remained at the water slab surface.
 The \suldiox~in the saturated system accumulated mostly at the surface, but some residual \suldiox~remained well into the bulk water.
@@
 In the simulated neat-water slab, the result that the \suldiox~molecular density distribution concentrates near the water surface indicates an affinity for the interfacial region.
-This is consistent with what is found experimentally in our laboratory using VSFS,\cite{Tarbuck2005,Tarbuck2006} and also supported by the computational simulations and spectral calculations of Baer et al.\cite{Baer2010} The experimental studies indicated that upon exposure of \suldiox~to a \wat~surface, a layer of \suldiox~hydrate complex forms, modifying the structure of water above the Gibbs dividing surface (GDS).
+During the course of the simulation, the \suldiox~does not venture into the bulk water nor does it escape the water surface into the gas phase, but remains located within 5-\angs~of the surface region.
+This is consistent with what is found experimentally in our laboratory using VSFS,\cite{Tarbuck2005,Tarbuck2006} and also supported by the computational simulations and spectral calculations of Baer et al.\cite{Baer2010} The experimental studies indicated that upon exposure of \suldiox~to a \wat~surface, a layer of solvated \suldiox~forms, modifying the structure of water above the Gibbs dividing surface (GDS).
 In the experimental system, the \suldiox~bonding interaction with the free OH oscillators is manifested in a red-shift of the free OH frequency indicative of a bonding interaction.
-In a later section of this article we look at the effect of \suldiox~on surface water orientation.
-During the course of the simulation, the \suldiox~does not venture into the bulk water nor does it escape the water surface into the gas phase, but remains located within 5-\angs~of the surface region.
 
@@
 Because of the limitations of the classical model in accurately reproducing the first hydration shell around the \suldiox,\cite{Baer2010} we do not draw specific conclusions about the hydrate complex's geometry.
-However, the surface affinity of the hydrated \suldiox~is well reproduced by both the classical and ab initio methods.
+However, the surface affinity of the hydrated \suldiox~is well reproduced by both the classical and ab initio methods.\cite{Baer2010}
 

### 2011-06-30 EquilibratedOrientation.tex
Draft 8

@@
 
-Geometric analyses were performed to characterize the net molecular orientation of \wat~and \suldiox~molecules at different depths from the water surface.
+Geometric analyses were performed to characterize the net molecular orientation of \wat~and \suldiox~molecules at different depths from the water surface location.
 At each distance from the surface location, an orientation profile was created for both the \wat~and \suldiox~molecules.
 The orientation distribution for the angles $\theta$ and $\phi$ at all depths were combined to form 2D intensity plots that show how the molecular orientation distributions change with distance to the surface location.
-These plots allow for a visual interpretation of how the net orientations are affected when moving from the gas phase to the surface, and then further into the aqueous bulk.
+These plots (Figures \ref{fig:water-orientation}, \ref{fig:so2-transit-angles}, and \ref{fig:so2-surface-angles}) allow for a visual interpretation of how the net orientations are affected when moving from the gas phase through the interfacial region and to the surface location, and then further into the aqueous interfacial region and bulk.
 Both the neat-water, with only a single \suldiox~introduced, and the high-concentration saturated system were analyzed.
-In the case of the neat-water system, the introduction of a single \suldiox~does not greatly affect the water surface, and the results of the water orientation are very similar to a neat-water system without any adsorbed solutes (not shown).
+In the case of the neat-water system, the introduction of a single \suldiox~does not greatly affect molecular orientation of water molecules in the interfacial region.
+These results of the water orientation are very similar to a neat-water system without any adsorbed solutes (not shown).
 
@@
 Likewise, areas of the plot with high intensity over a small orientational range are considered to exhibit an orientational preference at the given depth.
-The histograms are arranged with the plots of $\cos(\theta)$ on the left and $\cos(\phi)$ on the right columns.
+The histograms are arranged with the plots of $\cos(\theta)$ on the left and $\cos(\phi)$ on the right.
 The surface is located at 0\angs.
@@
 
-The orientation depth-profiles for \wat~are shown in figure \ref{fig:water-orientation} for both the neat-water (top) and saturated (bottom) systems during the equilibrium MD simulations.
+The orientation depth-profiles for \wat~are shown in Figure \ref{fig:water-orientation} for both the neat-water (top) and saturated (bottom) systems during the equilibrium MD simulations.
+The interfacial region for both these calculations and the VSF experiments is defined as the region where mlecular orientational anisotropy exists around the surface water location.
+Referring to Figure \ref{fig:water-orientation}, calculations indicate an interfacial width of approx.
+10\angs~for the neat-water system, and approx.
+16 \angs~for the saturated system.
 In both systems the strongest orientational preference is found at the slab surfaces (distances greater than 0\angs) where the water is furthest towards the gas phase.
 Previous work on orientational preference of water at air surfaces shows the same trend as our neat-water results.\cite{Walker2006b,Hore2008} The narrow "peninsula" of high intensity that extends in the $\cos(\theta)$ plots to the right of the -5\angs~location shows the overall preference of water to orient at the surface.
-The $\cos(\phi)$ plots are similar to each other with a narrow region of reorientation, but the effect of the surface is greater in the neat-water system as evidenced by the sharper transition in intensity from blue to red, compared to the saturated system that has a less pronounced intensity change.
+The $\cos(\phi)$ plots are similar to each other with a narrow region of reorientation, but the effect in the interfacial region is greater in the neat-water system as evidenced by the sharper transition in intensity from blue to red, compared to the saturated system that has a less pronounced intensity change.
 
-The bisector tilt of the water molecules, $\cos(\theta)$, concentrates around $\cos(\theta)=0$ within the first few\angs~above and below the surface, becoming progressively isotropic further into the water bulk of both systems.
+The bisector tilt of the water molecules, $\cos(\theta)$, concentrates around $\cos(\theta)=0$ within the first few\angs~above and below the water surface location, becoming progressively isotropic further through the interfacial region and into the water bulk of both systems.
 As the tilt nears $\cos(\theta)=0$ the \wat~bisector lies within the plane of the surface indicating a water orientation either flat on the surface, or with some amount of ``twist'' sending the OH bonds in towards, or out of the bulk.
 The value of $\phi$ determines the ``twist'' in this case.
-Both systems show a defined intensity concentration (dark red) in the distributions around $\cos(\phi)=1$ at the water surfaces.
+Both systems show a defined intensity concentration (dark red) in the distributions around $\cos(\phi)=1$ on the aqueous side of the interfacial region.
 This results from an orientation of the water's y-axis (normal to the molecular plane) aligned perpendicular to the plane of the water surface.
@@
 Although the plots show overall similarities for both the neat-water and saturated systems, the presence of a layer of adsorbed \suldiox~molecules alters the water orientation of the waters furthest into the gas phase.
-For the saturated solution the resulting orientation of waters above 0\angs, shown in the saturated plot of $\cos(\theta)$ of figure \ref{fig:water-orientation}, is with a bisector pointing further into the adsorbed \suldiox~gas layer, and both hydrogens pointing outward from the aqueous bulk.
-The effect is more pronounced further from the water phase, and above 5\angs~the $\cos(\theta)$ distribution is completely centered around $\cos(\theta)=+1$ (see figure \ref{fig:water-angles}).
+For the saturated solution the resulting orientation of waters above 0\angs, shown in the saturated plot of $\cos(\theta)$ of Figure \ref{fig:water-orientation}, is with a bisector pointing further into the adsorbed \suldiox~gas layer, and both hydrogens pointing outward from the aqueous bulk.
+The effect is more pronounced further from the water phase, and above 5\angs~the $\cos(\theta)$ distribution is completely centered around $\cos(\theta)=+1$ (see Figure \ref{fig:water-angles}).
 
@@
 
-The noise in the neat-water plots of figure \ref{fig:water-orientation} above 5\angs~(manifested as disconnected points of intensity throughout the range of orientations) is a result of fewer waters venturing beyond those extents and thus less data far from the surface.
+The noise in the neat-water plots of Figure \ref{fig:water-orientation} above 5\angs~(manifested as disconnected points of intensity throughout the range of orientations) is a result of fewer waters venturing beyond those extents and thus less data far from the water surface location.
 This is one indication that waters near a layer of adsorbed \suldiox~venture further above the interface relative to the low \suldiox~concentration, where they can have more \suldiox~interaction with the adsorbed gas molecules.
@@
 Orientation distributions of the adsorbed \suldiox~molecules were created during the equilibrium simulations for both the neat-water and saturated systems.
-Figure \ref{fig:so2-orientation} shows the distributions of $\cos(\theta)$ and $\cos(\phi)$ (arranged similarly to the water orientation distributions plots in figure \ref{fig:water-orientation}).
+Figure \ref{fig:so2-orientation} shows the distributions of $\cos(\theta)$ and $\cos(\phi)$ (arranged similarly to the water orientation distributions plots in Figure \ref{fig:water-orientation}).
 The \suldiox~orientation data set for the neat-water system is much smaller as only a single \suldiox~molecule was simulated in the bulk.
@@
 
-Within 0-5\angs~of the water surface (the approximate distance from the surface that \suldiox~begins interacting with the outer-most waters) the angular distribution of the single \suldiox~(in the neat-water system) is concentrated primarily in $\cos(\theta)>0$.
+In the interfacial region Within 0-5\angs~of the water surface location (the approximate distance from the surface that \suldiox~begins interacting with the outer-most waters) the angular distribution of the single \suldiox~(in the neat-water system) is concentrated primarily in $\cos(\theta)>0$.
 The peak of the distribution occurs at $\cos(\theta)=1$.
@@
 
-The plots indicate that \suldiox~orients similarly at both the low and high concentration water surface environments.
+The plots indicate that \suldiox~orients similarly at both the low and high concentration interfacial environments.
 The plots of $\phi$ for both concentrations exhibit isotropic distributions.
@@
 5\angs~of the surface.
-This behavior continues down to at least 5\angs~below the surface, notwithstanding any chemical reactions that may occur that are not simulated using the classical MD techniques utilized in this work.
+This behavior continues down to at least 5\angs~below the water surface location, notwithstanding any chemical reactions that may occur that are not simulated using the classical MD techniques utilized in this work.
 

### 2011-06-30 Introduction.tex
Draft 8

@@
 The doorway to the uptake of a gas by an aqueous solution is the water surface.
-Although we know much about the behavior of a gas on either side of that entrance, far less is known about how that surface acts to attract, facilitate, or thwart the transit of the gas between the two bulk phases.
+Although we know much about the behavior of a gas on either side of that entrance, far less is known about how that surface acts to attract, facilitate, or thwart the transit of a molecule between the two bulk phases.
+What is the interplay between the gas and surface water molecules and when does one begin to influence the behavior of the other?
 What species form during gas adsorption onto liquid surfaces, and what are the intermediary steps?
-To what extent does an adsorbing gas affect liquid water surface molecules, or the structure of the interface throughout an adsorption process?
+Is molecular orientation of either the gas or surface molecules a factor in the adsorption process?
 Are specific gas or liquid molecular orientations necessary for gaseous adsorption?
@@
 
-In this work we look at the orientational properties of sulfur dioxide gas adsorbing to a water surface as a model for the transit of gases across an aqueous surface.
-The interactions between sulfur species and aqueous systems are currently an active area of chemical research and general interest.\cite{Lattanzi2010,Shah2011,Tzivian2011,Johns2011,Faloona2009,Jurkat2010,Wu2011} Sulfur dioxide enters the environment as an important industrial product, and also naturally through terrestrial processes.
+An important gas for developing a picture of gaseous adsorption and entry into a wter surface is \suldiox.\cite{Lattanzi2010,Shah2011,Tzivian2011,Johns2011,Faloona2009,Jurkat2010,Wu2011} Sulfur dioxide enters the environment as an important industrial product, and also naturally through terrestrial processes.
 Atmospheric dust particles and gases have been implicated in the oxidation of \suldiox, and act as reaction surfaces for chemical mechanisms that are still poorly understood.\cite{Baltrusaitis2011,Rubasinghege2010,Li2007} \suldiox~acts as a major component of atmospheric pollution, and is a precursor to acid rain formation, and cloud nucleation.
 Its high solubility in water makes \suldiox~an integral compound in many aqueous atmospheric reactions, as well.
-We seek a fuller understanding of the interactions between \suldiox~and water at interfaces because they are central to many atmospheric particle and aerosol reactions.
+
+In this work we provide a molecular picture of \suldiox~adsorption on a water surface.
+The studies demonstrate the strong orientational effect that surface water molecules have on the adsorbing gas during the approach and entry into the surface region, and how surface water molecules rise towards the gas phase to assist in the capture of \suldiox~molecules at both high and low \suldiox~surface concentrations.
+
+These computational studies complement and significantly expand the picture developed in our recent experimental vibrational sum frequency spectroscopy (VSFS) studies of \suldiox~adsorption of aqueous solutions of various compositions and temperatures, and the subsequent studies by Baer et al.\cite{Baer2010} using both classical and ab initio simulations.
+The most recent experimental studies have shown that whereas the binding of gaseous \suldiox~to a water surface is greatly enhanced at cold temperatures, the reversibility of the adsorption process remains.
+Complimentary experiments showed that low pH aqueous environments inhibit the bulk reactions of \suldiox, but do not affect the surface binding or its reversibility.
+What is apparent in the VSF spectra obtained in all of these experiments is the tendency of water to reorient upon surface bonding, with the effect becoming more pronounced at high \suldiox~surface concentrations.
+Since the \suldiox~molecule was not specifically probed, conclusions on how \suldiox~bonding contributes to reorienting surface water molecules and the orientation of \suldiox~itself upon approach and surface bonding could only be inferred.
+Obtaining a more complete picture of this adsorption process is important for understanding gaseous adsorption of this environmentally important gas on water and aerosol surfaces as well as being a model system for understanding the more general nature of gases at aqueous interfaces.
 
@@
 
-Our latest VSFS experiments on \suldiox~behavior further expand our understanding of \suldiox~in atmospheric processes.\cite{Ota2011} A temperature dependent study has shown that the binding of gaseous \suldiox~to water surfaces is greatly enhanced at atmospherically relevant cold temperatures.
-The surface binding of unreacted \suldiox~complexes is found to be a completely reversible process when at cold temperatures.
-The same study also showed that low pH aqueous environments inhibit the bulk reaction of \suldiox, but do not affect the surface binding of the gas molecules.
-Furthermore, the resultant experimental spectra suggest that the surface waters reorient in response to the presence of adsorbed \suldiox, but the details about this reorientation can not be garnered from the experiments.
-Such details include knowing how surface water orientation changes upon approach of the gaseous \suldiox, how this varies upon \suldiox~surface bonding, and the orientation of \suldiox~throughout the surface bonding process.
-%Knowing that reorientation may occur due to \suldiox~adsorption, how can we confirm it?
-what is the nature of the reorienting waters?
-How do they reorient, and to what extent?
-Conclusive experimental evidence of the adsorbing gas or surface water molecular orientation is still missing, so we are moved to use computational methods for further understanding this phenomena.
-
-Here we provide such detailed information to complement and augment these experimental studies.
-Our studies involve computation of the orientational depth profiles of water and \suldiox~before and after the adsorption process.
-The molecular behaviors that occur upon introduction of \suldiox~to a \wat~surface are determined, and the orientational response of both the \suldiox~and \wat~molecules using equilibrium and steered (SMD) classical molecular dynamics simulations are analyzed.
+To fill this void, the computational studies described herein provide a detailed picture of the orientation of both \suldiox~and surface water molecules during the adsorption process.
+The depth profiling studies which examine the orientation of both species during the approach and entry of the gas into the interfacial region are obtained using equilibrium and steered (SMD) classical molecular dynamics simulations.
 The latter approach involves steering a gas molecule into the aqueous phase, and characterizing its molecular orientation as it transits through the interfacial region.
-This unique approach enables new insights into the behaviors of gas molecules as they move near to liquid water to be obtained.
-We also simulate how \suldiox~adsorption occurs on a water surface saturated with adsorbed \suldiox, analogous to the cold \suldiox~experimental studies recently performed.\cite{Ota2011} The results of this study provide an intimate perspective on the adsorption of \suldiox~at an aqueous surface, and a more complete picture of gaseous adsorption to liquid interfaces.
+This unique approach enables new insights into the behaviors of gas molecules as they move near to liquid water.
+We also simulate how \suldiox~adsorption occurs on a water surface saturated with adsorbed \suldiox, analogous to the conditions of the \suldiox~experimental studies recently performed.\cite{Ota2011} The results of this study provide an intimate perspective on the adsorption of \suldiox~at an aqueous surface, and a more complete picture of gaseous adsorption to liquid interfaces.

### 2011-06-30 Methods.tex
Draft 8

@@
 The simulations all employed periodic boundaries to create an ``infinite-slab'' geometry.
-After equilibrating the neat-\wat~slabs two types of systems were created by introducing \suldiox: a single-\suldiox~system, herein referred to as the ``neat-water'' system, and a saturated \suldiox~system with many gaseous surface and bulk-water \suldiox~molecules.
+After equilibrating the neat-\wat~slabs two types of systems were created by introducing \suldiox: a single-\suldiox~system, herein referred to as the ``neat-water'' system, and a ``saturated'' \suldiox~system with many gaseous surface and bulk-water \suldiox~molecules.
 
-The low and high concentration simulated systems, ``neat-water'' and saturated water, respectively, were created as follows: the neat-water simulation involved the addition of a single \suldiox~molecule either within the bulk of the water slab (for equilibrium MD), or above the slab surface (in the SMD simulations).
+The low and high concentration simulated systems, ``neat-water'' and ``saturated water'', respectively, were created as follows: the neat-water simulation involved the addition of a single \suldiox~molecule either within the bulk of the water slab (for equilibrium MD), or above the slab surface (in the SMD simulations).
 The single-\suldiox~neat-water system was then evolved for 2 ns to produce an equilibrated starting configuration.
@@
 This saturated system was exposed to an additional 50 \suldiox~in the gas phase above the water surface.
-After equilibration for 2 ns, both the low and high concentration systems were then evolved for a further 10 ns data collection using a time step of 0.5fs, with atomic coordinates recorded every 100 fs.
+After equilibration for 2 ns, both the low and high concentration systems were then evolved for a further 10 ns data collection using a time step of 0.5 fs, with atomic coordinates recorded every 100 fs.
 
@@
 Steered molecular dynamics (SMD) were then performed on the 50 system configurations (in both the neat-water and saturated configurations) to guide the \suldiox~down towards a tethered water near the water slab's center of mass by applying a small steering force to the \suldiox-sulfur atom.
-This steering technique has been previously developed and used to successfully model chemical events.\cite{Isralewitz2001} The \suldiox~thus passed through the continuum of environments from gas phase to (neat- and saturated) water surface adsorption, and finally absorption into the bulk of the \wat~slab.
+This steering technique has been previously developed and used to successfully model chemical events.\cite{Isralewitz2001,Giorgino2011,Bizzarri2011,Strzelecki2009,Patargias2009,Liu2006} The \suldiox~thus passed through the continuum of environments from gas phase to (neat- and saturated) water surface adsorption, and finally absorption into the bulk of the \wat~slab.
 Each of the SMD simulations were performed for a total of 200 ps, using a time step of 1 fs, and taking snapshots of the system every 25 fs.
@@
 
-	The first portion of the studies involved creating orientational depth-profiles of \suldiox~and \wat~upon exposure and adsorption of the gas.
-The reference point used for depth calculations is the location of the water surface.
-A liquid surface, however, is a dynamic boundary that is neither flat nor stationary.
-To define a water surface location, several previous studies have used the technique of fitting a line shape to the averaged density profile of the water, and extracting interfacial shape and location parameters.\cite{Shamay2010,Wick2006c,Chowdhary2006} Hyperbolic tangent functions have been used often, and values for the ``Gibb's dividing surface'' location, and interfacial width have thus been determined.\cite{Matsumoto1988} However, in long simulations the location and shape of the interface changes, and the motion of surface waters alters the interfacial width at any given time step.
+	The first portion of the studies involved creating orientational depth-profiles in the interfacial region comprised of \suldiox~and \wat~upon exposure and adsorption of the gas.
+To do this we must define a reference point in thie interfacial region which we refer to here as the water surface location.
+Recognizing that a liquid surface is a dynamic boundary that is neither flat nor stationary.
+Several previous studies have used the technique of fitting a line shape to the averaged density profile of the water, and extracting interfacial shape and location parameters to define the water surface location.\cite{Shamay2010,Wick2006c,Chowdhary2006} Hyperbolic tangent functions have been used often, and values for the ``Gibb's dividing surface'' location, and interfacial width have thus been determined.\cite{Matsumoto1988} However, in long simulations the location and shape of the interface changes, and the motion of surface waters alters the interfacial width at any given time step.
 Thus, the density profile fitting will capture averaged widths and locations, not instantaneous values.
 Similarly, the averaged values of location and width will obscure information about any drift or deformations the surface undergoes.
+The analysis presented here attempts to retain these subtleties through the use of a ``corrected'' coordinate system.
 
 Figure \ref{fig:density-flaw} demonstrates the problem of surface location drift during a simulation.
-Figure \ref{fig:density-flaw}a shows the density profile of water and \suldiox~over the course of one of the 10 ns trajectories used in this work using the original uncorrected coordinates of the system.
-The water density profile and location (dashed line in figure \ref{fig:density-flaw}a) is produced by averaging the instantaneous density profile at each time step in the simulation over all the time steps.
+Figure \ref{fig:density-flaw}A shows the density profile of water and \suldiox~over the course of one of the 10 ns trajectories used in this work using the original uncorrected coordinates of the system taken from the raw atomic positions of the molecular dynamics data output.
+The water density profile and location (thin grey line in Figure \ref{fig:density-flaw}A) is produced by averaging the instantaneous density profile at each time step in the simulation over all the time steps.
 The water profile was then fit to a $\tanh$ function (black line) to extract the position and width parameters of the water surface.
@@
 
-	To overcome this problem we define the surface location by calculating a reference location at each time step by averaging the positions of the waters contained in the topmost monolayer.
+	To overcome this problem we define the water surface location by calculating a reference location at each time step by averaging the positions of the waters contained in the topmost monolayer.
 This provides us with a consistent and intuitive reference point in the simulations to which we relate our analyses, but does not increase the computational burden.
-
-The location of the water surface is determined by the averaged positions of a set number of the outer-most waters in the simulated aqueous slabs.
 The number of waters included in the averaging is determined by taking a few issues into account.
@@
 Increasing the number of waters used in calculating the surface location diminishes the effects of the few waters that briefly rise above the surface into the gas phase, stabilizing both the surface position and thickness values.
-Taking the close-packed model as a maximum number of waters fit into a flat surface, the authors decided to use 70 water molecules for calculating instantaneous water surface locations.
+Taking the close-packed model as a maximum number of waters fit into a flat surface, 70 water molecules were used for calculating instantaneous water surface locations.
 
-Using this alternative method for finding the outer monolayer location was implemented, and the surface location is plotted as a function of simulation time in figure \ref{fig:density-flaw}b.
+This method for finding the outer monolayer location was implemented, and the surface location is plotted as a function of simulation time in Figure \ref{fig:density-flaw}B.
 It is apparent from the surface location plot that the water slab location, and thus the surface location, drifts over the 10 ns, spanning approx.
@@
 		\includegraphics[scale=1.0]{images/density/density-location.png}
-		\caption{(A) Density profiles of \wat~(dashed) and \suldiox~(red) from a 10 ns simulation of the neat-water system with a single sulfur dioxide.
+		\caption{(A) Density profiles of \wat~(grey) and \suldiox~(red) from a 10 ns simulation of the neat-water system with a single sulfur dioxide.
 The \wat~density profile was fit using a hyperbolic tangent function (black) to extract location and width data for the water surface.
+Position values are based on the atomic position data of the water oxygens and \suldiox~sulfurs from the raw molecular dynamics output.
 (B) The instantaneous location of the outer \wat~monolayer was calculated for each simulation time step (blue), and the surface location extracted from the density fitting was also shown for reference (horizontal dashed line).
@@
 
-The molecular orientation of both the \wat~and \suldiox~is a prerequisite for understanding the chemistry occurring during the \suldiox~adsorption process.
+Knowing the molecular orientation of both the \wat~and \suldiox~is a prerequisite for understanding the chemistry occurring during the \suldiox~adsorption process.
 With the surface location as defined above, the simulated systems were analyzed to characterize the orientation of \wat~and \suldiox~in various environments above, within, and below the aqueous surface region.
 The two molecules studied are similarly shaped with a C2$_v$ axis along their bisectors, and a molecular plane defined by three atoms.
-A body-fixed frame is defined for both \wat~and \suldiox~as shown in figure \ref{fig:molecular-frame}.
+A body-fixed frame is defined for both \wat~and \suldiox~as shown in Figure \ref{fig:molecular-frame}.
 In each analysis a space-fixed reference axis is used that corresponds to the long axis of the system's periodic cell normal to the plane of the water surface.
 The orientational analyses presented herein focus on two angles used to define molecular orientation.
-The molecular orientation angles $\theta$ and $\phi$ are determined from a set reference axis as shown in figure \ref{fig:water-angles}a.
+The molecular orientation angles $\theta$ and $\phi$ are determined from a set reference axis as shown in Figure \ref{fig:water-angles}A.
 
@@
 An angle of $\cos(\theta)=1$ indicates a molecule with its bisector aligned with the reference axis, while $\cos(\theta)=-1$ results from an anti-aligned configuration.
-Sample representations of different values of $\cos(\theta)$ are shown in figure \ref{fig:water-angles}b.
+Sample representations of different values of $\cos(\theta)$ are shown in Figure \ref{fig:water-angles}B.
 
@@
 the plane of the molecule is aligned with the plane of the surface), or if it is perpendicular.
-The values of $\cos(\phi)$ for different molecular orientations are shown in figure \ref{fig:water-angles}c.
+The values of $\cos(\phi)$ for different molecular orientations are shown in Figure \ref{fig:water-angles}C.
 %However, the symmetry of the \wat~and \suldiox~molecules create equivalence between the two H, or O, atoms (in the \wat~or\suldiox, respectively).
@@
 Previously published works have detailed graph theory for use in molecular systems, and have used graph theoretical methods to describe ice and water clusters, structural energies, as well as in the study of relationships in biological systems.\cite{Huber2007,Shi2005,Radhakrishnan1991,Anick2002} As such, we will not present here a review of the basic graph theoretical details that are well documented in the referenced works.
-As shown in figure \ref{fig:so2-complex-graph}, the atoms of a \suldiox~and its nearest neighbors are represented as a set of verrtices, with the covalent and hydrogen bonds between them forming the edges of a graph.
+As shown in Figure \ref{fig:so2-complex-graph}, the atoms of a \suldiox~and its nearest neighbors are represented as a set of verrtices, with the covalent and hydrogen bonds between them forming the edges of a graph.
 

### 2011-06-30 TransitOrientation.tex
Draft 8

@@
 	The orientation of \suldiox~molecules throughout the aqueous adsorption process was monitored during the transit SMD simulations.
-The angles $\theta$ and $\phi$ of the transiting \suldiox~(figure \ref{fig:water-angles}) were calculated for each timestep of the SMD simulations as the \suldiox~was pulled into the water slab from the gas phase, both in the neat-water and saturated slab systems.
-The angle cosine values were collected for the 50 simulations of both systems for each distance from the water surface, resulting in the 2-dimensional angle and depth-profile histograms shown in figure \ref{fig:so2-transit-angles}.
-Only the surface into which the transiting \suldiox~was steered is used in the analyses.
+The angles $\theta$ and $\phi$ of the transiting \suldiox~(Figure \ref{fig:water-angles}) were calculated for each timestep of the SMD simulations as the \suldiox~was pulled into the water slab from the gas phase, both in the neat-water and saturated slab systems.
+The angle cosine values were collected for the 50 simulations of both systems for each distance from the water surface location, resulting in the 2-dimensional angle and depth-profile histograms shown in Figure \ref{fig:so2-transit-angles}.
 In all the data sets, the populations of the angle histograms at each distance from the surface were normalized to aid in comparison of regions with differing \suldiox~residence times.
@@
 
-	From its starting position above the water surface, until the \suldiox~moves to within 10\angs~of the surface of both systems, the orientation is isotropic in $\theta$ and $\phi$ (not shown).
+	From its starting position above the water surface, until the \suldiox~moves to within 10\angs~of the water surface location of both systems, the orientation is isotropic in $\theta$ and $\phi$ (not shown).
 Isotropic orientation is manifested in the plots as mostly uniform coloration at a given distance from the surface independent of $\cos$ of $\phi$.
-Near and into the surface the bisector angle $\theta$ becomes more perpendicular to the water interface ($\cos(\theta)\approx1$) with the \suldiox~sulfur pointing into the water phase, consistent with the equilibrium MD simulation results above.
-At the point when the \suldiox~reaches the surface location (distance $=0$\angs), the bisector is perpendicular in both the neat-water and saturated systems.
+Near and into the interfacial region the bisector angle $\theta$ becomes more perpendicular to the water surface ($\cos(\theta)\approx1$) with the \suldiox~sulfur pointing into the water phase, consistent with the equilibrium MD simulation results above.
+At the point when the \suldiox~reaches the water surface location (distance $=0$\angs), the bisector is perpendicular to the interface in both the neat-water and saturated systems.
 Figure \ref{fig:so2-surface-angles}a depicts \suldiox~near the water surface with the bisector aligned perpendicular to the interface.
@@
   % talk about theta differences
-  Comparing in more detail the difference in these two systems upon \suldiox~approach, for the neat-water surface, a transition occurs at approximately 5\angs~from the surface.
-This coincides with the approximate distance of bonds formed to the outer-most water molecules.
-Below 5\angs above the surface, the orientation persists and the \suldiox~begins to interact with the topmost surface waters, and orients accordingly.
+  Comparing in more detail the difference in these two systems upon \suldiox~approach, for the neat-water surface, a transition occurs at approximately 5\angs~from the surface water location.
+Below 5\angs~above the surface, \suldiox~has a preferred net orientation and is close enough to the water surface that it begins to interact with the topmost surface waters.
 In the saturated system, the same trend occurs, however the onset of the perpendicular orientation begins at approximately 8\angs~above the surface.
@@
 In both systems the $\cos(\phi)$ profiles exhibit mostly isotropic distributions above 0\angs, with several regions of lighter coloration interspersed, but without a clearly formed orientational trend.
-At the neat-water surface and just below, from 0\angs~to -5\angs~in the water phase, the $\theta$ profile broadens down to $\cos(\theta)=0.5$, coinciding with a clear trend in the profile of $\phi$ near $\cos(\phi)=0$ shown in dark red at the bottom of the plot.
+At the neat-water surface and just below, from 0\angs~to -5\angs~(in the water phase), the $\theta$ profile broadens down to $\cos(\theta)=0.5$, coinciding with a clear trend in the profile of $\phi$ near $\cos(\phi)=0$ shown in dark red at the bottom of the plot.
 This corresponds to \suldiox~inclined up to 60 degrees from the surface normal.

### 2011-07-01 Article.tex
Fixed bibliography note issue, and made title page w/ abstract.
Lots of small corrections (typos!!) to draft 8 handed back for polishing.

--- 2011-06-28 Article.tex
@@
 
-\title{Sulfur Dioxide on Water's Surface}
+\title{Dancing on Water: The Choreography of Sulfur Dioxide Adsorption to Aqueous Surfaces}
 \author{Eric S. Shamay \and Kevin E. Johnson \and Geraldine L. Richmond}
@@
 
-%\maketitle
+\maketitle
 
@@
 \begin{abstract}
-Sulfur dioxide binding to water surfaces is a model system for the study of gaseous adsorption processes on aqueous systems.
-Equilibrium and adsorbing \suldiox~molecules in a pure water interfacial system are modeled by classical molecular dynamics simulations using fully polarizable models for both \suldiox~and \wat.
-Neat-water systems with a single \suldiox, and high \suldiox~concentration saturated water systems are modeled at equilibrium and using steered molecular dynamics (SMD) to ascertain the orientational profile of a transiting \suldiox~molecule.
-Orientational depth profiling reveals the preferred orientations of \suldiox~during adsorption, and the complementary behaviors of surface waters adsorbing the gaseous solute.
+One might expect the high surface tension of water to be a barrier to adsorption of a gas into the liquid phase, but we know that gaseous adsorption on and into a water surface is a common phenomenon on this planet.
+What is not commonly known is how an atmospheric gas such as \suldiox~and molecules at the water surface can overcome the barrier created by strong water-water surface bonding interactions.
+What this interplay looks like, the distances from the water surface at which these attractive interactions begin, and how they influence the orientational nature of both \suldiox~and surface water molecules is the focus of this computational study.
+The results fill a void in the information about this system existing from previous experimental studies by providing information about the dimensional nature of the gas-surface interactions, and the details of how the two species twist and turn orientationally with increased surface interactions.
+Classical molecular dynamics have been employed in both equilibrium and steered molecular dynamics (SMD) simulations for \suldiox~at a neat-water surface, and at a surface with high interfacial \suldiox~concentrations.
+The results provide new molecular insights for understanding the interaction of this prevalent gas on aerosols and other aqueous surfaces in the environment.
 \end{abstract}

### 2011-07-01 Conclusions.tex
Lots of small corrections (typos!!) to draft 8 handed back for polishing.

@@
 
-Through these simulations we are able to characterize the orientational behavior of \suldiox~during and after adsorption.
-Our equilibrium simulations show that a single \suldiox~molecule, representing a low concentration, has a high surface affinity.
-At a high \suldiox~concentration, unreacted \suldiox~molecules are found further out of the water phase as a bound layer of \suldiox~crowds the surface and binds all the surface waters.
+Through these simulations we characterize the orientational behavior of \suldiox~during and after adsorption.
+Our equilibrium neat-water simulations show that a single \suldiox~molecule, representing a low concentration, has a high surface affinity.
+At a high \suldiox~concentration in the saturated systems, \suldiox~molecules are also surface active, and are found further out of the water phase than at the lower concentration.
+These \suldiox~form a bound layer of molecules that crowd the surface and interact with the surface waters.
 The orientation of \suldiox~on the water surface was found to be similar for both low and high concentrations.
@@
 The results for the transit through the interface show that in both systems of low and high \suldiox~concentration an adsorbing \suldiox~has very similar orientation to those already bound to the water surface.
-The \suldiox~reorients as it makes its first contact with the water surface.
+The \suldiox~reorients as it makes its first contact with the water interface.
 Within 5 $\AA$ of the surface the \suldiox~is mostly oriented with its sulfur towards the water phase.
@@
 
-Having now examined this passageway of \suldiox~from the gas to adsorbed aqueous phases, we may turn to further filling in the details of adsorption and interfacial chemistry of gas molecules.
+Having now examined the transition of \suldiox~from the gas to adsorbed aqueous phases, we are better able to research the many questions remaining of adsorption and interfacial chemistry of gas molecules.
 What species form during the adsorption transition, and how do molecular properties affect the process?
 Can we form a theory that will more fully explain the transit of many small molecules of environmental and industrial interest?
-This study is one of several aiming to characterize \suldiox~adsorption and behavior on aqueous surfaces.
+This study is one of several characterizing \suldiox~adsorption and behavior on aqueous surfaces.
 We plan to report further results of ongoing computational simulation and experimental studies regarding temperature and chemical constituent effects on atmospherically relevant water surfaces.

### 2011-07-01 DensityResults.tex
Lots of small corrections (typos!!) to draft 8 handed back for polishing.

@@
 
-In the simulated neat-water slab, the result that the \suldiox~molecular density distribution concentrates near the water surface indicates an affinity for the interfacial region.
+In the simulated neat-water slab, the \suldiox~molecular density distribution concentrates near the water surface location, indicating an affinity for the interfacial region.
 During the course of the simulation, the \suldiox~does not venture into the bulk water nor does it escape the water surface into the gas phase, but remains located within 5-\angs~of the surface region.
-This is consistent with what is found experimentally in our laboratory using VSFS,\cite{Tarbuck2005,Tarbuck2006} and also supported by the computational simulations and spectral calculations of Baer et al.\cite{Baer2010} The experimental studies indicated that upon exposure of \suldiox~to a \wat~surface, a layer of solvated \suldiox~forms, modifying the structure of water above the Gibbs dividing surface (GDS).
+This is consistent with what is found experimentally in our laboratory using VSFS,\cite{Tarbuck2005,Tarbuck2006} and also supported by the computational simulations and spectral calculations of Baer et al.\cite{Baer2010} The experimental studies indicated that upon exposure of \suldiox~to a \wat~surface, a layer of solvated \suldiox~forms, modifying the structure of water in the upper surface region.
 In the experimental system, the \suldiox~bonding interaction with the free OH oscillators is manifested in a red-shift of the free OH frequency indicative of a bonding interaction.
@@
 The center of the \suldiox~density distribution is further into the gas phase than for the neat-water surface with a single-\suldiox~molecule.
-Because of the limitations of the classical model in accurately reproducing the first hydration shell around the \suldiox,\cite{Baer2010} we do not draw specific conclusions about the hydrate complex's geometry.
-However, the surface affinity of the hydrated \suldiox~is well reproduced by both the classical and ab initio methods.\cite{Baer2010}
+Because of the limitations of the classical model in accurately reproducing the first hydration shell around the \suldiox,\cite{Baer2010} we do not draw specific conclusions about the hydrated surface complex's geometry.
+However, the surface affinity of the hydrated \suldiox~is well reproduced by both the classical and ab initio methods.
 

### 2011-07-01 EquilibratedOrientation.tex
Lots of small corrections (typos!!) to draft 8 handed back for polishing.

@@
 The orientation depth-profiles for \wat~are shown in Figure \ref{fig:water-orientation} for both the neat-water (top) and saturated (bottom) systems during the equilibrium MD simulations.
-The interfacial region for both these calculations and the VSF experiments is defined as the region where mlecular orientational anisotropy exists around the surface water location.
-Referring to Figure \ref{fig:water-orientation}, calculations indicate an interfacial width of approx.
-10\angs~for the neat-water system, and approx.
-16 \angs~for the saturated system.
-In both systems the strongest orientational preference is found at the slab surfaces (distances greater than 0\angs) where the water is furthest towards the gas phase.
+The interfacial region for both these calculations and the VSF experiments is defined as the region where molecular orientational anisotropy exists around the surface water location.
+Referring to Figure \ref{fig:water-orientation}, calculations indicate an interfacial width of approximately 10\angs~for the neat-water system, and approximately 16 \angs~for the saturated system.
+In both systems the strongest orientational preference is found at the slab surfaces (positions above 0\angs) where the water is furthest towards the gas phase.
 Previous work on orientational preference of water at air surfaces shows the same trend as our neat-water results.\cite{Walker2006b,Hore2008} The narrow "peninsula" of high intensity that extends in the $\cos(\theta)$ plots to the right of the -5\angs~location shows the overall preference of water to orient at the surface.
@@
 
-Although the plots show overall similarities for both the neat-water and saturated systems, the presence of a layer of adsorbed \suldiox~molecules alters the water orientation of the waters furthest into the gas phase.
+Although the plots show overall similarities for both the neat-water and saturated systems, the presence of a layer of adsorbed \suldiox~molecules alters the orientation of those waters furthest into the gas phase.
 For the saturated solution the resulting orientation of waters above 0\angs, shown in the saturated plot of $\cos(\theta)$ of Figure \ref{fig:water-orientation}, is with a bisector pointing further into the adsorbed \suldiox~gas layer, and both hydrogens pointing outward from the aqueous bulk.
@@
 
-In the interfacial region Within 0-5\angs~of the water surface location (the approximate distance from the surface that \suldiox~begins interacting with the outer-most waters) the angular distribution of the single \suldiox~(in the neat-water system) is concentrated primarily in $\cos(\theta)>0$.
+In the interfacial region within 0-5\angs~of the water surface location (the approximate distance from the surface that \suldiox~begins interacting with the outer-most waters) the angular distribution of the single \suldiox~(in the neat-water system) is concentrated primarily in $\cos(\theta)>0$.
 The peak of the distribution occurs at $\cos(\theta)=1$.

### 2011-07-01 Introduction.tex
Lots of small corrections (typos!!) to draft 8 handed back for polishing.

@@
 Although we know much about the behavior of a gas on either side of that entrance, far less is known about how that surface acts to attract, facilitate, or thwart the transit of a molecule between the two bulk phases.
-What is the interplay between the gas and surface water molecules and when does one begin to influence the behavior of the other?
+What is the interplay between the gas and surface water molecules, and when does one begin to influence the behavior of the other?
 What species form during gas adsorption onto liquid surfaces, and what are the intermediary steps?
@@
 
-An important gas for developing a picture of gaseous adsorption and entry into a wter surface is \suldiox.\cite{Lattanzi2010,Shah2011,Tzivian2011,Johns2011,Faloona2009,Jurkat2010,Wu2011} Sulfur dioxide enters the environment as an important industrial product, and also naturally through terrestrial processes.
+An important gas for developing a picture of gaseous adsorption and entry into a water surface is \suldiox.\cite{Lattanzi2010,Shah2011,Tzivian2011,Johns2011,Faloona2009,Jurkat2010,Wu2011} Sulfur dioxide enters the environment as an important industrial product, and also naturally through terrestrial processes.
 Atmospheric dust particles and gases have been implicated in the oxidation of \suldiox, and act as reaction surfaces for chemical mechanisms that are still poorly understood.\cite{Baltrusaitis2011,Rubasinghege2010,Li2007} \suldiox~acts as a major component of atmospheric pollution, and is a precursor to acid rain formation, and cloud nucleation.
@@
 In this work we provide a molecular picture of \suldiox~adsorption on a water surface.
-The studies demonstrate the strong orientational effect that surface water molecules have on the adsorbing gas during the approach and entry into the surface region, and how surface water molecules rise towards the gas phase to assist in the capture of \suldiox~molecules at both high and low \suldiox~surface concentrations.
+This study demonstrates the strong orientational effect of surface water molecules on adsorbing gas during the approach and entry into the surface region at both high and low \suldiox~surface concentrations.
+Obtaining a more complete picture of this adsorption process is important for understanding gaseous adsorption of this environmentally important gas on water and aerosol surfaces as well as being a model system for understanding the more general nature of gases at aqueous interfaces.
 
-These computational studies complement and significantly expand the picture developed in our recent experimental vibrational sum frequency spectroscopy (VSFS) studies of \suldiox~adsorption of aqueous solutions of various compositions and temperatures, and the subsequent studies by Baer et al.\cite{Baer2010} using both classical and ab initio simulations.
+These computational studies complement and significantly expand the picture developed in our recent experimental vibrational sum frequency spectroscopy (VSFS) studies of \suldiox~adsorption of aqueous solutions of various compositions and temperatures,\cite{Tarbuck2005,Tarbuck2006} and the subsequent studies using both classical and ab initio simulations.
+These experimental studies showed that an \suldiox~surface hydrate complex forms when an aqueous surface is exposed to \suldiox~gas.
+The computational study by Baer et al.\cite{Baer2010} then made a series of predictions of the specific nature of the hydrated complex through classical and ab initio simulations.
+That work developed a detailed picture of the nature of the \suldiox~surface complex with water, and related it to the surface water OH vibrational IR spectra.
 The most recent experimental studies have shown that whereas the binding of gaseous \suldiox~to a water surface is greatly enhanced at cold temperatures, the reversibility of the adsorption process remains.
@@
 Since the \suldiox~molecule was not specifically probed, conclusions on how \suldiox~bonding contributes to reorienting surface water molecules and the orientation of \suldiox~itself upon approach and surface bonding could only be inferred.
-Obtaining a more complete picture of this adsorption process is important for understanding gaseous adsorption of this environmentally important gas on water and aerosol surfaces as well as being a model system for understanding the more general nature of gases at aqueous interfaces.
-
-Our recent vibrational sum frequency spectroscopy (VSFS) experiments have provided insight into the adsorption and reaction of gases in atmospheric aerosols.\cite{Tarbuck2005,Tarbuck2006} These studies showed that an \suldiox~surface hydrate complex forms when an aqueous surface is exposed to \suldiox~gas.
-A computational study by Baer et al.\cite{Baer2010} then made a series of predictions of the specific nature of the hydrated complex through classical and ab initio simulations.
-That work developed a detailed picture of the nature of the \suldiox~surface complex with water, and related it to the surface water OH vibrational IR spectra.
 

### 2011-07-01 Methods.tex
Lots of small corrections (typos!!) to draft 8 handed back for polishing.

@@
 A second set of simulations began with an equilibrated water slab as in the surface equilibrated method above.
-However, in both the equilibrated neat-water and saturated systems, a single \suldiox~was introduced 20\angs~above the water slab surface, with the sulfur atom tethered to its initial position.
-The systems were evolved for 1 ns, taking coordinate snapshots every 20 ps to create 50 starting points for further simulations.
+However, in both the neat-water and saturated starting systems, a single \suldiox~was introduced 20\angs~above the water slab surface, with the sulfur atom tethered to its initial position.
+The systems were then evolved for 1 ns, taking coordinate snapshots every 20 ps to create 50 starting points for further simulations.
 Steered molecular dynamics (SMD) were then performed on the 50 system configurations (in both the neat-water and saturated configurations) to guide the \suldiox~down towards a tethered water near the water slab's center of mass by applying a small steering force to the \suldiox-sulfur atom.
@@
 	The first portion of the studies involved creating orientational depth-profiles in the interfacial region comprised of \suldiox~and \wat~upon exposure and adsorption of the gas.
-To do this we must define a reference point in thie interfacial region which we refer to here as the water surface location.
-Recognizing that a liquid surface is a dynamic boundary that is neither flat nor stationary.
+Recognizing that a liquid surface is a dynamic boundary that is neither flat nor stationary, we must define a reference point in the interfacial region which we refer to here as the water surface location.
 Several previous studies have used the technique of fitting a line shape to the averaged density profile of the water, and extracting interfacial shape and location parameters to define the water surface location.\cite{Shamay2010,Wick2006c,Chowdhary2006} Hyperbolic tangent functions have been used often, and values for the ``Gibb's dividing surface'' location, and interfacial width have thus been determined.\cite{Matsumoto1988} However, in long simulations the location and shape of the interface changes, and the motion of surface waters alters the interfacial width at any given time step.
@@
 Figure \ref{fig:density-flaw}A shows the density profile of water and \suldiox~over the course of one of the 10 ns trajectories used in this work using the original uncorrected coordinates of the system taken from the raw atomic positions of the molecular dynamics data output.
-The water density profile and location (thin grey line in Figure \ref{fig:density-flaw}A) is produced by averaging the instantaneous density profile at each time step in the simulation over all the time steps.
+The water density profile and location (thin gray line in Figure \ref{fig:density-flaw}A) is produced by averaging the instantaneous density profile at each time step in the simulation over all the time steps.
 The water profile was then fit to a $\tanh$ function (black line) to extract the position and width parameters of the water surface.
@@
 Increasing the number of waters used in calculating the surface location diminishes the effects of the few waters that briefly rise above the surface into the gas phase, stabilizing both the surface position and thickness values.
-Taking the close-packed model as a maximum number of waters fit into a flat surface, 70 water molecules were used for calculating instantaneous water surface locations.
+Taking the close-packed model as a maximum number of waters fit into a flat surface, the topmost 70 water molecules were used for calculating instantaneous water surface locations for each simulation step.
 
@@
 		\includegraphics[scale=1.0]{images/density/density-location.png}
-		\caption{(A) Density profiles of \wat~(grey) and \suldiox~(red) from a 10 ns simulation of the neat-water system with a single sulfur dioxide.
+		\caption{(A) Density profiles of \wat~(Gray) and \suldiox~(red) from a 10 ns simulation of the neat-water system with a single sulfur dioxide.
 The \wat~density profile was fit using a hyperbolic tangent function (black) to extract location and width data for the water surface.

### 2011-07-01 Methods.tex
Starting draft 9

@@
 This method for finding the outer monolayer location was implemented, and the surface location is plotted as a function of simulation time in Figure \ref{fig:density-flaw}B.
-It is apparent from the surface location plot that the water slab location, and thus the surface location, drifts over the 10 ns, spanning approx.
-12 $\AA$.
+It is apparent from the surface location plot that the water slab location, and thus the surface location, drifts over the 10 ns, spanning approximately 12 $\AA$.
 However, the maximum standard deviation of the positions of the waters comprising the surface layer at each time step is only 1.85 $\AA$.

### 2011-07-04 Conclusions.tex
fixed x-axis location on figures. Redid conclusions. Minor edits for the rest.

@@
 
-Presented herein are the results of several classical molecular dynamic simulations that focus on understanding how surface water molecules and adsorbing \suldiox~gas molecules twist and turn as the gas transits the interfacial region.
+Presented herein are the results of several classical molecular dynamic simulations that focus on understanding how surface water molecules and adsorbing \suldiox~gas molecules twist and turn as the gas adsorbs and transits the interfacial region.
 The computational studies emulate and expand on the experimental spectroscopic studies from this laboratory which have found \suldiox~surface complexation at a water surface.\cite{Tarbuck2005,Tarbuck2006,Ota2011}  These spectroscopic studies show clear evidence of \suldiox-water surface complexation but details about this surface complex could only be inferred from spectral changes in the surface water spectrum since \suldiox~could not be monitored directly.
@@
 
-Through these simulations we characterize the orientational behavior of \suldiox~during and after adsorption.
+Through these simulations we also characterize the orientational behavior of \suldiox~during and after adsorption.
 Our equilibrium neat-water simulations show that a single \suldiox~molecule, representing a low concentration, has a high surface affinity.
 At a high \suldiox~concentration in the saturated systems, \suldiox~molecules are also surface active, and are found further out of the water phase than at the lower concentration.
-These \suldiox~form a bound layer of molecules that crowd the surface and interact with the surface waters.
+These \suldiox~molecules form a bound layer that crowds the surface and interacts with the surface waters.
 The orientation of \suldiox~on the water surface was found to be similar for both low and high concentrations.
-Those \suldiox~molecules at or below the surface strongly orient with the sulfur atom pointed in towards the water bulk, and the oxygen atoms out towards the gas phase.
-The \suldiox~slightly above the water surface loses the orientational preference within 10 $\AA$ and those further from the water are more isotropically oriented.
-Figure \ref{fig:so2-surface-cartoon} depicts what the neat-water and saturated surfaces might look like for both \suldiox~and \wat~orientations and locations.
+Those \suldiox~molecules at or below the surface water location strongly orient with the sulfur atom pointed in towards the water bulk, and the oxygen atoms out towards the gas phase.
+The \suldiox~molecules slightly above the water surface lose this net orientation within 10 \angs.
+Those  molecules further from the water are more isotropically oriented.
+Figure \ref{fig:so2-surface-cartoon} depicts what the neat-water and saturated surface molecules look like for both \suldiox~and \wat~orientations and locations based on the calculations.
 
@@
 Steered molecular dynamics simulations were used to model the behavior of an adsorbing \suldiox~as it moves from the gas phase above the water down through the surface and into the bulk.
-The results for the transit through the interface show that in both systems of low and high \suldiox~concentration an adsorbing \suldiox~has very similar orientation to those already bound to the water surface.
 The \suldiox~reorients as it makes its first contact with the water interface.
-Within 5 $\AA$ of the surface the \suldiox~is mostly oriented with its sulfur towards the water phase.
+Within 5 \angs~of the surface the \suldiox~is mostly oriented with its sulfur towards the water phase.
+The results for the transit through the interface show that in both systems of low and high \suldiox~concentration an adsorbing \suldiox~near the interface has very similar orientation to those molecules already bound to the water surface.
 The \suldiox~pulled further into the water bulk retains its orientation until it is past the interfacial region and then isotropically orients with the bulk water.
 
-Having now examined the transition of \suldiox~from the gas to adsorbed aqueous phases, we are better able to research the many questions remaining of adsorption and interfacial chemistry of gas molecules.
+These studies provide a starting point for future studies in this area that seek to understand how gases of different concentrations and chemical composition adsorb and transit across an aqueous/air interface.
+Obtaining such knowledge will be invaluable for understanding many environmental aerosol and land water systems where gaseous uptake at a water surface does not conform to expectations.
+In such cases, surface complexes are often invoked with little direct evidence of the formation of such complexes.
+Investigations of these and related low-temperature effects will be forthcoming in future publications.
+
+%Having now examined the transition of \suldiox~from the gas to adsorbed aqueous phases, we are better able to research the many questions remaining of adsorption and interfacial chemistry of gas molecules.
 What species form during the adsorption transition, and how do molecular properties affect the process?

### 2011-07-04 Introduction.tex
fixed x-axis location on figures. Redid conclusions. Minor edits for the rest.

@@
 
-An important gas for developing a picture of gaseous adsorption and entry into a water surface is \suldiox.\cite{Lattanzi2010,Shah2011,Tzivian2011,Johns2011,Faloona2009,Jurkat2010,Wu2011} Sulfur dioxide enters the environment as an important industrial product, and also naturally through terrestrial processes.
+An important gas for developing a picture of gaseous adsorption and entry into a water surface is sulfur dioxide.\cite{Lattanzi2010,Shah2011,Tzivian2011,Johns2011,Faloona2009,Jurkat2010,Wu2011} \suldiox~enters the environment as an important industrial product, and also naturally through terrestrial processes.
 Atmospheric dust particles and gases have been implicated in the oxidation of \suldiox, and act as reaction surfaces for chemical mechanisms that are still poorly understood.\cite{Baltrusaitis2011,Rubasinghege2010,Li2007} \suldiox~acts as a major component of atmospheric pollution, and is a precursor to acid rain formation, and cloud nucleation.
 Its high solubility in water makes \suldiox~an integral compound in many aqueous atmospheric reactions, as well.
+Obtaining a more complete picture of the \suldiox~adsorption process is important for understanding gaseous adsorption of this environmentally important gas on water and aerosol surfaces as well as being a model system for understanding the more general nature of gases at aqueous interfaces.
 
 In this work we provide a molecular picture of \suldiox~adsorption on a water surface.
-This study demonstrates the strong orientational effect of surface water molecules on adsorbing gas during the approach and entry into the surface region at both high and low \suldiox~surface concentrations.
-Obtaining a more complete picture of this adsorption process is important for understanding gaseous adsorption of this environmentally important gas on water and aerosol surfaces as well as being a model system for understanding the more general nature of gases at aqueous interfaces.
-
+This study demonstrates the strong orientational effect of surface water molecules on the adsorbing gas during the approach and entry into the surface region at both high and low \suldiox~surface concentrations.
 These computational studies complement and significantly expand the picture developed in our recent experimental vibrational sum frequency spectroscopy (VSFS) studies of \suldiox~adsorption of aqueous solutions of various compositions and temperatures,\cite{Tarbuck2005,Tarbuck2006} and the subsequent studies using both classical and ab initio simulations.
@@
 That work developed a detailed picture of the nature of the \suldiox~surface complex with water, and related it to the surface water OH vibrational IR spectra.
-The most recent experimental studies have shown that whereas the binding of gaseous \suldiox~to a water surface is greatly enhanced at cold temperatures, the reversibility of the adsorption process remains.
-Complimentary experiments showed that low pH aqueous environments inhibit the bulk reactions of \suldiox, but do not affect the surface binding or its reversibility.
+The most recent experimental studies have shown that whereas the binding of gaseous \suldiox~to a water surface is greatly enhanced at cold temperatures, the reversibility of the adsorption process remains.\cite{Ota2011}  Complementary experiments showed that low pH aqueous environments inhibit the bulk reactions of \suldiox, but do not affect the surface binding or its reversibility.
 What is apparent in the VSF spectra obtained in all of these experiments is the tendency of water to reorient upon surface bonding, with the effect becoming more pronounced at high \suldiox~surface concentrations.

### 2011-07-04 Methods.tex
fixed x-axis location on figures. Redid conclusions. Minor edits for the rest.

--- 2011-07-01 Methods.tex
@@
 	To overcome this problem we define the water surface location by calculating a reference location at each time step by averaging the positions of the waters contained in the topmost monolayer.
-This provides us with a consistent and intuitive reference point in the simulations to which we relate our analyses, but does not increase the computational burden.
+This provides a consistent and intuitive reference point in the simulations to which we relate our analyses, but does not increase the computational burden.
 The number of waters included in the averaging is determined by taking a few issues into account.
@@
 This method for finding the outer monolayer location was implemented, and the surface location is plotted as a function of simulation time in Figure \ref{fig:density-flaw}B.
-It is apparent from the surface location plot that the water slab location, and thus the surface location, drifts over the 10 ns, spanning approx.
-12 $\AA$.
-However, the maximum standard deviation of the positions of the waters comprising the surface layer at each time step is only 1.85 $\AA$.
+It is apparent from the surface location plot that the water slab location, and thus the surface location, drifts over the 10 ns, spanning approximately 12 \angs.
+However, the maximum standard deviation of the positions of the waters comprising the surface layer at each time step is only 1.85 \angs.
 Consequently, all depth locations in our analyses are calculated relative to the instantaneous surface location at the corresponding time step of the trajectories.
@@
 \end{figure}
-
-%\subsection{Hydrate Complex Analysis}
-
-%During the adsorption process of a gaseous \suldiox~moving in towards a water surface hydrate complexes are formed as the waters bind to the molecule.
-Computationally, there are various methods to determine hydrate complex formation.
-The most simple metric is to count the number of hydrogen bonds formed from the \suldiox~to neighboring molecules.
-This method can be furthered extended to also include the strength of those bonding interactions by additionally reporting the bondlengths to the hydrogen bonding partners.
-Longer bondlengths are typical of weaker bonds, and this method is used in this work to identify the bonding environment of the \suldiox~during solvation.
-
-%A simple count of hydrogen bonds is telling of the interactions and environments of an adsorbing \suldiox.
-However, a more sophisticated analysis is needed to fully characterize the geometry of any complexes that form.
-In this work a graph theoretical method has been used to identify hydrate complexes that form cyclic structures in addition to those that form more simple hydrogen bonding geometries.
-Previously published works have detailed graph theory for use in molecular systems, and have used graph theoretical methods to describe ice and water clusters, structural energies, as well as in the study of relationships in biological systems.\cite{Huber2007,Shi2005,Radhakrishnan1991,Anick2002} As such, we will not present here a review of the basic graph theoretical details that are well documented in the referenced works.
-As shown in Figure \ref{fig:so2-complex-graph}, the atoms of a \suldiox~and its nearest neighbors are represented as a set of verrtices, with the covalent and hydrogen bonds between them forming the edges of a graph.
-
-%We present here a technique for determining bonding cycles in a hydrate complex, but this technique is general enough to be applied to any graph representation of simple molecular systems.
-The first step in the determination of cycles is translating the atoms (vertices) and bonds (edges) to a graph.
-Figure \ref{fig:so2-complex-graph} shows the procedure visually, where here the root vertex of the graph is the \suldiox-sulfur, and each atom is numbered by the distance, in atoms, to the root vertex.
-We use only the closest neighboring water molecules that may take part in a cyclic complex in building the graph.
-After forming the graph, a breadth-first search (BFS) is performed starting at the root vertex.\cite{Knuth1997} The BFS algorithm uses a coloring technique where all unexamined vertices are colored white, queued vertices are gray, and examined vertices are black.
-During the BFS, encountering a gray vertex as a target of an adjacent edge during examination indicates that a cycle has been found in the graph.
-At each vertex dequeue for examination, the parent of every vertex is recorded such that the lineage from gray target to root vertex can be reconstructed to determine the participating atoms in the cycle.
-
-%\begin{figure}[h!]
-%\begin{center}
-%\includegraphics[scale=1.0]{images/Cycles/so2complexfigure.png}
-%\caption{A \suldiox~and its nearest neighbors are represented as a graph of vertices and edges.
-The bonds are all given equal weights and the graph is undirected.
-The numbering shows the relative distance from any atom to the root node (\suldiox-sulfur 1), and also enumerates the iteration through the BFS process at which the atom is discovered.
-A BFS on this graph would result in discovering the doubly-bound hydrogen (5) as a gray target, and the gray source would be either of the connected oxygens (4).
-At each iteration the predecessor in the BFS is recorded in order to reconstruct the connectivity of the cycle.
-This cycle is composed of 8 atoms, and 3 molecules}
-%\label{fig:so2-complex-graph}
-%\end{center}
-%\end{figure}

### 2011-07-04 Introduction.tex
Draft 9

@@
 
-An important gas for developing a picture of gaseous adsorption and entry into a water surface is sulfur dioxide.\cite{Lattanzi2010,Shah2011,Tzivian2011,Johns2011,Faloona2009,Jurkat2010,Wu2011} \suldiox~enters the environment as an important industrial product, and also naturally through terrestrial processes.
-Atmospheric dust particles and gases have been implicated in the oxidation of \suldiox, and act as reaction surfaces for chemical mechanisms that are still poorly understood.\cite{Baltrusaitis2011,Rubasinghege2010,Li2007} \suldiox~acts as a major component of atmospheric pollution, and is a precursor to acid rain formation, and cloud nucleation.
+An important gas for developing a picture of gaseous adsorption and entry into a water surface is sulfur dioxide.\cite{Lattanzi2010,Shah2011,Tzivian2011,Johns2011,Faloona2009,Jurkat2010,Wu2011,Jayne1990,Yang2002} \suldiox~enters the environment as an important industrial product, and also naturally through terrestrial processes.
+Atmospheric dust particles and gases have been implicated in the oxidation of \suldiox, and act as reaction surfaces for chemical mechanisms that are still poorly understood.\cite{Baltrusaitis2011,Rubasinghege2010,Li2007,Madsen2008,Boniface2000} \suldiox~acts as a major component of atmospheric pollution, and is a precursor to acid rain formation, and cloud nucleation.
 Its high solubility in water makes \suldiox~an integral compound in many aqueous atmospheric reactions, as well.

### 2011-07-04 Methods.tex
Draft 9

@@
 First, counting the waters found in the topmost cross-section of the water slab over several time steps indicated between 65-75 waters that established a full monolayer.
-This was done by a visual inspection of the slab using the VMD MD visualization package.\cite{HUMP96} Alternatively, assuming a spherical model of water with a radius of 2.2\angs~(the typical length of a bulk-water hydrogen bond), two layers of hexagonally tight-packed spheres yielded a similar number of surface water molecules.
+This was done by a visual inspection of the slab using the VMD MD visualization package.\cite{Humphrey1996} Alternatively, assuming a spherical model of water with a radius of 2.2\angs~(the typical length of a bulk-water hydrogen bond), two layers of hexagonally tight-packed spheres yielded a similar number of surface water molecules.
 Increasing the number of waters used in calculating the surface location diminishes the effects of the few waters that briefly rise above the surface into the gas phase, stabilizing both the surface position and thickness values.

### 2011-07-05 EquilibratedOrientation.tex
Starting edits for final draft

@@
 The plots of $\phi$ for both concentrations exhibit isotropic distributions.
-However, the $\theta$ values follow similar trends at both concentrations indicating that the \suldiox~sulfur orients down towards the water phase, with both oxygens pointing away from the surface once the \suldiox~is within approx.
-5\angs~of the surface.
+However, the $\theta$ values follow similar trends at both concentrations indicating that the \suldiox~sulfur orients down towards the water phase, with both oxygens pointing away from the surface once the \suldiox~is within approximately 5\angs~of the surface.
 This behavior continues down to at least 5\angs~below the water surface location, notwithstanding any chemical reactions that may occur that are not simulated using the classical MD techniques utilized in this work.

### 2011-07-05 Conclusions.tex
Final draft

@@
 
-The simulations described herein show that gaseous \suldiox~quickly adsorbs to the water surface and continues to bind until a complete surface coverage is reached.
+Our simulations show that gaseous \suldiox~quickly adsorbs to the water surface and continues to bind until a complete surface coverage is reached.
 Surface waters reorient in the presence of adsorbed \suldiox.
@@
 These studies provide a starting point for future studies in this area that seek to understand how gases of different concentrations and chemical composition adsorb and transit across an aqueous/air interface.
-Obtaining such knowledge will be invaluable for understanding many environmental aerosol and land water systems where gaseous uptake at a water surface does not conform to expectations.
-In such cases, surface complexes are often invoked with little direct evidence of the formation of such complexes.
-Investigations of these and related low-temperature effects will be forthcoming in future publications.
+Obtaining such knowledge will be invaluable for understanding many environmental aerosol and land water systems where gaseous uptake at a water surface does not conform to expectations.\cite{Jayne1990,Yang2002,Worsnop1989,Boniface2000} Investigations of these and related low-temperature effects will be forthcoming in future publications.
+
+% In such cases, surface complexes are often invoked with little direct evidence of the formation of such complexes.
 

### 2011-07-05 DensityResults.tex
Final draft

@@
 The center of the \suldiox~density distribution is further into the gas phase than for the neat-water surface with a single-\suldiox~molecule.
-Because of the limitations of the classical model in accurately reproducing the first hydration shell around the \suldiox,\cite{Baer2010} we do not draw specific conclusions about the hydrated surface complex's geometry.
-However, the surface affinity of the hydrated \suldiox~is well reproduced by both the classical and ab initio methods.
+Because of the limitations of the classical model in accurately reproducing the first hydration shell around the \suldiox, we do not draw specific conclusions about the hydrated surface complex's geometry.
+However, the surface affinity of the hydrated \suldiox~is well reproduced by both the classical and ab initio methods.\cite{Baer2010}
 

### 2011-07-05 EquilibratedOrientation.tex
Final draft

@@
 
-%The histograms show overall similarities for both systems in their shapes and intensities from approx.
-0\angs~and below.
+%The histograms show overall similarities for both systems in their shapes and intensities from approximately 0\angs~and below.
 The presence of a saturated layer of adsorbed \suldiox~molecules, however, alters the water orientation, but not necessarily the orientational depth of the interface.
@@
 The noise in the neat-water plots of Figure \ref{fig:water-orientation} above 5\angs~(manifested as disconnected points of intensity throughout the range of orientations) is a result of fewer waters venturing beyond those extents and thus less data far from the water surface location.
-This is one indication that waters near a layer of adsorbed \suldiox~venture further above the interface relative to the low \suldiox~concentration, where they can have more \suldiox~interaction with the adsorbed gas molecules.
-These results show that the interactions with neighboring \suldiox~molecules allow the waters above the surface to orient more perpendicularly to the interface.
+This is one indication that waters near a layer of adsorbed \suldiox~venture further above the water surface location relative to the low \suldiox~concentration, where they can have more \suldiox~interaction with the adsorbed gas molecules.
+These results show that the interactions with neighboring \suldiox~molecules allow the waters above the water surface location to orient more perpendicularly to the interface.
 This is consistent with our recent experimental VSFS studies which showed evidence for the reorienting behavior of water due to the \suldiox~interactions with the topmost surface waters.\cite{Ota2011}
@@
 Waters on the neat surface lie flatter, whereas the presence of the \suldiox~allows a greater range of ``twist'' for those waters in the plane of the interface.
-The $\phi$ distributions quickly become isotropic above the surface as the bisectors orient more perpendicularly, and below the surface as the bulk water loses any orientational preference.
+The $\phi$ distributions quickly become isotropic above the surface location as the bisectors orient more perpendicularly, and below the surface as the bulk water loses any orientational preference.
 

### 2011-07-05 TransitOrientation.tex
Lots of small corrections (typos!!) to draft 8 handed back for polishing.
Final draft

@@
 		\includegraphics[scale=1.0]{images/angle-cartoons/so2-surface.png}
-		\caption{(A) \suldiox~molecules above and below both neat-water and saturated surfaces orient with the bisectors perpendicular to the interface.
+		\caption{(A) \suldiox~molecules above and below both neat-water and saturated water surface location orient with the bisectors perpendicular to the interface.
 The \suldiox~sulfur points down towards the aqueous phase.
-(B) In the neat-water system, a transiting \suldiox~molecule located below the surface briefly reorients with one S-O bond more towards the interface above, and the other bond pointing further down into the water bulk.}
+(B) In the neat-water system, a transiting \suldiox~molecule located below the water surface briefly reorients with one S-O bond more towards the interface above, and the other bond pointing further down into the water bulk.}
 		\label{fig:so2-surface-angles}
@@
 The result is that \suldiox~in this region shows a preference for one of the S-O bonds to point more towards the aqueous bulk and the other to point away towards the interface.
-Figure \ref{fig:so2-surface-angles}b depicts the orientation of the \suldiox~slightly below the water surface.
-This is in contrast to \suldiox~molecules above the surface, oriented more perpendicularly.
-The behavior is likely due to the interaction between the waters and one of the S-O bonds leading to a higher solvation than above the water surface.
+Figure \ref{fig:so2-surface-angles}B depicts the orientation of the \suldiox~slightly below the water surface location.
+This is in contrast to \suldiox~molecules above the water surface location, oriented more perpendicularly.
+The behavior is likely due to the interaction between the waters and one of the S-O bonds leading to a higher solvation than above the water surface location.
 As the \suldiox~is solvated by more highly-coordinated bulk water, the S-O bonds experience less equal interaction environments.
@@
 It is notable that the same reorientation does not occur as strongly in the saturated system.
-The presence of the adsorbed \suldiox~layer apparently decreases the reorienting behavior likely because of the disrupting effect the higher \suldiox~concentration has on the water interactions at the surface.
+The presence of the adsorbed \suldiox~layer apparently decreases the reorienting behavior likely because of the disrupting effect the higher \suldiox~concentration has on the water interactions in the interfacial region.
 %Below this 5\angs~region under the surface, the \suldiox~resumes an isotropic orientation as it becomes fully solvated by bulk waters.

### 2011-07-05 submission/Article.tex
Ready for submission!

\documentclass{article}

\usepackage {amsmath}
\usepackage {setspace}
\usepackage [pdftex]{graphicx}
\usepackage[sort&compress]{achemso}

\oddsidemargin  0.0in
\evensidemargin 0.0in
\textwidth      6.5in

\title{Dancing on Water: The Choreography of Sulfur Dioxide Adsorption to Aqueous Surfaces}
\author{Eric S. Shamay \and Kevin E. Johnson \and Geraldine L. Richmond}

\begin{document}

\newcommand{\suldiox}{SO$_2$}
\newcommand{\ang}{\,$\textrm{\AA}$}
\newcommand{\angs}{\ang}
\newcommand{\wat}{H$_2$O}

\maketitle

\doublespacing

\section{Introduction}

The doorway to the uptake of a gas by an aqueous solution is the water surface. Although we know much about the behavior of a gas on either side of that entrance, far less is known about how that surface acts to attract, facilitate, or thwart the transit of a molecule between the two bulk phases. What is the interplay between the gas and surface water molecules, and when does one begin to influence the behavior of the other? What species form during gas adsorption onto liquid surfaces, and what are the intermediary steps?  Is molecular orientation of either the gas or surface molecules a factor in the adsorption process? Are specific gas or liquid molecular orientations necessary for gaseous adsorption?  Experimental studies to address such questions are valuable but do not provide the full resolution necessary to determine the geometries of adsorbing gases, or to determine the orientations of the molecules at the liquid surface near the adsorption site. This type of information can be determined computationally, and when coupled to the experimental studies can provide a more comprehensive picture of the gas-liquid surface adsorption process.

An important gas for developing a picture of gaseous adsorption and entry into a water surface is sulfur dioxide.\cite{Lattanzi2010,Shah2011,Tzivian2011,Johns2011,Faloona2009,Jurkat2010,Wu2011,Jayne1990,Yang2002} \suldiox~enters the environment as an important industrial product, and also naturally through terrestrial processes. Atmospheric dust particles and gases have been implicated in the oxidation of \suldiox, and act as reaction surfaces for chemical mechanisms that are still poorly understood.\cite{Baltrusaitis2011,Rubasinghege2010,Li2007,Madsen2008,Boniface2000} \suldiox~acts as a major component of atmospheric pollution, and is a precursor to acid rain formation, and cloud nucleation. Its high solubility in water makes \suldiox~an integral compound in many aqueous atmospheric reactions, as well.  Obtaining a more complete picture of the \suldiox~adsorption process is important for understanding gaseous adsorption of this environmentally important gas on water and aerosol surfaces as well as being a model system for understanding the more general nature of gases at aqueous interfaces.

In this work we provide a molecular picture of \suldiox~adsorption on a water surface.  This study demonstrates the strong orientational effect of surface water molecules on the adsorbing gas during the approach and entry into the surface region at both high and low \suldiox~surface concentrations.  These computational studies complement and significantly expand the picture developed in our recent experimental vibrational sum frequency spectroscopy (VSFS) studies of \suldiox~adsorption of aqueous solutions of various compositions and temperatures,\cite{Tarbuck2005,Tarbuck2006} and the subsequent studies using both classical and ab initio simulations.  These experimental studies showed that an \suldiox~surface hydrate complex forms when an aqueous surface is exposed to \suldiox~gas. The computational study by Baer et al.\cite{Baer2010} then made a series of predictions of the specific nature of the hydrated complex through classical and ab initio simulations. That work developed a detailed picture of the nature of the \suldiox~surface complex with water, and related it to the surface water OH vibrational IR spectra.  The most recent experimental studies have shown that whereas the binding of gaseous \suldiox~to a water surface is greatly enhanced at cold temperatures, the reversibility of the adsorption process remains.\cite{Ota2011}  Complementary experiments showed that low pH aqueous environments inhibit the bulk reactions of \suldiox, but do not affect the surface binding or its reversibility.  What is apparent in the VSF spectra obtained in all of these experiments is the tendency of water to reorient upon surface bonding, with the effect becoming more pronounced at high \suldiox~surface concentrations. Since the \suldiox~molecule was not specifically probed, conclusions on how \suldiox~bonding contributes to reorienting surface water molecules and the orientation of \suldiox~itself upon approach and surface bonding could only be inferred.

To fill this void, the computational studies described herein provide a detailed picture of the orientation of both \suldiox~and surface water molecules during the adsorption process.  The depth profiling studies which examine the orientation of both species during the approach and entry of the gas into the interfacial region are obtained using equilibrium and steered (SMD) classical molecular dynamics simulations. The latter approach involves steering a gas molecule into the aqueous phase, and characterizing its molecular orientation as it transits through the interfacial region.  This unique approach enables new insights into the behaviors of gas molecules as they move near to liquid water. We also simulate how \suldiox~adsorption occurs on a water surface saturated with adsorbed \suldiox, analogous to the conditions of the \suldiox~experimental studies recently performed.\cite{Ota2011} The results of this study provide an intimate perspective on the adsorption of \suldiox~at an aqueous surface, and a more complete picture of gaseous adsorption to liquid interfaces.

\section{Computational Approach}

Molecular dynamics simulations were performed using the Amber 11 software suite.\cite{Case2010} Polarizable models for the \wat~and \suldiox~molecules were used in the simulations, and have been used previously in studies on interfacial systems because they are known to more accurately reproduce interfacial structure and free energy profiles.\cite{Wick2007,Rivera2006,Dang1998} The \wat~model used is the POL3 water model,\cite{Caldwell1995} and for \suldiox~we used the model of Baer et al. that places a single polarizable center on the sulfur atom.\cite{Baer2010}

All simulations began with an equilibrated cube of 900 \wat~molecules, with sides of length 30\angs. The long axis of each simulation cell (the axis normal to the water surface) was then lengthened to 120\angs, and the systems were further equilibrated for 10 ns. The simulations all employed periodic boundaries to create an ``infinite-slab'' geometry. After equilibrating the neat-\wat~slabs two types of systems were created by introducing \suldiox: a single-\suldiox~system, herein referred to as the ``neat-water'' system, and a ``saturated'' \suldiox~system with many gaseous surface and bulk-water \suldiox~molecules.

The low and high concentration simulated systems, ``neat-water'' and ``saturated water'', respectively, were created as follows: the neat-water simulation involved the addition of a single \suldiox~molecule either within the bulk of the water slab (for equilibrium MD), or above the slab surface (in the SMD simulations). The single-\suldiox~neat-water system was then evolved for 2 ns to produce an equilibrated starting configuration. Because of the extremely low concentration of \suldiox~in the neat-water system the surface waters behave similarly to a true neat-water air-liquid interface. The neat-water system orientational results shown later in this work reproduce well the results of our previous orientational studies of surface water behavior.\cite{Walker2006b,Hore2008} The saturated system had 22 \suldiox~molecules introduced to the water slab bulk in order to saturate it to a level coinciding with the Henry's law constant for \suldiox~in water. Additionally, 50 \suldiox~molecules were introduced into the gas phase outside of the saturated water slab to simulate an added \suldiox~gas pressure. The saturated system with both bulk and gaseous \suldiox~was then evolved for 2 ns to produce a starting configuration for further saturated simulations.

\subsection{Equilibrium Molecular Dynamics Simulations}

Equilibrium simulations involved adding \suldiox~to a water slab and equilibrating as outlined above. The neat-water system had a single \suldiox~added to the center of the water box, representing a concentration of 0.06 M. The more concentrated ``saturated'' system consisted of 22 \suldiox~molecules in the bulk corresponding to a concentration 1.35 M. This saturated system was exposed to an additional 50 \suldiox~in the gas phase above the water surface. After equilibration for 2 ns, both the low and high concentration systems were then evolved for a further 10 ns data collection using a time step of 0.5 fs, with atomic coordinates recorded every 100 fs.

\subsection{Steered Molecular Dynamics Simulations}

A second set of simulations began with an equilibrated water slab as in the surface equilibrated method above. However, in both the neat-water and saturated starting systems, a single \suldiox~was introduced 20\angs~above the water slab surface, with the sulfur atom tethered to its initial position. The systems were then evolved for 1 ns, taking coordinate snapshots every 20 ps to create 50 starting points for further simulations. Steered molecular dynamics (SMD) were then performed on the 50 system configurations (in both the neat-water and saturated configurations) to guide the \suldiox~down towards a tethered water near the water slab's center of mass by applying a small steering force to the \suldiox-sulfur atom. This steering technique has been previously developed and used to successfully model chemical events.\cite{Isralewitz2001,Giorgino2011,Bizzarri2011,Strzelecki2009,Patargias2009,Liu2006} The \suldiox~thus passed through the continuum of environments from gas phase to (neat- and saturated) water surface adsorption, and finally absorption into the bulk of the \wat~slab. Each of the SMD simulations were performed for a total of 200 ps, using a time step of 1 fs, and taking snapshots of the system every 25 fs. Figure \ref{fig:starting-configurations} illustrates two sample starting configurations for the SMD simulations, showing both the neat-water slab and the saturated slab configurations before steering the \suldiox~towards the water bulk.

A separate set of SMD simulations were performed with tethering of one of the \suldiox-oxygens to the water slab center of mass. This was done to ensure that the orientation of the \suldiox~during the adsorption transit was not an artifact of the choice of atom used for tethering. The simulations produced the same results (not shown) for the orientational analyses, so the data from the original tethering scheme was used.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/startingconfigurations.png}
		\caption{Sample starting configurations for the two types of SMD simulations. The neat-water slab simulation introduces a single \suldiox~molecule that is then guided into the surface of the water and further into the bulk (left). The saturated slab simulation begins with a water system that has been loaded with \suldiox~to saturate the water phase, and also with a high pressure of \suldiox~gas (right). The single \suldiox~(shown at the top) is then steered through the surface region saturated with \suldiox~molecules, and into the water bulk.}
		\label{fig:starting-configurations}
	\end{center}
\end{figure}

\subsection {Water Surface Location}

	The first portion of the studies involved creating orientational depth-profiles in the interfacial region comprised of \suldiox~and \wat~upon exposure and adsorption of the gas. Recognizing that a liquid surface is a dynamic boundary that is neither flat nor stationary, we must define a reference point in the interfacial region which we refer to here as the water surface location. Several previous studies have used the technique of fitting a line shape to the averaged density profile of the water, and extracting interfacial shape and location parameters to define the water surface location.\cite{Shamay2010,Wick2006c,Chowdhary2006} Hyperbolic tangent functions have been used often, and values for the ``Gibb's dividing surface'' location, and interfacial width have thus been determined.\cite{Matsumoto1988} However, in long simulations the location and shape of the interface changes, and the motion of surface waters alters the interfacial width at any given time step. Thus, the density profile fitting will capture averaged widths and locations, not instantaneous values. Similarly, the averaged values of location and width will obscure information about any drift or deformations the surface undergoes. The analysis presented here attempts to retain these subtleties through the use of a ``corrected'' coordinate system.

Figure \ref{fig:density-flaw} demonstrates the problem of surface location drift during a simulation. Figure \ref{fig:density-flaw}A shows the density profile of water and \suldiox~over the course of one of the 10 ns trajectories used in this work using the original uncorrected coordinates of the system taken from the raw atomic positions of the molecular dynamics data output. The water density profile and location (thin gray line in Figure \ref{fig:density-flaw}A) is produced by averaging the instantaneous density profile at each time step in the simulation over all the time steps. The water profile was then fit to a $\tanh$ function (black line) to extract the position and width parameters of the water surface. The fitted Gibbs dividing surface has a width of 3.77\angs, which is comparable to values reported for similar neat-water systems.\cite{Dang1997,Hore2008} A bulge in the gas-phase ($\leq 52$\angs) side of the water density profile is indicative of the drift of the water slab over the course of the trajectory. Thus the calculated location and width from the $\tanh$ line fit are not accurate over long trajectories for defining a stationary reference point.

	To overcome this problem we define the water surface location by calculating a reference location at each time step by averaging the positions of the waters contained in the topmost monolayer. This provides a consistent and intuitive reference point in the simulations to which we relate our analyses, but does not increase the computational burden. The number of waters included in the averaging is determined by taking a few issues into account. First, counting the waters found in the topmost cross-section of the water slab over several time steps indicated between 65-75 waters that established a full monolayer. This was done by a visual inspection of the slab using the VMD MD visualization package.\cite{Humphrey1996} Alternatively, assuming a spherical model of water with a radius of 2.2\angs~(the typical length of a bulk-water hydrogen bond), two layers of hexagonally tight-packed spheres yielded a similar number of surface water molecules. Increasing the number of waters used in calculating the surface location diminishes the effects of the few waters that briefly rise above the surface into the gas phase, stabilizing both the surface position and thickness values. Taking the close-packed model as a maximum number of waters fit into a flat surface, the topmost 70 water molecules were used for calculating instantaneous water surface locations for each simulation step.

This method for finding the outer monolayer location was implemented, and the surface location is plotted as a function of simulation time in Figure \ref{fig:density-flaw}B. It is apparent from the surface location plot that the water slab location, and thus the surface location, drifts over the 10 ns, spanning approximately 12 \angs. However, the maximum standard deviation of the positions of the waters comprising the surface layer at each time step is only 1.85 \angs. Consequently, all depth locations in our analyses are calculated relative to the instantaneous surface location at the corresponding time step of the trajectories.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/density-location.png}
		\caption{(A) Density profiles of \wat~(Gray) and \suldiox~(red) from a 10 ns simulation of the neat-water system with a single sulfur dioxide. The \wat~density profile was fit using a hyperbolic tangent function (black) to extract location and width data for the water surface. Position values are based on the atomic position data of the water oxygens and \suldiox~sulfurs from the raw molecular dynamics output. (B) The instantaneous location of the outer \wat~monolayer was calculated for each simulation time step (blue), and the surface location extracted from the density fitting was also shown for reference (horizontal dashed line). Long simulations of liquid slabs result in drift of the slab location, and the consequent broadening of the water density profile if the surface location is not determined repeatedly throughout the simulation.}
		\label{fig:density-flaw}
	\end{center}
\end{figure}

\subsection{Molecular Orientation}

Knowing the molecular orientation of both the \wat~and \suldiox~is a prerequisite for understanding the chemistry occurring during the \suldiox~adsorption process. With the surface location as defined above, the simulated systems were analyzed to characterize the orientation of \wat~and \suldiox~in various environments above, within, and below the aqueous surface region. The two molecules studied are similarly shaped with a C2$_v$ axis along their bisectors, and a molecular plane defined by three atoms. A body-fixed frame is defined for both \wat~and \suldiox~as shown in Figure \ref{fig:molecular-frame}. In each analysis a space-fixed reference axis is used that corresponds to the long axis of the system's periodic cell normal to the plane of the water surface. The orientational analyses presented herein focus on two angles used to define molecular orientation. The molecular orientation angles $\theta$ and $\phi$ are determined from a set reference axis as shown in Figure \ref{fig:water-angles}A.

	The ``tilt'' angle, $\theta$, defines the angle formed between the molecular bisector vector (the molecular z-axis, pointing from the central atom in the direction of the other two atoms) and the positive system reference axis. Thus the value of $\theta$ falls within a range of $[0,\pi]$. An angle of $\cos(\theta)=1$ indicates a molecule with its bisector aligned with the reference axis, while $\cos(\theta)=-1$ results from an anti-aligned configuration. Sample representations of different values of $\cos(\theta)$ are shown in Figure \ref{fig:water-angles}B.

	A second angle, $\phi$, defines the molecular ``twist'' for certain configurations of the molecule. $\phi$ is the angle formed between the vector normal to the plane of the molecule (molecular y-axis) and the system reference axis. The values of $\phi$ fall in the interval $[0,\frac{\pi}{2}]$ because of the symmetry of \wat~and \suldiox~molecules with respect to twist about their bisector axes. For values of $\cos(\theta) \approx 0$, $\phi$ provides additional information about whether the molecular orientation is ``flat'' to the surface (e.g. the plane of the molecule is aligned with the plane of the surface), or if it is perpendicular. The values of $\cos(\phi)$ for different molecular orientations are shown in Figure \ref{fig:water-angles}C. %However, the symmetry of the \wat~and \suldiox~molecules create equivalence between the two H, or O, atoms (in the \wat~or\suldiox, respectively). $\phi=-\pi$ is equivalent to $\phi=\pi$, and both situations are equivalent to $\phi=0$. Because $\phi=-\phi$, the value is reported within the range of $[0,\frac \pi 2]$. 
	However, values of $\cos(\theta)$ close to 1 or -1 result in an isotropic distribution in $\cos(\phi)$ because of the symmetry of the plane of the surface in directions perpendicular to the reference axis.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/molecularframesmall.png}
		\caption{Molecular body-fixed axes are defined with the x-z plane formed by the three atoms, and the z-axis aligned to the molecular bisector. The y-axis is normal to the molecular plane, and one of the bonds points in the positive x direction.}
		\label{fig:molecular-frame}
	\end{center}
\end{figure}

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/molecular-angles.png}
		\caption{(A) The definition of the angles $\theta$ and $\phi$ used to define molecular orientation of \suldiox~and \wat~relative to a reference axis. (B) $\theta$ is the value of the ``tilt'' of the molecular bisector from the reference axis, with $\cos(\theta)$ values ranging from -1 to +1 indicating different bisector orientations. (C) $\phi$ is the ``twist'' angle formed between the vector normal to the molecular plane (the body-fixed y-axis) and the reference axis. For values of $\cos(\theta) \approx 0$, $\phi$ shows how flat the molecule is relative to the interfacial plane.}
		\label{fig:water-angles}
	\end{center}
\end{figure}

\subsection{Surface Density Distributions}

One measure of surface activity is the spatial distribution of molecules in the interfacial region. The density distributions of both \wat~and \suldiox~were calculated for the equilibrium MD simulations. The results presented in Figure \ref{fig:density} show both the water (black) and \suldiox~(red) density distributions across the two simulated slabs. As shown, the single \suldiox~in the neat-water system remained at the water slab surface. The \suldiox~in the saturated system accumulated mostly at the surface, but some residual \suldiox~remained well into the bulk water.

In the simulated neat-water slab, the \suldiox~molecular density distribution concentrates near the water surface location, indicating an affinity for the interfacial region. During the course of the simulation, the \suldiox~does not venture into the bulk water nor does it escape the water surface into the gas phase, but remains located within 5-\angs~of the surface region. This is consistent with what is found experimentally in our laboratory using VSFS,\cite{Tarbuck2005,Tarbuck2006} and also supported by the computational simulations and spectral calculations of Baer et al.\cite{Baer2010} The experimental studies indicated that upon exposure of \suldiox~to a \wat~surface, a layer of solvated \suldiox~forms, modifying the structure of water in the upper surface region. In the experimental system, the \suldiox~bonding interaction with the free OH oscillators is manifested in a red-shift of the free OH frequency indicative of a bonding interaction.

The saturated solution simulation results indicate that under the high concentration of \suldiox, the \suldiox~accumulation at the surface is increased. However, unlike the neat-water slab, the saturated slab has a non-zero bulk concentration of \suldiox. The added concentration of \suldiox~creates a layer of molecules bound to the top of the water surface. The center of the \suldiox~density distribution is further into the gas phase than for the neat-water surface with a single-\suldiox~molecule. Because of the limitations of the classical model in accurately reproducing the first hydration shell around the \suldiox, we do not draw specific conclusions about the hydrated surface complex's geometry. However, the surface affinity of the hydrated \suldiox~is well reproduced by both the classical and ab initio methods.\cite{Baer2010}

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/density.png}
		\caption{Molecular density distributions of \wat~(black) and \suldiox~(red) calculated along the long-axis of the simulated cells. The position axis shows the distance from the instantaneous surface location of the \wat~slab, with positive values located above the slab towards the gas phase, and negative values located in the water bulk. Distributions of both the neat-water simulation with a single \suldiox~(left, with the \suldiox~density scaled 10x for clarity), and the saturated system (right) are shown.}
		\label{fig:density}
	\end{center}
\end{figure}

\subsection{Equilibrated MD}

Geometric analyses were performed to characterize the net molecular orientation of \wat~and \suldiox~molecules at different depths from the water surface location. At each distance from the surface location, an orientation profile was created for both the \wat~and \suldiox~molecules. The orientation distribution for the angles $\theta$ and $\phi$ at all depths were combined to form 2D intensity plots that show how the molecular orientation distributions change with distance to the surface location. These plots (Figures \ref{fig:water-orientation}, \ref{fig:so2-transit-angles}, and \ref{fig:so2-surface-angles}) allow for a visual interpretation of how the net orientations are affected when moving from the gas phase through the interfacial region and to the surface location, and then further into the aqueous interfacial region and bulk. Both the neat-water, with only a single \suldiox~introduced, and the high-concentration saturated system were analyzed. In the case of the neat-water system, the introduction of a single \suldiox~does not greatly affect molecular orientation of water molecules in the interfacial region. These results of the water orientation are very similar to a neat-water system without any adsorbed solutes (not shown).

The 2D orientational depth profile plots show areas of low intensity in dark blue, and highest intensity in dark red. Regions of the plots where the intensity (coloration) is equally distributed from top to bottom along the entire orientational range are considered isotropic. Likewise, areas of the plot with high intensity over a small orientational range are considered to exhibit an orientational preference at the given depth.  The histograms are arranged with the plots of $\cos(\theta)$ on the left and $\cos(\phi)$ on the right. The surface is located at 0\angs. The angle distributions from both simulated slab surfaces were averaged for all the orientation analyses.

\subsubsection{\wat~Orientation}

The orientation depth-profiles for \wat~are shown in Figure \ref{fig:water-orientation} for both the neat-water (top) and saturated (bottom) systems during the equilibrium MD simulations. The interfacial region for both these calculations and the VSF experiments is defined as the region where molecular orientational anisotropy exists around the surface water location. Referring to Figure \ref{fig:water-orientation}, calculations indicate an interfacial width of approximately 10\angs~for the neat-water system, and approximately 16 \angs~for the saturated system. In both systems the strongest orientational preference is found at the slab surfaces (positions above 0\angs) where the water is furthest towards the gas phase. Previous work on orientational preference of water at air surfaces shows the same trend as our neat-water results.\cite{Walker2006b,Hore2008} The narrow "peninsula" of high intensity that extends in the $\cos(\theta)$ plots to the right of the -5\angs~location shows the overall preference of water to orient at the surface. The $\cos(\phi)$ plots are similar to each other with a narrow region of reorientation, but the effect in the interfacial region is greater in the neat-water system as evidenced by the sharper transition in intensity from blue to red, compared to the saturated system that has a less pronounced intensity change.

The bisector tilt of the water molecules, $\cos(\theta)$, concentrates around $\cos(\theta)=0$ within the first few\angs~above and below the water surface location, becoming progressively isotropic further through the interfacial region and into the water bulk of both systems. As the tilt nears $\cos(\theta)=0$ the \wat~bisector lies within the plane of the surface indicating a water orientation either flat on the surface, or with some amount of ``twist'' sending the OH bonds in towards, or out of the bulk. The value of $\phi$ determines the ``twist'' in this case. Both systems show a defined intensity concentration (dark red) in the distributions around $\cos(\phi)=1$ on the aqueous side of the interfacial region. This results from an orientation of the water's y-axis (normal to the molecular plane) aligned perpendicular to the plane of the water surface. Thus, for both the neat-water and saturated systems, the preferred orientation of water molecules at a distance of 0\angs~ is to lie mostly flat to the plane of the interface with a slight ``twist'' sending one OH bond further into the gas phase, and the other OH towards the water bulk. This result agrees somewhat with a recent air/water study by Fan et al. in which the surface orientations of several water models were analyzed.\cite{Fan2009} They simulated water using non-polarizable models, however, which alters the behavior of water at interfacial regions when compared to the polarizable POL3 model used in our work. The main conclusions are similar in that one OH tends to point further into the air phase than the other, but differs in that this effect is less pronounced with the polarizable model.

Although the plots show overall similarities for both the neat-water and saturated systems, the presence of a layer of adsorbed \suldiox~molecules alters the orientation of those waters furthest into the gas phase. For the saturated solution the resulting orientation of waters above 0\angs, shown in the saturated plot of $\cos(\theta)$ of Figure \ref{fig:water-orientation}, is with a bisector pointing further into the adsorbed \suldiox~gas layer, and both hydrogens pointing outward from the aqueous bulk. The effect is more pronounced further from the water phase, and above 5\angs~the $\cos(\theta)$ distribution is completely centered around $\cos(\theta)=+1$ (see Figure \ref{fig:water-angles}).

The noise in the neat-water plots of Figure \ref{fig:water-orientation} above 5\angs~(manifested as disconnected points of intensity throughout the range of orientations) is a result of fewer waters venturing beyond those extents and thus less data far from the water surface location. This is one indication that waters near a layer of adsorbed \suldiox~venture further above the water surface location relative to the low \suldiox~concentration, where they can have more \suldiox~interaction with the adsorbed gas molecules. These results show that the interactions with neighboring \suldiox~molecules allow the waters above the water surface location to orient more perpendicularly to the interface. This is consistent with our recent experimental VSFS studies which showed evidence for the reorienting behavior of water due to the \suldiox~interactions with the topmost surface waters.\cite{Ota2011}

The distribution of $\cos(\phi)$ is more sharply defined (i.e. less isotropic) for the neat-water system than for the saturated one. Waters on the neat surface lie flatter, whereas the presence of the \suldiox~allows a greater range of ``twist'' for those waters in the plane of the interface. The $\phi$ distributions quickly become isotropic above the surface location as the bisectors orient more perpendicularly, and below the surface as the bulk water loses any orientational preference.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/h2oangles.png}
		\caption{Molecular orientation histograms of \wat~throughout the surface equilibrated systems. The top surface is located at a distance of 0\angs~with negative distance values located in the bulk of the slab. Shown are the angle distributions for $\theta$ (left column) and $\phi$ (right column) in both the neat-\wat~system (top row) and the saturated system (bottom row). The distributions are normalized to account for the changing number of water molecules at different positions in the system. Regions of high intensity are dark red, and low intensity are dark blue. The scattered points of coloration to the far right of each plot indicates that few waters were located at those positions, and thus few data points were collected.}
		\label{fig:water-orientation}
	\end{center}
\end{figure}

\subsubsection{\suldiox~Orientation}

Orientation distributions of the adsorbed \suldiox~molecules were created during the equilibrium simulations for both the neat-water and saturated systems. Figure \ref{fig:so2-orientation} shows the distributions of $\cos(\theta)$ and $\cos(\phi)$ (arranged similarly to the water orientation distributions plots in Figure \ref{fig:water-orientation}). The \suldiox~orientation data set for the neat-water system is much smaller as only a single \suldiox~molecule was simulated in the bulk. The resulting distribution plots are thus noisier than the corresponding saturated plots with more scattered points of high intensity, but effective comparisons can still be drawn.

In the interfacial region within 0-5\angs~of the water surface location (the approximate distance from the surface that \suldiox~begins interacting with the outer-most waters) the angular distribution of the single \suldiox~(in the neat-water system) is concentrated primarily in $\cos(\theta)>0$. The peak of the distribution occurs at $\cos(\theta)=1$. This indicates that the \suldiox~bisector points out of the water surface, with the sulfur atom pointing towards the aqueous bulk, and the two oxygens pointing into the gas phase. This same distribution occurs in the saturated system for positions below 5\angs. Beyond 5\angs~above the surface both distributions become isotropic. Promixity to the water highly orients the \suldiox~with the sulfur atom pointing in towards the water bulk. Moving further away from the water surface, and interacting less with \wat~molecules allows for greater orientational freedom as exhibited in the isotropy of the distributions above 5\angs.

The plots of $\phi$ are both isotropic, although the neat-water system plot is quite noisy from the small data set. Because the \suldiox~near the surface is oriented perpendicularly to the interface, the $\cos(\phi)$ orientation is expected to be isotropic. Further from the water surface where the bisector orientation becomes isotropic, the $\cos(\phi)$ distribution also becomes isotropic. For the \suldiox~orientation, the $\phi$ angle does not provide further information regarding the surface behavior.

The plots indicate that \suldiox~orients similarly at both the low and high concentration interfacial environments. The plots of $\phi$ for both concentrations exhibit isotropic distributions. However, the $\theta$ values follow similar trends at both concentrations indicating that the \suldiox~sulfur orients down towards the water phase, with both oxygens pointing away from the surface once the \suldiox~is within approximately 5\angs~of the surface. This behavior continues down to at least 5\angs~below the water surface location, notwithstanding any chemical reactions that may occur that are not simulated using the classical MD techniques utilized in this work.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/so2-angles.png}
		\caption{Molecular orientation distributions for \suldiox~molecules adsorbed to the water slab surface. Distributions are shown for $\cos(\theta)$ (left column) and $\cos(\phi)$ (right column) for both the neat-water (top) and saturated (bottom) systems. For both systems the $\theta$ distributions show \suldiox~bound to the water surface with the sulfur pointing towards the water slab, and the oxygens pointing to the gas phase. In this configuration the $\phi$ distribution is isotropic because of the water slab's in-plane symmetry.}
		\label{fig:so2-orientation}
	\end{center}
\end{figure}

\subsection{Steered MD Transit Simulations}

\subsubsection {\suldiox~Orientation}

	The orientation of \suldiox~molecules throughout the aqueous adsorption process was monitored during the transit SMD simulations. The angles $\theta$ and $\phi$ of the transiting \suldiox~(Figure \ref{fig:water-angles}) were calculated for each timestep of the SMD simulations as the \suldiox~was pulled into the water slab from the gas phase, both in the neat-water and saturated slab systems. The angle cosine values were collected for the 50 simulations of both systems for each distance from the water surface location, resulting in the 2-dimensional angle and depth-profile histograms shown in Figure \ref{fig:so2-transit-angles}. In all the data sets, the populations of the angle histograms at each distance from the surface were normalized to aid in comparison of regions with differing \suldiox~residence times.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/so2-angles-transit.png}
		\caption{\suldiox~molecular orientation distributions during SMD transit simulations into an aqueous slab. Both the neat-water (top row) and saturated (bottom row) data sets were analyzed to determine the angles $\theta$ (left column), and $\phi$ (right column) of the transiting \suldiox. The distributions show angle cosines plotted against the distance of the \suldiox~to the location of the water surface. The data sets were averaged at each distance for the respective 50 simulations, and every distribution at each distance was normalized such that each histogram has a max population of 1 for purposes of comparison.}
		\label{fig:so2-transit-angles}
	\end{center}
\end{figure}

	From its starting position above the water surface, until the \suldiox~moves to within 10\angs~of the water surface location of both systems, the orientation is isotropic in $\theta$ and $\phi$ (not shown). Isotropic orientation is manifested in the plots as mostly uniform coloration at a given distance from the surface independent of $\cos$ of $\phi$. Near and into the interfacial region the bisector angle $\theta$ becomes more perpendicular to the water surface ($\cos(\theta)\approx1$) with the \suldiox~sulfur pointing into the water phase, consistent with the equilibrium MD simulation results above. At the point when the \suldiox~reaches the water surface location (distance $=0$\angs), the bisector is perpendicular to the interface in both the neat-water and saturated systems. Figure \ref{fig:so2-surface-angles}a depicts \suldiox~near the water surface with the bisector aligned perpendicular to the interface. In the absence of simulated ionic species that form through \suldiox-\wat~chemistry at the surface, it is clear that the adsorbing \suldiox~in the gas phase takes on a preferred orientation to adsorb on a water surface. The main difference between the neat and saturated water systems is where the point of the transition from isotropic to preferred orientations is found.

  Comparing in more detail the difference in these two systems upon \suldiox~approach, for the neat-water surface, a transition occurs at approximately 5\angs~from the surface water location. Below 5\angs~above the surface, \suldiox~has a preferred net orientation and is close enough to the water surface that it begins to interact with the topmost surface waters. In the saturated system, the same trend occurs, however the onset of the perpendicular orientation begins at approximately 8\angs~above the surface. The layer of adsorbed \suldiox~already present in the saturated system most likely interacts with the transiting \suldiox~molecule. Also, topmost water molecules from the surface move up to a few \angs~inside the adsorbed \suldiox~layer and interact with the transiting \suldiox~further from the surface than those in the neat-water system. It is remarkable that the orientational trend appears so strongly in the $\theta$ plots even with so little data as was collected from the single \suldiox~molecule of each simulation. From onset of orientation above the surface until 10\angs~below, the \suldiox~holds a preferred orientation.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/so2-surface.png}
		\caption{(A) \suldiox~molecules above and below both neat-water and saturated water surface location orient with the bisectors perpendicular to the interface. The \suldiox~sulfur points down towards the aqueous phase. (B) In the neat-water system, a transiting \suldiox~molecule located below the water surface briefly reorients with one S-O bond more towards the interface above, and the other bond pointing further down into the water bulk.}
		\label{fig:so2-surface-angles}
	\end{center}
\end{figure}

 With a mostly perpendicular bisector angle, it is expected that the values of $\phi$ for \suldiox~would be isotropic relative to the reference axis. This is the case in both systems, with only a few exceptions. In both systems the $\cos(\phi)$ profiles exhibit mostly isotropic distributions above 0\angs, with several regions of lighter coloration interspersed, but without a clearly formed orientational trend. At the neat-water surface and just below, from 0\angs~to -5\angs~(in the water phase), the $\theta$ profile broadens down to $\cos(\theta)=0.5$, coinciding with a clear trend in the profile of $\phi$ near $\cos(\phi)=0$ shown in dark red at the bottom of the plot. This corresponds to \suldiox~inclined up to 60 degrees from the surface normal. The result is that \suldiox~in this region shows a preference for one of the S-O bonds to point more towards the aqueous bulk and the other to point away towards the interface. Figure \ref{fig:so2-surface-angles}B depicts the orientation of the \suldiox~slightly below the water surface location. This is in contrast to \suldiox~molecules above the water surface location, oriented more perpendicularly. The behavior is likely due to the interaction between the waters and one of the S-O bonds leading to a higher solvation than above the water surface location. As the \suldiox~is solvated by more highly-coordinated bulk water, the S-O bonds experience less equal interaction environments. Baer et al. noted that their force field model for the \suldiox~does not reproduce well the first hydration shell geometries,\cite{Baer2010} so conclusions regarding the specific interactions and hydrate geometries between the \suldiox~and \wat~cannot be made here. It is notable that the same reorientation does not occur as strongly in the saturated system. The presence of the adsorbed \suldiox~layer apparently decreases the reorienting behavior likely because of the disrupting effect the higher \suldiox~concentration has on the water interactions in the interfacial region.

\section {Conclusions}

Gaseous adsorption on solid surfaces has been extensively studied over the past few decades with much learned about how molecular geometry and orientation of the adsorbate are influenced by the proximity of the solid slab.  For a liquid surface where the surface slab is no longer rigid but has molecules with considerable freedom of movement, the surface and approaching gas molecules can be active partners in attaining the optimal geometry and orientation necessary for adsorption and subsequent uptake.    And unlike the solid surface defined by a sharp plane, the interfacial region for the liquid-gas system is much broader, extends on either side of a defined center plane, and is host to a broad distribution of gas-liquid molecular geometries and orientations that change as the gas molecules transit through the interfacial region into the bulk liquid.  Although  our current molecular level understanding of  the complex dances that these molecules play in this fluid interfacial region is in its infancy, emerging studies such as these are beginning to provide unique new insights that are key to understanding many  environmentally important processes at aqueous surfaces.

Presented herein are the results of several classical molecular dynamic simulations that focus on understanding how surface water molecules and adsorbing \suldiox~gas molecules twist and turn as the gas adsorbs and transits the interfacial region. The computational studies emulate and expand on the experimental spectroscopic studies from this laboratory which have found \suldiox~surface complexation at a water surface.\cite{Tarbuck2005,Tarbuck2006,Ota2011}  These spectroscopic studies show clear evidence of \suldiox-water surface complexation but details about this surface complex could only be inferred from spectral changes in the surface water spectrum since \suldiox~could not be monitored directly.   These simulations do not have that limitation and hence can provide information about the behavior of both surface partners and in particular, how their proximity influences the orientation behavior of each other.  The orientational information obtained in these simulations are provided via calculated depth profiles which show the molecular distribution of orientations of the two different interfacial molecules throughout the dimensions of the interfacial region.

Our simulations show that gaseous \suldiox~quickly adsorbs to the water surface and continues to bind until a complete surface coverage is reached. Surface waters reorient in the presence of adsorbed \suldiox. The waters at the interface of a neat-water surface tend to lay flatter to the surface than when a saturating layer of \suldiox~is present. The waters interacting with the layer of adsorbed \suldiox~orient more perpendicularly to the interface, and further expose their ``free-OH'' uncoupled bonds for interactions with \suldiox, and hydrate complex formation. Furthermore, we have found that surface waters underneath a blanket layer of adsorbed \suldiox~will penetrate further into the gas phase, allowing for greater mobility of waters away from the aqueous bulk in the presence of \suldiox.

Through these simulations we also characterize the orientational behavior of \suldiox~during and after adsorption. Our equilibrium neat-water simulations show that a single \suldiox~molecule, representing a low concentration, has a high surface affinity. At a high \suldiox~concentration in the saturated systems, \suldiox~molecules are also surface active, and are found further out of the water phase than at the lower concentration. These \suldiox~molecules form a bound layer that crowds the surface and interacts with the surface waters. The orientation of \suldiox~on the water surface was found to be similar for both low and high concentrations. Those \suldiox~molecules at or below the surface water location strongly orient with the sulfur atom pointed in towards the water bulk, and the oxygen atoms out towards the gas phase. The \suldiox~molecules slightly above the water surface lose this net orientation within 10 \angs. Those  molecules further from the water are more isotropically oriented. Figure \ref{fig:so2-surface-cartoon} depicts what the neat-water and saturated surface molecules look like for both \suldiox~and \wat~orientations and locations based on the calculations.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/system-surface.png}
		\caption{\wat~and \suldiox~both exhibit preferred orientations in the region near the liquid-gas interface. The neat-water system (A) with a single \suldiox~molecule (low-concentration) has surface waters orienting mostly flat to the interface. When the \suldiox~concentration is increased, as in the saturated system (B), the waters at the surface behave similar to the neat-water interface, but waters that venture into the adsorbed \suldiox~gas layer orient strongly with their bisectors pointing out from the aqueous phase towards the gas. In both cases, the \suldiox~orients with its molecular bisector pointing out to the gas phase when it is near the surface, and isotropically further into the gas phase. Sulfur, oxygen and hydrogen atoms are colored yellow, red, and white, respectively.}
		\label{fig:so2-surface-cartoon}
	\end{center}
\end{figure}

Steered molecular dynamics simulations were used to model the behavior of an adsorbing \suldiox~as it moves from the gas phase above the water down through the surface and into the bulk. The \suldiox~reorients as it makes its first contact with the water interface. Within 5 \angs~of the surface the \suldiox~is mostly oriented with its sulfur towards the water phase. The results for the transit through the interface show that in both systems of low and high \suldiox~concentration an adsorbing \suldiox~near the interface has very similar orientation to those molecules already bound to the water surface. The \suldiox~pulled further into the water bulk retains its orientation until it is past the interfacial region and then isotropically orients with the bulk water.

These studies provide a starting point for future studies in this area that seek to understand how gases of different concentrations and chemical composition adsorb and transit across an aqueous/air interface. Obtaining such knowledge will be invaluable for understanding many environmental aerosol and land water systems where gaseous uptake at a water surface does not conform to expectations.\cite{Jayne1990,Yang2002,Worsnop1989,Boniface2000} Investigations of these and related low-temperature effects will be forthcoming in future publications.

\bibliography{citations}

\end{document}

### 2011-07-07 submission/Article.tex
added white background to fill in transparency on images, and finalized the submission zip file

@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/startingconfigurations.png}
+		\includegraphics[scale=1.0]{startingconfigurations.png}
 		\caption{Sample starting configurations for the two types of SMD simulations.
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/density-location.png}
+		\includegraphics[scale=1.0]{density-location.png}
 		\caption{(A) Density profiles of \wat~(Gray) and \suldiox~(red) from a 10 ns simulation of the neat-water system with a single sulfur dioxide.
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/molecularframesmall.png}
+		\includegraphics[scale=1.0]{molecularframesmall.png}
 		\caption{Molecular body-fixed axes are defined with the x-z plane formed by the three atoms, and the z-axis aligned to the molecular bisector.
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/molecular-angles.png}
+		\includegraphics[scale=1.0]{molecular-angles.png}
 		\caption{(A) The definition of the angles $\theta$ and $\phi$ used to define molecular orientation of \suldiox~and \wat~relative to a reference axis.
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/density.png}
+		\includegraphics[scale=1.0]{density.png}
 		\caption{Molecular density distributions of \wat~(black) and \suldiox~(red) calculated along the long-axis of the simulated cells.
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/h2oangles.png}
+		\includegraphics[scale=1.0]{h2oangles.png}
 		\caption{Molecular orientation histograms of \wat~throughout the surface equilibrated systems.
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/so2-angles.png}
+		\includegraphics[scale=1.0]{so2-angles.png}
 		\caption{Molecular orientation distributions for \suldiox~molecules adsorbed to the water slab surface.
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/so2-angles-transit.png}
+		\includegraphics[scale=1.0]{so2-angles-transit.png}
 		\caption{\suldiox~molecular orientation distributions during SMD transit simulations into an aqueous slab.
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/so2-surface.png}
+		\includegraphics[scale=1.0]{so2-surface.png}
 		\caption{(A) \suldiox~molecules above and below both neat-water and saturated water surface location orient with the bisectors perpendicular to the interface.
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/system-surface.png}
+		\includegraphics[scale=1.0]{system-surface.png}
 		\caption{\wat~and \suldiox~both exhibit preferred orientations in the region near the liquid-gas interface.
@@
 
+\section{Acknowledgements}
+
+The authors wish to thank Dr.
+Daniel Steck of the University of Oregon Department of Physics for the generous use of computing resources, the National Science Foundation (Grant CHE-1051215), and the US Department of Education Graduate Assistance in Areas of National Need (GAANN) program (Grant P200A070436) for support of this research.
+
 \bibliography{citations}

### 2011-10-12 EquilibratedOrientation.tex
Final images for the water and so2 equilibrated orientations

@@
 At each distance from the surface location, an orientation profile was created for both the \wat~and \suldiox~molecules.
-The orientation distribution for the angles $\theta$ and $\phi$ at all depths were combined to form 2D intensity plots that show how the molecular orientation distributions change with distance to the surface location.
+The bivariate orientation distributions for the angles $\theta$ and $\phi$ at various depths were combined to form the intensity plots that show how the molecular orientation distributions change with distance to the surface location.
 These plots (Figures \ref{fig:water-orientation}, \ref{fig:so2-transit-angles}, and \ref{fig:so2-surface-angles}) allow for a visual interpretation of how the net orientations are affected when moving from the gas phase through the interfacial region and to the surface location, and then further into the aqueous interfacial region and bulk.
@@
 
-The 2D orientational depth profile plots show areas of low intensity in dark blue, and highest intensity in dark red.
-Regions of the plots where the intensity (coloration) is equally distributed from top to bottom along the entire orientational range are considered isotropic.
+The depth profile plots are arranged as a grid of 2D histograms.
+Each histogram is calculated for all molecules falling within a particular depth in the water interfacial region.
+The depth of each plot is marked (in \ang) in the upper-right, with the water surface location set at 0\angs, positive depths lie on the gas-phase side of the surface, and negative depths are on the bulk water side.
+The horizontal axes of each hisogram represent $\theta$ values, and the vertical axes represent $\phi$ values.
+Areas of low intensity appear in dark blue, and highest intensity in dark red.
+Regions of the plots where the intensity (coloration) is equally distributed along either the vertical or horizontal axes are considered isotropic in $\phi$ or $\theta$, respectively.
 Likewise, areas of the plot with high intensity over a small orientational range are considered to exhibit an orientational preference at the given depth.
-The histograms are arranged with the plots of $\cos(\theta)$ on the left and $\cos(\phi)$ on the right.
-The surface is located at 0\angs.
 The angle distributions from both simulated slab surfaces were averaged for all the orientation analyses.
@@
 The interfacial region for both these calculations and the VSF experiments is defined as the region where molecular orientational anisotropy exists around the surface water location.
-Referring to Figure \ref{fig:water-orientation}, calculations indicate an interfacial width of approximately 10\angs~for the neat-water system, and approximately 16 \angs~for the saturated system.
-In both systems the strongest orientational preference is found at the slab surfaces (positions above 0\angs) where the water is furthest towards the gas phase.
-Previous work on orientational preference of water at air surfaces shows the same trend as our neat-water results.\cite{Walker2006b,Hore2008} The narrow "peninsula" of high intensity that extends in the $\cos(\theta)$ plots to the right of the -5\angs~location shows the overall preference of water to orient at the surface.
-The $\cos(\phi)$ plots are similar to each other with a narrow region of reorientation, but the effect in the interfacial region is greater in the neat-water system as evidenced by the sharper transition in intensity from blue to red, compared to the saturated system that has a less pronounced intensity change.
+Fitting the water density profiles we have calculated an interfacial width of approximately 10\angs~for the neat-water system, and approximately 16\angs~for the saturated system.
+In both systems the strongest orientational preference is found at the slab surfaces (positions near 0\angs).
+Previous work on orientational preference of water at air surfaces shows the same trend as our neat-water results.\cite{Walker2006b,Hore2008} The swaths of coloration indicating high intensity appearing in the neat-water plots from -4 to +4\angs, (and the corresponding regions in most of the saturated system plots) show the overall preference of water to orient at the surface.
+The plots of the neat-water and saturated systems are similar to each other with a narrow region of reorientation, but the effect in the interfacial region is greater in the neat-water system as evidenced by the sharper transition in intensity from blue to red, compared to the saturated system that has a less pronounced intensity change over larger areas of the histograms.
 
-The bisector tilt of the water molecules, $\cos(\theta)$, concentrates around $\cos(\theta)=0$ within the first few\angs~above and below the water surface location, becoming progressively isotropic further through the interfacial region and into the water bulk of both systems.
-As the tilt nears $\cos(\theta)=0$ the \wat~bisector lies within the plane of the surface indicating a water orientation either flat on the surface, or with some amount of ``twist'' sending the OH bonds in towards, or out of the bulk.
+The bisector tilt of the water molecules, $\theta$, concentrates around $\theta=$ 90\textdegrees within the first few\angs~above and below the water surface location, becoming progressively isotropic further through the interfacial region and into the water bulk of both systems.
+Above the surface location at positive distances, $\theta \leq$ 90\textdegrees~indicates that the water hydrogens tend to point towards the gas-phase side of the surface.
+As the tilt nears $\theta=$ 90\textdegrees~the \wat~bisector lies within the plane of the surface indicating a water orientation either flat on the surface, or with some amount of ``twist'' sending the OH bonds in towards, or out of the bulk.
 The value of $\phi$ determines the ``twist'' in this case.
-Both systems show a defined intensity concentration (dark red) in the distributions around $\cos(\phi)=1$ on the aqueous side of the interfacial region.
-This results from an orientation of the water's y-axis (normal to the molecular plane) aligned perpendicular to the plane of the water surface.
-Thus, for both the neat-water and saturated systems, the preferred orientation of water molecules at a distance of 0\angs~ is to lie mostly flat to the plane of the interface with a slight ``twist'' sending one OH bond further into the gas phase, and the other OH towards the water bulk.
-This result agrees somewhat with a recent air/water study by Fan et al.
+Both systems show a similar trend where waters at or just below the surface have values of $\phi$ near 90\textdegrees, and waters above the surface take on values of $\phi$ near 0\textdegrees.
+This jump in the angular distribution of $\phi$ indicates that waters at or below the surface lie mostly flat in the plane of the interface, and as they move above the surface towards the gas phase, they reorient with one OH bond pointing towards the water bulk, and one pointing out of the surface into the gas.
+This behavior is more strongly pronounced in the neat-water system where most of the surface waters are not interacting with an adsorbed layer of \suldiox~molecules.
+
+%This result agrees somewhat with a recent air/water study by Fan et al.
 in which the surface orientations of several water models were analyzed.\cite{Fan2009} They simulated water using non-polarizable models, however, which alters the behavior of water at interfacial regions when compared to the polarizable POL3 model used in our work.
@@
 Although the plots show overall similarities for both the neat-water and saturated systems, the presence of a layer of adsorbed \suldiox~molecules alters the orientation of those waters furthest into the gas phase.
-For the saturated solution the resulting orientation of waters above 0\angs, shown in the saturated plot of $\cos(\theta)$ of Figure \ref{fig:water-orientation}, is with a bisector pointing further into the adsorbed \suldiox~gas layer, and both hydrogens pointing outward from the aqueous bulk.
-The effect is more pronounced further from the water phase, and above 5\angs~the $\cos(\theta)$ distribution is completely centered around $\cos(\theta)=+1$ (see Figure \ref{fig:water-angles}).
+For the saturated solution, the resulting orientation of waters above 0\angs, shown in the bottom plot of Figure \ref{fig:water-orientation}, is nearly isotropic in $\phi$, and with $\theta leq$ 90\textdegrees.
+This results from waters with bisectors pointing further into the adsorbed \suldiox~gas layer, and both hydrogens pointing outward from the aqueous bulk.
+The effect is more pronounced as the waters move further from the water surface, and above 4\angs~the $\theta$ distribution is mostly centered around $\theta)=$ 0\textdegrees (see Figure \ref{fig:water-angles}).
 
@@
 
-The noise in the neat-water plots of Figure \ref{fig:water-orientation} above 5\angs~(manifested as disconnected points of intensity throughout the range of orientations) is a result of fewer waters venturing beyond those extents and thus less data far from the water surface location.
-This is one indication that waters near a layer of adsorbed \suldiox~venture further above the water surface location relative to the low \suldiox~concentration, where they can have more \suldiox~interaction with the adsorbed gas molecules.
-These results show that the interactions with neighboring \suldiox~molecules allow the waters above the water surface location to orient more perpendicularly to the interface.
+The angle distributions above 6\angs~in the neat-water plots of Figure \ref{fig:water-orientation} are mostly isotropic (manifested as uniform coloration throughout the range of orientations).
+Furthermore, there are few data points that make up the histograms, a result of fewer waters venturing beyond those extents.
+Conversely, waters near a layer of adsorbed \suldiox~venture further above the water surface location relative to the low \suldiox~concentration, where they can have interactions with the adsorbed \suldiox~gas molecules.
+The waters above the water surface location orient perpendicularly to the interface.
 This is consistent with our recent experimental VSFS studies which showed evidence for the reorienting behavior of water due to the \suldiox~interactions with the topmost surface waters.\cite{Ota2011}
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/h2o-angles/h2oangles.png}
+		\includegraphics[scale=1.0]{images/h2o-angles/theta-phi.png}
 		\caption{Molecular orientation histograms of \wat~throughout the surface equilibrated systems.
-The top surface is located at a distance of 0\angs~with negative distance values located in the bulk of the slab.
-Shown are the angle distributions for $\theta$ (left column) and $\phi$ (right column) in both the neat-\wat~system (top row) and the saturated system (bottom row).
-The distributions are normalized to account for the changing number of water molecules at different positions in the system.
-Regions of high intensity are dark red, and low intensity are dark blue.
-The scattered points of coloration to the far right of each plot indicates that few waters were located at those positions, and thus few data points were collected.}
+The depth (in \angs) of each histogram within the water interface is numbered in the upper-right of each respective plot.
+The water surface location is set at a position of 0\angs~with negative distance values located in the bulk of the slab, and positive distances towards the gas phase.
+Each histogram is a bivariate angle distributions for $\theta$ (horizontal axes) and $\phi$ (vertical axes) in both the neat-\wat~system (top) and the saturated system (bottom).
+Regions of high intensity are dark red, and low intensity are dark blue.}
 		\label{fig:water-orientation}
@@
 Orientation distributions of the adsorbed \suldiox~molecules were created during the equilibrium simulations for both the neat-water and saturated systems.
-Figure \ref{fig:so2-orientation} shows the distributions of $\cos(\theta)$ and $\cos(\phi)$ (arranged similarly to the water orientation distributions plots in Figure \ref{fig:water-orientation}).
+Figure \ref{fig:so2-orientation} shows the 2D distributions of $\theta$ and $\phi$ (arranged similarly to the water orientation distributions plots in Figure \ref{fig:water-orientation}).
 The \suldiox~orientation data set for the neat-water system is much smaller as only a single \suldiox~molecule was simulated in the bulk.
-The resulting distribution plots are thus noisier than the corresponding saturated plots with more scattered points of high intensity, but effective comparisons can still be drawn.
+The resulting distribution plots are thus representative of the single surface active \suldiox~molecule.
+The neat-water \suldiox~molecule remains within a narrower region of the interface than the saturated system \suldiox, but effective comparisons can still be drawn.
 
-In the interfacial region within 0-5\angs~of the water surface location (the approximate distance from the surface that \suldiox~begins interacting with the outer-most waters) the angular distribution of the single \suldiox~(in the neat-water system) is concentrated primarily in $\cos(\theta)>0$.
-The peak of the distribution occurs at $\cos(\theta)=1$.
+In the interfacial region the angular distribution of the single \suldiox~(in the neat-water system) is concentrated primarily in $\theta \lt$ 90 \textdegrees.
+The peak of the distribution occurs at $\theta)=$ 0\textdegrees.
 This indicates that the \suldiox~bisector points out of the water surface, with the sulfur atom pointing towards the aqueous bulk, and the two oxygens pointing into the gas phase.
-This same distribution occurs in the saturated system for positions below 5\angs.
-Beyond 5\angs~above the surface both distributions become isotropic.
-Promixity to the water highly orients the \suldiox~with the sulfur atom pointing in towards the water bulk.
-Moving further away from the water surface, and interacting less with \wat~molecules allows for greater orientational freedom as exhibited in the isotropy of the distributions above 5\angs.
+This same distribution occurs in the saturated system for depths below the surface location, $\lt 0$\angs.
+Beyond 4\angs~above the surface, both distributions become mostly isotropic, either because the \suldiox~does not venture into the gas in the neat-water system, or because of the nature of the adsorbed \suldiox~layer in the saturated system.
+Promixity to the water surface highly orients the \suldiox~bisector.
+% with the sulfur atom pointing in towards the water bulk.
+%Moving further away from the water surface, and interacting less with \wat~molecules allows for greater orientational freedom as exhibited in the isotropy of the distributions above 4\angs.
 %Because both the neat-water and saturated systems show rather similar distributions, it is possible that the concentration of \suldiox~does not strongly affect the orientation of the \suldiox, unlike the orientations of waters.
 
-The plots of $\phi$ are both isotropic, although the neat-water system plot is quite noisy from the small data set.
-Because the \suldiox~near the surface is oriented perpendicularly to the interface, the $\cos(\phi)$ orientation is expected to be isotropic.
-Further from the water surface where the bisector orientation becomes isotropic, the $\cos(\phi)$ distribution also becomes isotropic.
-For the \suldiox~orientation, the $\phi$ angle does not provide further information regarding the surface behavior.
+The distributions of $\phi$ are isotropic in both systems at all depths.
+Because the \suldiox~bisector near the surface is oriented perpendicularly to the interface, the isotropy in $\phi$ is expected.
+Further from the water surface where the bisector orientation becomes isotropic, the $\cos(\phi)$ distribution remains isotropic.
+For the \suldiox~orientation, the $\phi$ angle does not provide further information regarding the surface behavior or orientational preference.
 
-The plots indicate that \suldiox~orients similarly at both the low and high concentration interfacial environments.
-The plots of $\phi$ for both concentrations exhibit isotropic distributions.
-However, the $\theta$ values follow similar trends at both concentrations indicating that the \suldiox~sulfur orients down towards the water phase, with both oxygens pointing away from the surface once the \suldiox~is within approximately 5\angs~of the surface.
+%The plots indicate that \suldiox~orients similarly at both the low and high concentration interfacial environments.
+The $\theta$ values follow similar trends at both concentrations indicating that the \suldiox~sulfur orients down towards the water phase, with both oxygens pointing away from the surface once the \suldiox~is within approximately 5\angs~of the surface.
 This behavior continues down to at least 5\angs~below the water surface location, notwithstanding any chemical reactions that may occur that are not simulated using the classical MD techniques utilized in this work.
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/so2-angles/so2-angles.png}
+		\includegraphics[scale=1.0]{images/so2-angles/theta-phi.png}
 		\caption{Molecular orientation distributions for \suldiox~molecules adsorbed to the water slab surface.
-Distributions are shown for $\cos(\theta)$ (left column) and $\cos(\phi)$ (right column) for both the neat-water (top) and saturated (bottom) systems.
-For both systems the $\theta$ distributions show \suldiox~bound to the water surface with the sulfur pointing towards the water slab, and the oxygens pointing to the gas phase.
-In this configuration the $\phi$ distribution is isotropic because of the water slab's in-plane symmetry.}
+The angular distributions are shown for the neat-water (top) and saturated (bottom) systems, and are arranged similarly to the water orientation plots of Figure \ref{fig:water-orientation}.
+For both systems the distributions show \suldiox~bound to the water surface with the sulfur primarily pointing towards the water slab, and the oxygens pointing to the gas phase.
+In this configuration the $\phi$ distribution is isotropic because of the water slab's in-plane symmetry.
+Note that the range of surface depths in the low concentration system is half that of the saturated system.
+This is due to the narrower position range of the single \suldiox~molecule.}
 		\label{fig:so2-orientation}

### 2011-10-12 Methods.tex
Final images for the water and so2 equilibrated orientations

@@
 	The ``tilt'' angle, $\theta$, defines the angle formed between the molecular bisector vector (the molecular z-axis, pointing from the central atom in the direction of the other two atoms) and the positive system reference axis.
-Thus the value of $\theta$ falls within a range of $[0,\pi]$.
-An angle of $\cos(\theta)=1$ indicates a molecule with its bisector aligned with the reference axis, while $\cos(\theta)=-1$ results from an anti-aligned configuration.
-Sample representations of different values of $\cos(\theta)$ are shown in Figure \ref{fig:water-angles}B.
+Thus the value of $\theta$ falls within a range of [0\textdegree,180\textdegree].
+An angle of $\theta=$ 0\textdegree~indicates a molecule with its bisector aligned with the reference axis, while $\theta=$ 180\textdegree~results from an anti-aligned configuration.
+Sample representations of molecular orientations resulting from different values of $\theta$ are shown in Figure \ref{fig:water-angles}B.
 
-	A second angle, $\phi$, defines the molecular ``twist'' for certain configurations of the molecule.
-$\phi$ is the angle formed between the vector normal to the plane of the molecule (molecular y-axis) and the system reference axis.
-The values of $\phi$ fall in the interval $[0,\frac{\pi}{2}]$ because of the symmetry of \wat~and \suldiox~molecules with respect to twist about their bisector axes.
-For values of $\cos(\theta) \approx 0$, $\phi$ provides additional information about whether the molecular orientation is ``flat'' to the surface (e.g.
+	A second angle, $\phi$, defines the molecular ``twist'' of the molecule.
+$\phi$ is the angle of rotation around the molecular bisector axis that quantifies the rotation of the molecular plane with respect to a plane perpendicular to the water surface.
+The values of $\phi$ fall in the interval [0\textdegree,90\textdegree] because of the symmetry of \wat~and \suldiox~molecules with respect to twist about their bisector axes.
+For values of $\theta \approx$ 90\textdegree, $\phi$ provides additional information about whether the molecular orientation is ``flat'' to the surface (e.g.
 the plane of the molecule is aligned with the plane of the surface), or if it is perpendicular.
-The values of $\cos(\phi)$ for different molecular orientations are shown in Figure \ref{fig:water-angles}C.
+The values of $\phi$ for different molecular orientations are depicted in Figure \ref{fig:water-angles}C.
 %However, the symmetry of the \wat~and \suldiox~molecules create equivalence between the two H, or O, atoms (in the \wat~or\suldiox, respectively).
@@
 Because $\phi=-\phi$, the value is reported within the range of $[0,\frac \pi 2]$.
-	However, values of $\cos(\theta)$ close to 1 or -1 result in an isotropic distribution in $\cos(\phi)$ because of the symmetry of the plane of the surface in directions perpendicular to the reference axis.
+	Values of $\theta$ close to 0\textdegree~or 180\textdegree~result in an isotropic distribution in $\phi$ because of the symmetry of the plane of the surface in directions perpendicular to the surface normal reference axis.
 
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/angle-cartoons/molecular-angles.png}
+		\includegraphics[scale=1.0]{images/angle-cartoons/molecular-angles-small.png}
 		\caption{(A) The definition of the angles $\theta$ and $\phi$ used to define molecular orientation of \suldiox~and \wat~relative to a reference axis.
-(B) $\theta$ is the value of the ``tilt'' of the molecular bisector from the reference axis, with $\cos(\theta)$ values ranging from -1 to +1 indicating different bisector orientations.
-(C) $\phi$ is the ``twist'' angle formed between the vector normal to the molecular plane (the body-fixed y-axis) and the reference axis.
-For values of $\cos(\theta) \approx 0$, $\phi$ shows how flat the molecule is relative to the interfacial plane.}
+(B) $\theta$ is the value of the ``tilt'' of the molecular bisector from the reference axis, with $\theta$ values ranging from  0\textdegree~(aligned) to 180\textdegree~(anti-aligned).
+$\theta = $ 90\textdegree~places the bisector in the plane of the water surface.
+Several bisector orientations are shown depicting different $\theta$ values.
+(C) $\phi$ is the ``twist'' angle defined as a rotation around the molecular bisector axis.
+A value of 0\textdegree~aligns the plane of the molecule perpendicular to the plane of the water surface.
+A value of 90\textdegree~rotates the molecule to lie flat in the plane of the water surface.
+Because of the symmetries of \suldiox, \wat, and the plane of the water surface, the range of $\phi$ values is limited to 0\textdegree~$\leq \phi \leq$~90\textdegree.}
 		\label{fig:water-angles}

### 2011-10-18 Abstract.tex
working on review responses

One might expect the high surface tension of water to be a barrier to adsorption of a gas onto the liquid phase, but we know that gaseous adsorption on, and absorption into a water surface is a common phenomenon on this planet. What is not commonly known is how an atmospheric gas such as SO2 and molecules at the water surface can overcome the barrier created by strong water-water surface bonding interactions. What this interplay looks like, the distances from the water surface at which these attractive interactions begin, and how they influence the orientational nature of both SO2 and surface water molecules is the focus of this computational study. The results fill a void in the information about this system existing from previous experimental studies by providing information about the dimensional nature of the gassurface interactions, and the details of how the two species twist and turn orientationally with increased surface interactions. Classical molecular dynamics have been employed in both equilibrium and steered molecular dynamics (SMD) simulations for SO2 at a neat-water surface, and at a surface with high interfacial SO2 concentrations. The results provide new molecular insights for understanding the interaction of this prevalent gas on aerosols and other aqueous surfaces in the environment

### 2011-10-18 EquilibratedOrientation.tex
working on review responses

@@
 
-The bisector tilt of the water molecules, $\theta$, concentrates around $\theta=$ 90\textdegrees within the first few\angs~above and below the water surface location, becoming progressively isotropic further through the interfacial region and into the water bulk of both systems.
-Above the surface location at positive distances, $\theta \leq$ 90\textdegrees~indicates that the water hydrogens tend to point towards the gas-phase side of the surface.
-As the tilt nears $\theta=$ 90\textdegrees~the \wat~bisector lies within the plane of the surface indicating a water orientation either flat on the surface, or with some amount of ``twist'' sending the OH bonds in towards, or out of the bulk.
+The bisector tilt of the water molecules, $\theta$, concentrates around $\theta=$ 90\textdegree within the first few\angs~above and below the water surface location, becoming progressively isotropic further through the interfacial region and into the water bulk of both systems.
+Above the surface location at positive distances, $\theta \leq$ 90\textdegree~indicates that the water hydrogens tend to point towards the gas-phase side of the surface.
+As the tilt nears $\theta=$ 90\textdegree~the \wat~bisector lies within the plane of the surface indicating a water orientation either flat on the surface, or with some amount of ``twist'' sending the OH bonds in towards, or out of the bulk.
 The value of $\phi$ determines the ``twist'' in this case.
-Both systems show a similar trend where waters at or just below the surface have values of $\phi$ near 90\textdegrees, and waters above the surface take on values of $\phi$ near 0\textdegrees.
+Both systems show a similar trend where waters at or just below the surface have values of $\phi$ near 90\textdegree, and waters above the surface take on values of $\phi$ near 0\textdegree.
 This jump in the angular distribution of $\phi$ indicates that waters at or below the surface lie mostly flat in the plane of the interface, and as they move above the surface towards the gas phase, they reorient with one OH bond pointing towards the water bulk, and one pointing out of the surface into the gas.
@@
 Although the plots show overall similarities for both the neat-water and saturated systems, the presence of a layer of adsorbed \suldiox~molecules alters the orientation of those waters furthest into the gas phase.
-For the saturated solution, the resulting orientation of waters above 0\angs, shown in the bottom plot of Figure \ref{fig:water-orientation}, is nearly isotropic in $\phi$, and with $\theta leq$ 90\textdegrees.
+For the saturated solution, the resulting orientation of waters above 0\angs, shown in the bottom plot of Figure \ref{fig:water-orientation}, is nearly isotropic in $\phi$, and with $\theta leq$ 90\textdegree.
 This results from waters with bisectors pointing further into the adsorbed \suldiox~gas layer, and both hydrogens pointing outward from the aqueous bulk.
-The effect is more pronounced as the waters move further from the water surface, and above 4\angs~the $\theta$ distribution is mostly centered around $\theta)=$ 0\textdegrees (see Figure \ref{fig:water-angles}).
+The effect is more pronounced as the waters move further from the water surface, and above 4\angs~the $\theta$ distribution is mostly centered around $\theta)=$ 0\textdegree (see Figure \ref{fig:water-angles}).
 
@@
 
-In the interfacial region the angular distribution of the single \suldiox~(in the neat-water system) is concentrated primarily in $\theta \lt$ 90 \textdegrees.
-The peak of the distribution occurs at $\theta)=$ 0\textdegrees.
+In the interfacial region the angular distribution of the single \suldiox~(in the neat-water system) is concentrated primarily in $\theta \lt$ 90 \textdegree.
+The peak of the distribution occurs at $\theta)=$ 0\textdegree.
 This indicates that the \suldiox~bisector points out of the water surface, with the sulfur atom pointing towards the aqueous bulk, and the two oxygens pointing into the gas phase.

### 2011-10-18 Methods.tex
working on review responses

@@
 Because of the extremely low concentration of \suldiox~in the neat-water system the surface waters behave similarly to a true neat-water air-liquid interface.
-The neat-water system orientational results shown later in this work reproduce well the results of our previous orientational studies of surface water behavior.\cite{Walker2006b,Hore2008} The saturated system had 22 \suldiox~molecules introduced to the water slab bulk in order to saturate it to a level coinciding with the Henry's law constant for \suldiox~in water.
-Additionally, 50 \suldiox~molecules were introduced into the gas phase outside of the saturated water slab to simulate an added \suldiox~gas pressure.
+The neat-water system orientational results shown later in this work reproduce well the results of our previous orientational studies of surface water behavior.\cite{Walker2006b,Hore2008} The saturated system had 22 \suldiox~molecules introduced to the water slab bulk in order to saturate it to a level coinciding with the Henry's law constant for \suldiox~in water (k\textdegree$_H = 1.4$ mol/kg*bar).\cite{Lide2000} Additionally, 50 \suldiox~molecules were introduced into the gas phase outside of the saturated water slab to simulate an added \suldiox~gas pressure.
 The saturated system with both bulk and gaseous \suldiox~was then evolved for 2 ns to produce a starting configuration for further saturated simulations.

### 2011-10-19 DensityResults.tex
Responses to the reviewers

@@
 The density distributions of both \wat~and \suldiox~were calculated for the equilibrium MD simulations.
-The results presented in Figure \ref{fig:density} show both the water (black) and \suldiox~(red) density distributions across the two simulated slabs.
+The results presented in Figure \ref{fig:density} show both the water (black) and \suldiox~(red) density distributions, averaged over the two simulated interfaces of each slab.
 As shown, the single \suldiox~in the neat-water system remained at the water slab surface.
@@
 The center of the \suldiox~density distribution is further into the gas phase than for the neat-water surface with a single-\suldiox~molecule.
+Additionally, the water profile is broader in the saturated system.
+This indicates that water penetrates into the \suldiox~layer, and moves further into the gas phase than under the neat-water conditions.
 Because of the limitations of the classical model in accurately reproducing the first hydration shell around the \suldiox, we do not draw specific conclusions about the hydrated surface complex's geometry.

### 2011-10-19 Supplemental.tex
Responses to the reviewers

\section {Supplementary Materials}

The SO$_2$ model used in our simulations was originally parameterized for use with the Dang-Chang water model as used in the work of Baer et al (J. Phys. Chem. B 2010, 114, 7245–7249). In the present work, the POL3 water model was used because of the computational efficiency afforded by the rigid bonds, and the incorporation of the POL3 model into the Amber suite of MD tools that was used for carrying out the simulations. The radial distribution functions of the water-SO$_2$ have been calculated, along with the interaction energy of the H$_2$O-SO$_2$ system as a function of the S-O distance. These are directly comparable to the results of Baer et al. The results are in very good agreement with the Dang-Chang model used with the polarizable SO$_2$, and justify the use of the simpler POL3 model for these interfacial simulations.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/rdf/rdf.png}
	\end{center}
\end{figure}

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/energy/interaction-energy.png}
	\end{center}
\end{figure}

### 2011-10-19 EquilibratedOrientation.tex
merging

@@
 
-In the interfacial region the angular distribution of the single \suldiox~(in the neat-water system) is concentrated primarily in $\theta \lt$ 90 \textdegree.
+In the interfacial region the angular distribution of the single \suldiox~(in the neat-water system) is concentrated primarily in $\theta <$ 90 \textdegree.
 The peak of the distribution occurs at $\theta)=$ 0\textdegree.
 This indicates that the \suldiox~bisector points out of the water surface, with the sulfur atom pointing towards the aqueous bulk, and the two oxygens pointing into the gas phase.
-This same distribution occurs in the saturated system for depths below the surface location, $\lt 0$\angs.
+This same distribution occurs in the saturated system for depths below the surface location, $< 0$\angs.
 Beyond 4\angs~above the surface, both distributions become mostly isotropic, either because the \suldiox~does not venture into the gas in the neat-water system, or because of the nature of the adsorbed \suldiox~layer in the saturated system.

### 2011-10-19 ReviewResponses.tex

The authors greatly appreciate the detail of the reviews provided, and have made significant changes to the manuscript to incorporate the suggestions, and to clarify the various points raised.

Review #3

1. The suggested literature (Garrett et al. and Wilson) cover many aspects of molecular dynamics simulations and computational techniques used for analysis of gaseous uptake into aqueous surfaces. The work of Garrett, Schenter, and Morita address the landscape of current techniques and water models used, but do not directly address sulfur dioxide uptake in their included analyses. The referenced publications in that work have already been included in our manuscript (see Jayne et al, 1990). Wilson and Pohorille (J. Phys. Chem. B 1997, 101, 3130) look at the adsorption of alcohols to the liquid-vapor interface, and present a refined analysis of the solvation process, but do not address sulfur dioxide uptake or orientation. We believe that although dozens, if not hundreds, of very fine publications have been dedicated to uptake of various species at the air-water interface, referencing extraneous works that do not specifically address sulfur dioxide behavior would detract from our present manuscript. We did not aim for this manuscript to be an exhaustive review of the field of gaseous uptake and solvation processes via molecular dynamics simulations. We have thus limited the scope of our bibliography to include only the most pertinent publication relating to sulfur dioxide behavior.

We believe that our introduction to the manuscript specifically addresses what is and is not known, and how the present work adds to our knowledge of sulfur dioxide behavior. We emphasize here again that our study was not set to be a review of the field of gaseous uptake, but a narrowly defined study of sulfur dioxide orientational behavior on an aqueous surface. Thus we did not include a discussion relating to other gaseous molecules or related systems.

2. The suggestion to use bivariate distributions is appreciated. We have reworked our data and presented the relevant depth-profiles as bivariate plots P($\theta$, $\phi$) showing the specific orientations of water and sulfur dioxide at various depths (slices) throughout the interfacial region.

3. The "Computational Approach" section does not contain any text regarding results or analysis. The section is reserved to define the computational approach, and to define the terminology used regarding depth and orientational analysis.

4. The term "absorption" is used only once in the computational methods section. Regarding the abstract, we have submitted a corrected version.

5. To address the used of the POL3 model with the SO2 model developed by Baer et al, we have submitted a supplemental analysis to be included with the manuscript. We have calculated the radial distribution function of the water-SO2 system to compare the resultant structure and geometries of the interactions between the two types of molecules. Furthermore, we have calculated the SO2-H2O interaction as a function of the S-O separation. Both of these are directly comparable to the results of Baer et al. Our calculations show a nearly identical result to Baer's in both the interaction energy and liquid structure using the POL3 water model. Thus we feel this justifies the use of the more computationally efficient POL3 model.

6. The Henry's law constant has been added to the text.

7. The pressure in the gas region was set initially to 1 atm. The system was then equilibrated until all the gaseous SO2 adsorbed to the water surface, at which point another 1 atm of SO2 was introduced. This procedure was repeated until the gas phase SO2 no longer adsorbed to the surface. This required a total of 50 SO2 molecules introduced in the gas phase. Thus a constant pressure of 1 atm was used throughout equilibration and data collection simulations.

8. We have addressed the technical details in the text.

9. Sections for both the equilibrium MD and the steered MD were specifically included to resolve any doubt regarding the simulations of the two different systems. The text may appear to be redundant, but we are addressing two separate systems that are simulated slightly differently.

10. The terminology has been chosen as the most appropriate descriptor for the system to distinguish the equilibrated system from the steered MD system. System descriptors have been used consistently throughout the manuscript, and we feel that they are appropriate to the current work.

11a. The system was simulated using the NSCM flag in Amber, and the drift still appears. This is not a drift of the entire system, but a slight swelling or movement of the slab over time. The problem is not remedied by the suggested solution.

11b. The same calculation was applied to both surfaces. The surface location was determined as suggested by the reviewer: independently for both surfaces.

12. Text regarding the Gibb's dividing surface has been updated to correct the language used. The manuscript now referes to the width of the water density profile.

13. Hydrogen bond lengths were taken from the radial distribution functions of interfacial waters. The manuscript has been updated by removing the note regarding the "typical bond length".

14c. The profiles were indeed averaged over both interfaces of each simulated slab. It has been noted in the manuscript.

14d. The water penetration into the SO2 layer is discussed at length later in the manuscript. A few sentences regarding this behavior have been added to the section on the density profiles.

15. We have removed this section from the manuscript as we are attempting to limit the scope of our work to the study of SO2 behavior, and not as a review of water models used in various capacities. The results of the current work for the POL3 model used agree well with our previous MD studies employing POL3. We have added in the supplemental section our justification for the use of this model.

### 2011-10-19 Introduction.tex

@@
 
-An important gas for developing a picture of gaseous adsorption and entry into a water surface is sulfur dioxide.\cite{Lattanzi2010,Shah2011,Tzivian2011,Johns2011,Faloona2009,Jurkat2010,Wu2011,Jayne1990,Yang2002} \suldiox~enters the environment as an important industrial product, and also naturally through terrestrial processes.
+An important gas for developing a picture of gaseous adsorption and entry into a water surface is sulfur dioxide.\cite{Lattanzi2010,Shah2011,Tzivian2011,Johns2011,Faloona2009,Jurkat2010,Wu2011,Jayne1990,Jayne1990a,Yang2002} \suldiox~enters the environment as an important industrial product, and also naturally through terrestrial processes.
 Atmospheric dust particles and gases have been implicated in the oxidation of \suldiox, and act as reaction surfaces for chemical mechanisms that are still poorly understood.\cite{Baltrusaitis2011,Rubasinghege2010,Li2007,Madsen2008,Boniface2000} \suldiox~acts as a major component of atmospheric pollution, and is a precursor to acid rain formation, and cloud nucleation.

### 2011-10-19 Methods.tex

@@
 Molecular dynamics simulations were performed using the Amber 11 software suite.\cite{Case2010} Polarizable models for the \wat~and \suldiox~molecules were used in the simulations, and have been used previously in studies on interfacial systems because they are known to more accurately reproduce interfacial structure and free energy profiles.\cite{Wick2007,Rivera2006,Dang1998} The \wat~model used is the POL3 water model,\cite{Caldwell1995} and for \suldiox~we used the model of Baer et al.
-that places a single polarizable center on the sulfur atom.\cite{Baer2010}
+that places a single polarizable center on the sulfur atom.\cite{Baer2010} An intermolecular cutoff of 12\angs~was used for long-range electrostatic forces.
+The simulations were performed in the NVT ensemble using Langevin dynamics for temperature control.
+Induced dipoles were treated by the polarizable potential functions of the Amber molecular dynamics software.
 
@@
 The water profile was then fit to a $\tanh$ function (black line) to extract the position and width parameters of the water surface.
-The fitted Gibbs dividing surface has a width of 3.77\angs, which is comparable to values reported for similar neat-water systems.\cite{Dang1997,Hore2008} A bulge in the gas-phase ($\leq 52$\angs) side of the water density profile is indicative of the drift of the water slab over the course of the trajectory.
+The fitted surface water density profile has a width of 3.77\angs, which is comparable to values reported for similar neat-water systems.\cite{Dang1997,Hore2008} A bulge in the gas-phase ($\leq 52$\angs) side of the water density profile is indicative of the drift of the water slab over the course of the trajectory.
 Thus the calculated location and width from the $\tanh$ line fit are not accurate over long trajectories for defining a stationary reference point.
@@
 First, counting the waters found in the topmost cross-section of the water slab over several time steps indicated between 65-75 waters that established a full monolayer.
-This was done by a visual inspection of the slab using the VMD MD visualization package.\cite{Humphrey1996} Alternatively, assuming a spherical model of water with a radius of 2.2\angs~(the typical length of a bulk-water hydrogen bond), two layers of hexagonally tight-packed spheres yielded a similar number of surface water molecules.
+This was done by a visual inspection of the slab using the VMD MD visualization package.\cite{Humphrey1996} Alternatively, assuming a spherical model of water with a radius of 2.2\angs, two layers of hexagonally tight-packed spheres yielded a similar number of surface water molecules.
 Increasing the number of waters used in calculating the surface location diminishes the effects of the few waters that briefly rise above the surface into the gas phase, stabilizing both the surface position and thickness values.

### 2011-10-19 ReviewResponses.tex
Finishing up responses to reviews. Now have to finish altering the images of the transit orientation and make a TOC image.

@@
+\documentclass{article}
+
+\usepackage {setspace}
+\usepackage [pdftex]{graphicx}
+\usepackage[sort&compress]{achemso}
+
+\usepackage[
+top    = 1.0in,
+bottom = 1.0in]{geometry}
+
+\begin{document}
+
+\newcommand{\suldiox}{SO$_2$}
+\newcommand{\ang}{\,$\textrm{\AA}$}
+\newcommand{\angs}{\ang}
+\newcommand{\wat}{H$_2$O}
+
+%\doublespacing
+
 The authors greatly appreciate the detail of the reviews provided, and have made significant changes to the manuscript to incorporate the suggestions, and to clarify the various points raised.
+Below are responses to each of the issues raised by the reviewers.
 
-Review #3
+\subsection*{Review 2}
 
-1.
-The suggested literature (Garrett et al.
-and Wilson) cover many aspects of molecular dynamics simulations and computational techniques used for analysis of gaseous uptake into aqueous surfaces.
-The work of Garrett, Schenter, and Morita address the landscape of current techniques and water models used, but do not directly address sulfur dioxide uptake in their included analyses.
-The referenced publications in that work have already been included in our manuscript (see Jayne et al, 1990).
-Wilson and Pohorille (J.
+\begin{enumerate}
+
+\item The steered MD technique used in our study is one established previously in the publications referenced in the introduction of the manuscript (26-31).
+Specifically, the theoretical foundation of the technique was established by Isralewitz (Current opinion in structural biology [0959-440X] Isralewitz yr:2001 vol:11 iss:2 pg:224-230).
+The technique applies a constant force to any combination of atoms or atomic subgroups in the system in order to affect a motion.
+The amount of force applied is very small compared to the intermolecular forces of the aqueous surface, but the applied force is constant, and slowly moves the \suldiox~into the aqueous surface.
+We believe the method to be sound, and have experimented with several of the parameters to find a combination that does not artificially bias the geometries and orientations of the atoms involved.
+
+\item In this work we attempted to two distinct studies that involve bulk-solvated \suldiox~as a starting configuration.
+The first study looks at a single \suldiox~already adsorbed to a water surface (the neat-water system).
+The aqueous system was then equilibrated with \suldiox~already bound and solvated.
+Evolution of the system through an extended equilibration resulted in a very surface-active \suldiox, which is the expected result both from our own surface-specific experiments, and from the results of Baer et al, that simulate this system identically (save for the choice of water model used).
+Thus the initial artificially solvated \suldiox~evolved to a natural location near the aqueous surface.
+
+The second study ("saturated") used the solvated \suldiox~as a starting configuration as well, but with two distinctions from the "neat-water" system.
+First there were several more \suldiox~molecules in the bulk instead of a single molecule.
+Secondly, \suldiox~was introduced in the gas-phase as well, and subsequently adsorbed to the water surface during the equilibration phase.
+Although the starting conditions of bulk solvation in both the neat-water and saturated systems were not necessarily representative of the \suldiox~behavior, the extended equilibration phase of our simulations evolved the \suldiox~to a location and system geometry that is consistent with our experimental results (surface active), and with the results of simulations carried out previously by Baer et al (residual sub-surface \suldiox).
+Thus, our "surface equilibrated method" finds justification in previous published results.
+
+Regarding the steered MD (SMD), the systems were established and equilibrated in a manner identical to the surface equilibrated method.
+Although the equilibration starting configurations may not have been representative of an equilibrium system, the configurations at the start of our data collection were what we expected, and what previous studies dictated would occur in terms of \suldiox~location and behavior.
+The subsequent SMD then introduced \suldiox~from the gas phase, which does not introduce into the system artificial behavior, as \suldiox~gas is known to adsorb from the vapor (as pointed out in the suggested literature of Donaldson et al, below, and from our own experimental studies).
+
+Lastly, because the bond breaking and ionization of the \suldiox~molecules does not occur in classical MD, we do not draw conclusions regarding solvated \suldiox~below the interfacial region in the bulk.
+We limit the scope of our manuscript to orientation and behavior within the interfacial region.
+Because of the points listed above, we believe that the artificial introduction of \suldiox~into the bulk water phase is justified, considering the effects of the extensive equilibration process prior to our data collection steps.
+We have not taken further steps to alter the manuscript in this regard.
+
+\item The work of Donaldson et al has been added to the manuscript.
+This work had been previously referenced in publications of our group's experimental work with \suldiox~adsorption and hydration detected via SFG techniques.
+We appreciate the mention of this oversight in the current manuscript.
+
+\end{enumerate}
+
+\subsection*{Review 3}
+\begin{enumerate}
+\item The suggested literature review (Wilson and Pohorille, J.
 Phys.
 Chem.
-B 1997, 101, 3130) look at the adsorption of alcohols to the liquid-vapor interface, and present a refined analysis of the solvation process, but do not address sulfur dioxide uptake or orientation.
-We believe that although dozens, if not hundreds, of very fine publications have been dedicated to uptake of various species at the air-water interface, referencing extraneous works that do not specifically address sulfur dioxide behavior would detract from our present manuscript.
+B 1997, 101, 3130) covers many aspects of molecular dynamics simulations and computational techniques used for analysis of gaseous uptake into aqueous surfaces.
+The work of Garrett, Schenter, and Morita address the landscape of current techniques and water models used, but neither of them directly address sulfur dioxide uptake in the texts.
+The referenced publications in that work have already been included in our manuscript (see Jayne et al, 1990).
+Wilson and Pohorille look at the adsorption of alcohols to the liquid-vapor interface, and present a refined analysis of the solvation process, but do not address sulfur dioxide uptake or orientation.
+We believe that although dozens, if not hundreds, of very fine publications have been dedicated to uptake of various species at the air-water interface, referencing such works that do not specifically address sulfur dioxide behavior would distract from our present goals set forth in the manuscript.
 We did not aim for this manuscript to be an exhaustive review of the field of gaseous uptake and solvation processes via molecular dynamics simulations.
-We have thus limited the scope of our bibliography to include only the most pertinent publication relating to sulfur dioxide behavior.
+We have thus limited the scope of our bibliography to include only the most pertinent publication relating to sulfur dioxide behavior, and the relevant techniques.
 
-We believe that our introduction to the manuscript specifically addresses what is and is not known, and how the present work adds to our knowledge of sulfur dioxide behavior.
-We emphasize here again that our study was not set to be a review of the field of gaseous uptake, but a narrowly defined study of sulfur dioxide orientational behavior on an aqueous surface.
+We believe that our introduction to the manuscript specifically addresses what is and is not known, and how the present work adds to our knowledge of sulfur dioxide behavior in aqueous/vapor interfacial environments.
+We emphasize here again that our study was not set to be a review of the field of gaseous uptake, but a narrowly defined study of sulfur dioxide orientational behavior on an aqueous surface, and a complement to our experimental studies on the same system.
 Thus we did not include a discussion relating to other gaseous molecules or related systems.
+Those discussions have already been published in several of our previous experimental studies that we reference in the manuscript.
 
-2.
-The suggestion to use bivariate distributions is appreciated.
+\item The suggestion to use bivariate distributions is appreciated and well made.
 We have reworked our data and presented the relevant depth-profiles as bivariate plots P($\theta$, $\phi$) showing the specific orientations of water and sulfur dioxide at various depths (slices) throughout the interfacial region.
 
-3.
-The "Computational Approach" section does not contain any text regarding results or analysis.
-The section is reserved to define the computational approach, and to define the terminology used regarding depth and orientational analysis.
+\item The "Computational Approach" section does not contain any text regarding results or analysis.
+The section is reserved to define the computational approach used, and to define the terminology used regarding depth and orientational analysis.
 
-4. The term "absorption" is used only once in the computational methods section. Regarding the abstract, we have submitted a corrected version.
+\item The term "absorption" is used only once in the computational methods section. Regarding the abstract, we have submitted a corrected version and greatly appreciate mention of this error.
 
-5.
-To address the used of the POL3 model with the SO2 model developed by Baer et al, we have submitted a supplemental analysis to be included with the manuscript.
-We have calculated the radial distribution function of the water-SO2 system to compare the resultant structure and geometries of the interactions between the two types of molecules.
-Furthermore, we have calculated the SO2-H2O interaction as a function of the S-O separation.
-Both of these are directly comparable to the results of Baer et al.
+\item To address the used of the POL3 model with the \suldiox~model developed by Baer et al, we have submitted a supplemental analysis to be included with the manuscript.
+We have calculated the radial distribution functions of the \wat-\suldiox~system to compare the resultant structure and geometries of the interactions between the two types of molecules.
+Furthermore, we have calculated the \suldiox-\wat~interaction as a function of the S-O separation.
+The results of both of these analyses are directly comparable to the results of Baer et al.
 Our calculations show a nearly identical result to Baer's in both the interaction energy and liquid structure using the POL3 water model.
@@
 
-6. The Henry's law constant has been added to the text.
+\item The Henry's law constant has been added to the text.
 
-7.
-The pressure in the gas region was set initially to 1 atm.
-The system was then equilibrated until all the gaseous SO2 adsorbed to the water surface, at which point another 1 atm of SO2 was introduced.
-This procedure was repeated until the gas phase SO2 no longer adsorbed to the surface.
-This required a total of 50 SO2 molecules introduced in the gas phase.
+\item The pressure in the gas region was set initially to 1 atm.
+The system was then equilibrated until all the gaseous \suldiox~adsorbed to the water surface, at which point another 1 atm of \suldiox~was introduced.
+This procedure was repeated until the gas phase \suldiox~no longer adsorbed to the surface.
+This required a total of 50 \suldiox~molecules introduced in the gas phase.
 Thus a constant pressure of 1 atm was used throughout equilibration and data collection simulations.
+We do not feel that altering the text is a necessary step regarding this point as the bulk of \suldiox~molecules adsorb to the surface and do not remain in gas to contribute to the overall gas pressure.
 
-8. We have addressed the technical details in the text.
+\item We have addressed the technical details of the simulation in the text of the manuscript.
 
-9.
-Sections for both the equilibrium MD and the steered MD were specifically included to resolve any doubt regarding the simulations of the two different systems.
-The text may appear to be redundant, but we are addressing two separate systems that are simulated slightly differently.
+\item Two sections, one for each of the surface equilibrated MD and the steered MD, were specifically included to resolve any doubt regarding the simulation conditions of the two different techniques used.
+The text may appear to be redundant, but we are addressing two separate simulation techniques.
 
-10.
-The terminology has been chosen as the most appropriate descriptor for the system to distinguish the equilibrated system from the steered MD system.
+\item The terminology (e.g.
+"surface equilibrated method") has been chosen as the most appropriate descriptor for the system to clearly distinguish the equilibrated system from the steered MD system.
 System descriptors have been used consistently throughout the manuscript, and we feel that they are appropriate to the current work.
 
-11a.
-The system was simulated using the NSCM flag in Amber, and the drift still appears.
-This is not a drift of the entire system, but a slight swelling or movement of the slab over time.
+\item 
+\begin{enumerate}
+\item All simulations were performed using the NSCM flag in Amber, and the drift still occurs.
+This is not a drift of the entire system center of mass, but a slight swelling or movement of the slab over time.
 The problem is not remedied by the suggested solution.
 
-11b. The same calculation was applied to both surfaces. The surface location was determined as suggested by the reviewer: independently for both surfaces.
+\item In all analyses the calculations were applied to both slab surfaces. The surface locations were determined as suggested by the reviewer: independently for both surfaces.
+\end{enumerate}
 
-12. Text regarding the Gibb's dividing surface has been updated to correct the language used. The manuscript now referes to the width of the water density profile.
+\item Text regarding the Gibb's dividing surface has been updated to correct the language used.
+The manuscript now refers to the width of the water density profile instead of the width of the Gibb's dividing surface.
 
-13. Hydrogen bond lengths were taken from the radial distribution functions of interfacial waters. The manuscript has been updated by removing the note regarding the "typical bond length".
+\item Hydrogen bond lengths were taken from the radial distribution functions of interfacial waters. The manuscript has been updated by removing the note regarding the "typical bond length".
 
-14c. The profiles were indeed averaged over both interfaces of each simulated slab. It has been noted in the manuscript.
+\item 
+\begin{enumerate}
+\item All profiles (e.g. density, orientational, etc.) were indeed averaged over both slab surfaces of each simulated slab. It has been noted in the manuscript.
 
-14d. The water penetration into the SO2 layer is discussed at length later in the manuscript. A few sentences regarding this behavior have been added to the section on the density profiles.
+\item The water penetration into the \suldiox~layer is discussed at length later in the manuscript.
+A few sentences regarding this behavior have been added to the section on the density profiles as suggested.
+\end{enumerate}
 
-15.
-We have removed this section from the manuscript as we are attempting to limit the scope of our work to the study of SO2 behavior, and not as a review of water models used in various capacities.
+\item We have removed this section from the manuscript as we are attempting to limit the scope of our work to the study of \suldiox~behavior, and not as a review of water models used in various capacities.
 The results of the current work for the POL3 model used agree well with our previous MD studies employing POL3.
 We have added in the supplemental section our justification for the use of this model.
+
+\end{enumerate}
+
+\end{document}

### 2011-10-19 Introduction.tex
Finishing up responses to reviews. Now have to finish altering the images of the transit orientation and make a TOC image.

--- 2011-07-04 Introduction.tex
@@
 
-An important gas for developing a picture of gaseous adsorption and entry into a water surface is sulfur dioxide.\cite{Lattanzi2010,Shah2011,Tzivian2011,Johns2011,Faloona2009,Jurkat2010,Wu2011,Jayne1990,Yang2002} \suldiox~enters the environment as an important industrial product, and also naturally through terrestrial processes.
+An important gas for developing a picture of gaseous adsorption and entry into a water surface is sulfur dioxide.\cite{Donaldson1995,Lattanzi2010,Shah2011,Tzivian2011,Johns2011,Faloona2009,Jurkat2010,Wu2011,Jayne1990,Jayne1990a,Yang2002} \suldiox~enters the environment as an important industrial product, and also naturally through terrestrial processes.
 Atmospheric dust particles and gases have been implicated in the oxidation of \suldiox, and act as reaction surfaces for chemical mechanisms that are still poorly understood.\cite{Baltrusaitis2011,Rubasinghege2010,Li2007,Madsen2008,Boniface2000} \suldiox~acts as a major component of atmospheric pollution, and is a precursor to acid rain formation, and cloud nucleation.

### 2011-10-24 Article.tex
Final images for the water and so2 equilibrated orientations
merging
Polishing the review responses

@@
 \usepackage {amsmath}
+\usepackage {textcomp}
 \usepackage {setspace}
@@
 
+\usepackage[
+top    = 0.5in,
+bottom = 1.0in]{geometry}
+%left   = 2.50cm,
+%right  = 2.50cm]{geometry}
+
 % some formatting tags
-\oddsidemargin  0.0in
-\evensidemargin 0.0in
-\textwidth      6.5in
+%\oddsidemargin  0.0in
+%\evensidemargin 0.0in
+%\textwidth      6.5in
 
@@
 \newcommand{\wat}{H$_2$O}
+%\newcommand{\deg}{$\,^{\circ}$}
 
@@
 \begin{abstract}
-One might expect the high surface tension of water to be a barrier to adsorption of a gas into the liquid phase, but we know that gaseous adsorption on and into a water surface is a common phenomenon on this planet.
+One might expect the high surface tension of water to be a barrier to absorption of a gas into the liquid phase, but we know that gaseous adsorption onto, and subsequent absorption into a water surface is a common phenomenon on this planet.
 What is not commonly known is how an atmospheric gas such as \suldiox~and molecules at the water surface can overcome the barrier created by strong water-water surface bonding interactions.

### 2011-10-24 EquilibratedOrientation.tex
Polishing the review responses

@@
 At each distance from the surface location, an orientation profile was created for both the \wat~and \suldiox~molecules.
-The bivariate orientation distributions for the angles $\theta$ and $\phi$ at various depths were combined to form the intensity plots that show how the molecular orientation distributions change with distance to the surface location.
+The bi-variate orientation distributions for the angles $\theta$ and $\phi$ at various depths were combined to form the intensity plots that show how the molecular orientation distributions change with distance to the surface location.
 These plots (Figures \ref{fig:water-orientation}, \ref{fig:so2-transit-angles}, and \ref{fig:so2-surface-angles}) allow for a visual interpretation of how the net orientations are affected when moving from the gas phase through the interfacial region and to the surface location, and then further into the aqueous interfacial region and bulk.
@@
 The depth of each plot is marked (in \ang) in the upper-right, with the water surface location set at 0\angs, positive depths lie on the gas-phase side of the surface, and negative depths are on the bulk water side.
-The horizontal axes of each hisogram represent $\theta$ values, and the vertical axes represent $\phi$ values.
+The horizontal axes of each histogram represent $\theta$ values, and the vertical axes represent $\phi$ values.
 Areas of low intensity appear in dark blue, and highest intensity in dark red.
@@
 The water surface location is set at a position of 0\angs~with negative distance values located in the bulk of the slab, and positive distances towards the gas phase.
-Each histogram is a bivariate angle distributions for $\theta$ (horizontal axes) and $\phi$ (vertical axes) in both the neat-\wat~system (top) and the saturated system (bottom).
+Each histogram is a bi-variate angle distributions for $\theta$ (horizontal axes) and $\phi$ (vertical axes) in both the neat-\wat~system (top) and the saturated system (bottom).
 Regions of high intensity are dark red, and low intensity are dark blue.}

### 2011-10-25 resubmission/ReviewResponses.tex
Final resubmission to the journal

--- 2011-10-19 ReviewResponses.tex
@@
 Specifically, the theoretical foundation of the technique was established by Isralewitz (Current opinion in structural biology [0959-440X] Isralewitz yr:2001 vol:11 iss:2 pg:224-230).
-The technique applies a constant force to any combination of atoms or atomic subgroups in the system in order to affect a motion.
-The amount of force applied is very small compared to the intermolecular forces of the aqueous surface, but the applied force is constant, and slowly moves the \suldiox~into the aqueous surface.
+The technique applies a constant force or velocity to any combination of atoms or atomic subgroups in the system in order to affect a motion.
+In our SMD simulations, the amount of force applied is very small compared to the intermolecular forces of the aqueous surface, but the applied force is constant, and slowly moves the transiting \suldiox~into the aqueous surface from the gas.
 We believe the method to be sound, and have experimented with several of the parameters to find a combination that does not artificially bias the geometries and orientations of the atoms involved.
@@
 Evolution of the system through an extended equilibration resulted in a very surface-active \suldiox, which is the expected result both from our own surface-specific experiments, and from the results of Baer et al, that simulate this system identically (save for the choice of water model used).
-Thus the initial artificially solvated \suldiox~evolved to a natural location near the aqueous surface.
+Thus the initial artificially solvated \suldiox~evolved to a natural location near the aqueous surface following equilibration.
 
-The second study ("saturated") used the solvated \suldiox~as a starting configuration as well, but with two distinctions from the "neat-water" system.
+The second study (i.e.
+the "saturated" system) used the solvated \suldiox~as a starting configuration as well, but with two distinctions from the "neat-water" system.
 First there were several more \suldiox~molecules in the bulk instead of a single molecule.
@@
 Regarding the steered MD (SMD), the systems were established and equilibrated in a manner identical to the surface equilibrated method.
-Although the equilibration starting configurations may not have been representative of an equilibrium system, the configurations at the start of our data collection were what we expected, and what previous studies dictated would occur in terms of \suldiox~location and behavior.
+Although the starting point of the equilibration may not have been representative of an equilibrium system of solvated \suldiox, the configurations at the start of our data collection were what we expected, and what previous studies dictated would occur in terms of \suldiox~location and behavior.
 The subsequent SMD then introduced \suldiox~from the gas phase, which does not introduce into the system artificial behavior, as \suldiox~gas is known to adsorb from the vapor (as pointed out in the suggested literature of Donaldson et al, below, and from our own experimental studies).
@@
 
-\item The suggestion to use bivariate distributions is appreciated and well made.
-We have reworked our data and presented the relevant depth-profiles as bivariate plots P($\theta$, $\phi$) showing the specific orientations of water and sulfur dioxide at various depths (slices) throughout the interfacial region.
+\item The suggestion to use bi-variate distributions is appreciated and well made.
+We have significantly reanalyzed our data sets and presented the relevant depth-profiles as bi-variate plots P($\theta$, $\phi$) showing the specific orientations of water and sulfur dioxide at various depths (slices) throughout the interfacial region.
+Significant changes have been made to the text to accommodate the new plots.
 
@@
 
-\item The Henry's law constant has been added to the text.
+\item The Henry's law constant for \suldiox~in water has been added to the text.
 
@@
 We do not feel that altering the text is a necessary step regarding this point as the bulk of \suldiox~molecules adsorb to the surface and do not remain in gas to contribute to the overall gas pressure.
+A sentence regarding this was added to the methods section to clarify the procedure.
 
-\item We have addressed the technical details of the simulation in the text of the manuscript.
+\item We have addressed the technical details of the simulation conditions in the text of the manuscript.
 
-\item Two sections, one for each of the surface equilibrated MD and the steered MD, were specifically included to resolve any doubt regarding the simulation conditions of the two different techniques used.
+\item Two methods sections, one each for the surface equilibrated MD and the steered MD, were specifically included to resolve any doubt regarding the simulation conditions of the two different techniques used.
 The text may appear to be redundant, but we are addressing two separate simulation techniques.
@@
 \begin{enumerate}
-\item All simulations were performed using the NSCM flag in Amber, and the drift still occurs.
+\item All simulations were performed using the NSCM flag in Amber, and the surface location drift still occurs.
 This is not a drift of the entire system center of mass, but a slight swelling or movement of the slab over time.

### 2011-10-25 Conclusions.tex
Final resubmission to the journal

@@
 Surface waters reorient in the presence of adsorbed \suldiox.
-The waters at the interface of a neat-water surface tend to lay flatter to the surface than when a saturating layer of \suldiox~is present.
-The waters interacting with the layer of adsorbed \suldiox~orient more perpendicularly to the interface, and further expose their ``free-OH'' uncoupled bonds for interactions with \suldiox, and hydrate complex formation.
+The waters at and just below the interface of a neat-water surface tend to lay flatter to the surface than when a saturating layer of \suldiox~is present.
+The waters above the surface location or interacting with the layer of adsorbed \suldiox~orient more perpendicularly to the interface, and further expose their ``free-OH'' uncoupled bonds for interactions with \suldiox, and hydrate complex formation.
 Furthermore, we have found that surface waters underneath a blanket layer of adsorbed \suldiox~will penetrate further into the gas phase, allowing for greater mobility of waters away from the aqueous bulk in the presence of \suldiox.
@@
 Those \suldiox~molecules at or below the surface water location strongly orient with the sulfur atom pointed in towards the water bulk, and the oxygen atoms out towards the gas phase.
-The \suldiox~molecules slightly above the water surface lose this net orientation within 10 \angs.
+The \suldiox~molecules slightly above the water surface lose this net orientation within 6-8 \angs.
 Those  molecules further from the water are more isotropically oriented.
@@
 The \suldiox~reorients as it makes its first contact with the water interface.
-Within 5 \angs~of the surface the \suldiox~is mostly oriented with its sulfur towards the water phase.
+Within 4 \angs~of the surface the \suldiox~is mostly oriented with its sulfur towards the water phase.
 The results for the transit through the interface show that in both systems of low and high \suldiox~concentration an adsorbing \suldiox~near the interface has very similar orientation to those molecules already bound to the water surface.

### 2011-10-25 EquilibratedOrientation.tex
Final resubmission to the journal

@@
 The bi-variate orientation distributions for the angles $\theta$ and $\phi$ at various depths were combined to form the intensity plots that show how the molecular orientation distributions change with distance to the surface location.
-These plots (Figures \ref{fig:water-orientation}, \ref{fig:so2-transit-angles}, and \ref{fig:so2-surface-angles}) allow for a visual interpretation of how the net orientations are affected when moving from the gas phase through the interfacial region and to the surface location, and then further into the aqueous interfacial region and bulk.
+These plots (Figures \ref{fig:water-orientation}, \ref{fig:so2-orientation}, and \ref{fig:so2-transit-angles}) allow for a visual interpretation of how the net orientations are affected when moving from the gas phase through the interfacial region and to the surface location, and then further into the aqueous interfacial region and bulk.
 Both the neat-water, with only a single \suldiox~introduced, and the high-concentration saturated system were analyzed.
@@
 Although the plots show overall similarities for both the neat-water and saturated systems, the presence of a layer of adsorbed \suldiox~molecules alters the orientation of those waters furthest into the gas phase.
-For the saturated solution, the resulting orientation of waters above 0\angs, shown in the bottom plot of Figure \ref{fig:water-orientation}, is nearly isotropic in $\phi$, and with $\theta leq$ 90\textdegree.
+For the saturated solution, the resulting orientation of waters above 0\angs, shown in the bottom set of plots of Figure \ref{fig:water-orientation}, is nearly isotropic in $\phi$, and with $\theta \leq$ 90\textdegree.
 This results from waters with bisectors pointing further into the adsorbed \suldiox~gas layer, and both hydrogens pointing outward from the aqueous bulk.
-The effect is more pronounced as the waters move further from the water surface, and above 4\angs~the $\theta$ distribution is mostly centered around $\theta)=$ 0\textdegree (see Figure \ref{fig:water-angles}).
+The effect is more pronounced as the waters move further from the water surface, and above 4\angs~the $\theta$ distribution is mostly concentrated around $\theta=$ 0\textdegree (see Figure \ref{fig:water-angles}).
 
@@
 
-The distribution of $\cos(\phi)$ is more sharply defined (i.e.
+The distribution of $\phi$ is more sharply defined (i.e.
 less isotropic) for the neat-water system than for the saturated one.
-Waters on the neat surface lie flatter, whereas the presence of the \suldiox~allows a greater range of ``twist'' for those waters in the plane of the interface.
-The $\phi$ distributions quickly become isotropic above the surface location as the bisectors orient more perpendicularly, and below the surface as the bulk water loses any orientational preference.
+Waters on the neat surface lie flat or perpendicular to the surface if they are below or above the water surface location, respectively.
+The presence of the \suldiox~allows a greater range of ``twist'' for those waters in the plane of the interface.
+The $\phi$ distributions quickly become isotropic above the saturated water surface location, shown as a uniform coloration across $\phi$ for most values of $\theta$.
 
@@
 The neat-water \suldiox~molecule remains within a narrower region of the interface than the saturated system \suldiox, but effective comparisons can still be drawn.
+Note that the depth range of the plots in Figure \ref{fig:so2-orientation} is different for the neat-water and saturated systems reflecting the surface mobility of the \suldiox~molecules in the two systems.
 
 In the interfacial region the angular distribution of the single \suldiox~(in the neat-water system) is concentrated primarily in $\theta <$ 90 \textdegree.
-The peak of the distribution occurs at $\theta)=$ 0\textdegree.
+The peak of the distribution occurs at $\theta=$ 0\textdegree.
 This indicates that the \suldiox~bisector points out of the water surface, with the sulfur atom pointing towards the aqueous bulk, and the two oxygens pointing into the gas phase.
@@
 Because the \suldiox~bisector near the surface is oriented perpendicularly to the interface, the isotropy in $\phi$ is expected.
-Further from the water surface where the bisector orientation becomes isotropic, the $\cos(\phi)$ distribution remains isotropic.
-For the \suldiox~orientation, the $\phi$ angle does not provide further information regarding the surface behavior or orientational preference.
+Further from the water surface where the bisector orientation becomes isotropic, the $\phi$ distribution remains isotropic.
+For the surface \suldiox~orientation, the $\phi$ angle does not provide further information regarding the surface behavior or orientational preference.
 
@@
 In this configuration the $\phi$ distribution is isotropic because of the water slab's in-plane symmetry.
-Note that the range of surface depths in the low concentration system is half that of the saturated system.
+Note that the range of surface depths in the low concentration plot is half that of the saturated system.
 This is due to the narrower position range of the single \suldiox~molecule.}

### 2011-10-25 Methods.tex
Polishing the review responses
Final resubmission to the journal

@@
 
-Molecular dynamics simulations were performed using the Amber 11 software suite.\cite{Case2010} Polarizable models for the \wat~and \suldiox~molecules were used in the simulations, and have been used previously in studies on interfacial systems because they are known to more accurately reproduce interfacial structure and free energy profiles.\cite{Wick2007,Rivera2006,Dang1998} The \wat~model used is the POL3 water model,\cite{Caldwell1995} and for \suldiox~we used the model of Baer et al.
+Molecular dynamics simulations were performed using the Amber 11 software suite.\cite{Case2010} Polarizable models for the \wat~and \suldiox~molecules were used in the simulations, and have been used previously in studies on interfacial systems because they are known to more accurately reproduce interfacial structure and free energy profiles.\cite{Wick2007,Rivera2006,Dang1998} The \wat~model used is the POL3 water model (also discussed in the supplemental information),\cite{Caldwell1995} and for \suldiox~we used the model of Baer et al.
 that places a single polarizable center on the sulfur atom.\cite{Baer2010} An intermolecular cutoff of 12\angs~was used for long-range electrostatic forces.
@@
 
-The low and high concentration simulated systems, ``neat-water'' and ``saturated water'', respectively, were created as follows: the neat-water simulation involved the addition of a single \suldiox~molecule either within the bulk of the water slab (for equilibrium MD), or above the slab surface (in the SMD simulations).
+The low and high concentration simulated systems, ``neat-water'' and ``saturated'', respectively, were created as follows: the neat-water simulation involved the addition of a single \suldiox~molecule either within the bulk of the water slab (for equilibrium MD), or above the slab surface (in the SMD simulations).
 The single-\suldiox~neat-water system was then evolved for 2 ns to produce an equilibrated starting configuration.
@@
 The neat-water system orientational results shown later in this work reproduce well the results of our previous orientational studies of surface water behavior.\cite{Walker2006b,Hore2008} The saturated system had 22 \suldiox~molecules introduced to the water slab bulk in order to saturate it to a level coinciding with the Henry's law constant for \suldiox~in water (k\textdegree$_H = 1.4$ mol/kg*bar).\cite{Lide2000} Additionally, 50 \suldiox~molecules were introduced into the gas phase outside of the saturated water slab to simulate an added \suldiox~gas pressure.
+The additional gas in the vapor phase was added over the course of several ns to keep a constant 1 atm of \suldiox~pressure above the water surface as \suldiox~gas molecules adsorbed to the surface.
 The saturated system with both bulk and gaseous \suldiox~was then evolved for 2 ns to produce a starting configuration for further saturated simulations.
@@
 
-Figure \ref{fig:density-flaw} demonstrates the problem of surface location drift during a simulation.
+Figure \ref{fig:density-flaw} demonstrates the problem of surface location drift during a simulation (even with utilizing the Amber NSCM parameter).
 Figure \ref{fig:density-flaw}A shows the density profile of water and \suldiox~over the course of one of the 10 ns trajectories used in this work using the original uncorrected coordinates of the system taken from the raw atomic positions of the molecular dynamics data output.
@@
 (B) $\theta$ is the value of the ``tilt'' of the molecular bisector from the reference axis, with $\theta$ values ranging from  0\textdegree~(aligned) to 180\textdegree~(anti-aligned).
-$\theta = $ 90\textdegree~places the bisector in the plane of the water surface.
+$\theta = $ 90\textdegree~aligns the bisector parallel to the plane of the water surface.
 Several bisector orientations are shown depicting different $\theta$ values.

### 2011-10-25 TransitOrientation.tex
Final resubmission to the journal

@@
 The angles $\theta$ and $\phi$ of the transiting \suldiox~(Figure \ref{fig:water-angles}) were calculated for each timestep of the SMD simulations as the \suldiox~was pulled into the water slab from the gas phase, both in the neat-water and saturated slab systems.
-The angle cosine values were collected for the 50 simulations of both systems for each distance from the water surface location, resulting in the 2-dimensional angle and depth-profile histograms shown in Figure \ref{fig:so2-transit-angles}.
-In all the data sets, the populations of the angle histograms at each distance from the surface were normalized to aid in comparison of regions with differing \suldiox~residence times.
+The orientation depth-profiles were collected for the 50 simulations of both systems for various distances from the water surface location, resulting in the 2-dimensional angle and depth-profile histograms shown in Figure \ref{fig:so2-transit-angles}.
 
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{images/transit-so2-angles/so2-angles-transit.png}
-		\caption{\suldiox~molecular orientation distributions during SMD transit simulations into an aqueous slab.
-Both the neat-water (top row) and saturated (bottom row) data sets were analyzed to determine the angles $\theta$ (left column), and $\phi$ (right column) of the transiting \suldiox.
-The distributions show angle cosines plotted against the distance of the \suldiox~to the location of the water surface.
-The data sets were averaged at each distance for the respective 50 simulations, and every distribution at each distance was normalized such that each histogram has a max population of 1 for purposes of comparison.}
+		\includegraphics[scale=1.0]{images/so2-angles/theta-phi-transit.png}
+		\caption{Molecular orientation distributions of an \suldiox~at different interfacial depths of an aqueous slab during SMD transit simulations.
+Both the neat-water (top) and saturated (bottom) data sets were analyzed to determine the angles $\theta$ (horizontal axes) and $\phi$ (vertical axes) of the transiting \suldiox.
+The distributions shown are those for the single transiting \suldiox~molecule averaged over the 50 SMD simulations.}
 		\label{fig:so2-transit-angles}
@@
 
-	From its starting position above the water surface, until the \suldiox~moves to within 10\angs~of the water surface location of both systems, the orientation is isotropic in $\theta$ and $\phi$ (not shown).
-Isotropic orientation is manifested in the plots as mostly uniform coloration at a given distance from the surface independent of $\cos$ of $\phi$.
-Near and into the interfacial region the bisector angle $\theta$ becomes more perpendicular to the water surface ($\cos(\theta)\approx1$) with the \suldiox~sulfur pointing into the water phase, consistent with the equilibrium MD simulation results above.
-At the point when the \suldiox~reaches the water surface location (distance $=0$\angs), the bisector is perpendicular to the interface in both the neat-water and saturated systems.
-Figure \ref{fig:so2-surface-angles}a depicts \suldiox~near the water surface with the bisector aligned perpendicular to the interface.
+	From its starting position above the water surface, until the \suldiox~moves to within 6\angs~of the water surface location of both systems, the orientation is isotropic in $\theta$ and $\phi$.
+Isotropic orientation is manifested in the plots as mostly uniform coloration at a given distance from the surface independent of $\theta$ or $\phi$.
+Near and into the interfacial region the bisector angle $\theta$ becomes more perpendicular to the water surface ($\theta \approx 0$\textdegree) with the \suldiox~sulfur pointing into the water phase, consistent with the equilibrium MD simulation results above.
+At the point when the \suldiox~reaches the water surface location (0\angs), the bisector is perpendicular to the interface in both the neat-water and saturated systems.
+%Figure \ref{fig:so2-surface-angles}a depicts \suldiox~near the water surface with the bisector aligned perpendicular to the interface.
 	In the absence of simulated ionic species that form through \suldiox-\wat~chemistry at the surface, it is clear that the adsorbing \suldiox~in the gas phase takes on a preferred orientation to adsorb on a water surface.
@@
 
-  % talk about theta differences
-  Comparing in more detail the difference in these two systems upon \suldiox~approach, for the neat-water surface, a transition occurs at approximately 5\angs~from the surface water location.
-Below 5\angs~above the surface, \suldiox~has a preferred net orientation and is close enough to the water surface that it begins to interact with the topmost surface waters.
+  % theta differences
+  Comparing in more detail the difference in these two systems upon \suldiox~approach, for the neat-water surface, a transition occurs at approximately 4\angs~above the surface water location.
+Below 4\angs~above the surface, \suldiox~has a preferred net orientation and is close enough to the water surface that it begins to interact with the topmost surface waters.
 In the saturated system, the same trend occurs, however the onset of the perpendicular orientation begins at approximately 8\angs~above the surface.
@@
 It is remarkable that the orientational trend appears so strongly in the $\theta$ plots even with so little data as was collected from the single \suldiox~molecule of each simulation.
-From onset of orientation above the surface until 10\angs~below, the \suldiox~holds a preferred orientation.
+From onset of orientation above the surface until 10\angs~below (not shown), the \suldiox~holds a preferred orientation.
 
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[scale=1.0]{images/angle-cartoons/so2-surface.png}
-		\caption{(A) \suldiox~molecules above and below both neat-water and saturated water surface location orient with the bisectors perpendicular to the interface.
+%\begin{figure}[h!]
+	%\begin{center}
+		%\includegraphics[scale=1.0]{images/angle-cartoons/so2-surface.png}
+		%\caption{(A) \suldiox~molecules above and below both neat-water and saturated water surface location orient with the bisectors perpendicular to the interface.
 The \suldiox~sulfur points down towards the aqueous phase.
 (B) In the neat-water system, a transiting \suldiox~molecule located below the water surface briefly reorients with one S-O bond more towards the interface above, and the other bond pointing further down into the water bulk.}
-		\label{fig:so2-surface-angles}
-	\end{center}
-\end{figure}
+		%\label{fig:so2-surface-angles}
+	%\end{center}
+%\end{figure}
 
-  % talk about phi differences
+  % phi differences
 	%Indeed, the $\phi$ orientations appear mostly isotropic without any clear trends for preferred orientation in that angle.
@@
 This is the case in both systems, with only a few exceptions.
-In both systems the $\cos(\phi)$ profiles exhibit mostly isotropic distributions above 0\angs, with several regions of lighter coloration interspersed, but without a clearly formed orientational trend.
-At the neat-water surface and just below, from 0\angs~to -5\angs~(in the water phase), the $\theta$ profile broadens down to $\cos(\theta)=0.5$, coinciding with a clear trend in the profile of $\phi$ near $\cos(\phi)=0$ shown in dark red at the bottom of the plot.
-This corresponds to \suldiox~inclined up to 60 degrees from the surface normal.
-The result is that \suldiox~in this region shows a preference for one of the S-O bonds to point more towards the aqueous bulk and the other to point away towards the interface.
-Figure \ref{fig:so2-surface-angles}B depicts the orientation of the \suldiox~slightly below the water surface location.
-This is in contrast to \suldiox~molecules above the water surface location, oriented more perpendicularly.
-The behavior is likely due to the interaction between the waters and one of the S-O bonds leading to a higher solvation than above the water surface location.
+In both systems the $\phi$ profiles exhibit mostly isotropic distributions above 0\angs, with several regions of lighter coloration interspersed, but without a clearly formed orientational trend.
+At the neat-water surface and just above (from 0-2\angs~above the water phase), the $\theta$ profile broadens to $\theta = 90$\textdegree, near $\phi = 90$\textdegree~appearing as a shoulder of light coloration in the bottom-left of the 0-2\angs~axes.
+This indicates that \suldiox~inclined up to 90\textdegree~from the surface normal will have a preferred $\phi$ orientation lying more flat to the water surface.
+This is in contrast to \suldiox~molecules above the water surface location, oriented more perpendicularly and without a preference for a particular range of values in $\phi$.
+The behavior is likely due to the interaction between the waters and the S-O bonds leading to a higher solvation than above the water surface location.
 As the \suldiox~is solvated by more highly-coordinated bulk water, the S-O bonds experience less equal interaction environments.

### 2011-10-25 resubmission/Article.tex
Final resubmission to the journal

--- 2011-07-07 submission/Article.tex
@@
 \usepackage[sort&compress]{achemso}
+%\usepackage{lineno}
 
-\oddsidemargin  0.0in
-\evensidemargin 0.0in
-\textwidth      6.5in
+\usepackage[
+top    = 0.5in,
+bottom = 1.0in]{geometry}
+%left   = 2.50cm,
+%right  = 2.50cm]{geometry}
+
+% some formatting tags
+%\oddsidemargin  0.0in
+%\evensidemargin 0.0in
+%\textwidth      6.5in
 
@@
 
+%\onehalfspacing
+%\linenumbers 
 \doublespacing
+
+\begin{abstract}
+One might expect the high surface tension of water to be a barrier to absorption of a gas into the liquid phase, but we know that gaseous adsorption onto, and subsequent absorption into a water surface is a common phenomenon on this planet.
+What is not commonly known is how an atmospheric gas such as \suldiox~and molecules at the water surface can overcome the barrier created by strong water-water surface bonding interactions.
+What this interplay looks like, the distances from the water surface at which these attractive interactions begin, and how they influence the orientational nature of both \suldiox~and surface water molecules is the focus of this computational study.
+The results fill a void in the information about this system existing from previous experimental studies by providing information about the dimensional nature of the gas-surface interactions, and the details of how the two species twist and turn orientationally with increased surface interactions.
+Classical molecular dynamics have been employed in both equilibrium and steered molecular dynamics (SMD) simulations for \suldiox~at a neat-water surface, and at a surface with high interfacial \suldiox~concentrations.
+The results provide new molecular insights for understanding the interaction of this prevalent gas on aerosols and other aqueous surfaces in the environment.
+\end{abstract}
 
@@
 
-Molecular dynamics simulations were performed using the Amber 11 software suite.\cite{Case2010} Polarizable models for the \wat~and \suldiox~molecules were used in the simulations, and have been used previously in studies on interfacial systems because they are known to more accurately reproduce interfacial structure and free energy profiles.\cite{Wick2007,Rivera2006,Dang1998} The \wat~model used is the POL3 water model,\cite{Caldwell1995} and for \suldiox~we used the model of Baer et al.
-that places a single polarizable center on the sulfur atom.\cite{Baer2010}
+Molecular dynamics simulations were performed using the Amber 11 software suite.\cite{Case2010} Polarizable models for the \wat~and \suldiox~molecules were used in the simulations, and have been used previously in studies on interfacial systems because they are known to more accurately reproduce interfacial structure and free energy profiles.\cite{Wick2007,Rivera2006,Dang1998} The \wat~model used is the POL3 water model (also discussed in the supplemental information),\cite{Caldwell1995} and for \suldiox~we used the model of Baer et al.
+that places a single polarizable center on the sulfur atom.\cite{Baer2010} An intermolecular cutoff of 12\angs~was used for long-range electrostatic forces.
+The simulations were performed in the NVT ensemble using Langevin dynamics for temperature control.
+Induced dipoles were treated by the polarizable potential functions of the Amber molecular dynamics software.
 
@@
 
-The low and high concentration simulated systems, ``neat-water'' and ``saturated water'', respectively, were created as follows: the neat-water simulation involved the addition of a single \suldiox~molecule either within the bulk of the water slab (for equilibrium MD), or above the slab surface (in the SMD simulations).
+The low and high concentration simulated systems, ``neat-water'' and ``saturated'', respectively, were created as follows: the neat-water simulation involved the addition of a single \suldiox~molecule either within the bulk of the water slab (for equilibrium MD), or above the slab surface (in the SMD simulations).
 The single-\suldiox~neat-water system was then evolved for 2 ns to produce an equilibrated starting configuration.
 Because of the extremely low concentration of \suldiox~in the neat-water system the surface waters behave similarly to a true neat-water air-liquid interface.
-The neat-water system orientational results shown later in this work reproduce well the results of our previous orientational studies of surface water behavior.\cite{Walker2006b,Hore2008} The saturated system had 22 \suldiox~molecules introduced to the water slab bulk in order to saturate it to a level coinciding with the Henry's law constant for \suldiox~in water.
-Additionally, 50 \suldiox~molecules were introduced into the gas phase outside of the saturated water slab to simulate an added \suldiox~gas pressure.
+The neat-water system orientational results shown later in this work reproduce well the results of our previous orientational studies of surface water behavior.\cite{Walker2006b,Hore2008} The saturated system had 22 \suldiox~molecules introduced to the water slab bulk in order to saturate it to a level coinciding with the Henry's law constant for \suldiox~in water (k\textdegree$_H = 1.4$ mol/kg*bar).\cite{Lide2000} Additionally, 50 \suldiox~molecules were introduced into the gas phase outside of the saturated water slab to simulate an added \suldiox~gas pressure.
+The additional gas in the vapor phase was added over the course of several ns to keep a constant 1 atm of \suldiox~pressure above the water surface as \suldiox~gas molecules adsorbed to the surface.
 The saturated system with both bulk and gaseous \suldiox~was then evolved for 2 ns to produce a starting configuration for further saturated simulations.
@@
 
+% need to add in the specifics about the SMD routine, force constants, calculations, etc.
 A second set of simulations began with an equilibrated water slab as in the surface equilibrated method above.
@@
 
-\subsection {Water Surface Location}
+\subsection {Aqueous Surface Location}
 
@@
 
+  %Taking one of the simulations of the neat-water slabs, a series of calculations were performed to find the averaged positions and the standard deviations of varying numbers of the topmost waters.
+It was found that using 70 waters produced a standard deviation of the water positions corresponding to a width of 3.2\angs.
+as reported by Baer et al.when fitting the density profile with a hyperbolic tangent lineshape.
+To match the interfacial width of Baer using the standard deviation of our simulations, 70 water molecules would have to be included in the top monolayer.
+Thus, the calculations and analyses presented in this work incorporate 70 water molecules when calculating the surface location.
+
 \subsection{Molecular Orientation}
@@
 Because $\phi=-\phi$, the value is reported within the range of $[0,\frac \pi 2]$.
-	However, values of $\cos(\theta)$ close to 1 or -1 result in an isotropic distribution in $\cos(\phi)$ because of the symmetry of the plane of the surface in directions perpendicular to the reference axis.
+	Values of $\theta$ close to 0\textdegree~or 180\textdegree~result in an isotropic distribution in $\phi$ because of the symmetry of the plane of the surface in directions perpendicular to the surface normal reference axis.
+
+%	The two molecules studied, \wat~and \suldiox, and similarly shaped with a C2$_v$ axis along their bisectors, and a molecular plane is defined by their three atoms.
+The orientational analyses presented herein focus on two angles used to define the molecular orientation in space.
+In each analysis a reference axis is used to define molecular angles, and is the z-axis unless otherwise specified.
+A ``tilt'' angle, $\theta$, defines the angle formed between a vector (generally the bisector vector pointing from the central atom in the direction of the other two atoms) and the positive reference axis.
+Thus the value of $\theta$ falls within a range of $[0,\pi]$.
+A second angle, $\phi$, defines the molecular ``twist'' around the reference axis.
+$\phi$ is formed between the projection of a vector onto the x-y reference plane, and the x-axis.
+$\phi$ is thus defined relative to the x-axis, and its values are in the interval $[-\pi,\pi]$.
+However, because of the symmetry of the \wat~and \suldiox~molecules and the equivalence of the two H or O atoms (in the \wat~or\suldiox, respectively), $\phi=-\pi$ is equivalent to $\phi=\pi$, and both situations are equivalent to $\phi=0$.
+Because $\phi=-\phi$, the value is reported within the range of $[0,\frac \pi 2]$.
+Figure \ref{fig:spherical-angle} shows the angle definitions using the z-axis as the reference axis.
 
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{molecular-angles.png}
+		\includegraphics[scale=1.0]{molecular-angles-small.png}
 		\caption{(A) The definition of the angles $\theta$ and $\phi$ used to define molecular orientation of \suldiox~and \wat~relative to a reference axis.
-(B) $\theta$ is the value of the ``tilt'' of the molecular bisector from the reference axis, with $\cos(\theta)$ values ranging from -1 to +1 indicating different bisector orientations.
-(C) $\phi$ is the ``twist'' angle formed between the vector normal to the molecular plane (the body-fixed y-axis) and the reference axis.
-For values of $\cos(\theta) \approx 0$, $\phi$ shows how flat the molecule is relative to the interfacial plane.}
+(B) $\theta$ is the value of the ``tilt'' of the molecular bisector from the reference axis, with $\theta$ values ranging from  0\textdegree~(aligned) to 180\textdegree~(anti-aligned).
+$\theta = $ 90\textdegree~aligns the bisector parallel to the plane of the water surface.
+Several bisector orientations are shown depicting different $\theta$ values.
+(C) $\phi$ is the ``twist'' angle defined as a rotation around the molecular bisector axis.
+A value of 0\textdegree~aligns the plane of the molecule perpendicular to the plane of the water surface.
+A value of 90\textdegree~rotates the molecule to lie flat in the plane of the water surface.
+Because of the symmetries of \suldiox, \wat, and the plane of the water surface, the range of $\phi$ values is limited to 0\textdegree~$\leq \phi \leq$~90\textdegree.}
 		\label{fig:water-angles}
@@
 The density distributions of both \wat~and \suldiox~were calculated for the equilibrium MD simulations.
-The results presented in Figure \ref{fig:density} show both the water (black) and \suldiox~(red) density distributions across the two simulated slabs.
+The results presented in Figure \ref{fig:density} show both the water (black) and \suldiox~(red) density distributions, averaged over the two simulated interfaces of each slab.
 As shown, the single \suldiox~in the neat-water system remained at the water slab surface.
 The \suldiox~in the saturated system accumulated mostly at the surface, but some residual \suldiox~remained well into the bulk water.
+
+%The density distributions of the \suldiox~were fit to gaussian lineshapes as shown in equation \ref{eq:gaussian} using the common definitions of the variables $A$, $\mu$, and $\sigma$, and given as a function of the distance to the top surface, $z$.
+
+%\begin{equation}
+%  \rho(z)=\frac{A}{\sqrt{2 \pi \sigma^2}} e^{-\frac{(z-\mu)^2}{2\sigma^2}}
+%  \label{eq:gaussian}
+%\end{equation}
 
@@
 At each distance from the surface location, an orientation profile was created for both the \wat~and \suldiox~molecules.
-The orientation distribution for the angles $\theta$ and $\phi$ at all depths were combined to form 2D intensity plots that show how the molecular orientation distributions change with distance to the surface location.
-These plots (Figures \ref{fig:water-orientation}, \ref{fig:so2-transit-angles}, and \ref{fig:so2-surface-angles}) allow for a visual interpretation of how the net orientations are affected when moving from the gas phase through the interfacial region and to the surface location, and then further into the aqueous interfacial region and bulk.
+The bi-variate orientation distributions for the angles $\theta$ and $\phi$ at various depths were combined to form the intensity plots that show how the molecular orientation distributions change with distance to the surface location.
+These plots (Figures \ref{fig:water-orientation}, \ref{fig:so2-orientation}, and \ref{fig:so2-transit-angles}) allow for a visual interpretation of how the net orientations are affected when moving from the gas phase through the interfacial region and to the surface location, and then further into the aqueous interfacial region and bulk.
 Both the neat-water, with only a single \suldiox~introduced, and the high-concentration saturated system were analyzed.
@@
 
-The 2D orientational depth profile plots show areas of low intensity in dark blue, and highest intensity in dark red.
-Regions of the plots where the intensity (coloration) is equally distributed from top to bottom along the entire orientational range are considered isotropic.
+The depth profile plots are arranged as a grid of 2D histograms.
+Each histogram is calculated for all molecules falling within a particular depth in the water interfacial region.
+The depth of each plot is marked (in \ang) in the upper-right, with the water surface location set at 0\angs, positive depths lie on the gas-phase side of the surface, and negative depths are on the bulk water side.
+The horizontal axes of each histogram represent $\theta$ values, and the vertical axes represent $\phi$ values.
+Areas of low intensity appear in dark blue, and highest intensity in dark red.
+Regions of the plots where the intensity (coloration) is equally distributed along either the vertical or horizontal axes are considered isotropic in $\phi$ or $\theta$, respectively.
 Likewise, areas of the plot with high intensity over a small orientational range are considered to exhibit an orientational preference at the given depth.
-The histograms are arranged with the plots of $\cos(\theta)$ on the left and $\cos(\phi)$ on the right.
-The surface is located at 0\angs.
 The angle distributions from both simulated slab surfaces were averaged for all the orientation analyses.
@@
 The interfacial region for both these calculations and the VSF experiments is defined as the region where molecular orientational anisotropy exists around the surface water location.
-Referring to Figure \ref{fig:water-orientation}, calculations indicate an interfacial width of approximately 10\angs~for the neat-water system, and approximately 16 \angs~for the saturated system.
-In both systems the strongest orientational preference is found at the slab surfaces (positions above 0\angs) where the water is furthest towards the gas phase.
-Previous work on orientational preference of water at air surfaces shows the same trend as our neat-water results.\cite{Walker2006b,Hore2008} The narrow "peninsula" of high intensity that extends in the $\cos(\theta)$ plots to the right of the -5\angs~location shows the overall preference of water to orient at the surface.
-The $\cos(\phi)$ plots are similar to each other with a narrow region of reorientation, but the effect in the interfacial region is greater in the neat-water system as evidenced by the sharper transition in intensity from blue to red, compared to the saturated system that has a less pronounced intensity change.
+Fitting the water density profiles we have calculated an interfacial width of approximately 10\angs~for the neat-water system, and approximately 16\angs~for the saturated system.
+In both systems the strongest orientational preference is found at the slab surfaces (positions near 0\angs).
+Previous work on orientational preference of water at air surfaces shows the same trend as our neat-water results.\cite{Walker2006b,Hore2008} The swaths of coloration indicating high intensity appearing in the neat-water plots from -4 to +4\angs, (and the corresponding regions in most of the saturated system plots) show the overall preference of water to orient at the surface.
+The plots of the neat-water and saturated systems are similar to each other with a narrow region of reorientation, but the effect in the interfacial region is greater in the neat-water system as evidenced by the sharper transition in intensity from blue to red, compared to the saturated system that has a less pronounced intensity change over larger areas of the histograms.
 
-The bisector tilt of the water molecules, $\cos(\theta)$, concentrates around $\cos(\theta)=0$ within the first few\angs~above and below the water surface location, becoming progressively isotropic further through the interfacial region and into the water bulk of both systems.
-As the tilt nears $\cos(\theta)=0$ the \wat~bisector lies within the plane of the surface indicating a water orientation either flat on the surface, or with some amount of ``twist'' sending the OH bonds in towards, or out of the bulk.
+The bisector tilt of the water molecules, $\theta$, concentrates around $\theta=$ 90\textdegree within the first few\angs~above and below the water surface location, becoming progressively isotropic further through the interfacial region and into the water bulk of both systems.
+Above the surface location at positive distances, $\theta \leq$ 90\textdegree~indicates that the water hydrogens tend to point towards the gas-phase side of the surface.
+As the tilt nears $\theta=$ 90\textdegree~the \wat~bisector lies within the plane of the surface indicating a water orientation either flat on the surface, or with some amount of ``twist'' sending the OH bonds in towards, or out of the bulk.
 The value of $\phi$ determines the ``twist'' in this case.
-Both systems show a defined intensity concentration (dark red) in the distributions around $\cos(\phi)=1$ on the aqueous side of the interfacial region.
-This results from an orientation of the water's y-axis (normal to the molecular plane) aligned perpendicular to the plane of the water surface.
-Thus, for both the neat-water and saturated systems, the preferred orientation of water molecules at a distance of 0\angs~ is to lie mostly flat to the plane of the interface with a slight ``twist'' sending one OH bond further into the gas phase, and the other OH towards the water bulk.
-This result agrees somewhat with a recent air/water study by Fan et al.
+Both systems show a similar trend where waters at or just below the surface have values of $\phi$ near 90\textdegree, and waters above the surface take on values of $\phi$ near 0\textdegree.
+This jump in the angular distribution of $\phi$ indicates that waters at or below the surface lie mostly flat in the plane of the interface, and as they move above the surface towards the gas phase, they reorient with one OH bond pointing towards the water bulk, and one pointing out of the surface into the gas.
+This behavior is more strongly pronounced in the neat-water system where most of the surface waters are not interacting with an adsorbed layer of \suldiox~molecules.
+
+%This result agrees somewhat with a recent air/water study by Fan et al.
 in which the surface orientations of several water models were analyzed.\cite{Fan2009} They simulated water using non-polarizable models, however, which alters the behavior of water at interfacial regions when compared to the polarizable POL3 model used in our work.
@@
 Although the plots show overall similarities for both the neat-water and saturated systems, the presence of a layer of adsorbed \suldiox~molecules alters the orientation of those waters furthest into the gas phase.
-For the saturated solution the resulting orientation of waters above 0\angs, shown in the saturated plot of $\cos(\theta)$ of Figure \ref{fig:water-orientation}, is with a bisector pointing further into the adsorbed \suldiox~gas layer, and both hydrogens pointing outward from the aqueous bulk.
-The effect is more pronounced further from the water phase, and above 5\angs~the $\cos(\theta)$ distribution is completely centered around $\cos(\theta)=+1$ (see Figure \ref{fig:water-angles}).
+For the saturated solution, the resulting orientation of waters above 0\angs, shown in the bottom set of plots of Figure \ref{fig:water-orientation}, is nearly isotropic in $\phi$, and with $\theta \leq$ 90\textdegree.
+This results from waters with bisectors pointing further into the adsorbed \suldiox~gas layer, and both hydrogens pointing outward from the aqueous bulk.
+The effect is more pronounced as the waters move further from the water surface, and above 4\angs~the $\theta$ distribution is mostly concentrated around $\theta=$ 0\textdegree (see Figure \ref{fig:water-angles}).
 
-The noise in the neat-water plots of Figure \ref{fig:water-orientation} above 5\angs~(manifested as disconnected points of intensity throughout the range of orientations) is a result of fewer waters venturing beyond those extents and thus less data far from the water surface location.
-This is one indication that waters near a layer of adsorbed \suldiox~venture further above the water surface location relative to the low \suldiox~concentration, where they can have more \suldiox~interaction with the adsorbed gas molecules.
-These results show that the interactions with neighboring \suldiox~molecules allow the waters above the water surface location to orient more perpendicularly to the interface.
+%The histograms show overall similarities for both systems in their shapes and intensities from approximately 0\angs~and below.
+The presence of a saturated layer of adsorbed \suldiox~molecules, however, alters the water orientation, but not necessarily the orientational depth of the interface.
+In both $\theta$ distributions, the orientations of waters in the bulk region are isotropic until 5\angs~below the surface location.
+At 5\angs~below the surface the water molecules begin to orient with their bisectors within the plane of the interface (perpendicular to the surface normal, $\cos(\theta)\approx 0$).
+The corresponding location in the plot of $\phi$ shows that those molecules are also mostly flat to the surface ($\cos(\phi)\approx 1$).
+Moving further out from the bulk and into the gas phase, the distributions show $\cos(\theta)$ increasing.
+Waters further out from the bulk have fewer bonding interactions, and orient with their hydrogens more towards the gas phase.
+The bisector pointing further into the gas phase leads to isotropy in the values of $\phi$.
+
+The angle distributions above 6\angs~in the neat-water plots of Figure \ref{fig:water-orientation} are mostly isotropic (manifested as uniform coloration throughout the range of orientations).
+Furthermore, there are few data points that make up the histograms, a result of fewer waters venturing beyond those extents.
+Conversely, waters near a layer of adsorbed \suldiox~venture further above the water surface location relative to the low \suldiox~concentration, where they can have interactions with the adsorbed \suldiox~gas molecules.
+The waters above the water surface location orient perpendicularly to the interface.
 This is consistent with our recent experimental VSFS studies which showed evidence for the reorienting behavior of water due to the \suldiox~interactions with the topmost surface waters.\cite{Ota2011}
 
-The distribution of $\cos(\phi)$ is more sharply defined (i.e.
+The distribution of $\phi$ is more sharply defined (i.e.
 less isotropic) for the neat-water system than for the saturated one.
-Waters on the neat surface lie flatter, whereas the presence of the \suldiox~allows a greater range of ``twist'' for those waters in the plane of the interface.
-The $\phi$ distributions quickly become isotropic above the surface location as the bisectors orient more perpendicularly, and below the surface as the bulk water loses any orientational preference.
+Waters on the neat surface lie flat or perpendicular to the surface if they are below or above the water surface location, respectively.
+The presence of the \suldiox~allows a greater range of ``twist'' for those waters in the plane of the interface.
+The $\phi$ distributions quickly become isotropic above the saturated water surface location, shown as a uniform coloration across $\phi$ for most values of $\theta$.
+
+%Peaks in the distributions of the neat-\wat~system are more clearly pronounced as their intensities are more concentrated and larger than the surrounding area of the profiles.
+This difference indicates that the transition from the preferred orientation at the water surface has a sharper distinction from the isotropic bulk than in the system with the saturated \suldiox~surface.
+It appears that the same orientation trend is present in both systems, but the presence of the \suldiox~at the water surface decreases the degree of water orientation at the interface.
 
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{h2oangles.png}
+		\includegraphics[scale=1.0]{theta-phi-h2o.png}
 		\caption{Molecular orientation histograms of \wat~throughout the surface equilibrated systems.
-The top surface is located at a distance of 0\angs~with negative distance values located in the bulk of the slab.
-Shown are the angle distributions for $\theta$ (left column) and $\phi$ (right column) in both the neat-\wat~system (top row) and the saturated system (bottom row).
-The distributions are normalized to account for the changing number of water molecules at different positions in the system.
-Regions of high intensity are dark red, and low intensity are dark blue.
-The scattered points of coloration to the far right of each plot indicates that few waters were located at those positions, and thus few data points were collected.}
+The depth (in \angs) of each histogram within the water interface is numbered in the upper-right of each respective plot.
+The water surface location is set at a position of 0\angs~with negative distance values located in the bulk of the slab, and positive distances towards the gas phase.
+Each histogram is a bi-variate angle distributions for $\theta$ (horizontal axes) and $\phi$ (vertical axes) in both the neat-\wat~system (top) and the saturated system (bottom).
+Regions of high intensity are dark red, and low intensity are dark blue.}
 		\label{fig:water-orientation}
@@
 Orientation distributions of the adsorbed \suldiox~molecules were created during the equilibrium simulations for both the neat-water and saturated systems.
-Figure \ref{fig:so2-orientation} shows the distributions of $\cos(\theta)$ and $\cos(\phi)$ (arranged similarly to the water orientation distributions plots in Figure \ref{fig:water-orientation}).
+Figure \ref{fig:so2-orientation} shows the 2D distributions of $\theta$ and $\phi$ (arranged similarly to the water orientation distributions plots in Figure \ref{fig:water-orientation}).
 The \suldiox~orientation data set for the neat-water system is much smaller as only a single \suldiox~molecule was simulated in the bulk.
-The resulting distribution plots are thus noisier than the corresponding saturated plots with more scattered points of high intensity, but effective comparisons can still be drawn.
+The resulting distribution plots are thus representative of the single surface active \suldiox~molecule.
+The neat-water \suldiox~molecule remains within a narrower region of the interface than the saturated system \suldiox, but effective comparisons can still be drawn.
+Note that the depth range of the plots in Figure \ref{fig:so2-orientation} is different for the neat-water and saturated systems reflecting the surface mobility of the \suldiox~molecules in the two systems.
 
-In the interfacial region within 0-5\angs~of the water surface location (the approximate distance from the surface that \suldiox~begins interacting with the outer-most waters) the angular distribution of the single \suldiox~(in the neat-water system) is concentrated primarily in $\cos(\theta)>0$.
-The peak of the distribution occurs at $\cos(\theta)=1$.
+In the interfacial region the angular distribution of the single \suldiox~(in the neat-water system) is concentrated primarily in $\theta <$ 90 \textdegree.
+The peak of the distribution occurs at $\theta=$ 0\textdegree.
 This indicates that the \suldiox~bisector points out of the water surface, with the sulfur atom pointing towards the aqueous bulk, and the two oxygens pointing into the gas phase.
-This same distribution occurs in the saturated system for positions below 5\angs.
-Beyond 5\angs~above the surface both distributions become isotropic.
-Promixity to the water highly orients the \suldiox~with the sulfur atom pointing in towards the water bulk.
-Moving further away from the water surface, and interacting less with \wat~molecules allows for greater orientational freedom as exhibited in the isotropy of the distributions above 5\angs.
+This same distribution occurs in the saturated system for depths below the surface location, $< 0$\angs.
+Beyond 4\angs~above the surface, both distributions become mostly isotropic, either because the \suldiox~does not venture into the gas in the neat-water system, or because of the nature of the adsorbed \suldiox~layer in the saturated system.
+Promixity to the water surface highly orients the \suldiox~bisector.
+% with the sulfur atom pointing in towards the water bulk.
+%Moving further away from the water surface, and interacting less with \wat~molecules allows for greater orientational freedom as exhibited in the isotropy of the distributions above 4\angs.
+%Because both the neat-water and saturated systems show rather similar distributions, it is possible that the concentration of \suldiox~does not strongly affect the orientation of the \suldiox, unlike the orientations of waters.
 
-The plots of $\phi$ are both isotropic, although the neat-water system plot is quite noisy from the small data set.
-Because the \suldiox~near the surface is oriented perpendicularly to the interface, the $\cos(\phi)$ orientation is expected to be isotropic.
-Further from the water surface where the bisector orientation becomes isotropic, the $\cos(\phi)$ distribution also becomes isotropic.
-For the \suldiox~orientation, the $\phi$ angle does not provide further information regarding the surface behavior.
+The distributions of $\phi$ are isotropic in both systems at all depths.
+Because the \suldiox~bisector near the surface is oriented perpendicularly to the interface, the isotropy in $\phi$ is expected.
+Further from the water surface where the bisector orientation becomes isotropic, the $\phi$ distribution remains isotropic.
+For the surface \suldiox~orientation, the $\phi$ angle does not provide further information regarding the surface behavior or orientational preference.
 
-The plots indicate that \suldiox~orients similarly at both the low and high concentration interfacial environments.
-The plots of $\phi$ for both concentrations exhibit isotropic distributions.
-However, the $\theta$ values follow similar trends at both concentrations indicating that the \suldiox~sulfur orients down towards the water phase, with both oxygens pointing away from the surface once the \suldiox~is within approximately 5\angs~of the surface.
+%The plots indicate that \suldiox~orients similarly at both the low and high concentration interfacial environments.
+The $\theta$ values follow similar trends at both concentrations indicating that the \suldiox~sulfur orients down towards the water phase, with both oxygens pointing away from the surface once the \suldiox~is within approximately 5\angs~of the surface.
 This behavior continues down to at least 5\angs~below the water surface location, notwithstanding any chemical reactions that may occur that are not simulated using the classical MD techniques utilized in this work.
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{so2-angles.png}
+		\includegraphics[scale=1.0]{theta-phi-so2.png}
 		\caption{Molecular orientation distributions for \suldiox~molecules adsorbed to the water slab surface.
-Distributions are shown for $\cos(\theta)$ (left column) and $\cos(\phi)$ (right column) for both the neat-water (top) and saturated (bottom) systems.
-For both systems the $\theta$ distributions show \suldiox~bound to the water surface with the sulfur pointing towards the water slab, and the oxygens pointing to the gas phase.
-In this configuration the $\phi$ distribution is isotropic because of the water slab's in-plane symmetry.}
+The angular distributions are shown for the neat-water (top) and saturated (bottom) systems, and are arranged similarly to the water orientation plots of Figure \ref{fig:water-orientation}.
+For both systems the distributions show \suldiox~bound to the water surface with the sulfur primarily pointing towards the water slab, and the oxygens pointing to the gas phase.
+In this configuration the $\phi$ distribution is isotropic because of the water slab's in-plane symmetry.
+Note that the range of surface depths in the low concentration plot is half that of the saturated system.
+This is due to the narrower position range of the single \suldiox~molecule.}
 		\label{fig:so2-orientation}
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{so2-angles-transit.png}
-		\caption{\suldiox~molecular orientation distributions during SMD transit simulations into an aqueous slab.
-Both the neat-water (top row) and saturated (bottom row) data sets were analyzed to determine the angles $\theta$ (left column), and $\phi$ (right column) of the transiting \suldiox.
-The distributions show angle cosines plotted against the distance of the \suldiox~to the location of the water surface.
-The data sets were averaged at each distance for the respective 50 simulations, and every distribution at each distance was normalized such that each histogram has a max population of 1 for purposes of comparison.}
+		\includegraphics[scale=1.0]{theta-phi-transit.png}
+		\caption{Molecular orientation distributions of an \suldiox~at different interfacial depths of an aqueous slab during SMD transit simulations.
+Both the neat-water (top) and saturated (bottom) data sets were analyzed to determine the angles $\theta$ (horizontal axes) and $\phi$ (vertical axes) of the transiting \suldiox.
+The distributions shown are those for the single transiting \suldiox~molecule averaged over the 50 SMD simulations.}
 		\label{fig:so2-transit-angles}
@@
 
-  Comparing in more detail the difference in these two systems upon \suldiox~approach, for the neat-water surface, a transition occurs at approximately 5\angs~from the surface water location.
-Below 5\angs~above the surface, \suldiox~has a preferred net orientation and is close enough to the water surface that it begins to interact with the topmost surface waters.
+  % theta differences
+  Comparing in more detail the difference in these two systems upon \suldiox~approach, for the neat-water surface, a transition occurs at approximately 4\angs~above the surface water location.
+Below 4\angs~above the surface, \suldiox~has a preferred net orientation and is close enough to the water surface that it begins to interact with the topmost surface waters.
 In the saturated system, the same trend occurs, however the onset of the perpendicular orientation begins at approximately 8\angs~above the surface.
@@
 It is remarkable that the orientational trend appears so strongly in the $\theta$ plots even with so little data as was collected from the single \suldiox~molecule of each simulation.
-From onset of orientation above the surface until 10\angs~below, the \suldiox~holds a preferred orientation.
+From onset of orientation above the surface until 10\angs~below (not shown), the \suldiox~holds a preferred orientation.
 
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[scale=1.0]{so2-surface.png}
-		\caption{(A) \suldiox~molecules above and below both neat-water and saturated water surface location orient with the bisectors perpendicular to the interface.
+%\begin{figure}[h!]
+	%\begin{center}
+		%\includegraphics[scale=1.0]{images/angle-cartoons/so2-surface.png}
+		%\caption{(A) \suldiox~molecules above and below both neat-water and saturated water surface location orient with the bisectors perpendicular to the interface.
 The \suldiox~sulfur points down towards the aqueous phase.
 (B) In the neat-water system, a transiting \suldiox~molecule located below the water surface briefly reorients with one S-O bond more towards the interface above, and the other bond pointing further down into the water bulk.}
-		\label{fig:so2-surface-angles}
-	\end{center}
-\end{figure}
+		%\label{fig:so2-surface-angles}
+	%\end{center}
+%\end{figure}
+
+  % phi differences
+	%Indeed, the $\phi$ orientations appear mostly isotropic without any clear trends for preferred orientation in that angle.
+This appears as a mostly uniform color across the entire plot, with minor variations that are disjointed throughout the distance range.
 
@@
 The presence of the adsorbed \suldiox~layer apparently decreases the reorienting behavior likely because of the disrupting effect the higher \suldiox~concentration has on the water interactions in the interfacial region.
+%Below this 5\angs~region under the surface, the \suldiox~resumes an isotropic orientation as it becomes fully solvated by bulk waters.
+
+  %Upon approaching the saturated surface at approx.
+5\angs, the \suldiox~$\phi$ orientation distribution begins to cluster near $\cos(\phi)=1$, with the $\theta$ distribution remaining more isotropic, but with a trend of $\cos(\theta)>0$.
+At 5\angs the \suldiox~begins interacting with the other \suldiox~molecules that sit atop the water, and the resulting orientation distribution suggests that contact with the \suldiox~layer causes the adsorbing molecule to lie more flat to the surface.
+Within this same region from 0-5\angs above the surface, the $\phi$ distribution of the neat-\wat~system remains isotropic.
+The intensity of the distribution (darker blue) also indicates that the \suldiox~spends little time in this region, as it is quickly adsorbed further into the water bulk.
+
+	%Just below the water surface as the distance moves into negative values, both systems show clear orientational preferences.
+In both systems the bisector orientation clusters such that $\cos(\theta)<0.5$, with most of the distribution intensity around $\cos(\theta)=1$.
+This suggests that \suldiox~in a water surface orients with the sulfur pointing into the water bulk, and the oxygens pointing towards the surface water molecules.
+
+	%Differences ocurr between the system orientation profiles once the \suldiox~has moved further than 10\angs below the surface.
+While both $\theta$ and $\phi$ hold the same trend from just under the surface through to the bulk of the slab in the saturated system, the orientation profile becomes much more isotropic past 10\angs under the surface of the neat-\wat~system.
+The orientation effect is shallower under the neat-\wat~surface.
+However, the orientation distribution peaks most intensely, and more tightly clustered with the \suldiox~bisector aligning with the surface normal, just underneath the neat-\wat~surface than the saturated one.
+The distributions thus show an orienting force extending deeper into the surface of the saturated system than the neat-\wat~system.
+In terms of how deeply the \suldiox~molecule is oriented, the interfacial region of the saturated system is larger and extends further into the aqueous bulk.
 
@@
 
-\section{Acknowledgements}
+% In such cases, surface complexes are often invoked with little direct evidence of the formation of such complexes.
 
-The authors wish to thank Dr.
-Daniel Steck of the University of Oregon Department of Physics for the generous use of computing resources, the National Science Foundation (Grant CHE-1051215), and the US Department of Education Graduate Assistance in Areas of National Need (GAANN) program (Grant P200A070436) for support of this research.
+%Having now examined the transition of \suldiox~from the gas to adsorbed aqueous phases, we are better able to research the many questions remaining of adsorption and interfacial chemistry of gas molecules.
+What species form during the adsorption transition, and how do molecular properties affect the process?
+Can we form a theory that will more fully explain the transit of many small molecules of environmental and industrial interest?
+This study is one of several characterizing \suldiox~adsorption and behavior on aqueous surfaces.
+We plan to report further results of ongoing computational simulation and experimental studies regarding temperature and chemical constituent effects on atmospherically relevant water surfaces.
 
-\bibliography{citations}
+\bibliography{bib/SulfurDioxideText}
 
